### 2018-02-11 writeup/Outline/main.tex
start running models

\documentclass[11pt]{article}

\input{tex/header/macros.tex}
\input{tex/header/packages.tex}

\title{Climate Change Adaptation: Uncertainty Associated with Sequential Finite Period Decisions under Nonstationarity}
\author[1,2]{James Doss-Gollin}
\author[1,2]{Upmanu Lall}
\author[1,2]{David Farnham}
\affil[1]{Columbia Water Center, Columbia University}
\affil[2]{Department of Earth and Environmental Engineering, Columbia University}

\input{tex/header/biblatex.tex}
\addbibresource{../library.bib}

% -----------------------------------------------------------------------------
% AUTHOR AND TITLE AND TITLEPAGE
% -----------------------------------------------------------------------------

\begin{document}
\maketitle
%\RaggedRight

\begin{abstract}
  Hydroclimatic systems exhibit organized low-frequency and regime-like variability at multiple time scales, causing the risk associated with climate extremes such as floods and droughts to vary at multiple timescales.
  Despite broad recognition of this nonstationarity, there has been little theoretical development of ideas for the design and operation of infrastructure considering the regime structure of such changes and their potential predictability.
  This theory is well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies with $M$.
  In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show that \texttt{RESULT}.
  This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
\end{abstract}

\section{Introduction}

Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep{Milly2008,Merz2014,Hurst1951,Sveinsson2005,Hodgkins2017,Matalas2012}.
This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
Yet the same memory and low-frequency variability that complicate projecting flood sequences far into the future may also impart some predictability at short timescales \citep{Jain2001}.
The existence of this information gap suggests that approaches for adapting to hydroclimate risks, such as floods, which have short project operation periods, may be preferential in some cases to large and permanent projects with long project operations.

Classical methods for design of infrastructure first specify a return period $T$ (\ie 100 years), and then design a project that protects against this $T$-year event.
Much of the flood frequency analysis (FFA) literature has tended to focus on the details of flood frequency estimation and on trend detection associated with climate change and land use cover \citep{Merz2008,Kidson2016,Merz2014}.
The mathematical problem of interest then becomes estimation of the probability distribution $f(p_T)$, where $p_T$ is binary event that the threshold $Q_T$ is exceeded:
\begin{equation*}
  p_T(t) \equiv \probability \qty[ Q(t) \geq Q_T] \qqtext{for} t = \qty[t_0 + 1, \ldots, t_0 + M]
\end{equation*}
and where $t_0$ is the first year forecast and $M$ is the project operation period.
Although in a stationary setting an unbiased estimator of $Pr \qty[ Q(t) \geq Q_T]$ converges asymtotically to $1 - \frac{1}{T}$, the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of $Q_T$ is used for design \citep{Stedinger1997}\footnote{Vogel198x citation}.

Alternatively, risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design  may itself be a decision variable.
\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of $p$ (equivalent to different values of $T$), and also for different sample sizes for estimation $N$ and different project operation periods $M$.
The duration $M$ may reflect the physical life of a proposed infrastructure element, or its economic life.
In the context of financial instruments which may be used for risk mitigation, $M$ specifies the length of the contract and $Q_T$ the threshold at which a payout is triggered.
In such a setting, the mathematical problem is to estimate the future distribution of $Q(t)$ rather than just the future distribution of the binary outcome $p_T$.

Under classical assumptions of stationarity, all estimates are assumed independent of time.
Yet even under stationarity, \citet{Jain2001} showed that low-frequency variability can lead to a high probability of surprise -- particularly when the length of the record used for estimation $N$ is short.
This finding is particularly important to keep in mind given that low-frequency change has been found to be large as compared to detected trends in the observed record \citep{Hodgkins2017}.

Of course, the hydroclimate systems are not stationary, due to climate, river modification, and land use change \citep{Milly2008,Merz2014}.
However, predicting future hydrologic behavior in a particular location is inherently difficult given limited data and complex, highly nonlinear physics.
Physical scaling arguments for an intensified hydrologic cycle under climate change \citep[see][]{Muller2011,OGorman2015} do not in general hold over land \citep{Byrne2015,Shaw2016}.
Attempts to use a ``model chain'' approach emcompassing:
\begin{enumerate*}[label=(\roman*)]
  \item emission scenario;
  \item general circulation model (GCM);
  \item downscaling;
  \item hydrological catchment model; and
  \item flood frequency analysis,
\end{enumerate*}
typically with bias correction applied at several steps, can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
In particular, the downscaling and bias correction methods typically applied assume stationary relationships (\ie between GCM rainfall and observed rainfall) which have no physical basis, particularly in a changing climate.
Recent work has also sought to shorten the above model chain by modeling $Q(X(t))$, where $X$ is a set of climate state variables from a GCM run, and where the relationship between $X$ and $Q$ can more reasonably be assumed stationary \citep{Hall2014,Delgado2014,Silva2016}.
Alternatively, purely statistical approaches have extended classical flood frequency analysis by incorporating a time trend in the parameters of statistical distributions \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}.
In this case, however, the analyst can choose not only the distribution but also  the time parameterization, if any, for each parameter, leading to many researcher degrees of freedom \citep[``forking paths'' or ``multiple comparisons'';][]{Gelman2013} and exacerbating the problems caused by lack of theory for model choice even under nonstationarity \citep{Kidson2016}.

The wide variety of methods, each with its own advantages and disadvantages, for estimating the future distribution of flooding $p \qty(Q(t))$ illustrates the theoretical need for approaches to incorporate uncertain estimates of future floods into a decision framework.
As we have argued, however, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times.
There is thus \emph{a critical theoretical need for decision frameworks which can compare projects of different operation period} using an arbitrary choice of model(s) of future flood behavior.
Sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}\footnote{need some citations for sequential decisions}.

In this context, we consider financial risk mitigation instruments in which the price is a function of the amount of coverage $C$, the expected probability that the insurance is triggered $\expectation \qty(p_T) \equiv \frac{1}{T}$, and the variance (or more generally full uncertainty distribution) associated with this estimate $\variance \qty(p_T)$.
Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure $p_T$ for the next $M$ years, as well as the uncertainty and bias in this estimate.
However, to properly analyse it we need detailed information on complex cost and loss structures; we thus focus on the illustrative example of financial instruemnts for clarity.

In the following sections we illustrate some very basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary environment.
This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
We proceed as follows $\ldots$.

\section{Methods} \label{sec:methods}

We use a highly idealized model to generate sequences of annual-maximum floods using a known generating mechanism, then fit these sequences in a Bayesian framework to three flood models (stationary, nonstationary, and hidden markov model) to evalaluate their performance under different scenarios (defined by observational record length $N$ and project planning period $M$) and criteria.

\subsection{Synthetic Flood Sequence Generation}

Extensive discussion in the literature has considered the appropriate choice of distribution for annual-maximum flood sequences.
Although three-paremeter models are better able to represent extreme values than two parameter models \citep{IACWD1982,Vogel1996}, we assume that flood sequences are \emph{conditionally} lognormal, given a mean and standard deviation which vary in time:
\begin{equation} \label{eq:lognormal}
  \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
\end{equation}
where $\normal$ denotes a normal distribution and $\mu(t), \sigma(t)$ are the location and scale parameters, respectively.

In large river basins, extreme river flooding typically requires the large-scale transport and convergence of moisture,\footnote{Is this worth adding in} which often requires specific climate mechanisms.
For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent large-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
This is consistent with findings that have linked flood risk around the world with well-known climate mechanisms such as El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Ward2014,Emerton2017}.
To capture this influence, some recent studies have incorporated a climate index or set of climate indices $X$ as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
For a more comprehensive review, we refer the reader to \citet{Hall2014}.\footnote{Farnham et al}

ENSO \citep[see][for a comprehensive review]{Sarachik2010} varies in a quasi-oscillatory manner with variability on the order of \SIrange{4}{6}{year} modulated by low-frequency behavior.
Though the time scales are specific to ENSO, this pattern of quasi-oscillatory behavior modulated by low-frequency behavior is typical of many other climate phenomena.
A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index \footnote{We greatfully acknowledge the contribution of Nandini Ramesh} with stationary forcing.
To create an annual time series, we average the May-July months to create a time series $X(t)$ where $t$ is taken in annual steps $1, \ldots, \num{100000}$.

Though statistical models relating climate indices and streamflow maxima typically require strong assumptions (\ie a stationary relationship between predictor and predictand; see \texttt{Farnham et al}), this framework is the starting point of our synthetic flood generation approach.
We partition $\mu(t)$ into a dynamical component, driven by the variability of the climate predictor $X(t)$, and a trend component, represented by a linear increase in time.
This trend component may represent changes in land use, global temperature change, or river modification \citep[see][]{Merz2014}.
Though this is a highly simplified model neglecting many important mechanisms, it enables us to generate many synthetic sequences of $\qty[M + N]$ years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
Because the computational cost is small, it is easy to generate sets of sequences following the procedure
\begin{enumerate}
  \item Choose the parameters $M,N,\alpha,\beta,\gamma$
  \item Select the ``zero'' time of analysis $t_0 \in \qty[N, \ldots, \num{100000} - M]$ so that the time used is $t_{\text{use}} = t_0 - N + 1, \ldots, t_0 - M$ (note that this indexing treats $t_0$ as part of the $N$-year observed record)
  \item Subset $x(t) = X(t \in t_{\text{use}})$
  \item Let $\mu(t) = \beta x(t) + \gamma \qty(t - t_0)$
  \item Assume a constant coefficient of variation so that $\sigma(t) = \alpha \mu(t)$
  \item Draw $\log Q(t) \sim \normal \qty(\mu(t), \sigma(t))$
\end{enumerate}

In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of $M,N,\alpha,\beta,\gamma$ and discuss the processes which they can reasonably represent.

\subsection{Distributional Model \label{sec:estimation}}

Because we do not, in general, know the true distribution of annual-maximum floods, we compare a stationary and non-stationary parameterizations of the generalized extreme value (GEV) statistical distribution (eq.~\ref{eq:GEV}) for estimating the distribution of $Q(t)$ for the future $M$-year period.
\begin{equation}
  \label{eq:GEV}
  f(x \big| \mu, \sigma, \xi) = \frac{1}{\sigma}
    \begin{cases}
      \qty[1 + \xi \frac{x - \mu}{\sigma}] ^ {(-\flatfrac{1}{\xi})-1} \exp \qty[-\qty(1+\xi \frac{x - \mu}{\sigma})^{\flatfrac{-1}{\xi}}], & \xi \neq 0 \\
      \exp \qty[-\frac{x-\mu}{\sigma}] \exp \qty[- \exp \qty(- \frac{x-\mu}{\sigma})], & \xi = 0
    \end{cases}
\end{equation}
Where $\mu$ is the location parameter, $\sigma$ is the shape parameter, and $\xi$ is the scale parameter of the GEV distribution.
For each simulation, the first $N$ years $\qty(t=[t_0-N+1, \ldots, t_0])$ are treated as observations.
Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty.
Computation is carried out in the \texttt{python} environment, with simulation using the \texttt{numpy} package \citep{vanderWalt2011} and fitting using the No-U-Turn Sampler (NUTS) \citep{Hoffman2014} as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
Weak (uninformative) priors are chosen based on recommendations from \citet{Martins2000}; the exact parameterizations are described in following sections which describe specific experiments.

\subsection{Hidden Markov Model \label{sec:HMM}}

A Hidden Markov Model (HMM) is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved (\ie hidden) states \citep{Rabiner1986}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (\ie streamflow) depends only on the observed state.
As such, HMMs are a structured prediction method which extend general mixture models to sequences of data, where position in the sequence is relevant.
Because this structure allows the hidden state to represent hydroclimatological ``regimes'' \citep{Reinhold1982,Michelangeli1995,Merz2014}, HMMs have have been used extensively to model hydrological variables such as streamflow \citep{Bracken2016}.
HMMs have also been used to model the evolution of climate states such as ENSO.\footnote{Julian's work}

To supplement the distributional models described in \cref{sec:estimation}, we also fit observed streamflow sequences using a HMM as implemented in the python \texttt{pomegranate} package \citep{Schreiber2016}.
The model is fit to the $N$ observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
To choose the number of hidden states, a log probability score is used.
Future floods are estimated by simulating future states from the resulting matrix and then drawing floods log-normally using the estimated parameters.

\begin{table}[b]
  \begin{center}
    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
      \toprule
        Model Name & Description & Relevant Section \\
      \midrule
        stationary & $Q \sim \GEV \qty(\mu, \sigma, \xi)$ & \cref{sec:estimation} \\
        location trend & $Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)$ & \cref{sec:estimation} \\
        %full trend & $Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma_0 + \beta_\sigma \qty(t-t_0), \xi_0 + \beta_\xi \qty(t-t_0))$ \\
        HMM & log-normal Hidden Markov Model & \cref{sec:HMM} \\
      \bottomrule
    \end{tabular}
  \end{center}
  \caption{Summary of models used for fitting \label{tab:model-fitting}}
\end{table}

\subsection{Flood Estimation and Uncertainty}

Finally, we consider a contract in which an insurer agrees to pay the entire coverage cost $C$ every time that the trigger threshold $Q_T$ is exceeded.
Although $Q_T$ is in general a decision variable, we treat it as fixed.
Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that $Q(t) \geq Q_T$ for $t \in \vb{t}$, which is equivalent to $p_T(t)$.

For a particular year, the fair price for the insurance premium is $\expectation \qty(p_T) C$.
However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger $Q_T$ is specified in the contract and thus not uncertain).
This risk premium is proportional to the variance of $p_T$, or more generally to its uncertainty distribution.
The insurance premium for a particular year is thus
\begin{equation} \label{eq:premium}
  Z(t) = C \qty[\expectation \qty(p_T(t)) + R \variance \qty(p_T(t))]
\end{equation}
where $R \geq 0$ is a coefficient describing the cost of tue uncertainty in the estimate of $p_T(t)$.

In this paper we consider the price of a contract of $M$ years to be simply the summed price of the contract for each individual year.
We assume that the discount rate is equal to inflation (\ie the inflation-adjusted discount rate is zero) for simplicity.
A more comprehensive pricing mechanism for the financial instrument could incorporate more or different information about the uncertainty in the estimate, and should consider the price of the $M$ years as a portfolio (\ie should consider correlation between the different years).
Such an approach, however, is not necessary for this conceptual study.

\section{Are Generated Sequences Realistic? \label{sec:sequence-realistic}}

In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc and compare to some that have been identified in the literature.

\section{Stationary, Static Risk}

Consider static risk and a stationary process, and partiuclarly implications of $M$ and $N$.
Much of this is already laid out in \citet{Lall1987}.

\section{Stationary, Static Risk}

Consider static risk and a non-stationary process.
Estimate using drift or nonstationary models using full $N$ of the most recent $N$ (\ie what Vogel is proposing).
Illustrate via simulation for different signal to noise ratio of nonstationary terms and $M$ and $N$.

\section{Dynamic Risk}

Consider dynamic risk, \ie updating and sequential decisions, a nonstationary generative process, and $M$ and $N$ considerations

\section{Discussion}

Hopefully something to discuss?

\section{Summary}

Some interesting conclusions

% -----------------------------------------------------------------------------
% END HERE
% -----------------------------------------------------------------------------
\clearpage
\printbibliography

\end{document}

### 2018-02-11 writeup/Outline/misc.tex
start running models

\documentclass[11pt]{article}

\input{tex/header/macros.tex}
\input{tex/header/packages.tex}

\title{Climate Change Adaptation: Uncertainty Associated with Sequential Finite Period Decisions under Nonstationarity}
\author[1,2]{James Doss-Gollin}
\author[1,2]{Upmanu Lall}
\author[1,2]{David Farnham}
\affil[1]{Columbia Water Center, Columbia University}
\affil[2]{Department of Earth and Environmental Engineering, Columbia University}

\input{tex/header/biblatex.tex}
\addbibresource{../library.bib}

% -----------------------------------------------------------------------------
% AUTHOR AND TITLE AND TITLEPAGE
% -----------------------------------------------------------------------------

\begin{document}
\RaggedRight
\section{Manu's Ideas}

Let us tie in financial instruments as well as structural design -- financial instruments are probably better suited for our argument and also this is an active area whereas infrastructure stories are kind of not moving  -- but should cover both. Most importantly let us put the climate adaptation or climate risk mitigation spin on it since that is the most pertinent case -- so in the intro we should develop that idea

I think we need to have the focus on illustrating that for sequential decisions on offers or purchases of insurance like instruments or on the construction of infrastructure one needs to consider the risk as a function of estimation uncertainty of the quantile ($N$ for stationary, and $N-p$ for nonstationary -- where $N-p$ is the effective degrees of freedom) as well as the uncertainty associated with the time period of length $M$.

Where, the process is stationary and sequential decisions are considered, we explore whether the choice of the design quantile or the trigger for insurance like instruments depends on $N$ and $M$ in addition to the cost and loss parameters.

% THIS IS GOOD FOR DISCUSSION
For the nonstationary case, the projection of the risk on the period of length $M$ based on the estimation using a period of length $M$ also depends on the form of the underlying nonstationarity and the degree to which the ``prediction'' is likely to be successful.
This is a more interesting and realistic case, since directional information as to the risk to be faced in the M period may depend on the duration $M$ and the uncertainty may also correspondingly change.
Thus, exploring how the optimal quantile level or trigger for design changes under such conditions is of interest.
Here, we consider the identification and assessment of these uncertainties for various combinations of $M$, $N$, quantile or trigger level, and form of nonstationarity.
For simplicity, we consider a monotonic trend in the location and scale prameters, as well as a periodic trend in the location and scale parameters of the underlying climate extreme process, and consider varying signal to noise ratios for these parameters.
Models that are properly specified, \ie, period to estimate periodic trends, and linear for linear trends are considered.
Since our intention is to help develop intuition as to how these cases may influence decision making, we consider only these idealized cases, and do not consider model mis-specification for now.

% CONSIDER THIS FOR METHODS
The fair price for the premium is $p \times \kappa$, where $\kappa$ is the coverage and $p = p(X > X^*)$.
So if $p=0.01$ and coverage = \SI{1e9}{USD} and $M=\SI{1}{year}$ then this would be $0.01 \times 1e9$.
If M were 10 years and you got paid per event then one would need to compute the payoff probability for 1 or more events  (complement of 0 events) over that time corresponding to annual $p=0.01$.
Or to simplify we could make the example to be a payoff on the 1st event and no further events (though if we want to illustrate time clustering with time changing probabilities, the 1 or more game is attractive)
Now we can derive the fair premium for the $M=m$ game
However, there is a risk premium associated with the policy that prices the uncertainty in $p$ -- since $x_T$ the trigger is specified in the contract and is not uncertain. This risk premium is proportional to the $V(p)$ or more generally related to the uncertainty distribution of $p$.

Now we come to the $N$ side -- and the $V(p \big| N, M)$ will depend on the underlying model for the process, the estimation scheme and $M$ and $N$, so under the assumption of these things we can derive the risk premium and help illustrate how it changes by condition and hence identify a tradeoff point for dynamic vs static risk given these parameters and $M/P$ position during decision making.

% FOR INTRODUCTION
The insurance example is a way to illustrate the general principle.
If we are designing infrastructure we are faced with the same uncertainty on $p$.
However, to properly analyse it we need detailed information on losses that we may incur and costs.
These are themselves highly uncertain and would need to be assessed but the underlying principle that the risk due to the uncertainty and bias in $p$ has to be assessed is not any different and the parametric insurance example is then used to get the concept across -- this is how one should present it.
Of course as discussed separately, it is not just $N$ and $M$ but the static and dynamic risk considerations which imply predictive uncertainty and bias as well and this is why we demonstrate these ideas via simulation

% FOR METHODS / RESULTS SETUP
building blocks
\begin{enumerate}
  \item static risk, stationary process implications of $M$ and $N$ = basically my old paper provides $V$
  \item static risk, non-stationary process (drift and periodic terms, estimation using full $N$ or partial, most recent $N$ -- as some people recommend and Vogel argues recent $N$ is better in his forthcoming paper) illustrate via simulation for different signal to noise ratio of nonstationary terms and $M$ and $N$
  \item dynamic risk, \ie updating and sequential decisions, nonstationary, $M$ and $N$ considerations -- ok this is the last case
\end{enumerate}
So for AGU if you can even wrap up 1 and 2 and then by end of year paper with all 3 I think we are good for the 1st paper.
We need enough to motivate the real options paper which is then introduced as the decision making framework considering multiple options.

% FOR DISCUSSION -- THESE ARE IMPLICATIONS
Let us frame this more precisely in the context of climate change adaptation by giving examples
1) consider a place where under IPCC scenarios there is a wide variation in the projected probability distributions when all scenarios are considered and this uncertainty increases as one looks out into the future.
2) We have regime like behavior that varies stoachastically -- could be quasi-periodic or modeled using HMM etc and in this case, essentially we have a structured long-memory process with shifts in all parameters with time, but the underlying dynamics may be stationary.
Now from a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection. How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the f(Qp(t)) given a methodology used for prediction -- e..g the model chain using gcms and such and models fitted to a historical data length m.
WHat is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions. How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.

I think we need to have the focus on illustrating that for sequential decisions on offers or purchases of insurance like instruments or on the construction of infrastructure one needs to consider the risk as a function of estimation uncertainty of the quantile (N for stationary, and N-p for nonstationary -- where N-p is the effective degrees of freedom) as well as the uncertainty associated with the time period of length M. Where, the process is stationary and sequential decisions are considered, we explore whether the choice of the design quantile or the trigger for insurance like instruments depends on N and M in addition to the cost and loss parameters. For the nonstationary case, the projection of the risk on the period of length M based on the estimation using a period of length M also depends on the form of the underlying nonstationarity and the degree to which the "prediction" is likely to be successful. This is a more interesting and realistic case, since directional information as to the risk to be faced in the M period may depend on the duration M and the uncertainty may also correspondingly change. Thus, exploring how the optimal quantile level or trigger for design changes under such conditions is of interest. Here, we consider the identification and assessment of these uncertainties for various combinations of M, N, quantile or trigger level, and form of nonstationarity. For simplicity, we consider a monotonic trend in the location and scale prameters, as well as a periodic trend in the location and scale parameters of the underlying climate extreme process, and consider varying signal to noise ratios for these parameters. Models that are properly specified, i.e., period to estimate periodic trends, and linear for linear trends are considered. SInce our intention is to help develop intuition as to how these cases may influence decision making, we consider only these idealized cases, and do not consider model mis-specification for now.

\section{Extra Lit Review}

The first element of the model is the distribution function.
Although Bulletin 17-B \citep{IACWD1982} mandates the use of the log-Pearson type III (LP3) model of annual-maximum floods in the United States, annual-maximum floods are also modeled three-parameter distributions such as the generalized extreme value distribution and two-parameter models such as the lognormal (LN2) \citep{Vogel1996}.
Alternatively, peaks-over-threshold floods are typically modeled with the Generalized Pareto distribution \citep{Silva2016,Jain2001}.
Many other approaches have been utilized in the literature, as the choice of model is generally made for practical rather than theoretical reasons \citep{Kidson2016}.

The second element is an approach for estimating the parameters of the model.
Popular approaches include the delta, bootstrap, and profile likelihood approahces \citep{Obeysekera2014}.
More recently Bayesian approaches have gained popularity due to their ability to fully quantify uncertainty.
Another advantage of Bayesian approaches is the straightforward approach to integrating data from multiple data sources \citep[\ie][]{Sun2014,Lima2016,Bracken2016,Steinschneider2015b}, filling a key need identified in previous literature reviews \citep{Merz2014,Merz2008}.

Finally, FFA requires a parameterization of the time evolution of the model parameters.
Under assumptions of stationarity used in classical flood frequency analysis, this quantity is assumed constant in time: $p(Q(t)) \equiv p(Q)$.
One approach to nonstationary FFA is to parameterize the distribution of $Q$ as a linear, logarithmic, or polynomial function of time itself \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}.
A problem with these time-varying parameters is that the different forms of the trends might be almost indistinguishable for the observational period, but lead to different future behavior at the end of the project operation period \citep{Rootzen2013,Serinaldi2015}.
Additionally, as these statistical approaches condition on more variables, the number of parameters estimated increases, leading to problems of high uncertainty and overfitting \citep{Serinaldi2015}.
Some studies address this problem by using null hypothesis significance testing (NHST) to detect nonstationarity and fitting time-varying parameters only if nonstationarity is rejected \citep[\ie][]{Obeysekera2014,Luke2017}.
This approach, however, translates poorly to decision-making settings \citep{Rosner2014,Vogel2013}.
This is in line with recent criticims in the statistics literature of the use of NHST in situations where the null hypothesis, such as that the flood time series is fully stationary, is implausible \citep{gelman2016problems,Gelman2014a,McShane2017}.
\footnote{Manu: introduce idea that nonstationarity in terms of all parameters of these models may exist and be monotinic -- important to link to mechanism -- or quasi-periodic. Use appropriate refs. Then introduce the literature htat has tried to make estimates in these settings.}

An alternative approach to purely statistical fits of $Q(t)$ is to numerically simulate the physical processes that lead to floods.
Typically this is done through a long model chain emcompassing:
\begin{enumerate*}[label=(\arabic*)]
  \item emission scenario
  \item general circulation model (GCM)
  \item downscaling
  \item hydrological catchment model; and
  \item flood frequency analysis.
\end{enumerate*}
Typically bias correction is applied at several steps of this chain, which can complicate interpretability of results.
While this approach allows for estimates of flooding at high spatial and temporal resolution, results are sensitive to the choice of a model at each step of this chain and uncertainty propagation is difficult to characterize \citep{Dankers2009,Ott2013,Dittes2017}.
Ideally, one could run large ensembles of models that capture the full distribution of uncertainty in each step of the model chain and use ane exhaustive set of physical parameterizations, but in practice the computational cost of such an approach is prohibitively expensive.
Further, there is a need for theory for combining model results which are not idependent and identically distributed, complicating the task of flood frequency analysis.
More recently, studies have attempted to shorten the model chain by modeling $Q(X(t))$, where $X$ is a set of climate state variables from a GCM run \citep{Delgado2014,Silva2016,Griffis2007,Villarini2009,Villarini2010,Hall2014}.
Though this approach does not resolve the difficulties of GCM bias and herding\footnote{there must be a better way to say this -- I mean it in the Nate Silver sense}, by ``shortening'' the model chain it allows for a more interpretable characterization of $p(Q(t))$ and reduces uncertainty associated with subsequent steps.

\begin{enumerate}
  \item High government indebtedness levels in many if not most emerging-market countries no longer allow public-debtd riven delivery as a scalable alternative to build urgently needed infrastructure \citep{WEF-instruments-2016}.
  \item In specific cases, investors' risk appetite and the risk level of infrastructure projects can be bridged through credit and project guarantees, insurance and other credit-enhancement schemes, also known as Risk Mitigation instruments \citep{WEF-instruments-2016}.
  \item The sustainable management of water resources requires a long-term perspective.
  Yet looking to the future reveals a host of major uncertainties, with respect to pressures from climate change, demographic change, land-use changes and other socio-economic drivers \citep{Hall2012}.
  \item Despite widespread interest in index insurance for industries such as agriculture \citep{Clarke2013} and navigation \citep{Meyer2016}, there has been little exploration of financial instruments for hydroclimate risks beyond a year or two.
  \item The insurability of risks depends on a number of factors, including:
  \begin{enumerate*}[label=(\arabic*)]
    \item mutuality (that a large pool of risk can be created);
    \item quantifiable loss;
    \item randomness of the insured event; and
    \item economic viability (that the premium is sufficient to cover losses and affordable to the policyholders).
  \end{enumerate*} \citep{Wolfrom2016}.
\end{enumerate}

% ------------------------------------------------------------------------------
\section{Physical Drivers of Floods and Flooding}
% ------------------------------------------------------------------------------

Lagrangian tracking studies \citep{Gimeno2010} show that extreme rainfall in many mid-latitude areas originates as moisture in a distant oceanic source.
Mid-latitude extreme rainfall is often associated with large-scale circulation features which connect the tropics with the extratropics, as shown in the US Midwest \citep{Dirmeyer2010}.
In the Ohio River Basin, my own work establishes that moisture from the Gulf of Mexico and Caribbean transported via frontal systems leads to {pir} in the Ohio River Basin, in all seasons [in prep.].

In 2014, floods occurred in the Balkans when three consecutive cyclones of the same class followed near-identical trajectories, aligned with a strong jet.
These cyclones were steered by an amplified and persistent wavenumber $k \approx 6$ circulation \citep{Stadtherr2016} associated with a circumglobal wavetrain \citep{Branstator2002}.
Other case studies \citep{Grams2014,Nakamura2012} and theoretical analyses \citep{Hoskins2015} support the hypothesis that blocking or quasi-stationary jet sinuosity can lead to {pir}, and planetary-scale analyses demonstrate links between monthly-mean jet stream activity, wave resonance, slow wave speeds,  and mid-latitude surface extremes \citep{Screen2014,Coumou2014}.

Mid-latitude waves and eddies result from baroclinic instabilities propagating from regions of high baroclinicity towards the subtopics and poles \citep{Shaw2016}.
The equator-to-pole temperature gradient and ocean-land temperature contrast modulate the jet by changing the thermal wind balance and by modifying baroclinicity.
These gradients are projected to decrease in the future \citep{Cohen2014}, but jet activity is also forced by tropical convection anomalies, including the El Ni\~{n}o Southern Oscillation and the Madden-Julian Oscillation \citep{Karamperidou2012}, which generate wavetrains that can reach the mid-latitudes \citep{Roundy2012}.
In fact, modeling studies suggest that Rossby waves are triggered more readily by low-latitude deep convection than high-latitude boundary layer processes \citep{Hoskins1981}.

% BEGIN: COPY VERBATIM FROM LULU'S THESIS
% THIS IS NOT OK TO LEAVE AS-IS BUT IS MERELY USED AS A STARTING POINT
% FOR SOME FURTHER READING
\begin{quotation}
	Nonstationarity of flood risk has emerged as an important issue ([Olsen et al., 1999; Jain and Lall, 2000, 2001; Renard et al., 2006; Villarini et al., 2009; Katz, 2010; Lima and Lall, 2010b; Massei and Fournier, 2012]) and progress in addressing this concern can only come from an improved understanding of the associated climate dynamics.
	Various climate change projections [Trenberth et al., 2003; Held and Soden, 2006; Allan and Soden, 2008] suggest an intensification of precipitation in the future, in terms of both frequency and magnitude.
	The intensity of extreme precipitation is projected to increase under global warming in many parts of the world, even in the regions where mean precipitation may decrease [e.g., Kharin and Zwiers, 2000, 2005; Semenov and Bengtsson, 2002; Voss et al., 2002; Wilby and Wigley, 2002; Wehner, 2004].

	However, these arguments are driven largely by considerations of the moisture holding capacity as a function of temperature, as indicated by the Clausius-Clapeyron equation [Muller et al., 2011; Romps, 2011].
	Outside the tropics the change in water holding capacity could well be below or above Clausius-Clapeyron (CC) scaling whereas in the tropics it has been shown to obey Clausius-Clapeyron (CC) scaling [Muller et al., 2011; Romps, 2011].
	We postulate that in the midlatitudes it is important to consider the attendant atmospheric circulation and moisture transport dynamics that lead to persistent extreme precipitation and subsequent flooding as evidenced in the recent major floods cited earlier, and identified as important in Nakamura et al. [2012]’s analysis of 21 Ohio River floods that exceed the 10 year return period and in Lavers et al.[2011a]’s demonstration of the association between ARs and 10 largest winter floods events since 1970 in Britain.
	An understanding of the dynamical mechanisms and statistics associated with the frequency and structure of such events can aid exploration of their representation in ocean-atmosphere circulation models used for weather prediction, seasonal climate forecasting and projections of climate change.

	Although the climate mechanisms governing precipitation vary by location, several researchers indicate that extreme precipitation events in the mid-latitudes are typically associated with anomalous atmospheric moisture from warmer tropical or subtropical oceanic areas.
	Bao et al. [2006] show that enhanced Integrated Water Vapor (IWV) bands, also known Atmospheric Rivers (ARs) [Ralph and Dettinger, 2011], are associated with direct poleward transport of tropical moisture along the IWV bands from the Tropics all the way to the extratropics.
	Zhu and Newell [1998] showed that for meridional transport at middle latitudes, ARs account for a substantial part of the moisture transport.
	There are four or five narrow ribbons across the mid-latitudes, covering less than 10\% of the Earth's circumference, where the majority of the mid-latitude moisture fluxes occurred in filamentary features.
	The AR concept indicated a direction to track the moisture from warmer oceanic source to the heavy precipitated regions.
	Schubert et al. [2011] noted that stationary Rossby waves account for a substantial fraction of summertime monthly mean surface temperature and precipitation variability over a number of regions of the Northern Hemisphere middle latitudes.
	Further Knippertz and Wernli [2010] and Nakamura et al. [2012] note that Tropical moisture exports (TMEs) to the Northern Hemispheric extratropics are an important feature of the general circulation of the atmosphere and link tropical moisture sources with extratropical precipitation and occasionally with explosive cyclogenesis.
	Lavers et al. [2011] presented evidence that winter flood events in the UK are connected to ARs, which transport moisture from the subtropical N. Atlantic Ocean to the mid-latitudes.
	The penetration of tropical moisture to the higher latitudes may have considerable impacts on extreme precipitation especially poleward of 30N [Knippertz and Wernli, 2010].
\end{quotation}
% END: COPY VERBATIM FROM LULU'S THESIS

% ------------------------------------------------------------------------------
\section{Estimating Future Floods}
% ------------------------------------------------------------------------------

\subsection{Classical Methods for Flood Adaptation}

Much of the existing decision literature in the context of climate and flood adaptation bifircates between ``classical'' (stationary) and new, nonstationary methods.
However, it is straightforward to join these two frameworks together with some careful definitions, which follow.

Let $Q(t)$ be the annual-maximum streamflow for which a record of $N$ years exists.
Further, define $M$ as the future project operation period, for a single decision.
This question assumes that an estimate of the probability distribution function $p(Q(t))$ exists.
Although it is widely recognized that nonstationary techniques seek to model $Q(t)$, implicitly the stationary estimates achieve the same goal.
Thus, while many framings for flood risk or decision-making apply, all require an estimate of $p \qty(Q(t))$ for $t = 1, \ldots, M$.

Next, consider the $T$-year event to be the event that is exceeded with probability $1/T$ in a particular year; thus $T=T(t)$ and the corresponding event is denoted $Q_{T(t)}$.
If we average over the $M$-year project period, we can define a $T$-year event which is constant over this $M$-year period.
Let this event be denoted by $Q^*$, and is the event which is exceeded \emph{on average} $\frac{M}{T}$ times in $M$ years.
Under nonstationary conditions, the probability of exceeding this event $Q^*$ will not be constant in time -- unlike $Q_{T(t)}$ which is defined such that its probability of exceedance is constant in time but which is itself time-varying.

One common approach to designing infrastructure of policy for flood management is to consider the risk that $Q^*$ is exceeded \emph{at least once} in the $M$-year project lifetime.
Although it is commonly denoted as risk, this is a word with specific \citep[see][for further discussion]{Merz2014}.
Instead, we define the reliability $R$ by \cref{eq:iid-nonstationary}.
\begin{equation}
	R = 1 - \prod_{t=1}^M \qty[1 - P(Q(t) < Q^*)] \label{eq:iid-nonstationary}
\end{equation}
In the classical (stationary) setting, where $P(Q_t \geq Q^*)$ is assumed constant in time, then $P(Q(t) < Q^*)$ is constant and  \cref{eq:iid-nonstationary} simplifies to \cref{eq:iid-stationary}.
\begin{equation}
	R = 1 - \qty[1 - P \qty(Q \geq Q^*)]^M = \qty[1 - T^{-1}]^M \label{eq:iid-stationary}
\end{equation}

One important feature of decision analysis is risk: what its components are and how they interact.
\begin{itemize}
  \item Today it is recognized that all risk elements (hazard, exposure, vulnerability) are dynamic through time, not only hazard \citep{IPCC2012,Jongman2012,Merz2014}
  \item Flood protection by dikes aimed at reducing the flood hazard can lead to increased development behind dikes, thus increasing exposure, the so-called levee effect \citep{Tobin1995}
\end{itemize}

Risk-based design
\begin{itemize}
  \item \citet{Rosner2014}: the goal of RBDM is to choose a level of infrastructure protection that minimizes the total expected cost by calculating net benefits: the expected cost of damages avoided, less the cost of infrastructure. While protection against the $T=100$ year flood is the most common design target under traditional analysis, a RBDM process may lead to a protection target either smaller or larger than the 100 year flood, depending on the probability and the consequences of the flood as well as the costs of the needed infrastructure
\end{itemize}

Year-to-year variability, which is often regime-like and cyclical, tends to dominate climate change signals in historical records \citep{Merz2014,Hodgkins2017} and even in many projections of future extreme floods \citep{Dittes2017}.
Given that the classical challenges of persistence and short-term memory remain complex and challenging \citep{Matalas2012}, incorporating climate change, land use change, and river modification into projections of future flood risk is a daunting challenge.

\subsection{Fully Empirical Methods for Future Flood Estimation}

There has been enormous debate over the best choice of distribution and best way to estimate parameters.
\citet{Kidson2016} Table 2 provides a comprehensive listing of probability density functions used for FFA; see also \citet{Renard2006} for a more detailed discussion of several of these.
Regarding some debate as to the most appropriate model:
\begin{itemize}
  \item \citet{Raff2009} describe that in the document Bulletin 17-B (1982), which provides key guidance across the united states as to flood control, the need for non-stationarity is mentioned in the context of future study.
  The methodology of Bulletin 17-B is to use a log-Pearson type III (LP3) flood model to describe annual-maximum floods, with a balance given between local (gauge) and regional estimates of the skew parameter.
  Significant work since this study has gone into the problem of developing improved methods for using data to estimate local flood risks under the nonstationary {iid} assumptions \citep[see][for citations]{Raff2009}.
  \item Three parameter distributions such as the generalized extreme value (GEV), LP3 and three parameter lognormal (LN3) distributions tend to be best but the LN2 was considered to be the best two-parameter approximation to annual maximum flood series \citep{Vogel1996}
  \item In general, it is difficult to know which is the ``right'' model for the data (though this may not be a helpful way to think about it). \citet{Adlouni2008} provides a review of methods, largely graphical, to discriminate between different distributions, with the aim of identifying the ``most correct'' one.
  \item \citet{Kidson2016} argues for use of power law rather than curve-fitting, though notes that ``Many studies of self-similar processes are multifractal, \ie, exhibit different PL scalinge xponents over different scale ranges.''
  \item ``The recent literature of FFA has been characterized by: (1) a proliferation of mathematical models, lacking theoretical hydrologic justification, but used to extrapolate the return periods of floods beyond the gauged record; (2) official mandating of particular models, which has resulted in (3) research focused on increasingly reductionist and statistically sophisticated procedures for parameter fitting to these models from the limited gauged data.'' \citep{Kidson2016}
  \item ``One consequence of the adoption of a single standard model is the ‘one size fits nobody’ scenario, where optimal fitting to a specific catchment is precluded by a (possibly unsuited) national model–this may result in mediocre accuracy on flood prediction in any given basin.'' \citep{Kidson2016}
  \item ``The proliferation of models is itself symptomatic of the weak theoretical basis in hydrology for the application of these FFA models'' \citep{Kidson2016}
  \item \citet{Rootzen2013} argue that the concept of a return period is insufficient and argue that one should specify both a period of time, the design life period, and a probability of failure.
\end{itemize}

One approach is to consider linear trends in the parameters of the distribution of choice
\begin{itemize}
  \item \citet{Obeysekera2014} derives properties of a {gev} distribution and associated uncertainties using three different methods (delta, bootstrap, and profile likelihood) for a 3-parameter {gev} with a linear time trend in the location parameter. Like other studies, this study uses linear trends to consider whether it is necessary to model nonstationarity in the location parameter.
  \item \citep{Vogel2011} proposes a two-parameter log-normal (LN2) model with a time-varying mean parameter $\mu = \alpha + \beta t$ and finds that ``This nonstationary model of the mean value of y, conditioned on time, t, offers a simple, practical and useful method for modeling the change in the mean of the distribution of flood magnitudes.''
  \item \citet{Serinaldi2015} presents a simple Gumbel model with linear trends in some parameters as [my words] a straw man for non-stationary models. The authors then make several observations and assertions, some of which are quite relevant:
  \begin{itemize}
    \item even for a simple model, the number of parameters which must be estimated increases.
    \item the assumptions of linear trend in $\mu$ and $\delta$ means the linear trend must hold for the entire design life
    \item Any model which uses covariates other than time must sufficiently control for anything that might change so that the relationship can be assumed stationary -- if the relationship is nonstationary, the model is theoretically intractable
    \item When predicting out of sample, poorly-constructed nonstationary estimates give bad results
    \item Even if we have evidence for nonstationarity, stationary models should be used as benchmark for every more complex competitor
  \end{itemize}
  \item \citet{Strupczewski2001} consider Normal, LN2, LN3, LP2, LP3, and Gumbel distributions. They consider trends using: linear mean, parabolic mean, linear SD, parabolic SD, linear coefficient of variation impacting mean and SD simultaneously, parabolic coefficient of variation impacting mean and SD simultaneously, and unrelated linear trend in mean and standard deviation. They then use  Akaike Information Criterion  to identify the optimum distribution and trend function, which enabled an identification of the optimum non-stationary FFM in a class of 56 competing models. The maximum  likelihood  method was used to estimate the parameters of the identieed model using annual peak discharge series.
\end{itemize}

Another alternative is to consider changes in the floods themselves (rather than the underlying parameters of a probability distribution function)
\begin{itemize}
  \item \citet{Villarini2009} use splines to model annual peak flows as a function of time; these get good results but the authors note that these tend to blow up in out-of-sample prediction
\end{itemize}
Another option is to look at probability distribution functions that use a mixture model to address regime behavior
\begin{itemize}
  \item \citet{Waylen1986} build a 3-family mixture model for floods in Peru where a Gumbel distribution was fit to data during El Ni\~no, Neutral, and La Ni\~na years.
  \item \citet{Sveinsson2005} use a stochastic switching model, to find that the distribution of the estimated return period can be highly skewed:
  \begin{quotation}
    The process of interest  $X_t$  is assumed to be a sum of two independent random variables $Y_t$ and $Z_t$, where the $Y_t$s are {iid} variables and the $Z_t$s are assumed to represent departure of each “stationary” state from the long term mean of the process. During each “stationary state” the Zts remain fixed at a value referred to as a noise level. In the SM-2 model two consecutive stationary states always have noise levels of opposite signs, while in the SM-1 model the noise levels are allowed to fluctuate in a random manner. In this paper only the positive geometric distribution is considered for modeling the length the process spends in each stationary state.
  \end{quotation}
  \item \citet{Griffis2007} also propose a mixture model for El Ni\~no flood risk with the LP3 model, considering (i) separate categories based on the state of ENSO or (ii) a parametric relationship between climate indicies and flood behavior
\end{itemize}

Many studies bifurcate their choice of model by first attempting to detect nonstationarity (typically linear trends), and then choose the model based on this detection; this is a form of the more general problem of model selection
\begin{itemize}
  \item Analysis of trends should be approached with great caution: \citet{Cohn2005} argues that the statistical significance of apparent trends, sometimes cited to bolster scientific and political argument, is highly uncertain because significance depends critically on the null hypothesis which in turn reflects jective notions about what one expects to see. From a practical standpoint, the article argues, it may be preferable to acknowledge that the concept of statistical significance is meaningless when discussing poorly understood systems.
  \item \citet{Cohn2005} shows that when a hydrological system exhibits long-term persistence, spurious trends are often detected that may provide a good explanation of observed data but will not provide useful prediction.
  \item \citet{Madsen2014a} review a very large number of studies of changes in rainfall and streamflow in Europe, largely within a trend-detection and trend-projection framework
  \item \citet{Vogel2013,Rosner2014} argue that trying to determining whether there is a linear trend before making a decision about whether to adapt is not relevan, particularly given recent pushback against NHST \citep[\ie][]{Gelman2016,Gelman2014a}. However, the statistical studies mentioned would probably consider their discussion of ``Type I'' and ``Type II'' errors as unhelpful (Gelman has blogged about this extensively)
  \item  The different forms of the trends might be almost indistinguishable for the observational period, but lead to rather different future behavior \citep{Rootzen2013}.
  \item \citet{Qi2017} use trend test to justify linear trend in {gev} parameters
\end{itemize}

Bayesian statistics offer an alternative way to combine information from different sources or with different non-stationarity models
\begin{itemize}
  \item \citet{Renard2006} describes a Bayesian framework for considering several probabilistic models (stationary, step-change and linear trend models) and four extreme values distributions (exponential, generalized Pareto, Gumbel and GEV). The framework can be expanded to other models for trend and extreme values distributions. They provide guidance on use of prior information and sampling, though it is worth noting that Bayesian computation has evolved stantially since the article was written.
  \item Following the call of \citet{Merz2008,Merz2008a}, \citep{Viglione2013} presents a Bayesian approach to estimating the parameters of a stationary {gev} model where jective expert analysis is used as a prior, while historical records of extreme floods beyond the gauge data and information from other catchments are also integrated.
\end{itemize}

\subsection{Numerical Estimation of Future Floods}

A traditional approach to modeling the future $Q(t)$ is to use a model chain approach
\begin{itemize}
  \item The typical approach for deriving future flood hazard scenarios under climate change is to implement model chains consisting of the following elements: emission scenario; general circulation model (GCM); downscaling, possibly including bias correction; hydrological catchment model; flood frequency analysis \citep{Merz2014}
  \item \citet{Dankers2009} use climate simulations from two RCMs to drive a LISFLOOD hydrological model to assess changes in flood frequency in Europe. They find that at the local or river basin scale, the choice for a particular regional and global climate model or emissions scenario results in large differences in the simulated response of extreme river discharge and sometimes even in a climate signal of opposite sign. Floods associated with different mechanisms are affected in different ways.
  \item \citet{Kay2008} examine future flood frequency at two catchments in England by coupling several GCMs and hydrological models. The authors find large uncertainty, which comes from uncertainty in future winter rainfall in England and from other sources. They also find that natural variability is very important and must be considered.
  \item\citet{Ott2013} use a GCM-RCM-HM model chain to investigate the climate change signal on medium catchments in Germany. The ensemble spread in the climate change signal is large and varies with catchment and season, and the results show that most of the uncertainty of the change signal arises from the natural variability in winter and from the RCMs in summer.
\end{itemize}

\subsection{Alternative Methods of Future Flood Estimation}

One way to model $Q(t)$ is to consider the mechanisms that lead to extreme flooding and to model $Q(t)$ conditional on them (model them as $X$ and consider $p(Q \big| X(t))$)
\begin{itemize}
  \item ``We see a great potential in such low-dimensionality models for understanding flood changes because they force the modeller to identify the dominant processes and they offer the possibility for establishing direct causality links for observed and projected flood changes'' \citep{Hall2014}
  \item \citet{Delgado2014} model Mekong River floods conditional on the Western Pacific monsoon using a shortened model chain: ``emission scenario – global climate model – non-stationary flood frequency model''
  \item \citet{Silva2016} develop a peaks-over-threshold model where the parameters of the arrival time and conditional distribution depend on ENSO. An excellent review of POT literature is also proposed. N.B.: their use of polynomials looks highly suspect (\ie fig. 7), causing them to draw some physical conclusions that at best aren't strongly supported by the data. However, still useful as the literature review and derivations are good.
  \item \citet{Sun2014} use a Bayesian approach to build a piecewise linear model for rainfall conditional on the SOI, and incorporate regional inforamtion pooling using a Gaussian Copula
  \item \citep{Griffis2007} propose using some covariates, such as ENSO and PDO, to represent possible future climate states and to regress the parameters of the LP3 disgribution on those covariates.
  \item \citet{Villarini2009} use GAMLSS to model time trends in the annual maximum daily rainfall data and annual flood peak observations which are found to exhibit a striking increase in flood magnitudes. However, the GAMLLS approach works well for curve-fitting but blows up when predicting future floods
  \item \citet{Villarini2010} also use the GAMLSS and assume a parametric distribution for the response variable $Y$ (seasonal rainfall, minimum and maximum temperatures), and model the parameters of the distribution as functions of time $t$ (the explanatory variable) using cubic spline smoothing functions.
\end{itemize}

% -----------------------------------------------------------------------------
% END HERE
% -----------------------------------------------------------------------------
\clearpage
\printbibliography

\end{document}

### 2018-02-11 writeup/Outline/tex/sections/05-Results.tex
start running models

\section{Synthetic Flood Frequency Estimation}

\begin{enumerate}
  \item Points to hit in this section:
  \begin{enumerate}
    \item How do different estimation models perform for different $M,N$?
    \item What is the inherent uncertainty (\ie using correct model) in estimation for different $M,N$?
  \end{enumerate}
  \item First experiment: visualizing distributions
  \begin{enumerate}
    \item Points to hit with this experiment:
    \begin{enumerate}
      \item Bad estimation models lead to biased results that don't correctly hit the uncertainty
      \item There is a bias-variance tradeoff associated with making predictions at different $M$ for different lengths of records -- and in a setting with a lot of anthropogenic and low-frequency variability with how long your record $N$ is.
    \end{enumerate}
    \item Procedure for this experiment:
    \begin{enumerate}
      \item For several (two or three) values of $M,N$, create many sequences of $(M+N)$ years from the ``true'' model
      \item For each sequence:
      \begin{enumerate}
        \item Fit the model in a Bayesian framework
        \item For each draw from the posterior, calculate the probability of exceeding some threshold $\tau$ in the future $M$ years. Can consider subtracting out the value estimated from the ``true'' sequence for normalization purposes
        \item Calculate the expected value of exceeding this threshold and some metric of the spread (95\% confidence interval width?) associated with these MCMC draws of $p(y > \tau)$
      \end{enumerate}
      \item Plot these two distributions (over many sequences), separately for the two or three values of $M$ and $N$.
      \item We can repeat for a couple of estimation strategies.
    \end{enumerate}
  \end{enumerate}
  \item Second experiment: focusing on $M,N$
  \begin{enumerate}
    \item Points to hit with this experiment:
    \begin{enumerate}
      \item How does varying $M$ and $N$ affect the results above?
      \item What are implications of different $M$ and $N$ for estimation procedures used?
    \end{enumerate}
    \item Procedure for this experiment:
    \begin{enumerate}
      \item Create a grid of $M$ and $N$. For each grid cell, create many $(M+N)$-year sequences from the ``true'' model. For each sequence:
      \begin{enumerate}
        \item Fit the model in a Bayesian framework
        \item For each draw from the posterior, calculate the probability of exceeding some threshold $\tau$ in the future $M$ years. Can consider subtracting out the value estimated from the ``true'' sequence for normalization purposes
        \item Calculate the expected value of exceeding this threshold
      \end{enumerate}
      \item Calculate the expected value by averaging across all sequences for the given $M,N$
      \item Repeat over all grid cells
      \item Plot the grid
      \item Repeat for several estimation strategies
    \end{enumerate}
  \end{enumerate}
\end{enumerate}

% -----------------------------------------------------------------------------
% Loss Function
% -----------------------------------------------------------------------------

\section{Synthetic Utility Estimation}

\begin{enumerate}
  \item Points to hit in this section
  \begin{enumerate}
    \item Uncertainty decreases with $N$, if you are fitting using a reasonable model
    \item For short $M$ the uncertainty is less in a world with strong memory than in a world with more short-term variance and long-term climate change signal.
    \item For long $M$ the uncertainty in a world with strong long-term climate change signal may be less than in a world with strong persistence.
  \end{enumerate}
  \item Experiment: comparing two projects with different $M$. Procedure:
  \begin{enumerate}
    \item Define a cost function for the proposed decisions. Can be simplified -- assume exposure is constant(-ish?)
    \item Define a few representative values of $N$ (\SIlist{10;30;50;100}{year})
    \item For each decision, and each value of $N$:
    \begin{enumerate}
      \item create many $(M+N)$-year sequences from the ``true'' model. For each sequence:
      \begin{enumerate}
        \item Fit the estimation model in a Bayesian framework
        \item For each draw from the posterior, calculate the expected utility
      \end{enumerate}
      \item Plot the distribution for this decision and value of $N$.
    \end{enumerate}
  \end{enumerate}
  \item Experiment: time-varying flood exposure. Procedure: repeat above analysis except rather than varying $N$ (pick a standard like 30 or 50 years), vary the flood exposure (\ie between 7\% growth per year and 2\% decline per year). Thus show that even with a strong climate change signal, the distribution of future utility is extremely variable for a large (costly) project with long $M$ -- this should address ``our friends who want to build Trump walls for 2100 GCM projections''
\end{enumerate}

% -----------------------------------------------------------------------------
% Some Results
% -----------------------------------------------------------------------------

\section{Preliminary Results}

Here are some results for a few combinations of true models and fits.
For each model:
\begin{enumerate}
  \item For each (of 1000) sample of the parameters from the MCMC simulation and each (of 1000) sequences of streamflow that were synthetically generated, the $T$-year event. Here $T=100$. Plot distribution of $T$-year event.
  \item Look at exceedances in the true data of the estimated $T$-year event
  \item Calculate and plot expected number of exceedances per year
  \item Calculate and plot probability of at least one exceedance
\end{enumerate}

\subsection{Stationary Data fit to Stationary Model}
I set $\mu=10,\sigma=2$
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{stationary_stationary_x_star_distribution.pdf}
  \caption{Distribution of $100$-year event $X^*$ for stationary model with stationary fit}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{stationary_stationary_expected_exceedances.pdf}
  \caption{Exceedance per year for stationary model with stationary fit}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{stationary_stationary_risk_distribution.pdf}
  \caption{Probability of at least one exceedance for stationary model with stationary fit}
\end{figure}

\subsection{Low-Frequency Variability fit to Stationary Model}
I set $\mu_10$, $\sigma=1$, $A=[1, 0.25]$, $T=[40, 5]$.
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{lowfreq_stationary_x_star_distribution.pdf}
  \caption{Distribution of $100$-year event $X^*$ for low-frequency model with stationary fit}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{lowfreq_stationary_expected_exceedances.pdf}
  \caption{Exceedance per year for low-frequency model with stationary fit}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=5.5in]{lowfreq_stationary_risk_distribution.pdf}
  \caption{Probability of at least one exceedance for low-frequency model with stationary fit}
\end{figure}

\subsection{Remaining}

\begin{enumerate}
  \item Fitting simulated data:
  \begin{enumerate}
    \item I have working code to generate fit
    \item How to think about $X^*$ ($T$-year event): can we calculate it analytically as function of $\hat{\mu_0},\hat{\beta},\hat{\sigma},M$ or resolve using monte carlo simulation?
    \item Build \texttt{PyMC3} model for low-frequency variability with known $T$ (easy)
    \item Add linear trends (easy)
  \end{enumerate}
  \item Loss function
  \begin{enumerate}
    \item We've considered the problem of insuring a dam -- need to implement
  \end{enumerate}
\end{enumerate}

### 2018-02-11 writeup/Outline/tex/sections/07-discussion.tex
start running models

\section{Discussion}

\begin{enumerate}
  \item Using the true generating function under-states our uncertainty -- can connect to the idea that using model chaining also under-states our uncertainty
  \item These are idealized scenarios but are motivated by real-world problems in flood adaptation (begin to build theory which is applicable to old cities in U.S., fast-growing cities around the world, and more)
  \item More carefully critique the current literature in light of these results
  \item Future steps
  \begin{enumerate}
    \item Better loss estimation
    \item Real options model
    \item Better ways to downscale climate projections for estimating flood loss
    \item Extend beyond annual-maximum
  \end{enumerate}
\end{enumerate}

### 2018-02-11 experiments/stationary.py

@@ -1,0 +1,38 @@
+"""Stationary
+"""
+import numpy as np
+import xarray as xr
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2
+from floodsampling.fit import TrendLN2Stan
+from floodsampling.util import get_data_path
+
+# Define Parameters
+N_try = np.array([10, 50, 250])
+M_try = np.array([5, 10, 50])
+n_seq = 250
+n_sim = 100
+threshold = 2500
+
+# Initialize the bias and variance
+empty = np.zeros(shape=(N_try.size, M_try.size)) # indexed (N, M)
+bias = xr.DataArray(data=empty, coords={'N': N_try, 'M': M_try}, dims=['N', 'M'])
+variance = xr.DataArray(data=empty, coords={'N': N_try, 'M': M_try}, dims=['N', 'M'])
+
+M = 50
+N = 100
+
+gen_fun = CZNINO3LN2(
+    N=N, M=M, t0=0, mu_0=6.5, beta_mu=0.75, gamma=0,
+    coeff_var=0.075, sigma_min=0.01, n_seq=n_seq,
+)
+fit_fun = TrendLN2Stan(
+    sflow=gen_fun, n_sim=n_sim,
+    chains=1, warmup=1000
+)
+print(fit_fun.get_data())

### 2018-02-11 launch-notebook.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+#SBATCH --account=cwc
+#SBATCH -J notebook
+#SBATCH --time=6:00:00
+
+
+# Setup Environment
+module load anaconda
+source activate floodsampling
+
+export XDG_RUNTIME_DIR=""
+
+jupyter notebook --no-browser --ip "*" --notebook-dir $HOME

### 2018-02-13 writeup/Outline/main.tex
update files

@@
 \maketitle
-%\RaggedRight
-
+\RaggedRight{}
 \begin{abstract}
@@
   Despite broad recognition of this nonstationarity, there has been little theoretical development of ideas for the design and operation of infrastructure considering the regime structure of such changes and their potential predictability.
-  This theory is well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies with $M$.
+  This theory is well-suited to analysis of flood adaptation strategies with fixed project operation period \(M\), but not for analysis of sequential decision strategies.
   In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
@@
 
-Classical methods for design of infrastructure first specify a return period $T$ (\ie 100 years), and then design a project that protects against this $T$-year event.
+Classical methods for design of infrastructure first specify a return period \(T\) (\ie{} 100 years), and then design a project that protects against this \(T\)-year event.
 Much of the flood frequency analysis (FFA) literature has tended to focus on the details of flood frequency estimation and on trend detection associated with climate change and land use cover \citep{Merz2008,Kidson2016,Merz2014}.
-The mathematical problem of interest then becomes estimation of the probability distribution $f(p_T)$, where $p_T$ is binary event that the threshold $Q_T$ is exceeded:
+The mathematical problem of interest then becomes estimation of the probability distribution \(f(p_T)\), where \(p_T\) is binary event that the threshold \(Q_T\) is exceeded:
 \begin{equation*}
@@
 \end{equation*}
-and where $t_0$ is the first year forecast and $M$ is the project operation period.
-Although in a stationary setting an unbiased estimator of $Pr \qty[ Q(t) \geq Q_T]$ converges asymtotically to $1 - \frac{1}{T}$, the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of $Q_T$ is used for design \citep{Stedinger1997}\footnote{Vogel198x citation}.
+and where \(t_0\) is the first year forecast and \(M\) is the project operation period.
+Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \(Q_T\) is used for design \citep{Stedinger1997}\footnote{Vogel198x citation}.
 
 Alternatively, risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design  may itself be a decision variable.
-\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of $p$ (equivalent to different values of $T$), and also for different sample sizes for estimation $N$ and different project operation periods $M$.
-The duration $M$ may reflect the physical life of a proposed infrastructure element, or its economic life.
-In the context of financial instruments which may be used for risk mitigation, $M$ specifies the length of the contract and $Q_T$ the threshold at which a payout is triggered.
-In such a setting, the mathematical problem is to estimate the future distribution of $Q(t)$ rather than just the future distribution of the binary outcome $p_T$.
+\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of \(p\) (equivalent to different values of \(T\)), and also for different sample sizes for estimation \(N\) and different project operation periods \(M\).
+The duration \(M\) may reflect the physical life of a proposed infrastructure element, or its economic life.
+In the context of financial instruments which may be used for risk mitigation, \(M\) specifies the length of the contract and \(Q_T\) the threshold at which a payout is triggered.
+In such a setting, the mathematical problem is to estimate the future distribution of \(Q(t)\) rather than just the future distribution of the binary outcome \(p_T\).
 
 Under classical assumptions of stationarity, all estimates are assumed independent of time.
-Yet even under stationarity, \citet{Jain2001} showed that low-frequency variability can lead to a high probability of surprise -- particularly when the length of the record used for estimation $N$ is short.
+Yet even under stationarity, \citet{Jain2001} showed that low-frequency variability can lead to a high probability of surprise, particularly when the length of the record used for estimation \(N\) is short.
 This finding is particularly important to keep in mind given that low-frequency change has been found to be large as compared to detected trends in the observed record \citep{Hodgkins2017}.
@@
 Attempts to use a ``model chain'' approach emcompassing:
-\begin{enumerate*}[label=(\roman*)]
+\begin{enumerate*}[label= (\roman*) ]
   \item emission scenario;
@@
 typically with bias correction applied at several steps, can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
-In particular, the downscaling and bias correction methods typically applied assume stationary relationships (\ie between GCM rainfall and observed rainfall) which have no physical basis, particularly in a changing climate.
-Recent work has also sought to shorten the above model chain by modeling $Q(X(t))$, where $X$ is a set of climate state variables from a GCM run, and where the relationship between $X$ and $Q$ can more reasonably be assumed stationary \citep{Hall2014,Delgado2014,Silva2016}.
+In particular, the downscaling and bias correction methods typically applied assume stationary relationships (\eg{} between GCM rainfall and observed rainfall) which have no physical basis, particularly in a changing climate.
+Recent work has also sought to shorten the above model chain by modeling \( Q(X(t)) \), where\( X \) is a set of climate state variables from a GCM run, and where the relationship between \(X\) and \(Q\) can more reasonably be assumed stationary \citep{Hall2014,Delgado2014,Silva2016}.
 Alternatively, purely statistical approaches have extended classical flood frequency analysis by incorporating a time trend in the parameters of statistical distributions \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}.
@@
 
-The wide variety of methods, each with its own advantages and disadvantages, for estimating the future distribution of flooding $p \qty(Q(t))$ illustrates the theoretical need for approaches to incorporate uncertain estimates of future floods into a decision framework.
+The wide variety of methods, each with its own advantages and disadvantages, for estimating the future distribution of flooding \(p \qty(Q(t))\) illustrates the theoretical need for approaches to incorporate uncertain estimates of future floods into a decision framework.
 As we have argued, however, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times.
-There is thus \emph{a critical theoretical need for decision frameworks which can compare projects of different operation period} using an arbitrary choice of model(s) of future flood behavior.
+There is thus \emph{a critical theoretical need for decision frameworks which can compare projects of different operation period} using an arbitrary choice of model of future flood behavior.
 Sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
@@
 
-In this context, we consider financial risk mitigation instruments in which the price is a function of the amount of coverage $C$, the expected probability that the insurance is triggered $\expectation \qty(p_T) \equiv \frac{1}{T}$, and the variance (or more generally full uncertainty distribution) associated with this estimate $\variance \qty(p_T)$.
-Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure $p_T$ for the next $M$ years, as well as the uncertainty and bias in this estimate.
+Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure \(p_T\) for the next \(M\) years, as well as the uncertainty and bias in this estimate.
 However, to properly analyse it we need detailed information on complex cost and loss structures; we thus focus on the illustrative example of financial instruemnts for clarity.
@@
 This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
-We proceed as follows $\ldots$.
+We proceed as follows \( \ldots \).
 
-\section{Methods} \label{sec:methods}
+\section{Methods}\label{sec:methods}
 
-We use a highly idealized model to generate sequences of annual-maximum floods using a known generating mechanism, then fit these sequences in a Bayesian framework to three flood models (stationary, nonstationary, and hidden markov model) to evalaluate their performance under different scenarios (defined by observational record length $N$ and project planning period $M$) and criteria.
+We use a highly idealized model to generate sequences of annual-maximum floods using a known generating mechanism, then fit these sequences in a Bayesian framework to three flood models (stationary, nonstationary, and hidden markov model) to evalaluate their performance under different scenarios (defined by observational record length \(N\) and project planning period \(M\)) and criteria.
 
 \subsection{Synthetic Flood Sequence Generation}
-
-Extensive discussion in the literature has considered the appropriate choice of distribution for annual-maximum flood sequences.
-Although three-paremeter models are better able to represent extreme values than two parameter models \citep{IACWD1982,Vogel1996}, we assume that flood sequences are \emph{conditionally} lognormal, given a mean and standard deviation which vary in time:
-\begin{equation} \label{eq:lognormal}
-  \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
-\end{equation}
-where $\normal$ denotes a normal distribution and $\mu(t), \sigma(t)$ are the location and scale parameters, respectively.
 
@@
 This is consistent with findings that have linked flood risk around the world with well-known climate mechanisms such as El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Ward2014,Emerton2017}.
-To capture this influence, some recent studies have incorporated a climate index or set of climate indices $X$ as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
+To capture this influence, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
 For a more comprehensive review, we refer the reader to \citet{Hall2014}.\footnote{Farnham et al}
@@
 Though the time scales are specific to ENSO, this pattern of quasi-oscillatory behavior modulated by low-frequency behavior is typical of many other climate phenomena.
-A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index \footnote{We greatfully acknowledge the contribution of Nandini Ramesh} with stationary forcing.
-To create an annual time series, we average the May-July months to create a time series $X(t)$ where $t$ is taken in annual steps $1, \ldots, \num{100000}$.
+A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index\footnote{We greatfully acknowledge the contribution of Nandini Ramesh} with stationary forcing.
+To create an annual time series, we average the May-July months to create a time series \( X(t) \) where \(t\) is taken in annual steps \(1, \ldots, \num{100000}\).
 
-Though statistical models relating climate indices and streamflow maxima typically require strong assumptions (\ie a stationary relationship between predictor and predictand; see \texttt{Farnham et al}), this framework is the starting point of our synthetic flood generation approach.
-We partition $\mu(t)$ into a dynamical component, driven by the variability of the climate predictor $X(t)$, and a trend component, represented by a linear increase in time.
+Though statistical models relating climate indices and streamflow maxima typically require strong assumptions (\ie{} a stationary relationship between predictor and predictand; see \texttt{Farnham et al}), this framework is the starting point of our synthetic flood generation approach.
+
+\begin{equation} \label{eq:lognormal}
+  \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
+\end{equation}
+The parameters of the distribution vary in time according to the sampled NINO3 index denoted \( X(t) \) as well as time itself, representing secular trends:
+\begin{equation}
+  \mu(t) = \mu_0 + \beta X(t) + \gamma \qty(t - t_0)
+\end{equation}
+and a constant coefficient of variation
+\begin{equation}
+  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq 0.01 \\ 0.01 & \qqtext{else} \end{cases}
+\end{equation}
+
+We partition \( \mu(t) \) into a dynamical component, driven by the variability of the climate predictor \(X(t)\), and a trend component, represented by a linear increase in time.
 This trend component may represent changes in land use, global temperature change, or river modification \citep[see][]{Merz2014}.
-Though this is a highly simplified model neglecting many important mechanisms, it enables us to generate many synthetic sequences of $\qty[M + N]$ years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
-Because the computational cost is small, it is easy to generate sets of sequences following the procedure
-\begin{enumerate}
-  \item Choose the parameters $M,N,\alpha,\beta,\gamma$
-  \item Select the ``zero'' time of analysis $t_0 \in \qty[N, \ldots, \num{100000} - M]$ so that the time used is $t_{\text{use}} = t_0 - N + 1, \ldots, t_0 - M$ (note that this indexing treats $t_0$ as part of the $N$-year observed record)
-  \item Subset $x(t) = X(t \in t_{\text{use}})$
-  \item Let $\mu(t) = \beta x(t) + \gamma \qty(t - t_0)$
-  \item Assume a constant coefficient of variation so that $\sigma(t) = \alpha \mu(t)$
-  \item Draw $\log Q(t) \sim \normal \qty(\mu(t), \sigma(t))$
-\end{enumerate}
+Though this is a highly simplified model neglecting many important mechanisms, it enables us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
+Because the computational cost is small, it is easy to generate sets of sequences.
 
-In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of $M,N,\alpha,\beta,\gamma$ and discuss the processes which they can reasonably represent.
+In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of parameters and discuss the processes which they can reasonably represent.
 
-\subsection{Distributional Model \label{sec:estimation}}
+\subsection{Distributional Model\label{sec:estimation}}
 
-Because we do not, in general, know the true distribution of annual-maximum floods, we compare a stationary and non-stationary parameterizations of the generalized extreme value (GEV) statistical distribution (eq.~\ref{eq:GEV}) for estimating the distribution of $Q(t)$ for the future $M$-year period.
+Because we do not, in general, know the true distribution of annual-maximum floods, we compare a stationary and non-stationary parameterizations of the generalized extreme value (GEV) statistical distribution (eq.~\ref{eq:GEV}) for estimating the distribution of \(Q(t)\) for the future \(M\)-year period.
 \begin{equation}
@@
 \end{equation}
-Where $\mu$ is the location parameter, $\sigma$ is the shape parameter, and $\xi$ is the scale parameter of the GEV distribution.
-For each simulation, the first $N$ years $\qty(t=[t_0-N+1, \ldots, t_0])$ are treated as observations.
+Where \(\mu{}\) is the location parameter, \(\sigma{}\) is the shape parameter, and \(\xi{}\) is the scale parameter of the GEV distribution.
+For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations.
 Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty.
@@
 
-\subsection{Hidden Markov Model \label{sec:HMM}}
+\subsection{Hidden Markov Model\label{sec:HMM}}
 
-A Hidden Markov Model (HMM) is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved (\ie hidden) states \citep{Rabiner1986}.
-The (unobserved) states evolve following a first-order Markov process, and the observed variable (\ie streamflow) depends only on the observed state.
+A Hidden Markov Model (HMM) is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved (\ie{} hidden) states \citep{Rabiner1986}.
+The (unobserved) states evolve following a first-order Markov process, and the observed variable (\ie{} streamflow) depends only on the observed state.
 As such, HMMs are a structured prediction method which extend general mixture models to sequences of data, where position in the sequence is relevant.
@@
 To supplement the distributional models described in \cref{sec:estimation}, we also fit observed streamflow sequences using a HMM as implemented in the python \texttt{pomegranate} package \citep{Schreiber2016}.
-The model is fit to the $N$ observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
+The model is fit to the \(N\) observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
 To choose the number of hidden states, a log probability score is used.
@@
       \midrule
-        stationary & $Q \sim \GEV \qty(\mu, \sigma, \xi)$ & \cref{sec:estimation} \\
-        location trend & $Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)$ & \cref{sec:estimation} \\
-        %full trend & $Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma_0 + \beta_\sigma \qty(t-t_0), \xi_0 + \beta_\xi \qty(t-t_0))$ \\
+        stationary & \(Q \sim \GEV \qty(\mu, \sigma, \xi)\) & \cref{sec:estimation} \\
+        location trend & \(Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)\) & \cref{sec:estimation} \\
         HMM & log-normal Hidden Markov Model & \cref{sec:HMM} \\
@@
   \end{center}
-  \caption{Summary of models used for fitting \label{tab:model-fitting}}
+  \caption{Summary of models used for fitting\label{tab:model-fitting}}
 \end{table}
@@
 
-Finally, we consider a contract in which an insurer agrees to pay the entire coverage cost $C$ every time that the trigger threshold $Q_T$ is exceeded.
-Although $Q_T$ is in general a decision variable, we treat it as fixed.
-Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that $Q(t) \geq Q_T$ for $t \in \vb{t}$, which is equivalent to $p_T(t)$.
+Finally, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
+Although \(Q_T\) is in general a decision variable, we treat it as fixed.
+Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that \(Q(t) \geq Q_T\) for \(t \in \vb{t}\), which is equivalent to \(p_T(t)\).
 
-For a particular year, the fair price for the insurance premium is $\expectation \qty(p_T) C$.
-However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger $Q_T$ is specified in the contract and thus not uncertain).
-This risk premium is proportional to the variance of $p_T$, or more generally to its uncertainty distribution.
+For a particular year, the fair price for the insurance premium is \(\expectation \qty(p_T) C\).
+However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger \(Q_T\) is specified in the contract and thus not uncertain).
+This risk premium is proportional to the variance of \(p_T\), or more generally to its uncertainty distribution.
 The insurance premium for a particular year is thus
@@
 \end{equation}
-where $R \geq 0$ is a coefficient describing the cost of tue uncertainty in the estimate of $p_T(t)$.
+where \(R \geq 0\) is a coefficient describing the cost of tue uncertainty in the estimate of \(p_T(t)\).
 
-In this paper we consider the price of a contract of $M$ years to be simply the summed price of the contract for each individual year.
-We assume that the discount rate is equal to inflation (\ie the inflation-adjusted discount rate is zero) for simplicity.
-A more comprehensive pricing mechanism for the financial instrument could incorporate more or different information about the uncertainty in the estimate, and should consider the price of the $M$ years as a portfolio (\ie should consider correlation between the different years).
+In this paper we consider the price of a contract of \(M\) years to be simply the summed price of the contract for each individual year.
+We assume that the discount rate is equal to inflation (\ie{} the inflation-adjusted discount rate is zero) for simplicity.
+A more comprehensive pricing mechanism for the financial instrument could incorporate more or different information about the uncertainty in the estimate, and should consider the price of the \(M\) years as a portfolio (\ie{} should consider correlation between the different years).
 Such an approach, however, is not necessary for this conceptual study.
 
-\section{Are Generated Sequences Realistic? \label{sec:sequence-realistic}}
+\section{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
 
-In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc and compare to some that have been identified in the literature.
+In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc{} and compare to some that have been identified in the literature.
 
@@
 
-Consider static risk and a stationary process, and partiuclarly implications of $M$ and $N$.
+Consider static risk and a stationary process, and partiuclarly implications of \(\qty{M, N}\).
 Much of this is already laid out in \citet{Lall1987}.
@@
 Consider static risk and a non-stationary process.
-Estimate using drift or nonstationary models using full $N$ of the most recent $N$ (\ie what Vogel is proposing).
-Illustrate via simulation for different signal to noise ratio of nonstationary terms and $M$ and $N$.
+Estimate using drift or nonstationary models using full \(N\) of the most recent \(N\) (\ie{} what Vogel is proposing).
+Illustrate via simulation for different signal to noise ratio of nonstationary terms and \(\qty{M, N}\).
 
@@
 
-Consider dynamic risk, \ie updating and sequential decisions, a nonstationary generative process, and $M$ and $N$ considerations
+Consider dynamic risk, \ie{} updating and sequential decisions, a nonstationary generative process, and \(M\) and \(N\) considerations.
 
@@
 \clearpage
-\printbibliography
+\printbibliography{}
 

### 2018-02-13 experiments/plot_stationary_sequences.py

@@ -17,7 +17,7 @@ gen_fun = CZNINO3LN2(
 sequences = np.log(gen_fun.get_data())
 
 n_seq_plot = 50
-threshold = np.exp(8)
+threshold = 5000
 
 fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
     gridspec_kw={'width_ratios': [4, 1]})
@@ -42,4 +42,4 @@ ax.legend()
 fig.tight_layout()
 fig.subplots_adjust(top=0.9)
 fig.suptitle('{} Synthetic Streamflow Sequences'.format(n_seq_plot))
-plt.show()
+plt.savefig('figs/stationary_sequences.pdf')

### 2018-02-13 experiments/plot_trend_sequences.py

@@ -17,7 +17,7 @@ gen_fun = CZNINO3LN2(
 sequences = np.log(gen_fun.get_data())
 
 n_seq_plot = 50
-threshold = np.exp(8)
+threshold = 5000
 
 fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
     gridspec_kw={'width_ratios': [4, 1]})
@@ -42,4 +42,4 @@ ax.legend()
 fig.tight_layout()
 fig.subplots_adjust(top=0.9)
 fig.suptitle('{} Synthetic Streamflow Sequences'.format(n_seq_plot))
-plt.show()
+plt.savefig('figs/trend_sequences.pdf')

### 2018-02-13 experiments/stationary.py

@@ -14,11 +14,15 @@ from floodsampling.streamflow import CZNINO3LN2
 from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 
 # Define Parameters
-N_try = np.array([10, 30, 50, 150, 500])
-M_try = np.array([5, 10, 30, 50, 100])
+N_try = np.array([10, 30, 50, 150])
+M_try = np.array([5, 10, 30, 50])
 n_seq = 1000
 n_sim = 1000
 threshold = 5000
+
+def some_function(x: int, y: int) -> None:
+    "Some cool docstring"
+    return x * y
 
 # Initialize the bias and variance
 model_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
@@ -30,41 +34,46 @@ empty = xr.DataArray(
 )
 fit = xr.Dataset({'bias': empty, 'variance': empty}).to_dataframe()
 
-for M in M_try:
-    for N in N_try:
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, mu_0=6.5, beta_mu=0.75, gamma=0,
-            coeff_var=0.075, sigma_min=0.01, n_seq=n_seq,
-        )
-
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-        bias_i, variance_i = fit_fun.decompose_variance(threshold=threshold)
-        fit.loc[M, N, 'ln2_stationary']['bias'] = bias_i
-        fit.loc[M, N, 'ln2_stationary']['variance'] = variance_i
-
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-        bias_i, variance_i = fit_fun.decompose_variance(threshold=threshold)
-        fit.loc[M, N, 'ln2_trend']['bias'] = bias_i
-        fit.loc[M, N, 'ln2_trend']['variance'] = variance_i
-
-        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
-        bias_i, variance_i = fit_fun.decompose_variance(threshold=threshold)
-        fit.loc[M, N, 'hmm']['bias'] = bias_i
-        fit.loc[M, N, 'hmm']['variance'] = variance_i
+for N in N_try:
+    # We don't need to re-fit everything every time we increase M.
[…]

### 2018-02-13 writeup/Outline/main.tex
start re-organizing the introduction

@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
 
@@
   Hydroclimatic systems exhibit organized low-frequency and regime-like variability at multiple time scales, causing the risk associated with climate extremes such as floods and droughts to vary at multiple timescales.
-  Despite broad recognition of this nonstationarity, there has been little theoretical development of ideas for the design and operation of infrastructure considering the regime structure of such changes and their potential predictability.
-  This theory is well-suited to analysis of flood adaptation strategies with fixed project operation period \(M\), but not for analysis of sequential decision strategies.
+  Despite broad recognition of this nonstationarity, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
+  Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period \(M\), but not for analysis of sequential decision strategies.
   In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
-  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show that \texttt{RESULT}.
+  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show that \(\ldots{}\).
   This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
@@
 Classical methods for design of infrastructure first specify a return period \(T\) (\ie{} 100 years), and then design a project that protects against this \(T\)-year event.
-Much of the flood frequency analysis (FFA) literature has tended to focus on the details of flood frequency estimation and on trend detection associated with climate change and land use cover \citep{Merz2008,Kidson2016,Merz2014}.
-The mathematical problem of interest then becomes estimation of the probability distribution \(f(p_T)\), where \(p_T\) is binary event that the threshold \(Q_T\) is exceeded:
+The mathematical problem of interest then becomes estimation of the probability distribution \( f(p_T) \), where \( p_T \) is binary event that the threshold \( Q_T \) is exceeded by the annual-maxiumum flood series \( Q(t) \):
 \begin{equation*}
@@
 \end{equation*}
-and where \(t_0\) is the first year forecast and \(M\) is the project operation period.
-Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \(Q_T\) is used for design \citep{Stedinger1997}\footnote{Vogel198x citation}.
+where \( t_0 \) is the first year forecast and \( M \) is the project operation period.
+Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \( Q_T \) is used for design \citep{Stedinger1997}\footnote{Manu has mentioned a Vogel198x citation but I haven't successfully found it.}.
 
-Alternatively, risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design  may itself be a decision variable.
-\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of \(p\) (equivalent to different values of \(T\)), and also for different sample sizes for estimation \(N\) and different project operation periods \(M\).
-The duration \(M\) may reflect the physical life of a proposed infrastructure element, or its economic life.
-In the context of financial instruments which may be used for risk mitigation, \(M\) specifies the length of the contract and \(Q_T\) the threshold at which a payout is triggered.
-In such a setting, the mathematical problem is to estimate the future distribution of \(Q(t)\) rather than just the future distribution of the binary outcome \(p_T\).
+\subsection{Physical Reasoning}
+
+In order to make probabilistic predictions of future flooding in a particular location, it is helpful to begin by considering the specific physical mechanisms which may lead to flooding in particular location.
+
+In large river basins, extreme river flooding typically requires the large-scale transport and convergence of moisture\citep{Lu2013,Merz2014,Dacre2014}\footnote{Farnham et al}, which often requires specific climate mechanisms.
+For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent large-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
+This is consistent with findings that have linked flood risk around the world with well-known climate mechanisms such as El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Ward2014,Emerton2017}.
+
+Nonstationarity in hydroclimate systems can occur due to climate change, river modification, and land use change \citep{Milly2008,Merz2014}.
+Thermodynamic scaling arguments predict an intensified hydrologic cycle under global warming due to the Clausius-Clapeyron scaling of the atmosphere's moisture-holding capacity with temperature \citep[see][]{Muller2011,OGorman2015}.
+However, these scaling arguments do not in general hold over land \citep{Byrne2015,Shaw2016}.
+In addition to thermodynamical changes, climate extremes are highly sensitive to changes in future climate dynamics, which are themselves imperfectly understood due to complex and non-linear physics \citep{Palmer2013}.
+Consequently, predicting future flood risk in a particular location involves making estimates of future climate circulation patterns, the associated thermodynamics, as well as any catchment- and river-scale changes which may affect rainfall-runoff response.
+
+\subsection{Estimating Future Flood Risk}
 
@@
 
-Of course, the hydroclimate systems are not stationary, due to climate, river modification, and land use change \citep{Milly2008,Merz2014}.
-However, predicting future hydrologic behavior in a particular location is inherently difficult given limited data and complex, highly nonlinear physics.
-Physical scaling arguments for an intensified hydrologic cycle under climate change \citep[see][]{Muller2011,OGorman2015} do not in general hold over land \citep{Byrne2015,Shaw2016}.
 Attempts to use a ``model chain'' approach emcompassing:
@@
 typically with bias correction applied at several steps, can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
+
 In particular, the downscaling and bias correction methods typically applied assume stationary relationships (\eg{} between GCM rainfall and observed rainfall) which have no physical basis, particularly in a changing climate.
@@
 
-The wide variety of methods, each with its own advantages and disadvantages, for estimating the future distribution of flooding \(p \qty(Q(t))\) illustrates the theoretical need for approaches to incorporate uncertain estimates of future floods into a decision framework.
-As we have argued, however, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times.
+To capture this influence, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
+For a more comprehensive review, we refer the reader to \citet{Hall2014}.\footnote{Farnham et al}
+
+\subsection{Decision Approaches}
+
+Incorporating uncertain projections of physical variables into useful decision models is further hampered by the diffulties inherent to forecasting the needs of society\footnote{Cite needed}, which may even be coupled to the natural system.\footnote{Insert citation on socio-hydrology}
+Risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design  may itself be a decision variable.
+\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of \(p\) (equivalent to different values of \(T\)), and also for different sample sizes for estimation \(N\) and different project operation periods \(M\).
+Some prominent methodologies for risk-based design under nonstationary in the water resources literature include robust decision making (RDM)\footnote{(Lempert et al.
+2006; Lempert and Groves 2010; Matrosov et al.
+2013)}, scenario-neutral planning\footnote{(Prudhomme et al.
+2010)}, infogap analysis\footnote{(Ben-Haim 2006; Korteling et al.
+2013)}, and decision scaling\footnote{(Brown et al.
+2011, 2012; Moody and Brown 2013; Turner et al.
+2014)}.
+
+Alternatively, sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
+Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}
+\texttt{need some citations for sequential decisions}.
+Given that, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times, models which can capture this uncertainty may be particularly advantageous for sequential models.
+
 There is thus \emph{a critical theoretical need for decision frameworks which can compare projects of different operation period} using an arbitrary choice of model of future flood behavior.
-Sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
-Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}\footnote{need some citations for sequential decisions}.
+In the following sections we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary environment.
+Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure \(p_T\) for the next \(M\) years, as well as the uncertainty and bias in this estimate.
+This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
 
-Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure \(p_T\) for the next \(M\) years, as well as the uncertainty and bias in this estimate.
-However, to properly analyse it we need detailed information on complex cost and loss structures; we thus focus on the illustrative example of financial instruemnts for clarity.
-
-In the following sections we illustrate some very basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary environment.
-This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
-We proceed as follows \( \ldots \).
+We proceed as follows \(\ldots{}\)
 
@@
 
-We use a highly idealized model to generate sequences of annual-maximum floods using a known generating mechanism, then fit these sequences in a Bayesian framework to three flood models (stationary, nonstationary, and hidden markov model) to evalaluate their performance under different scenarios (defined by observational record length \(N\) and project planning period \(M\)) and criteria.
+We use an idealized model to generate sequences of annual-maximum floods, fit these sequences in a probabilistic framework to three flood models that incorporate time in different ways, and evalaluate the performance of these models for multiple observational record lengths \(N\) and project planning periods \(M\).
 
 \subsection{Synthetic Flood Sequence Generation}
-
-In large river basins, extreme river flooding typically requires the large-scale transport and convergence of moisture,\footnote{Is this worth adding in} which often requires specific climate mechanisms.
-For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent large-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
-This is consistent with findings that have linked flood risk around the world with well-known climate mechanisms such as El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Ward2014,Emerton2017}.
-To capture this influence, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
-For a more comprehensive review, we refer the reader to \citet{Hall2014}.\footnote{Farnham et al}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\textwidth]{../../figs/stationary_sequences.pdf}
+  \caption{50 simulated sequences created from the Cane-Zebiak model  \(N=50, M=50, t_0=0, \mu_0=6.5, \beta_\mu=0.5, \gamma=0, \alpha=0.075, \sigma_\text{min}=0.01\).
+49 of the sequences are shown in gray and a single randomly chosen sequence is highlighted in blue for emphasis.
+(L) shows the time series of log streamflow as a function of year, while (R) shows the distribution of 500 sequences over the historical and future periods.}\label{fig:stationary-sequences}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{../../figs/trend_sequences.pdf}
+  \caption{As \cref{fig:stationary-sequences} except \(\gamma=0.01\).}
+\end{figure}
+
 \section{Stationary, Static Risk}
@@
 
+\section{Notation}
+
+\begin{description}
+  \item[$Q$] Annual-maximum flood series
+  \item[$Q_T$] Flood threshold
+\end{description}
+
 % -----------------------------------------------------------------------------

### 2018-02-13 experiments/plot_stationary_sequences.py

@@ -40,6 +40,4 @@ ax.set_xticks([])
 ax.legend()
 
 fig.tight_layout()
-fig.subplots_adjust(top=0.9)
-fig.suptitle('{} Synthetic Streamflow Sequences'.format(n_seq_plot))
 plt.savefig('figs/stationary_sequences.pdf')

### 2018-02-13 experiments/plot_trend_sequences.py

@@ -40,6 +40,4 @@ ax.set_xticks([])
 ax.legend()
 
 fig.tight_layout()
-fig.subplots_adjust(top=0.9)
-fig.suptitle('{} Synthetic Streamflow Sequences'.format(n_seq_plot))
 plt.savefig('figs/trend_sequences.pdf')

### 2018-02-14 writeup/Outline/main.tex
Update intro some more

@@
 
-In large river basins, extreme river flooding typically requires the large-scale transport and convergence of moisture\citep{Lu2013,Merz2014,Dacre2014}\footnote{Farnham et al}, which often requires specific climate mechanisms.
-For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent large-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
-This is consistent with findings that have linked flood risk around the world with well-known climate mechanisms such as El Ni\~{n}o-Southern Oscillation (ENSO) \citep{Ward2014,Emerton2017}.
+Floods may be associated with a single rainfall event or be composed of a sequence of multiple, distinct events over a week or a season, such that there is pre-conditioning of the land surface soil moisture and natural and man-made stores that leads to a high flooding potential.
 
-Nonstationarity in hydroclimate systems can occur due to climate change, river modification, and land use change \citep{Milly2008,Merz2014}.
+In large river basins, Lagrangian tracking studies \citep{Gimeno2010} indicate that river flooding often requires the large-scale transport and convergence of moisture  which are driven by specific large-scale climate mechanisms \citep{Dacre2014,Gimeno2014,Lu2013,Dirmeyer2010}.\footnote{Farnham et al}
+For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent continental-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
+Other case studies \citep{Grams2014,Stadtherr2016}\footnote{Add my work} and theoretical analyses \citep{Hoskins2015,Screen2014,Coumou2014} support the hypothesis that specific climate mechanisms such as blocking or jet sinuosity  can lead to persistent, intense rainfall.
+
+At the seasonal and sub-seasonal scales, surface temperature boundary conditions are well known to significantly change the mean, variance, persistence, recurrence and spatial structure of atmospheric circulation patterns, as well as moisture availability.
+Strong links between climate phenomena such as El Ni\~no-Southern Oscillation (ENSO) and flood risk have been well established in both the water resources \citep[\ie{}][]{Ward2014} and atmospheric science \citep{Tedeschi2013} literatures.
+
+On longer time scales, changes to the physical processes that govern hydroclimate systems, known generically as nonstationarity \citep{Milly2008} can occur due to climate change, river modification, and land use change \citep{Merz2014}.
 Thermodynamic scaling arguments predict an intensified hydrologic cycle under global warming due to the Clausius-Clapeyron scaling of the atmosphere's moisture-holding capacity with temperature \citep[see][]{Muller2011,OGorman2015}.
-However, these scaling arguments do not in general hold over land \citep{Byrne2015,Shaw2016}.
-In addition to thermodynamical changes, climate extremes are highly sensitive to changes in future climate dynamics, which are themselves imperfectly understood due to complex and non-linear physics \citep{Palmer2013}.
-Consequently, predicting future flood risk in a particular location involves making estimates of future climate circulation patterns, the associated thermodynamics, as well as any catchment- and river-scale changes which may affect rainfall-runoff response.
+However, these scaling arguments do not in general hold over land \citep{Byrne2015} and typically describe short-duration rainfall maxima rather than persistent and intense rainfall sequences.
+As compared with thermodynamic changes there is less certainty as to how large-scale dynamical circulations may respond to warming signals \citep{Shaw2016,Barnes2015}, as the dynamics governing these mechanisms are highly non-linear \citep{Palmer2013} and in general not well resolved by climate models\footnote{Farnham 2018}.
 
-\subsection{Estimating Future Flood Risk}
+\subsection{Nonstationary Flood Frequency Analysis}
 
-Under classical assumptions of stationarity, all estimates are assumed independent of time.
-Yet even under stationarity, \citet{Jain2001} showed that low-frequency variability can lead to a high probability of surprise, particularly when the length of the record used for estimation \(N\) is short.
-This finding is particularly important to keep in mind given that low-frequency change has been found to be large as compared to detected trends in the observed record \citep{Hodgkins2017}.
+Under classical assumptions of stationarity, the parameters describing the hydroclimate system are assumed independent of time.
+Even when this assumption is satisfied, low-frequency variability can lead to a high probability of surprise, particularly when the length of the record used for estimation \(N\) is short \citep{Jain2001,Matalas2012}.
+However, widespread acceptance of nonstationarity as a principle has not led to easily generalizable methodologies for nonstationary flood frequency analysis.
 
-Attempts to use a ``model chain'' approach emcompassing:
+One approach is to extend the purely statistical models currently used in the United States \citep{IACWD1982}.
+Like the stationary case, there is no consensus as to the proper statistical distribution to use.
+Although Bulletin 17-B \citep{IACWD1982} mandates the use of the log-Pearson type III (LP3) model of annual-maximum floods in the United States, annual-maximum floods are also modeled three-parameter distributions such as the generalized extreme value distribution and two-parameter models such as the lognormal (LN2) \citep{Vogel1996}.
+Many studies have extended stationary statistical models to include a time trend on one or more parameters \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}, an approach which is conceptually and computationally simple but that lacks physical justification.
+Rather than conditioning parameters on time, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
+Other studies have used Markov or mixture models \citep{Waylen1986,Sveinsson2005,Griffis2007}, Bayesian approaches for combining information \citep{Lima2016,Bracken2017}, and other more sophisticated statistical models.
+
+An alternative approach is to use numerical models to represent the physical processes at all scales.
+Because models are not able to simultaneously represent scales ranging from global-scale climate phenomena to river hydrology, a ``model chain'' approach is used.
+A typical model chain might encompass:
 \begin{enumerate*}[label= (\roman*) ]
@@
   \item hydrological catchment model; and
-  \item flood frequency analysis,
+  \item flood frequency analysis.
 \end{enumerate*}
-typically with bias correction applied at several steps, can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
-
-In particular, the downscaling and bias correction methods typically applied assume stationary relationships (\eg{} between GCM rainfall and observed rainfall) which have no physical basis, particularly in a changing climate.
-Recent work has also sought to shorten the above model chain by modeling \( Q(X(t)) \), where\( X \) is a set of climate state variables from a GCM run, and where the relationship between \(X\) and \(Q\) can more reasonably be assumed stationary \citep{Hall2014,Delgado2014,Silva2016}.
-Alternatively, purely statistical approaches have extended classical flood frequency analysis by incorporating a time trend in the parameters of statistical distributions \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}.
-In this case, however, the analyst can choose not only the distribution but also  the time parameterization, if any, for each parameter, leading to many researcher degrees of freedom \citep[``forking paths'' or ``multiple comparisons'';][]{Gelman2013} and exacerbating the problems caused by lack of theory for model choice even under nonstationarity \citep{Kidson2016}.
-
-To capture this influence, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
-For a more comprehensive review, we refer the reader to \citet{Hall2014}.\footnote{Farnham et al}
+Typically bias correction is applied at several or steps of the procedure, which can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
 
@@
 
-Incorporating uncertain projections of physical variables into useful decision models is further hampered by the diffulties inherent to forecasting the needs of society\footnote{Cite needed}, which may even be coupled to the natural system.\footnote{Insert citation on socio-hydrology}
+Despite the abundance of approaches for estimating future flood hazard, each of which may give different estimates of future flood risk, the ultimate goal of flood frequency analysis is to inform risk management decisions.
+
 Risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design may itself be a decision variable.
@@
 This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
-
 We proceed as follows \(\ldots{}\)
@@
 
-\subsection{Synthetic Flood Sequence Generation}
+\subsection{Generating Synthetic Flood Sequences}
 
-ENSO \citep[see][for a comprehensive review]{Sarachik2010} varies in a quasi-oscillatory manner with variability on the order of \SIrange{4}{6}{year} modulated by low-frequency behavior.
-Though the time scales are specific to ENSO, this pattern of quasi-oscillatory behavior modulated by low-frequency behavior is typical of many other climate phenomena.
-A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index\footnote{We greatfully acknowledge the contribution of Nandini Ramesh} with stationary forcing.
-To create an annual time series, we average the May-July months to create a time series \( X(t) \) where \(t\) is taken in annual steps \(1, \ldots, \num{100000}\).
-
-Though statistical models relating climate indices and streamflow maxima typically require strong assumptions (\ie{} a stationary relationship between predictor and predictand; see \texttt{Farnham et al}), this framework is the starting point of our synthetic flood generation approach.
-
+We generate synthetic flood sequences from a simple conceptual model where the expected flood varies as a function of a single climate index plus time following the log-normal distribution:
 \begin{equation} \label{eq:lognormal}
@@
 \end{equation}
-The parameters of the distribution vary in time according to the sampled NINO3 index denoted \( X(t) \) as well as time itself, representing secular trends:
+
+ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a quasi-oscillatory manner with variability on the order of \SIrange{4}{6}{year} modulated by low-frequency oscillation \citep[see][for a comprehensive review]{Sarachik2010}.
+A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citep{Ramesh2017}.
+To create an annual time series, we average the October-December months to create a time series \( x(t) \) where \( t \) is taken in annual steps \(1, \ldots, \num{100000} \).
+The parameters of the distribution vary in time according to the sampled NINO3 index denoted \( x(t) \) as well as time itself, representing secular trends.
+First a starting point in the \SI{100000}{year} ENSO was randomly chosen, then \(\mu\) is determined by
 \begin{equation}
-  \mu(t) = \mu_0 + \beta X(t) + \gamma \qty(t - t_0)
+  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0)
 \end{equation}
@@
 \begin{equation}
-  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq 0.01 \\ 0.01 & \qqtext{else} \end{cases}
+  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq \sigma_\text{min} \\ \sigma_\text{min} & \qqtext{else} \end{cases}
 \end{equation}
+where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
 
-We partition \( \mu(t) \) into a dynamical component, driven by the variability of the climate predictor \(X(t)\), and a trend component, represented by a linear increase in time.
-This trend component may represent changes in land use, global temperature change, or river modification \citep[see][]{Merz2014}.
 Though this is a highly simplified model neglecting many important mechanisms, it enables us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
-Because the computational cost is small, it is easy to generate sets of sequences.
 
@@
 
-\subsection{Distributional Model\label{sec:estimation}}
+\subsection{Fitting Streamflow Sequences\label{sec:estimation}}
 
@@
 
+For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years. 
+While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.\footnote{Send this to discussion}
+
 \section{Summary}
 
-Some interesting conclusions
+The main point you are hitting is that the very nature of climate has long memory and some people point this out, many model it, and most future projections don't show this.
+All corrections (even if we like them) address the marginal distribution.
+So while the debate you refer to is interesting to a degree, the point we are making is that in terms of risk analysis it is not even tangentially relevant to reality and what we bring up -- our focus -- is strictly the risk based decision process and we are agnostic to the details of the models, other than the long memory features and spatial teleconnections associated with them that are seen as marks of the system behavior.
+We actually demonstrate that the potential loss in a 2100 scenario has to be catastrophic and really out of the range of any historical variability for it to be relevant. 
+If we accept this proposition then we have a strong case for the mitigation of climate change but a much weaker case for adaptation, since if we are not able to mitigate that scenario then building a giant wall to control the 2100 scenario today will be economically and socially infeasible anyway.
+So the next obvious question is in which situation would waiting and updating the decision make sense and this can be set up and explored or just discussed.
 
@@
 \begin{description}
-  \item[$Q$] Annual-maximum flood series
-  \item[$Q_T$] Flood threshold
+  \item[\(Q(t)\)] Annual-maximum flood series
+  \item[\(Q_T\)] Flood threshold
 \end{description}

### 2018-02-14 experiments/plot_stationary_sequences.py

@@ -19,6 +19,7 @@ sequences = np.log(gen_fun.get_data())
 n_seq_plot = 50
 threshold = 5000
 
+# PLOT LOG STREAMFLOW
 fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
     gridspec_kw={'width_ratios': [4, 1]})
 ax = axes[0]
@@ -36,6 +37,32 @@ fut_sequences = np.log(gen_fun.get_data('future').values.flatten())
 sns.distplot(hist_sequences, vertical=True, ax=ax, label='historical')
 sns.distplot(fut_sequences, vertical=True, ax=ax, label='future')
 ax.axhline(np.log(threshold), c='black', linewidth=0.4)
+ax.set_xticks([])
+ax.legend()
+
+fig.tight_layout()
+plt.savefig('figs/log_stationary_sequences.pdf')
+
+# PLOT ACTUAL STREAMFLOW
+sequences = gen_fun.get_data()
+fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
+    gridspec_kw={'width_ratios': [4, 1]})
+ax = axes[0]
+for i in np.arange(n_seq_plot-1):
+    sequences.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+sequences.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
+ax.set_ylabel('Log Streamflow')
+ax.set_title('')
+ax.axhline(threshold, c='black', linewidth=0.6)
+ax.axvline(0, c='black', linewidth=0.6)
+ax.set_ylim([0, 10000])
+
+ax = axes[1]
+hist_sequences = gen_fun.get_data('historical').values.flatten()
+fut_sequences = gen_fun.get_data('future').values.flatten()
+sns.distplot(hist_sequences, vertical=True, ax=ax, label='historical')
+sns.distplot(fut_sequences, vertical=True, ax=ax, label='future')
+ax.axhline(threshold, c='black', linewidth=0.4)
 ax.set_xticks([])
 ax.legend()
 

### 2018-02-14 experiments/plot_trend_sequences.py

@@ -19,6 +19,7 @@ sequences = np.log(gen_fun.get_data())
 n_seq_plot = 50
 threshold = 5000
 
+# PLOT LOG STREAMFLOW
 fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
     gridspec_kw={'width_ratios': [4, 1]})
 ax = axes[0]
@@ -36,6 +37,33 @@ fut_sequences = np.log(gen_fun.get_data('future').values.flatten())
 sns.distplot(hist_sequences, vertical=True, ax=ax, label='historical')
 sns.distplot(fut_sequences, vertical=True, ax=ax, label='future')
 ax.axhline(np.log(threshold), c='black', linewidth=0.4)
+ax.set_xticks([])
+ax.legend()
+
+fig.tight_layout()
+plt.savefig('figs/log_trend_sequences.pdf')
+
+
+# PLOT ACTUAL STREAMFLOW
+sequences = gen_fun.get_data()
+fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
+    gridspec_kw={'width_ratios': [4, 1]})
+ax = axes[0]
+for i in np.arange(n_seq_plot-1):
+    sequences.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+sequences.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
+ax.set_ylabel('Log Streamflow')
+ax.set_title('')
+ax.axhline(threshold, c='black', linewidth=0.6)
+ax.axvline(0, c='black', linewidth=0.6)
+ax.set_ylim([0, 10000])
+
+ax = axes[1]
+hist_sequences = gen_fun.get_data('historical').values.flatten()
+fut_sequences = gen_fun.get_data('future').values.flatten()
+sns.distplot(hist_sequences, vertical=True, ax=ax, label='historical')
+sns.distplot(fut_sequences, vertical=True, ax=ax, label='future')
+ax.axhline(threshold, c='black', linewidth=0.4)
 ax.set_xticks([])
 ax.legend()
 

### 2018-02-14 writeup/Outline/main.tex
write up the methods

@@
 
-\subsection{Nonstationary Flood Frequency Analysis}
+\subsection{Nonstationary Flood Frequency Analysis\label{sec:ffa}}
 
@@
 \end{equation}
+The parameter \( \mu(t) \) is generated using two different approaches, summarized below, while for both we assume a constant coefficient of variation:
+\begin{equation}
+  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq \sigma_\text{min} \\ \sigma_\text{min} & \qqtext{else} \end{cases}
+\end{equation}
+where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
+Though both approaches used to generate sequences \( \mu(t) \) use ighly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
+
+In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of parameters and discuss the processes which they can reasonably represent.
+
+\subsubsection{Synthetic ENSO Model\label{methods-enso}}
 
@@
 \end{equation}
-and a constant coefficient of variation
-\begin{equation}
-  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq \sigma_\text{min} \\ \sigma_\text{min} & \qqtext{else} \end{cases}
-\end{equation}
 where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
 
-Though this is a highly simplified model neglecting many important mechanisms, it enables us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
+\subsubsection{Markov Chain}
 
-In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of parameters and discuss the processes which they can reasonably represent.
+Alternatively, we consider a 2-state Markov chain with a constant transition probability which is the same for both states, such that the transition matrix is
+\begin{equation*}
+  P = \mqty[\pi & 1-\pi \\ 1-\pi & \pi]
+\end{equation*}
+This is first used to generate a sequence of states \(S(t)\).
+The value \( \mu(t) \) depends only on the \(S(t)\) and the time itself:
+\begin{equation*}
+  \mu(t) = \begin{cases}
+    \mu_{0,1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
+    \mu_{0,2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
+  \end{cases}
+\end{equation*}
+For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
 
-\subsection{Fitting Streamflow Sequences\label{sec:estimation}}
+\subsection{Estimating Future Streamflow Sequences\label{sec:estimation}}
+
+We do not, in general, know the true distribution of annual-maximum floods, and as discussed in \cref{sec:ffa} a plethora of approaches have been suggested in the literature.
+This paper does not propose new models or suggest a general ``best'' model; instead it aims to describe the total uncertainty in estimates using easily methods which have been well studied in the literature.
+
+In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 1000\footnote{Check this parameter} sequences of future streamflow are generated using Monte Carlo simulation.
+These estimates are then used to calculate \( \hat{p_T} \), and the estimates are evaluated as described in \cref{sec:methods-uncertainty}.
+Three models are assessed for estimating future streamflow; they are summarized in \cref{tab:model-fitting} and described in greater detail below.
+
+\begin{table}[bht]
+  \begin{center}
+    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
+      \toprule
+        Model Name & Description & Relevant Section \\
+      \midrule
+        stationary & \(Q \sim \GEV \qty(\mu, \sigma, \xi)\) & \cref{sec:method-stationary} \\
+        location trend & \(Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)\) & \cref{sec:method-trend} \\
+        HMM & log-normal Hidden Markov Model & \cref{sec:method-HMM} \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{Summary of models used for fitting\label{tab:model-fitting}}
+\end{table}
+
+\subsubsection{Stationary Model\label{sec:method-stationary}}
 
@@
 
-\subsection{Hidden Markov Model\label{sec:HMM}}
+\subsubsection{Trend Model\label{sec:method-trend}}
+
+As \cref{sec:method-stationary} except that
+\begin{equation*}
+  \mu = \mu_0 + \beta_\mu \qty(t - t_0)
+\end{equation*}
+and so the parameters \( \mu_0, \beta_\mu \) must be estimated rather than just \(\mu\).
+
+\subsubsection{Hidden Markov Model\label{sec:method-HMM}}
 
@@
 
-\begin{table}[b]
-  \begin{center}
-    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
-      \toprule
-        Model Name & Description & Relevant Section \\
-      \midrule
-        stationary & \(Q \sim \GEV \qty(\mu, \sigma, \xi)\) & \cref{sec:estimation} \\
-        location trend & \(Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)\) & \cref{sec:estimation} \\
-        HMM & log-normal Hidden Markov Model & \cref{sec:HMM} \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \caption{Summary of models used for fitting\label{tab:model-fitting}}
-\end{table}
-
-\subsection{Flood Estimation and Uncertainty}
+\subsection{Flood Estimation and Uncertainty\label{sec:methods-uncertainty}}
 
@@
 
-\section{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
+\section{Results}
+
+\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
 
@@
 
-\section{Stationary, Static Risk}
+\subsection{Stationary Case}
 
-Consider static risk and a stationary process, and partiuclarly implications of \(\qty{M, N}\).
-Much of this is already laid out in \citet{Lall1987}.
+\subsection{Trend Case}
 
-\section{Stationary, Static Risk}
+\section{Case Study}
 
-Consider static risk and a non-stationary process.
-Estimate using drift or nonstationary models using full \(N\) of the most recent \(N\) (\ie{} what Vogel is proposing).
-Illustrate via simulation for different signal to noise ratio of nonstationary terms and \(\qty{M, N}\).
-
-\section{Dynamic Risk}
-
-Consider dynamic risk, \ie{} updating and sequential decisions, a nonstationary generative process, and \(M\) and \(N\) considerations.
+We still need to find a good case study -- it can be a highly simplified one.
+We can come up with some numbers and a single streamflow sequence that has been observed.
 
@@
 
-Hopefully something to discuss?
-
 For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
-While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.\footnote{Send this to discussion}
+While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
 
@@
 
-\section{Notation}
+\section*{Notation}
 
 \begin{description}
-  \item[\(Q(t)\)] Annual-maximum flood series
-  \item[\(Q_T\)] Flood threshold
+  \item[\( Q(t) \)] Annual-maximum flood series
+  \item[\( Q_T \)] Flood threshold
+  \item[\( M \)] the project planning period
+  \item[\( N \)] the length of the observational record, in years
+  \item[\( p_T \)] the true probability that \( Q(t) \geq Q_T \) in the future \(M\)-year period
+  \item[\( \hat{p}_T \)] the estimated probability that \( Q(t) \geq Q_T \) in the future \(M\)-year period
 \end{description}

### 2018-02-16 writeup/Outline/main.tex
plot the stationary and trend sequences

@@
 
-We generate synthetic flood sequences from a simple conceptual model where the expected flood varies as a function of a single climate index plus time following the log-normal distribution:
+We generate synthetic flood sequences from two different simple conceptual models that represent chaotic dynamics and long memory.
+
+For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution:
 \begin{equation} \label{eq:lognormal}
@@
 \end{equation}
-The parameter \( \mu(t) \) is generated using two different approaches, summarized below, while for both we assume a constant coefficient of variation:
+where the parameters \( \mu(t), \sigma(t) \) are allowed to vary in time following the two approaches outlined below.
+For simplicitiy we assume a constant coefficient of variation:
 \begin{equation}
@@
 where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
-Though both approaches used to generate sequences \( \mu(t) \) use ighly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing many essential features of hydrological variability.
 
-In \cref{sec:sequence-realistic} we show sequences of streamflow generated by this model for some choices of parameters and discuss the processes which they can reasonably represent.
+Though both approaches used to generate sequences \( \mu(t) \) useh highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
+
+In \cref{sec:sequence-realistic} we show sequences of streamflow generated by both models for some choices of parameters and discuss the processes which they can reasonably represent.
 
@@
 ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a quasi-oscillatory manner with variability on the order of \SIrange{4}{6}{year} modulated by low-frequency oscillation \citep[see][for a comprehensive review]{Sarachik2010}.
-A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citep{Ramesh2017}.
+A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
+A \SI{1000}{year} subset of this data is shown in \cref{fig:enso-ts} and its spectrum in \cref{fig:enso-spectrum}.
 To create an annual time series, we average the October-December months to create a time series \( x(t) \) where \( t \) is taken in annual steps \(1, \ldots, \num{100000} \).
-The parameters of the distribution vary in time according to the sampled NINO3 index denoted \( x(t) \) as well as time itself, representing secular trends.
-First a starting point in the \SI{100000}{year} ENSO was randomly chosen, then \(\mu\) is determined by
+
+First a sequence of length \(M+N\) is randomly chosen from the \SI{100000}{year} ENSO sequence.
+Next, \(\mu\) is calculated following by
 \begin{equation}
@@
 
-\subsubsection{Markov Chain}
+\subsubsection{Markov Chain\label{methods-markov}}
 
-Alternatively, we consider a 2-state Markov chain with a constant transition probability which is the same for both states, such that the transition matrix is
+To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate \(\mu(t\).
+
+This Markov chain is given a constant probability of persistence which is the same for both states, such that the transition matrix is
 \begin{equation*}
@@
 For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
+We also assume \( \mu_{0, 1} \gg \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
 
@@
 
-Because we do not, in general, know the true distribution of annual-maximum floods, we compare a stationary and non-stationary parameterizations of the generalized extreme value (GEV) statistical distribution (eq.~\ref{eq:GEV}) for estimating the distribution of \(Q(t)\) for the future \(M\)-year period.
+We do not, in general, know the true distribution of annual-maximum floods, and there has been substantial discussion in the literature as to the choice of distribution to use for stationary flood frequency analysis.
+A reasonable choice is the generalized extreme value (GEV) statistical distribution:
 \begin{equation}
@@
       \qty[1 + \xi \frac{x - \mu}{\sigma}] ^ {(-\flatfrac{1}{\xi})-1} \exp \qty[-\qty(1+\xi \frac{x - \mu}{\sigma})^{\flatfrac{-1}{\xi}}], & \xi \neq 0 \\
-      \exp \qty[-\frac{x-\mu}{\sigma}] \exp \qty[- \exp \qty(- \frac{x-\mu}{\sigma})], & \xi = 0
+      \exp \qty[-\frac{x-\mu}{\sigma}] \exp \qty[- \exp \qty(- \frac{x-\mu}{\sigma})], & \xi = 0.
     \end{cases}
 \end{equation}
-Where \(\mu{}\) is the location parameter, \(\sigma{}\) is the shape parameter, and \(\xi{}\) is the scale parameter of the GEV distribution.
+We will refer to \(\mu{}\) is the location parameter, \(\sigma{}\) as the shape parameter, and \( \xi{} \) as the parameter.
+
 For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations.
-Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty.
-Computation is carried out in the \texttt{python} environment, with simulation using the \texttt{numpy} package \citep{vanderWalt2011} and fitting using the No-U-Turn Sampler (NUTS) \citep{Hoffman2014} as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
+Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
 Weak (uninformative) priors are chosen based on recommendations from \citet{Martins2000}; the exact parameterizations are described in following sections which describe specific experiments.
@@
 
-As \cref{sec:method-stationary} except that
+One approach for incorporating secular change into the GEV model is to add a time trend to the location parameter \citep[\ie{}][]{Bracken2017} such that
 \begin{equation*}
-  \mu = \mu_0 + \beta_\mu \qty(t - t_0)
+  \mu = \mu(t) = \mu_0 + \beta_\mu \qty(t - t_0).
 \end{equation*}
-and so the parameters \( \mu_0, \beta_\mu \) must be estimated rather than just \(\mu\).
+We fit the trend GEV model in exactly the same way as the stationary GEV model, except that the parameters  \( \mu_0, \beta_\mu \) must be estimated rather than just \( \mu \).
 
@@
 
-\subsection{Flood Estimation and Uncertainty\label{sec:methods-uncertainty}}
+\texttt{How much more do we need to say about HMM?}
 
-Finally, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
+\subsection{Reproducability}
+
+All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries.
+To carry implement the flood sequence generating and estimating functions, a new package called \texttt{floodsampling} was developed and is available at \url{github.com/jdossgollin/floodsampling}.
+The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.
+
+\section{Results}
+
+\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
+
+In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc{} and compare to some that have been identified in the literature.
+
+\begin{figure}[ht]
+  \caption{50 streamflow sequences from (Top) the ENSO-based model described in \cref{methods-enso} and (Bottom) the Markov chain modle described in \cref{methods-markov} with \(\gamma=0\), implying stationarity\label{fig:stationary-sequences}}
+\end{figure}
+
+\begin{figure}[ht]
+  \caption{50 streamflow sequences from (Top) the ENSO-based model described in \cref{methods-enso} and (Bottom) the Markov chain modle described in \cref{methods-markov}.
+To add secular variability, \(\gamma=0.01\) for the ENSO-based model, implying a linear trend, and for the Markov model \(\gamma_1 = 0.01\), implying a linear trend only in the ``wet'' state.\label{fig:trend-sequences}}
+\end{figure}
+
+\subsection{Stationary Case}
+
+\begin{figure}[ht]
+  \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\begin{figure}[ht]
+  \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\subsection{Trend Case}
+
+\begin{figure}[ht]
+  \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\begin{figure}[ht]
+  \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\section{Case Study}
+
+We still need to find a good case study; it can be a highly simplified one as long as there is a long streamflow record available and building levees is an option that has been proposed.
+If there's not an obvious candidate, we can use one of our models above.
+We can make some other strong assumptions
+\begin{itemize}
+  \item Risk exposure constant in time, which allows us to focus on the estimation rather than comprehensive risk manageement
+  \item Constant future discounting rate
+  \item Flood damages are linearly related to streamflow over some amount:
+  \begin{equation}
+    \mathcal{D}(t) = k \qty( Q(t) - Q_T)
+  \end{equation}
+\end{itemize}
+Then our two options are to build a levee, which we can assume to have a capital cost plus operation and maintenance cost which is proportional to the capital cost, which increases \(Q_T\) by some fixed amount, or to buy insurance.
+For the insurance, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
 Although \(Q_T\) is in general a decision variable, we treat it as fixed.
 Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that \(Q(t) \geq Q_T\) for \(t \in \vb{t}\), which is equivalent to \(p_T(t)\).
-
 For a particular year, the fair price for the insurance premium is \(\expectation \qty(p_T) C\).
@@
 
-In this paper we consider the price of a contract of \(M\) years to be simply the summed price of the contract for each individual year.
-We assume that the discount rate is equal to inflation (\ie{} the inflation-adjusted discount rate is zero) for simplicity.
-A more comprehensive pricing mechanism for the financial instrument could incorporate more or different information about the uncertainty in the estimate, and should consider the price of the \(M\) years as a portfolio (\ie{} should consider correlation between the different years).
-Such an approach, however, is not necessary for this conceptual study.
+Then at each point in time, calculate:
+\begin{itemize}
+  \item Expected cost plus damage per year for the levee using perfect knowledge of the next 101 years
+  \item Expected cost plus damage per year for the levee using MCMC projections of the next 101 years from all 3 models
+  \item Expected cost plus damage per year for a 5-year insurance contract using perfect knowledge of the next 5 years
+  \item Expected cost plus damage per year for a 5-year insurance contract using MCMC projections of the next 5 years from all 3 models
+\end{itemize}
 
-\section{Results}
-
-\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
-
-In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc{} and compare to some that have been identified in the literature.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/stationary_sequences.pdf}
-  \caption{50 simulated sequences created from the Cane-Zebiak model  \(N=50, M=50, t_0=0, \mu_0=6.5, \beta_\mu=0.5, \gamma=0, \alpha=0.075, \sigma_\text{min}=0.01\).
-49 of the sequences are shown in gray and a single randomly chosen sequence is highlighted in blue for emphasis.
-(L) shows the time series of log streamflow as a function of year, while (R) shows the distribution of 500 sequences over the historical and future periods.}\label{fig:stationary-sequences}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/trend_sequences.pdf}
-  \caption{As \cref{fig:stationary-sequences} except \(\gamma=0.01\).}
-\end{figure}
-
-\subsection{Stationary Case}
-
-\subsection{Trend Case}
-
-\section{Case Study}
-
-We still need to find a good case study -- it can be a highly simplified one.
-We can come up with some numbers and a single streamflow sequence that has been observed.
+Can add appropriate plot or plots.
 
@@
 While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
+
+From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
+How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length \(M\).
+WHat is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
+
+How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
 
@@
 
-\section*{Notation}
+% -----------------------------------------------------------------------------
+% END HERE
+% -----------------------------------------------------------------------------
+\appendix
+\printbibliography{}
+
+\section{Notation}
 
@@
 
-% -----------------------------------------------------------------------------
-% END HERE
-% -----------------------------------------------------------------------------
-\clearpage
-\printbibliography{}
+\section{Supplemental Figures}
+
+\begin{figure}
+  \caption{Insert here a 1000-year plot of the ENSO time series\label{fig:enso-ts}}
+\end{figure}
+
+\begin{figure}
+  \caption{Insert here a global wavelet or spectral plot of the ENSO time series\label{fig:enso-spectrum}}
+\end{figure}
 

### 2018-02-16 experiments/plot_stationary_sequences.py

@@ -8,63 +8,83 @@ import os
 import sys
 sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
 
-from floodsampling.streamflow import CZNINO3LN2
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 
-gen_fun = CZNINO3LN2(
-    N=50, M=50, t0=0, mu_0=6.5, beta_mu=0.5, gamma=0,
-    coeff_var=0.075, sigma_min=0.01, n_seq=500,
+time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
+plot_log = False # plot log streamflow?
+n_seq_plot = 50 # how many sequences to plot
+threshold = 2500
+if plot_log:
+    threshold = np.log(threshold)
+
+enso_gen = CZNINO3LN2(
+    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+    mu_0=6, beta_mu=0.5, gamma=0,
+    coeff_var=0.1, sigma_min=0.01, 
+    n_seq=time_param['n_seq'],
 )
-sequences = np.log(gen_fun.get_data())
+markov_gen = TwoStateSymmetricMarkovLN2(
+    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+    mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
+    coeff_var = 0.1, sigma_min = 0.01,
+    n_seq=time_param['n_seq'],
+)
+enso_seq = enso_gen.get_data()
+markov_seq = markov_gen.get_data()
 
-n_seq_plot = 50
-threshold = 5000
+if plot_log:
+    enso_seq = np.log(enso_seq)
+    markov_seq = np.log(markov_seq)
 
-# PLOT LOG STREAMFLOW
-fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
-    gridspec_kw={'width_ratios': [4, 1]})
-ax = axes[0]
+# Setup the figure
+fig, axes = plt.subplots(
+    nrows=2, ncols=2, sharey=True, figsize=(11, 5),
+    gridspec_kw={'width_ratios': [4, 1]},
+    sharex='col'
+)
+
+ax = axes[0, 0]
 for i in np.arange(n_seq_plot-1):
-    sequences.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
[…]

### 2018-02-16 experiments/plot_trend_sequences.py

@@ -8,64 +8,83 @@ import os
 import sys
 sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
 
-from floodsampling.streamflow import CZNINO3LN2
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 
-gen_fun = CZNINO3LN2(
-    N=50, M=50, t0=0, mu_0=6.5, beta_mu=0.5, gamma=0.01,
-    coeff_var=0.075, sigma_min=0.01, n_seq=500,
+time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
+plot_log = False # plot log streamflow?
+n_seq_plot = 50 # how many sequences to plot
+threshold = 2500
+if plot_log:
+    threshold = np.log(threshold)
+
+enso_gen = CZNINO3LN2(
+    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+    mu_0=6, beta_mu=0.5, gamma=0.01,
+    coeff_var=0.1, sigma_min=0.01, 
+    n_seq=time_param['n_seq'],
 )
-sequences = np.log(gen_fun.get_data())
+markov_gen = TwoStateSymmetricMarkovLN2(
+    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+    mu_1=6.75, mu_0 = 6, gamma_1=0.01, gamma_2=0, pi=0.9,
+    coeff_var = 0.1, sigma_min = 0.01,
+    n_seq=time_param['n_seq'],
+)
+enso_seq = enso_gen.get_data()
+markov_seq = markov_gen.get_data()
 
-n_seq_plot = 50
-threshold = 5000
+if plot_log:
+    enso_seq = np.log(enso_seq)
+    markov_seq = np.log(markov_seq)
 
-# PLOT LOG STREAMFLOW
-fig, axes = plt.subplots(nrows=1, ncols=2, sharey=True, figsize=(10, 4),
-    gridspec_kw={'width_ratios': [4, 1]})
-ax = axes[0]
+# Setup the figure
+fig, axes = plt.subplots(
+    nrows=2, ncols=2, sharey=True, figsize=(11, 5),
+    gridspec_kw={'width_ratios': [4, 1]},
+    sharex='col'
+)
+
+ax = axes[0, 0]
 for i in np.arange(n_seq_plot-1):
-    sequences.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
[…]

### 2018-02-16 experiments/stationary.py
code for running and plotting

@@ -1,79 +1,0 @@
-"""Stationary
-"""
-import numpy as np
-import xarray as xr
-import seaborn as sns
-import pandas as pd
-import matplotlib.pyplot as plt
-
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
-from floodsampling.streamflow import CZNINO3LN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-
-# Define Parameters
-N_try = np.array([10, 30, 50, 150])
-M_try = np.array([5, 10, 30, 50])
-n_seq = 1000
-n_sim = 1000
-threshold = 5000
-
-def some_function(x: int, y: int) -> None:
-    "Some cool docstring"
-    return x * y
-
-# Initialize the bias and variance
-model_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
-empty = np.zeros(shape=(N_try.size, M_try.size, model_names.size))
-empty = xr.DataArray(
-    data=empty,
-    coords={'N': N_try, 'M': M_try, 'model': model_names},
-    dims=['N', 'M', 'model']
-)
-fit = xr.Dataset({'bias': empty, 'variance': empty}).to_dataframe()
-
-for N in N_try:
-    # We don't need to re-fit everything every time we increase M.
-    # Just fit on the largest M then sub-set the data we want.
-    # TODO: Add a M argument to the get_bias_variance
-    gen_fun = CZNINO3LN2(
-        N=N, M=M_try.max(), t0=0, mu_0=6.5, beta_mu=0.75, gamma=0,
-        coeff_var=0.075, sigma_min=0.01, n_seq=n_seq,
-    )
-    stationary_fit = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-    trend_fit = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-    hmm_fit = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
-    for M in M_try:
-        # Loop through to get the data
-        obs_dat = stationary_fit.get_data('future').sel(year = slice(1, M))
-        stationary_dat = stationary_fit.get_data('future').sel(year = slice(1, M))
[…]

### 2018-02-16 run.sh

@@ -1,0 +1,8 @@
+python src/plot_stationary_sequences.py
+python src/plot_trend_sequences.py
+
+python src/calc_stationary.py
+python src/calc_trend.py
+
+python src/plot_stationary.py
+python src/plot_trend.py

### 2018-02-16 src/calc_stationary.py

@@ -1,0 +1,91 @@
+"""Stationary
+"""
+import numpy as np
+import xarray as xr
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+# Define Parameters
+N_try = np.array([10, 20, 30, 40, 50, 75, 100, 150])
+M_try = np.array([3, 5, 10, 20, 30, 50, 75])
+n_seq = 100 # PITIFULLY FEW!
+n_sim = 1000 # PITIFULLY SMALL!
+threshold = 5000
+figsize = (9, 6)
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+# Indexing: [N, gen_fun, fit_fun, M] THIS IS IMPORTANT
+bias = np.zeros(shape=(N_try.size, gen_names.size, fit_names.size, M_try.size)) 
+variance = np.zeros(shape=(N_try.size, gen_names.size, fit_names.size, M_try.size))
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M_try.max(), t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M_try.max(), t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
[…]

### 2018-02-16 src/calc_trend.py

@@ -1,0 +1,91 @@
+"""Stationary
+"""
+import numpy as np
+import xarray as xr
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+# Define Parameters
+N_try = np.array([10, 20, 30, 40, 50, 75, 100, 150])
+M_try = np.array([3, 5, 10, 20, 30, 50, 75])
+n_seq = 100 # PITIFULLY FEW!
+n_sim = 1000 # PITIFULLY SMALL!
+threshold = 5000
+figsize = (9, 6)
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+# Indexing: [N, gen_fun, fit_fun, M] THIS IS IMPORTANT
+bias = np.zeros(shape=(N_try.size, gen_names.size, fit_names.size, M_try.size)) 
+variance = np.zeros(shape=(N_try.size, gen_names.size, fit_names.size, M_try.size))
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M_try.max(), t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0.01,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M_try.max(), t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0.01, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
[…]

### 2018-02-16 src/plot_stationary.py

@@ -1,0 +1,50 @@
+import xarray as xr
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+fit = xr.open_dataset('data/stationary.nc')
+fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
+
+# PLOT THE BIAS
+fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
+for col,f in enumerate(fit['fit_fun'].values):
+    for row,g in enumerate(fit['gen_fun'].values):
+        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
+        ax=axes[row, col]
+        sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar=False, linewidths=0.1)
+        if row == 0:
+            ax.set_title(f)
+        if col  == (fit['fit_fun'].size - 1):
+            ax.set_ylabel(g)
+            ax.yaxis.set_label_position('right')
+
+fig.tight_layout()
+fig.subplots_adjust(right=0.85)
+cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
+sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
+
+fig.savefig('figs/stationary_bias.pdf')
+
+
+
+# PLOT THE VARIANCE
+fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
+for col,f in enumerate(fit['fit_fun'].values):
+    for row,g in enumerate(fit['gen_fun'].values):
+        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
+        ax=axes[row, col]
+        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.02, cmap='viridis', cbar=False, linewidths=0.1)
+        if row == 0:
+            ax.set_title(f)
+        if col  == (fit['fit_fun'].size - 1):
[…]

### 2018-02-16 src/plot_trend.py

@@ -1,0 +1,50 @@
+import xarray as xr
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+fit = xr.open_dataset('data/trend.nc')
+fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
+
+# PLOT THE BIAS
+fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
+for col,f in enumerate(fit['fit_fun'].values):
+    for row,g in enumerate(fit['gen_fun'].values):
+        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
+        ax=axes[row, col]
+        sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar=False, linewidths=0.1)
+        if row == 0:
+            ax.set_title(f)
+        if col  == (fit['fit_fun'].size - 1):
+            ax.set_ylabel(g)
+            ax.yaxis.set_label_position('right')
+
+fig.tight_layout()
+fig.subplots_adjust(right=0.85)
+cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
+sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
+
+fig.savefig('figs/trend_bias.pdf')
+
+
+
+# PLOT THE VARIANCE
+fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
+for col,f in enumerate(fit['fit_fun'].values):
+    for row,g in enumerate(fit['gen_fun'].values):
+        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
+        ax=axes[row, col]
+        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.02, cmap='viridis', cbar=False, linewidths=0.1)
+        if row == 0:
+            ax.set_title(f)
+        if col  == (fit['fit_fun'].size - 1):
[…]

### 2018-02-17 run.sh
plot some examples

@@ -1,5 +1,6 @@
 python src/plot_stationary_sequences.py
 python src/plot_trend_sequences.py
+python src/plot_enso.py
 
 python src/calc_stationary.py
 python src/calc_trend.py

### 2018-02-17 src/calc_trend.py

@@ -54,7 +54,7 @@ def get_fit_fun(index, gen_fun):
     if index == 0:
         fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
     elif index == 1:
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
+        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
     elif index == 2:
         fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
     else:
@@ -71,10 +71,10 @@ for n, N in enumerate(N_try):
                 # Loop through to get the data
                 gen_dat = gen_fun.get_data('future').sel(year = slice(1, M))
                 fit_dat = fit_fun.get_data('future').sel(year = slice(1, M))
-                p_T = np.nanmean(gen_dat > threshold)
-                p_T_hat = np.nanmean(fit_dat > threshold)
-                bias[n, g, f, m] = p_T_hat - p_T
-                variance[n, g, f, m] = np.nanvar(p_T_hat)
+                exceed_gen = gen_dat > threshold
+                exceed_fit = fit_dat > threshold
+                bias[n, g, f, m] = exceed_fit.mean() - exceed_gen.mean()
+                variance[n, g, f, m] = exceed_fit.std(dim='sim').mean()
 
 bias = xr.DataArray(
     bias, 

### 2018-02-17 src/plot_enso.py

@@ -1,0 +1,7 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+enso = pd.read_csv('~/Documents/GitHub/floodsampling/floodsampling/data/ramesh2017.csv', index_col='year')
+
+enso.loc[10000:12500].plot(figsize=(10, 5), linewidth=0.5)
+plt.savefig('figs/enso.pdf', bbox_inches='tight')

### 2018-02-17 src/plot_example_stationary_long.py

@@ -1,0 +1,74 @@
+import numpy as np
+import xarray as xr
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+M = 150
+N = 150
+n_seq = 1
+n_sim = 100
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
+    elif index == 1:
+        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
+    elif index == 2:
+        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
+    else:
+        raise ValueError('Invalid Index')
+    return fit_fun
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
+for g, gen_fun_name in enumerate(gen_names):
+    gen_fun = get_gen_fun(g)
+    for f, fit_fun_name in enumerate(fit_names):
[…]

### 2018-02-17 src/plot_example_stationary_short.py

@@ -1,0 +1,74 @@
+import numpy as np
+import xarray as xr
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+M = 20
+N = 30
+n_seq = 1
+n_sim = 100
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
+    elif index == 1:
+        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
+    elif index == 2:
+        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
+    else:
+        raise ValueError('Invalid Index')
+    return fit_fun
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
+for g, gen_fun_name in enumerate(gen_names):
+    gen_fun = get_gen_fun(g)
+    for f, fit_fun_name in enumerate(fit_names):
[…]

### 2018-02-17 src/plot_example_trend_long.py

@@ -1,0 +1,75 @@
+import numpy as np
+import xarray as xr
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+M = 150
+N = 150
+n_seq = 1
+n_sim = 100
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0.02,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0.02, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
+    elif index == 1:
+        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
+    elif index == 2:
+        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
+    else:
+        raise ValueError('Invalid Index')
+    return fit_fun
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
+for g, gen_fun_name in enumerate(gen_names):
+    gen_fun = get_gen_fun(g)
[…]

### 2018-02-17 src/plot_example_trend_short.py

@@ -1,0 +1,75 @@
+import numpy as np
+import xarray as xr
+import pandas as pd
+import matplotlib.pyplot as plt
+
+import os
+import sys
+sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+
+M = 20
+N = 30
+n_seq = 1
+n_sim = 100
+
+def get_gen_fun(index):
+    """Get a generating function as a function of the index 0 or 1
+    """
+    if index == 0:
+        gen_fun = CZNINO3LN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_0=6, beta_mu=0.5, gamma=0.02,
+            coeff_var=0.1, sigma_min=0.01
+        )
+    elif index == 1:
+        gen_fun = TwoStateSymmetricMarkovLN2(
+            N=N, M=M, t0=0, n_seq=n_seq,
+            mu_1=6.75, mu_0 = 6, gamma_1=0.02, gamma_2=0, pi=0.9,
+            coeff_var = 0.1, sigma_min = 0.01
+        )
+    else:
+        raise ValueError('Invalid Index')
+    return gen_fun
+
+def get_fit_fun(index, gen_fun):
+    """Get a fit function as a function of the index 0 or 1 or 2
+    """
+    if index == 0:
+        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
+    elif index == 1:
+        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
+    elif index == 2:
+        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
+    else:
+        raise ValueError('Invalid Index')
+    return fit_fun
+
+# Initialize the bias and variance
+gen_names = np.array(['enso', 'markov'])
+fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+
+fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
+for g, gen_fun_name in enumerate(gen_names):
+    gen_fun = get_gen_fun(g)
[…]

### 2018-02-17 src/plot_stationary.py

@@ -13,7 +13,7 @@ for col,f in enumerate(fit['fit_fun'].values):
     for row,g in enumerate(fit['gen_fun'].values):
         sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
         ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar=False, linewidths=0.1)
+        sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
         if row == 0:
             ax.set_title(f)
         if col  == (fit['fit_fun'].size - 1):
@@ -23,7 +23,7 @@ for col,f in enumerate(fit['fit_fun'].values):
 fig.tight_layout()
 fig.subplots_adjust(right=0.85)
 cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
+sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
 
 fig.savefig('figs/stationary_bias.pdf')
 
@@ -35,7 +35,7 @@ for col,f in enumerate(fit['fit_fun'].values):
     for row,g in enumerate(fit['gen_fun'].values):
         sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
         ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.02, cmap='viridis', cbar=False, linewidths=0.1)
+        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.15, cmap='viridis', cbar=False, linewidths=0.1)
         if row == 0:
             ax.set_title(f)
         if col  == (fit['fit_fun'].size - 1):
@@ -45,6 +45,6 @@ for col,f in enumerate(fit['fit_fun'].values):
 fig.tight_layout()
 fig.subplots_adjust(right=0.85)
[…]

### 2018-02-17 src/plot_stationary_sequences.py

@@ -87,4 +87,4 @@ for ax in axes.flat:
 
 
 fig.tight_layout()
-plt.savefig('figs/stationary_sequences.pdf')
+plt.savefig('figs/stationary_sequences.pdf', bbox_inches='tight')

### 2018-02-17 src/plot_trend.py

@@ -13,7 +13,7 @@ for col,f in enumerate(fit['fit_fun'].values):
     for row,g in enumerate(fit['gen_fun'].values):
         sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
         ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar=False, linewidths=0.1)
+        sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
         if row == 0:
             ax.set_title(f)
         if col  == (fit['fit_fun'].size - 1):
@@ -23,7 +23,7 @@ for col,f in enumerate(fit['fit_fun'].values):
 fig.tight_layout()
 fig.subplots_adjust(right=0.85)
 cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-sns.heatmap(sub, ax=ax, vmin=-0.15, vmax=0.15, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
+sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'Bias'})
 
 fig.savefig('figs/trend_bias.pdf')
 
@@ -35,7 +35,7 @@ for col,f in enumerate(fit['fit_fun'].values):
     for row,g in enumerate(fit['gen_fun'].values):
         sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
         ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.02, cmap='viridis', cbar=False, linewidths=0.1)
+        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.15, cmap='viridis', cbar=False, linewidths=0.1)
         if row == 0:
             ax.set_title(f)
         if col  == (fit['fit_fun'].size - 1):
@@ -45,6 +45,7 @@ for col,f in enumerate(fit['fit_fun'].values):
 fig.tight_layout()
 fig.subplots_adjust(right=0.85)
[…]

### 2018-02-17 src/plot_trend_sequences.py

@@ -87,4 +87,4 @@ for ax in axes.flat:
 
 
 fig.tight_layout()
-plt.savefig('figs/trend_sequences.pdf')
+plt.savefig('figs/trend_sequences.pdf', bbox_inches='tight')

### 2018-02-18 writeup/Outline/main.tex
plot some examples
continue writing

@@
 
-Classical methods for design of infrastructure first specify a return period \(T\) (\ie{} 100 years), and then design a project that protects against this \(T\)-year event.
-The mathematical problem of interest then becomes estimation of the probability distribution \( f(p_T) \), where \( p_T \) is binary event that the threshold \( Q_T \) is exceeded by the annual-maxiumum flood series \( Q(t) \):
-\begin{equation*}
-  p_T(t) \equiv \probability \qty[ Q(t) \geq Q_T] \qqtext{for} t = \qty[t_0 + 1, \ldots, t_0 + M]
-\end{equation*}
-where \( t_0 \) is the first year forecast and \( M \) is the project operation period.
-Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \( Q_T \) is used for design \citep{Stedinger1997}\footnote{Manu has mentioned a Vogel198x citation but I haven't successfully found it.}.
+At present, most approaches to nonstationary flood risk management, or climate change adaptation more broadly, have framed the problem first as one of detecting historical change or modeling future change, and then assessing the implications of this change for future infrastructure.
+Though more recent studies have pointed out that the paradigm of requiring proof of the existence of trends before adapting to nonstationarity does not lead to optimal decisionmaking \citep{Vogel2013,Rosner2014}, such discussion has also focused on linear time trends and on choosing whether or not to implement a particular adaptation policy with a fixed planning period.
+In contrast, public and private actors seeking to adapt to changing flood or climate risk typically have access to a suite of options including large flood control infrastructure, insurance contracts and other financial instruments, vulnerability reduction (\ie{} disaster response training), and exposure reduction (\ie{} moving infrastructure out of the floodplain).
+In this context, understanding under what conditions a large, permanent project or small, easily modified adaptation measure are optimal becomes an important question.
+
+In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period \(M\) and the length of the available data record \(N\).
+This project planning period can be interpreted as the design life of a physical structure, or as the contract length of a financial instrument.
+Our aim is not to identify the ``best'' model, but rather to build some theory as to how the underlying nature of the natural variability and nonstationarity and the choice of model used to model future flood risk map onto estimation bias and variance in the context of many different \(M\) and \(N\).
+This estimation bias and variance may map directly onto the price of an insurance contract\footnote{cite} or the possibility of over/under-design of a flood control structure.
+Though we focus on flood risk, this type of analysis may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricane, wind, \etc{}.
 
@@
 Floods may be associated with a single rainfall event or be composed of a sequence of multiple, distinct events over a week or a season, such that there is pre-conditioning of the land surface soil moisture and natural and man-made stores that leads to a high flooding potential.
-
-In large river basins, Lagrangian tracking studies \citep{Gimeno2010} indicate that river flooding often requires the large-scale transport and convergence of moisture  which are driven by specific large-scale climate mechanisms \citep{Dacre2014,Gimeno2014,Lu2013,Dirmeyer2010}.\footnote{Farnham et al}
+In large river basins, Lagrangian tracking studies \citep{Gimeno2010} indicate that river flooding often requires the large-scale transport and convergence of moisture  which are driven by specific large-scale climate mechåanisms \citep{Dacre2014,Gimeno2014,Lu2013,Dirmeyer2010}.\footnote{Farnham et al}
 For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent continental-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
-Other case studies \citep{Grams2014,Stadtherr2016}\footnote{Add my work} and theoretical analyses \citep{Hoskins2015,Screen2014,Coumou2014} support the hypothesis that specific climate mechanisms such as blocking or jet sinuosity  can lead to persistent, intense rainfall.
+Other case studies of  major floods in the Balkans \citep{Stadtherr2016}, Central Europe \citep{Grams2014}, and Paraguay\footnote{Add my work}, and theoretical analyses \citep{Hoskins2015,Screen2014,Coumou2014}, support the hypothesis that persistent, intense rainfall in a particular place is associated with specific and potentially identifiable climate mechanisms.
 
 At the seasonal and sub-seasonal scales, surface temperature boundary conditions are well known to significantly change the mean, variance, persistence, recurrence and spatial structure of atmospheric circulation patterns, as well as moisture availability.
-Strong links between climate phenomena such as El Ni\~no-Southern Oscillation (ENSO) and flood risk have been well established in both the water resources \citep[\ie{}][]{Ward2014} and atmospheric science \citep{Tedeschi2013} literatures.
-
 On longer time scales, changes to the physical processes that govern hydroclimate systems, known generically as nonstationarity \citep{Milly2008} can occur due to climate change, river modification, and land use change \citep{Merz2014}.
@@
 \subsection{Nonstationary Flood Frequency Analysis\label{sec:ffa}}
+
+Classical methods for design of infrastructure first specify a return period \(T\) (\ie{} 100 years), and then design a project that protects against this \(T\)-year event.
+The mathematical problem of interest then becomes estimation of the probability distribution \( f(p_T) \), where \( p_T \) is binary event that the threshold \( Q_T \) is exceeded by the annual-maxiumum flood series \( Q(t) \):
+\begin{equation}
+  p_T(t) \equiv \probability \qty[ Q(t) \geq Q_T] \qqtext{for} t = \qty[t_0 + 1, \ldots, t_0 + M]
+\end{equation}
+where \( t_0 \) is the first year forecast and \( M \) is the project operation period.
+Averaging over the \(M\)-year period, we can write for convenience
+\begin{equation}
+  p_T = \frac{1}{M} \sum_{t=1}^{M} p_T(t)
+\end{equation}
+Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \( Q_T \) is used for design \citep{Stedinger1997}\footnote{Manu has mentioned a Vogel198x citation but I haven't successfully found it.}.
 
@@
 Alternatively, sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
-Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}
-\texttt{need some citations for sequential decisions}.
+Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}.\footnote{need some citations for sequential decisions}
 Given that, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times, models which can capture this uncertainty may be particularly advantageous for sequential models.
@@
 
-\subsection{Generating Synthetic Flood Sequences}
+\subsection{Generating Synthetic Flood Sequences\label{sec:methods-generating}}
 
@@
 where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
+The two models are summarized in \cref{tab:model-generating}.
 
@@
 
-\subsubsection{Synthetic ENSO Model\label{methods-enso}}
+\begin{table}[bht]
+  \begin{center}
+    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
+      \toprule
+        Model Name & Description & Relevant Section \\
+      \midrule
+        NINO3 & \(\mu(t)\) proportional to synthetic NINO3 time series & \cref{sec:methods-enso} \\
+        Markov & \(\mu(t)\) depends only on sequence of latent states generated by Markov chain & \cref{sec:methods-markov} \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{Summary of models used for generating synthetic flood sequences\label{tab:model-generating}}
+\end{table}
+
+\subsubsection{Synthetic ENSO Model\label{sec:methods-enso}}
 
@@
 A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
-A \SI{1000}{year} subset of this data is shown in \cref{fig:enso-ts} and its spectrum in \cref{fig:enso-spectrum}.
+A \SI{1000}{year} subset of this data is shown in \cref{fig:enso-ts}.
 To create an annual time series, we average the October-December months to create a time series \( x(t) \) where \( t \) is taken in annual steps \(1, \ldots, \num{100000} \).
@@
 
-\subsubsection{Markov Chain\label{methods-markov}}
+\subsubsection{Markov Chain\label{sec:methods-markov}}
 
@@
   \begin{center}
-    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
+    \begin{tabular}{L{1.25in} L{3.0in} L{1.25in}}
       \toprule
@@
       \midrule
-        stationary & \(Q \sim \GEV \qty(\mu, \sigma, \xi)\) & \cref{sec:method-stationary} \\
-        location trend & \(Q \sim \GEV \qty(\mu_0 + \beta_\mu \qty(t-t_0), \sigma, \xi)\) & \cref{sec:method-trend} \\
-        HMM & log-normal Hidden Markov Model & \cref{sec:method-HMM} \\
+        \texttt{ln2\_stationary} & \( \log Q \sim N \qty(\mu, \sigma) \) & \cref{sec:method-stationary} \\
+        \texttt{ln2\_trend} & \(\log Q \sim N \qty(\mu_0 + \beta_\mu \qty[t - t_0], \sigma_0 + \beta_\sigma \qty[t - t_0] ) \) & \cref{sec:method-trend} \\
+        \texttt{HMM} & Hidden Markov Model & \cref{sec:method-HMM} \\
       \bottomrule
@@
 We do not, in general, know the true distribution of annual-maximum floods, and there has been substantial discussion in the literature as to the choice of distribution to use for stationary flood frequency analysis.
-A reasonable choice is the generalized extreme value (GEV) statistical distribution:
-\begin{equation}
-  \label{eq:GEV}
-  f(x \big| \mu, \sigma, \xi) = \frac{1}{\sigma}
-    \begin{cases}
-      \qty[1 + \xi \frac{x - \mu}{\sigma}] ^ {(-\flatfrac{1}{\xi})-1} \exp \qty[-\qty(1+\xi \frac{x - \mu}{\sigma})^{\flatfrac{-1}{\xi}}], & \xi \neq 0 \\
-      \exp \qty[-\frac{x-\mu}{\sigma}] \exp \qty[- \exp \qty(- \frac{x-\mu}{\sigma})], & \xi = 0.
-    \end{cases}
-\end{equation}
-We will refer to \(\mu{}\) is the location parameter, \(\sigma{}\) as the shape parameter, and \( \xi{} \) as the parameter.
+We consider the log-normal model (\cref{eq:lognormal}) for stationary flood distribution; while this model is less flexible than three-parameter models such as the GEV or LP3 \citep{Vogel1996}, it is computationally convenient and easy to understand.
+More importantly, since we know the true distribution of these synthetically generated flood, using the log-normal for both generating and fitting supports interpreting results as a lower bound on total uncertainty.
+In the stationary context, we assume \(\mu(t) = \mu\) and \(\sigma(t)=\sigma\).
 
@@
 Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
-Weak (uninformative) priors are chosen based on recommendations from \citet{Martins2000}; the exact parameterizations are described in following sections which describe specific experiments.
+Weak (uninformative) priors are chosen; in the context of a synthetic data experiment using informative priors seems inappropriate but other studies\footnote{cite} have used scaling information and regional information to inform prior distributions and reduce posterior variance.
 
@@
 
-One approach for incorporating secular change into the GEV model is to add a time trend to the location parameter \citep[\ie{}][]{Bracken2017} such that
-\begin{equation*}
-  \mu = \mu(t) = \mu_0 + \beta_\mu \qty(t - t_0).
-\end{equation*}
-We fit the trend GEV model in exactly the same way as the stationary GEV model, except that the parameters  \( \mu_0, \beta_\mu \) must be estimated rather than just \( \mu \).
+Many studies\footnote{cite} have used time parameterizations for fitting.
+We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \):
+\begin{align}
+  \mu(t) = \mu_0 + \beta_\mu \qty(t - t_0) \\
+  \sigma(t) = \sigma_0 + \beta_\sigma \qty(t - t_0)
+\end{align}
+
+As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors.
+The influence of priors in this case is stronger than in the stationary case, and a reasonable choice of priors might strongly favor \(\beta_mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
 
@@
 
+\subsection{Evaluating Fitting Models}
+
+For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
+For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
+\(K\) posterior simulations of future rainfall are then drawn from the fitted model.
+We then perform a standard decomposition of uncertainty:
+\begin{equation}
+  \mathbb{E} \qty[ \qty( \hat{p}_T - p_T )^2 ] = \qty( \mathbb{E} \qty[ \hat{p}_T ] - p_T )^2 + \mathbb{E} \qty[ \qty( \hat{p}_T - \mathbb{E} \qty[ \hat{p}_T ] )^2]
+\end{equation}
+where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
+
 \subsection{Reproducability}
 
-All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries.
+All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
 To carry implement the flood sequence generating and estimating functions, a new package called \texttt{floodsampling} was developed and is available at \url{github.com/jdossgollin/floodsampling}.
@@
 
+We begin by showing the types of synthetic streamflow sequence that are by the two models discussed \cref{sec:methods-generating}.
+We then examine how different statistical models for estimating future flood sequences perform under stationary conditions and when a secular trend is introduced.
+
 \subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
 
-In this section we discuss the sequences of streamflow that are generated and look at their ACF, PACF, distribution, wavelet spectrum, \etc{} and compare to some that have been identified in the literature.
+The models used to generate synthetic flood sequences used simplistic dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
 
-\begin{figure}[ht]
-  \caption{50 streamflow sequences from (Top) the ENSO-based model described in \cref{methods-enso} and (Bottom) the Markov chain modle described in \cref{methods-markov} with \(\gamma=0\), implying stationarity\label{fig:stationary-sequences}}
+To verify that these synthetic streamflow sequences sufficiently represent these behaviors, we plot several sequences.
+The ENSO model used a long run of the Cane-Zebiak model \citep{Zebiak1987,Ramesh2017} to determine \(\mu(t)\); a \SI{2500}{year} sub-set of the annualized time series is shown in \cref{fig:enso-ts}.
+As described in the literature,\footnote{citations, and maybe some paleo stuff as well} this model shows quasi-periodic oscillations in the \SIrange{3}{7}{year} band but with substantial low-frequency modulation of extreme activity.
+The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
+
+\Cref{fig:stationary-sequences} shows 50 streamflow sequences generated with \(M=50\) and \(N=100\).
+The ENSO sequences were generated with \(\gamma=0, \beta_\mu=0.5, \mu_0=6\), which gives a strong (but still noisy) relationship between the NINO3 index \(x(t)\)) and the annual-maximum streamflow \(Q(t)\).
+The Markov sequences were generated with \(\mu_1 = 6.55, \mu_2=5.7, \gamma_1=0, \gamma_2=0, \pi=0.9\), which gives a large difference between the two states.
+For both, the coefficient of variation \(\alpha=0.1\) and the minimum value of sigma \(\sigma_\text{min}=0.01\).
+As the time series plot shows, both generating models exhit substantial persistence in flood probability from one year to the next, though the ENSO model shows negative autocorrelation at lag 1 while the Markov model has positive autocorrelation at lag 1 (not shown).\footnote{Need to include this?}
+
+\begin{figure}[b]
+  \includegraphics[width=\textwidth]{../../figs/stationary_sequences.pdf}  
+  \caption{
+    50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-enso} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
+    Both models use a stationary parameterization.
+    49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
+    The right hand panel shows the distribution over the historical and future periods for 500 sequences from the same model, visualized with a histogram and kernel density estimate.
+    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.014133333333333333} by the NINO3 sequences and \num{0.013546666666666667} by the Markov sequences.\label{fig:stationary-sequences}
+  }
 \end{figure}
 
-\begin{figure}[ht]
-  \caption{50 streamflow sequences from (Top) the ENSO-based model described in \cref{methods-enso} and (Bottom) the Markov chain modle described in \cref{methods-markov}.
-To add secular variability, \(\gamma=0.01\) for the ENSO-based model, implying a linear trend, and for the Markov model \(\gamma_1 = 0.01\), implying a linear trend only in the ``wet'' state.\label{fig:trend-sequences}}
+To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
+For the ENSO time series, this mean setting \(\gamma=0.02\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.02\).
+The resulting sequences are shown in \cref{fig:trend-sequences}.
+The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
+\begin{figure}[b]
+  \includegraphics[width=\textwidth]{../../figs/trend_sequences.pdf}
+  \caption{
+    As \cref{fig:stationary-sequences} but a trend term has been added.
+    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.023306666666666667} by the NINO3 sequences and \num{0.024493333333333332} by the Markov sequences.\label{fig:trend-sequences}
+  }
 \end{figure}
@@
 \begin{figure}[ht]
+  \includegraphics[width=\textwidth]{../../figs/stationary_bias.pdf}
   \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
+  \includegraphics[width=\textwidth]{../../figs/stationary_variance.pdf}
   \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
+  \includegraphics[width=\textwidth]{../../figs/trend_bias.pdf}
   \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
+  \includegraphics[width=\textwidth]{../../figs/trend_variance.pdf}
   \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
   \item[\( Q(t) \)] Annual-maximum flood series
+  \item[\( T \)] Return period, in years
   \item[\( Q_T \)] Flood threshold
@@
   \item[\( N \)] the length of the observational record, in years
+  \item[\( x(t) \)] the annualized time series of the NINO3 index
   \item[\( p_T \)] the true probability that \( Q(t) \geq Q_T \) in the future \(M\)-year period
@@
 \begin{figure}
-  \caption{Insert here a 1000-year plot of the ENSO time series\label{fig:enso-ts}}
+  \includegraphics[width=\textwidth]{../../figs/enso.pdf}
+  \caption{
+    A \SI{2500}{year} sub-set of the synthetic time series described in \cref{sec:methods-enso}.\label{fig:enso-ts}
+  }
 \end{figure}
@@
 \begin{figure}
-  \caption{Insert here a global wavelet or spectral plot of the ENSO time series\label{fig:enso-spectrum}}
+  \includegraphics[width=\textwidth]{../../figs/example_stationary_long.pdf}
+  \caption{
+    A single sequence of streamflow generated by each of two generating functions (columns), each fit to three different fitting models (rows) under a stationary case (\(\gamma=0\)).
+Shown is a long series \(M=N=150\).\label{fig:stationary-long-eg}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{../../figs/example_stationary_short.pdf}
+  \caption{
+    As \cref{fig:stationary-long-eg} except \(N=30\).\label{fig:stationary-short-eg}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{../../figs/example_trend_long.pdf}
+  \caption{
+    A single sequence of streamflow generated by each of two generating functions (columns), each fit to three different fitting models (rows) under a trend case.
+For the ENSO model \(\gamma=0.02\) and for the Markov model \(\gamma_1 = 0.02\) and \(\gamma_2=0\).
+Shown is a long series \(M=N=150\).\label{fig:trend-long-eg}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{../../figs/example_trend_short.pdf}
+  \caption{
+    As \cref{fig:trend-long-eg} except \(N=30\).\label{fig:trend-short-eg}
+  }
 \end{figure}

### 2018-02-18 src/plot_example_stationary_short.py

@@ -10,7 +10,7 @@ sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 
-M = 20
+M = 150
 N = 30
 n_seq = 1
 n_sim = 100

### 2018-02-18 src/plot_stationary_sequences.py

@@ -25,7 +25,7 @@ enso_gen = CZNINO3LN2(
 )
 markov_gen = TwoStateSymmetricMarkovLN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
+    mu_1=6.55, mu_0=5.7, gamma_1=0, gamma_2=0, pi=0.9,
     coeff_var = 0.1, sigma_min = 0.01,
     n_seq=time_param['n_seq'],
 )
@@ -45,11 +45,11 @@ fig, axes = plt.subplots(
 
 ax = axes[0, 0]
 for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
 enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
 ax.axvline(0, c='black', linewidth=0.6)
 ax.set_xlabel('')
-ax.set_title('Synthetic ENSO-Based Sequences')
+ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
 if plot_log:
     ax.set_ylabel('Log Streamflow')
 else:
@@ -64,10 +64,10 @@ ax.legend()
 
 ax = axes[1, 0]
 for i in np.arange(n_seq_plot-1):
-    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
 markov_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
 ax.axvline(0, c='black', linewidth=0.6)
-ax.set_title('Markov Chain-Based Sequences')
+ax.set_title('{} Markov Chain-Based Sequences'.format(n_seq_plot))
 if plot_log:
     ax.set_ylabel('Log Streamflow')
 else:
@@ -88,3 +88,8 @@ for ax in axes.flat:
 
 fig.tight_layout()
 plt.savefig('figs/stationary_sequences.pdf', bbox_inches='tight')
+
+p_T_enso = (enso_seq > threshold).values.mean()
[…]

### 2018-02-18 src/plot_trend_sequences.py

@@ -19,13 +19,13 @@ if plot_log:
 
 enso_gen = CZNINO3LN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_0=6, beta_mu=0.5, gamma=0.01,
+    mu_0=6, beta_mu=0.5, gamma=0.02,
     coeff_var=0.1, sigma_min=0.01, 
     n_seq=time_param['n_seq'],
 )
 markov_gen = TwoStateSymmetricMarkovLN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_1=6.75, mu_0 = 6, gamma_1=0.01, gamma_2=0, pi=0.9,
+    mmu_1=6.55, mu_0=5.7, gamma_1=0.02, gamma_2=0, pi=0.9,
     coeff_var = 0.1, sigma_min = 0.01,
     n_seq=time_param['n_seq'],
 )
@@ -45,11 +45,11 @@ fig, axes = plt.subplots(
 
 ax = axes[0, 0]
 for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
 enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
 ax.axvline(0, c='black', linewidth=0.6)
 ax.set_xlabel('')
-ax.set_title('Synthetic ENSO-Based Sequences')
+ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
 if plot_log:
     ax.set_ylabel('Log Streamflow')
 else:
@@ -64,10 +64,10 @@ ax.legend()
 
 ax = axes[1, 0]
 for i in np.arange(n_seq_plot-1):
-    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.15, c='gray')
+    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
 markov_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
 ax.axvline(0, c='black', linewidth=0.6)
-ax.set_title('Markov Chain-Based Sequences')
+ax.set_title('{} Markov Chain-Based Sequences'.format(n_seq_plot))
 if plot_log:
[…]

### 2018-02-18 writeup/Outline/main.tex
better example plots

@@
   \begin{equation}
-    \mathcal{D}(t) = k \qty( Q(t) - Q_T)
+    \mathcal{D}(t) = k \qty[ Q(t) - Q_T ]
   \end{equation}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_stationary_long.pdf}
+  \includegraphics[width=\textwidth]{../../figs/example_long.pdf}
   \caption{
-    A single sequence of streamflow generated by each of two generating functions (columns), each fit to three different fitting models (rows) under a stationary case (\(\gamma=0\)).
-Shown is a long series \(M=N=150\).\label{fig:stationary-long-eg}
+    A single sequence of streamflow generated by the NINO3 function with (L) no trend and (R) \(\gamma=0.01\). 
+    The sequence is fit to each f`it to three different fitting models (rows). 
+    Generated with \(M=N=150\).\label{fig:example-long}
   }
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_stationary_short.pdf}
+  \includegraphics[width=\textwidth]{../../figs/example_short.pdf}
   \caption{
-    As \cref{fig:stationary-long-eg} except \(N=30\).\label{fig:stationary-short-eg}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_trend_long.pdf}
-  \caption{
-    A single sequence of streamflow generated by each of two generating functions (columns), each fit to three different fitting models (rows) under a trend case.
-For the ENSO model \(\gamma=0.02\) and for the Markov model \(\gamma_1 = 0.02\) and \(\gamma_2=0\).
-Shown is a long series \(M=N=150\).\label{fig:trend-long-eg}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_trend_short.pdf}
-  \caption{
-    As \cref{fig:trend-long-eg} except \(N=30\).\label{fig:trend-short-eg}
+    As \cref{fig:example-short} except \(N=30\).\label{fig:example-short}
   }

### 2018-02-18 src/plot_example_long.py

@@ -21,14 +21,14 @@ def get_gen_fun(index):
     if index == 0:
         gen_fun = CZNINO3LN2(
             N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0.02,
+            mu_0=6, beta_mu=0.5, gamma=0,
             coeff_var=0.1, sigma_min=0.01
         )
     elif index == 1:
-        gen_fun = TwoStateSymmetricMarkovLN2(
+        gen_fun = CZNINO3LN2(
             N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0.02, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01
+            mu_0=6, beta_mu=0.5, gamma=0.01,
+            coeff_var=0.1, sigma_min=0.01
         )
     else:
         raise ValueError('Invalid Index')
@@ -48,8 +48,8 @@ def get_fit_fun(index, gen_fun):
     return fit_fun
 
 # Initialize the bias and variance
-gen_names = np.array(['enso', 'markov'])
-fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
+gen_names = np.array(['NINO3 Stationary', 'NINO3 Trend'])
+fit_names = np.array(['LN2 Stationary', 'LN2 Trend', 'HMM'])
 
 fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
 for g, gen_fun_name in enumerate(gen_names):
@@ -61,15 +61,15 @@ for g, gen_fun_name in enumerate(gen_names):
         ax = axes[f, g]
         for i in np.arange(n_sim):
             ax.plot(fit_dat['year'], fit_dat.sel(sim=i).values.ravel(), c='gray', linewidth=0.5)
-        ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue')
-        ax.set_ylim([10, 100000])
+        ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue', linewidth=1)
         ax.semilogy()
+        ax.grid()
+        ax.set_ylim([10, 10000])
         if f == 0:
             ax.set_title(gen_fun_name)
         if g == len(gen_names)-1:
             ax.set_ylabel(fit_fun_name)
[…]

### 2018-02-18 src/plot_example_short.py

@@ -25,10 +25,10 @@ def get_gen_fun(index):
             coeff_var=0.1, sigma_min=0.01
         )
     elif index == 1:
-        gen_fun = TwoStateSymmetricMarkovLN2(
+        gen_fun = CZNINO3LN2(
             N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01
+            mu_0=6, beta_mu=0.5, gamma=0.01,
+            coeff_var=0.1, sigma_min=0.01
         )
     else:
         raise ValueError('Invalid Index')
@@ -48,7 +48,7 @@ def get_fit_fun(index, gen_fun):
     return fit_fun
 
 # Initialize the bias and variance
-gen_names = np.array(['enso', 'markov'])
+gen_names = np.array(['NINO3 Stationary', 'NINO3 Trend'])
 fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
 
 fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
@@ -61,9 +61,10 @@ for g, gen_fun_name in enumerate(gen_names):
         ax = axes[f, g]
         for i in np.arange(n_sim):
             ax.plot(fit_dat['year'], fit_dat.sel(sim=i).values.ravel(), c='gray', linewidth=0.5)
-        ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue')
-        ax.set_ylim([10, 100000])
+        ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue', linewidth=1)
         ax.semilogy()
+        ax.grid()
+        ax.set_ylim([10, 100000])
         if f == 0:
             ax.set_title(gen_fun_name)
         if g == len(gen_names)-1:
@@ -71,4 +72,4 @@ for g, gen_fun_name in enumerate(gen_names):
             ax.yaxis.set_label_position('right')
 
 fig.tight_layout()
-plt.savefig('figs/example_stationary_short.pdf')
+plt.savefig('figs/example_short.pdf')

### 2018-02-18 src/plot_example_stationary_long.py

@@ -1,74 +1,0 @@
-import numpy as np
-import xarray as xr
-import pandas as pd
-import matplotlib.pyplot as plt
-
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-
-M = 150
-N = 150
-n_seq = 1
-n_sim = 100
-
-def get_gen_fun(index):
-    """Get a generating function as a function of the index 0 or 1
-    """
-    if index == 0:
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0,
-            coeff_var=0.1, sigma_min=0.01
-        )
-    elif index == 1:
-        gen_fun = TwoStateSymmetricMarkovLN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01
-        )
-    else:
-        raise ValueError('Invalid Index')
-    return gen_fun
-
-def get_fit_fun(index, gen_fun):
-    """Get a fit function as a function of the index 0 or 1 or 2
-    """
-    if index == 0:
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-    elif index == 1:
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
-    elif index == 2:
-        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
-    else:
-        raise ValueError('Invalid Index')
-    return fit_fun
-
-# Initialize the bias and variance
-gen_names = np.array(['enso', 'markov'])
-fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
-
-fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
-for g, gen_fun_name in enumerate(gen_names):
-    gen_fun = get_gen_fun(g)
-    for f, fit_fun_name in enumerate(fit_names):
[…]

### 2018-02-18 src/plot_example_trend_short.py

@@ -1,75 +1,0 @@
-import numpy as np
-import xarray as xr
-import pandas as pd
-import matplotlib.pyplot as plt
-
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-
-M = 150
-N = 30
-n_seq = 1
-n_sim = 100
-
-def get_gen_fun(index):
-    """Get a generating function as a function of the index 0 or 1
-    """
-    if index == 0:
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0.02,
-            coeff_var=0.1, sigma_min=0.01
-        )
-    elif index == 1:
-        gen_fun = TwoStateSymmetricMarkovLN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0.02, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01
-        )
-    else:
-        raise ValueError('Invalid Index')
-    return gen_fun
-
-def get_fit_fun(index, gen_fun):
-    """Get a fit function as a function of the index 0 or 1 or 2
-    """
-    if index == 0:
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-    elif index == 1:
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
-    elif index == 2:
-        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
-    else:
-        raise ValueError('Invalid Index')
-    return fit_fun
-
-# Initialize the bias and variance
-gen_names = np.array(['enso', 'markov'])
-fit_names = np.array(['ln2_stationary', 'ln2_trend', 'hmm'])
-
-fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
-for g, gen_fun_name in enumerate(gen_names):
-    gen_fun = get_gen_fun(g)
[…]

### 2018-02-18 src/plot_stationary_sequences.py

@@ -2,7 +2,6 @@
 """
 import numpy as np
 import matplotlib.pyplot as plt
-import seaborn as sns
 
 import os
 import sys
@@ -11,11 +10,8 @@ sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 
 time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
-plot_log = False # plot log streamflow?
 n_seq_plot = 50 # how many sequences to plot
 threshold = 2500
-if plot_log:
-    threshold = np.log(threshold)
 
 enso_gen = CZNINO3LN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
@@ -32,59 +28,30 @@ markov_gen = TwoStateSymmetricMarkovLN2(
 enso_seq = enso_gen.get_data()
 markov_seq = markov_gen.get_data()
 
-if plot_log:
-    enso_seq = np.log(enso_seq)
-    markov_seq = np.log(markov_seq)
-
 # Setup the figure
-fig, axes = plt.subplots(
-    nrows=2, ncols=2, sharey=True, figsize=(11, 5),
-    gridspec_kw={'width_ratios': [4, 1]},
-    sharex='col'
-)
+fig, axes = plt.subplots(nrows=2, ncols=1, sharey=True, figsize=(12, 5), sharex='col')
 
-ax = axes[0, 0]
+ax = axes[0]
 for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
+    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
 enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
 ax.axvline(0, c='black', linewidth=0.6)
 ax.set_xlabel('')
 ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
-if plot_log:
-    ax.set_ylabel('Log Streamflow')
-else:
-    ax.set_ylabel('Streamflow')
 
-ax = axes[0, 1]
-hist_sequences = enso_seq.sel(year=slice(time_param['t0'] - time_param['N'], time_param['t0'])).values.flatten()
[…]

### 2018-02-18 src/plot_trend_sequences.py

@@ -2,7 +2,6 @@
 """
 import numpy as np
 import matplotlib.pyplot as plt
-import seaborn as sns
 
 import os
 import sys
@@ -11,80 +10,47 @@ sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 
 time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
-plot_log = False # plot log streamflow?
 n_seq_plot = 50 # how many sequences to plot
 threshold = 2500
-if plot_log:
-    threshold = np.log(threshold)
 
 enso_gen = CZNINO3LN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_0=6, beta_mu=0.5, gamma=0.02,
+    mu_0=6, beta_mu=0.5, gamma=0.01,
     coeff_var=0.1, sigma_min=0.01, 
     n_seq=time_param['n_seq'],
 )
 markov_gen = TwoStateSymmetricMarkovLN2(
     N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mmu_1=6.55, mu_0=5.7, gamma_1=0.02, gamma_2=0, pi=0.9,
+    mmu_1=6.55, mu_0=5.7, gamma_1=0.01, gamma_2=0, pi=0.9,
     coeff_var = 0.1, sigma_min = 0.01,
     n_seq=time_param['n_seq'],
 )
 enso_seq = enso_gen.get_data()
 markov_seq = markov_gen.get_data()
 
-if plot_log:
-    enso_seq = np.log(enso_seq)
-    markov_seq = np.log(markov_seq)
-
 # Setup the figure
-fig, axes = plt.subplots(
-    nrows=2, ncols=2, sharey=True, figsize=(11, 5),
-    gridspec_kw={'width_ratios': [4, 1]},
-    sharex='col'
-)
+fig, axes = plt.subplots(nrows=2, ncols=1, sharey=True, figsize=(12, 5), sharex='col')
 
-ax = axes[0, 0]
+ax = axes[0]
 for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray')
+    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
[…]

### 2018-02-19 src/plot_enso.py
More clear ENSO plot

@@ -1,7 +1,17 @@
 import pandas as pd
 import matplotlib.pyplot as plt
+import numpy as np
 
 enso = pd.read_csv('~/Documents/GitHub/floodsampling/floodsampling/data/ramesh2017.csv', index_col='year')
 
-enso.loc[10000:12500].plot(figsize=(10, 5), linewidth=0.5)
+# Plot Time Series
+enso.loc[10000:12500].plot(
+    figsize=(9, 5),
+    linewidth=0.5,
+    c='blue',
+    legend=False,
+    grid=True
+)
+plt.ylabel('NINO3')
+
 plt.savefig('figs/enso.pdf', bbox_inches='tight')

### 2018-02-19 writeup/main.tex
reorganize writeup
Add Manu's comments entirely

--- 2018-02-18 writeup/Outline/main.tex
@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 
-\input{tex/header/macros.tex}
-\input{tex/header/packages.tex}
+\input{header/macros.tex}
+\input{header/packages.tex}
 
-\title{Climate Change Adaptation: Uncertainty Associated with Sequential Finite Period Decisions under Nonstationarity}
+\title{Nonstationary, Quasi-Periodic Risk of Climate Extremes, and Decision Analysis}
 \author[1,2]{James Doss-Gollin}
@@
 
-\input{tex/header/biblatex.tex}
-\addbibresource{../library.bib}
+\input{header/biblatex.tex}
+\addbibresource{library.bib}
+\graphicspath{{../figs/}}
 
@@
 \section{Introduction}
+
+The 20th century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies.
+Event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.  
+Under the assumption that the historical or simulated record of events constitutes an independent and identically distributed process, and using an appropriate statistical inference model, an estimate of this relationship and often its uncertainty based on a finite \(N\) year record is developed.
+This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation, typically using an expected net present value criteria applied to the costs and residual losses expected from the mitigation instrument.
+Such models arguably arose out of statistical convenience and the need for a standardized approach across an institution. 
+The development of such catastrophic risk or flood risk models is now a veritable cottage industry, and the better practitioners, also consider sampling aspects associated with the data used in analysis and will compute some measure of informational uncertainty that is related to the effective sample size of the data used for developing the event magnitude-probability of exceedance relationship.
+To summarize, the approach uses historical or modeled data to establish an event magnitude vs probability of exceedance relationship and may also include an estimate of uncertainty around it. 
+Expected net present value analyses are then used for decisions on the level of risk mitigation to implement using a structural or financial solution. 
+This leads to the selection of an event magnitude and the corresponding probability of exceedance that is selected as the appropriate protection level.
+The uncertainty associated with the model may be used to develop a risk premium for insurance, and/or for assessing the potential over/under design for infrastructure elements, recognizing that the sampling distribution associated with the probability of exceedance associated with a specific design level tends to be asymmetric around the mean value of the estimate.
+
+Over time, it has become clear that even mild nonstationarities (\ie{} temporal changes in the mean, variance or other statistics in the underlying climate process can lead to significant trends in the estimated return periods or probability of exceedance.
+This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it through covariates or other means to develop time varying or state varying (\eg{} relative to the magnitude of a slowly varying climate indicator such as the El Ni\~{n}o-Southern Oscillation) estimates of the desired probability of exceedance.
+Despite such efforts, how one should use these nonstationarity estimates for the design of risk mitigation infrastructure, or the pricing of insurance like instruments has been much less studied.
+
+There are two kinds of nonstationairities of concern.
+The first may be related to anthropogenic climate change.
+The uncertainties associated with this are not purely informational (at least in the sense that they can be inferred from a historical sample or model).
+Nature as well as human actions over the course of this century will determine the outcomes, and neither is obviously predictable. 
+The approach being taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to Greenhouse gas emissions and then use climate models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
+While a lot of work has been done in this direction, we are concerned that even the physical projections of changing climate risk, especially with respect to floods are not credible given their relatively poor performance relative to the 20th century data, thus requiring ``bias correction methods'' whose applicability under extrapolation does not inspire confidence.
+A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
+Would it be over- or under- designed?
+Would the high degree of bias and uncertainty in its estimation lead to highly suboptimal investment strategies in risk mitigation?
+These questions go well beyond what can be reliably answered at this time. 
+This kind of epistemic uncertainty clearly grows as we look out into the distant future, and the importance of investment decisions that may engage us in considerations of designs that go far out into the future, \eg{} 2100AD, diminish if any positive discount rate is used for the future.
+While this consideration offsets the impact of the growing uncertainty for longer project life, it also suggests that it may be more effective to understand the consequences for analysis that may focus on shorter lived mitigation alternatives, ones that may last for 1, 5, 10, 20,  30 or 50 years.
+Sequences of such projects may better reflect climate adaptation goals in any case, especially if we believe that major climate changes may only happen beyond these horizons, or if we are optimistic and believe that humans may avert much of the potential catastrophe by the end of the longer of these horizons.
+
+Focusing on these shorter time horizons brings up the concern associated with the second type of nonstationarity in climate: the quasi-periodic inter-annual, decadal and multidecadal modes of climate variability that have extensive spatial teleconnections and appear to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed draws of events that is central to the typical analysis.
+\citet{Jain2001} were perhaps the first to explore how such processes impact the underlying risk in the context of floods, and the influence of ENSO. 
+They considered a novel experiment design, where they used a the largest event from a block of \(K\) years sampled from the historical record as the target design level, as an empirical estimate of the event magnitude that would be exceeded on average once in \(K\) years.
+The number of times this magnitude was exceeded in the subsequent block of \(K\) years was then recorded. 
+A corresponding randomized experiment was also conducted where the \(K\) years in each step were selected at random consistent with an \iid process hypothesis.
+The data used was an 88 year long record of annual maximum floods from the state of Washington in the USA, and a 1000 year record of a climate model generated El Nino index. 
+The remarkable finding was that the bootstrap frequency distribution associated with event magnitude exceedance process for the structured blocks was dramatically fatter tailed than the one for the randomized blocks.
+Blocks of \(K\) years in which many more exceedances (under-design) as well years in which many fewer (over-design)  happened than expected by chance were very common, demonstrating a high degree of clustering relative to the \iid assumption.
+This would translate into a practical observation at some location where over a period of a few years, several 100 year events may take place, leading people to assert that climate is changing.
+Remarkably, the persistent absence of the extreme events over a similar subsequent batch of years is usually not noticed or reported with equal ardor. 
+If there were a 20 year climate cycle then it would make sense that a batch of 10 years would be much wetter and another much drier, such that if we had 200 years of data, \ie{}, 10 cycles of this phenomenon, then one may still obtain relatively robust estimates of the event with the desired design annual probability of exceedance (e.g., 1/10 or 1/100).
+However, even so, the results of Jain and Lall (2001) say that one would get clustered outcomes reflecting the quasi-periodic or long memory climate behavior, and indeed they note that for the ENSO model data, even a K=200 year block size translates into a significant clustering effect for the next 200 years.
+Bonnafous et al (2017a,b) illustrate how such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies.
+The concern is even higher for portfolios of mines reflecting the space and time clustering of extremes across the portfolio.
+
+Given the \iid assumptions in her analysis of climate extremes, a typical risk analyst uses annualized cost-benefit analyses over the projected life of a structural risk mitigation project, and the expected loss in a given year from the exceedance of the nominal design level is computed as the potential loss* probability of exceedance.
+Clearly, in a nonstationary world, the probability of exceedance could be considered to be temporally variable over the project life. 
+This would however, need to be a prediction of that probability ex ante, using appropriate modeling tools applied to the historical data or to climate model based forecasts. 
+Faced with nonstationarity, some analysts propose using only the most recent subset of the data, e.g., the last 20 or 30 years to build the event magnitude-probability of exceedance relationship. 
+This can be thought of as a ``persistence'' model, which may be ``better than nothing'', and may or may not include a trend in the parameters of the probability distribution. 
+How many years should one retain from the recent past? 
+Too few and the estimation uncertainty dominates.
+Too many and the bias may dominate. 
+Of course, if the underlying climate is quasi-periodic then this choice can be arbitrarily poor, since the last batch of K years may be of exactly the opposite phase of the oscillation as what comes next.
+
+In summary, we need to carefully consider the number of years of historical data, \(N\), and the number of future years \(M\) over which the instrument of interest will be evaluated, as well as efforts at using the data from the \(N\) year period to predict what will happen in the \(M\) years.
+The risk estimation from the \(N\) years needs to explicitly inform a risk prediction over the \(M\) years, and depending on the approach and the form of the underlying nonstationarity, we can expect both bias and uncertainty associated with the prediction, and we need temporally varying probabilities to address the inevitable risk clustering.
+This is then the statement of the risk mitigation design problem that we need to consider for nonstationary climate. 
+This paper is the first to formally develop these ideas in the context of a decision analytic framework.
+
+We focus on structured experiments with synthetic and actual data on extremes to illustrate how one could approach the analyses and what the implications are of varying \(N\) and \(M\) on the ability to do effective inference and on the ability to judge operational performance.
+We consider idealized cases for the structural and financial instruments to convey the ideas, and defer more complex sequential decision making tasks to a subsequent paper. 
+The idea that one needs to consider the joint uncertainty in the estimation and occurrence of climate extremes for the \(N\) and \(M\) periods for stationary conditions was originally considered by \citet{Lall1987}.
+That analysis corresponds to what one would expect under \iid conditions, and would hence provide a baseline to compare the uncertainty distribution associated with a portfolio whose elements were designed using \(N\) years of data with an application to \(M\) future years at a common design probability of exceedance for the elements.
 
@@
 \begin{figure}[b]
-  \includegraphics[width=\textwidth]{../../figs/stationary_sequences.pdf}  
+  \includegraphics[width=\textwidth]{stationary_sequences.pdf}  
   \caption{
@@
 \begin{figure}[b]
-  \includegraphics[width=\textwidth]{../../figs/trend_sequences.pdf}
+  \includegraphics[width=\textwidth]{trend_sequences.pdf}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/stationary_bias.pdf}
+  \includegraphics[width=\textwidth]{stationary_bias.pdf}
   \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/stationary_variance.pdf}
+  \includegraphics[width=\textwidth]{stationary_variance.pdf}
   \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/trend_bias.pdf}
+  \includegraphics[width=\textwidth]{trend_bias.pdf}
   \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/trend_variance.pdf}
+  \includegraphics[width=\textwidth]{trend_variance.pdf}
   \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
@@
 
+\section{Achknowledgements}
+
+The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{1000000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh2017}.
+Kyle Johnson of Columbia University for his suggestions for improving reliability and documentation of the \texttt{floodsampling} package.
+
 \section{Notation}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/enso.pdf}
+  \includegraphics[width=\textwidth]{enso.pdf}
   \caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_long.pdf}
+  \includegraphics[width=\textwidth]{example_long.pdf}
   \caption{
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/example_short.pdf}
+  \includegraphics[width=\textwidth]{example_short.pdf}
   \caption{

### 2018-02-21 writeup/header/packages.tex
add slurm

% Packages With Options
\usepackage[letterpaper, margin=1in]{geometry} % page
\usepackage[english]{babel} % language
\usepackage[utf8]{inputenc} % input
\usepackage[inline]{enumitem} 
\usepackage[section]{placeins} % floats appear in the section they were defined in

% Other Package calls
\usepackage{
  array, % for fixed-width tables
  amssymb, amsmath, % for math
  authblk, % for authors
	booktabs, % for nice tables
  csquotes, % better biblatex
  graphicx, % for figures
  indentfirst, % indent the first line
  physics, % for physics notation
  ragged2e, % for better alignment of text
  siunitx, % for SI notation
  nth, % for 1st, 2nd, 3rd, nth
}

% PACKAGE SETTINGS
\setlength{\RaggedRightParindent}{\parindent} % fix ragged right
\graphicspath{{fig/}{../../fig/}}
\sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
\allowdisplaybreaks{} % let the align environment span multiple pages

% arrays with fixed width
\newcolumntype{L}[1]{>{\raggedright\let\newline \\ \arraybackslash\hspace{0pt}}m{#1}}

### 2018-02-21 writeup/main.tex
add slurm

--- 2018-02-18 writeup/Outline/main.tex
@@
 
-\input{tex/header/macros.tex}
-\input{tex/header/packages.tex}
+\input{header/macros.tex}
+\input{header/packages.tex}
 
-\title{Climate Change Adaptation: Uncertainty Associated with Sequential Finite Period Decisions under Nonstationarity}
+\title{Nonstationary, Quasi-Periodic Risk of Climate Extremes, and Decision Analysis}
 \author[1,2]{James Doss-Gollin}
@@
 \begin{abstract}
-  Hydroclimatic systems exhibit organized low-frequency and regime-like variability at multiple time scales, causing the risk associated with climate extremes such as floods and droughts to vary at multiple timescales.
-  Despite broad recognition of this nonstationarity, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
+  Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep[``nonstationarity'';][]{Milly2008,Merz2014,Hodgkins2017}.
+  This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
+  Despite broad recognition of nonstationarity's importance, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
   Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period \(M\), but not for analysis of sequential decision strategies.
   In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
-  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show that \(\ldots{}\).
+  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show \(\ldots{}\).
   This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
@@
 
-Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep{Milly2008,Merz2014,Hurst1951,Sveinsson2005,Hodgkins2017,Matalas2012}.
-This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
-Yet the same memory and low-frequency variability that complicate projecting flood sequences far into the future may also impart some predictability at short timescales \citep{Jain2001}.
-The existence of this information gap suggests that approaches for adapting to hydroclimate risks, such as floods, which have short project operation periods, may be preferential in some cases to large and permanent projects with long project operations.
+The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies \citep[\ie{} Bulletin 17-B for flood frequency analysis;][]{IACWD1982}.
+In this framework event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
+Under the assumption that the historical or simulated record of events constitutes an independent and identically distributed process, and using an appropriate statistical inference model, an estimate of this relationship and often its uncertainty based on a finite \(N\) year record is developed.
+This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation, typically using an expected net present value criteria applied to the costs and residual losses expected from the mitigation instrument.
 
-At present, most approaches to nonstationary flood risk management, or climate change adaptation more broadly, have framed the problem first as one of detecting historical change or modeling future change, and then assessing the implications of this change for future infrastructure.
-Though more recent studies have pointed out that the paradigm of requiring proof of the existence of trends before adapting to nonstationarity does not lead to optimal decisionmaking \citep{Vogel2013,Rosner2014}, such discussion has also focused on linear time trends and on choosing whether or not to implement a particular adaptation policy with a fixed planning period.
-In contrast, public and private actors seeking to adapt to changing flood or climate risk typically have access to a suite of options including large flood control infrastructure, insurance contracts and other financial instruments, vulnerability reduction (\ie{} disaster response training), and exposure reduction (\ie{} moving infrastructure out of the floodplain).
-In this context, understanding under what conditions a large, permanent project or small, easily modified adaptation measure are optimal becomes an important question.
+Over time it has become clear that even small nonstationarity (temporal changes in dynamics of the underlying hydroclimatalogical process) can lead to significant trends in the estimated \(T\)-year return periods or probability of exceedance \( p_T \) \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.
+This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last \(N\) years of the data.
+Each model will necessarily give different estimates, particularly as the model is forecast further into the future.
+Despite the wide suite of available models, the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied.
 
-In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period \(M\) and the length of the available data record \(N\).
+There are two kinds of nonstationairities of concern.
+The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Merz2014}.
+The uncertainties associated with this are not purely informational; both nature and future human actions will determine the outcomes, and neither is obviously predictable.
+The approach taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions and then use climate models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
+
+Despite continuing improvements in physical understanding and computational capacity, there is cause for concern with dependence on using physics-based numerical models to project future climate risk.
+Use of a ``model chain'' typically encompasses:
+\begin{enumerate*}[label= (\roman*) ]
+  \item emission scenario;
+  \item general circulation model (GCM);
+  \item downscaling;
+  \item hydrological catchment model; and
+  \item flood frequency analysis.
+\end{enumerate*}
+Given poor performance of flood projections relative to \nth{20} century data, ``bias correction methods'' whose applicability under extrapolation does not inspire confidence are typically applied \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
+Would it be over- or under- designed?
+Would the high degree of bias and uncertainty in its estimation lead to suboptimal investment strategies in risk mitigation?
+These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.
+
+If a positive discount rate is used for the future, as is typical in public and private sector planning models,\footnote{CITE} the investment returns on major public and private investment may diminish substantially far into the future.
+This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
+Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a paticular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
+
+Consideration of these shorter time horizons brings up the concern associated with the second type of nonstationarity in climate: the quasi-periodic inter-annual, decadal and multidecadal modes of climate variability that have extensive spatial teleconnections and appear to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed draws of events that is central to the classical analysis \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015,Hannachi2017}.
+Using a bootstrap approach with a long, stationary ENSO data series, \citet{Jain2001} demonstrated that the using the past \(K\) years to represent the next \(K\) years led to large surprise relative to when the data was fully randomized and the temporal structure removed.
+Even for a very long sequence (\ie{} \(K=200\) years), the past was found to be a misleading representation of the future.
+Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies,\footnote{Bonnafous et al (2017a,b)} who face spatial as well as temporal clustering.
+
+Clearly, in a nonstationary world the probability of exceedance is temporally variable over the project life.
+In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period \(M\) and the length of the historical record \(N\).
 This project planning period can be interpreted as the design life of a physical structure, or as the contract length of a financial instrument.
@@
 This estimation bias and variance may map directly onto the price of an insurance contract\footnote{cite} or the possibility of over/under-design of a flood control structure.
-Though we focus on flood risk, this type of analysis may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricane, wind, \etc{}.
+Though the specific examples target flood risk, this method and framework may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricanes, and wind.
+
+\section{Methods}\label{sec:methods}
+
+We use a highly idealized modeling framework to generate sequences of annual-maximum floods, fit these sequences in a probabilistic framework to three flood models that incorporate time in different ways, and evalaluate the performance of these models for multiple observational record lengths \(N\) and project planning periods \(M\).
+
+\subsection{Generating Synthetic Flood Sequences\label{sec:methods-generating}}
+
+We generate synthetic flood sequences from two different simple conceptual models that represent chaotic dynamics and long memory.
+For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution, conditional on an expected value \(\mu(t)\) which depends on individual model:
+\begin{equation} \label{eq:lognormal}
+  \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
+\end{equation}.
+For simplicitiy we assume a constant coefficient of variation:
+\begin{equation}\label{eq:calc-sigma}
+  \sigma(t) = 
+  \begin{cases} 
+    \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq \sigma_\text{min} \\ 
+    \sigma_\text{min} & \qqtext{else}
+  \end{cases}
+\end{equation}
+where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
+
+The procedure to generate a sequence of streamflow is thus:
+\begin{enumerate}
+  \item Choose \(t_0\), \(M\), and \(N\) so that the time is \(\vb{t}=t_0-N+1, \ldots, t_0+M\).
+  \item Use one of the models described below to simulate \(\mu(t)\)
+  \item Use \cref{eq:calc-sigma} to calculate \(\sigma(t)\)
+  \item For each time \(t \in \vb{t}\), draw \(Q(t)\) following \cref{eq:lognormal}.
+\end{enumerate}
+Though both approaches used to generate sequences \( \mu(t) \) use highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
+In \cref{sec:sequence-realistic} we show sequences of streamflow generated by both models for some choices of parameters and discuss the processes which they can reasonably represent.
+
+\subsubsection{Synthetic NINO3 Model\label{sec:methods-nino}}
+
+ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a quasi-oscillatory manner with variability on the order of 3 to 7 years with further low-frequency modulation \citep{Sarachik2010}.
+A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
+A subset of this data is shown in \cref{fig:enso-ts}.
+To create an annual time series, we average the October-December values of the NINO3 index for each year.
+
+To generate \(\mu(t)\), a sequence of length \(M+N\) is randomly chosen from the \SI{100000}{year} NINO3 sequence.
+Next, \(\mu\) is calculated following by
+\begin{equation}
+  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0)
+\end{equation}
+where parameter \(\gamma\) controls the magnitude of the secular trend (\(\gamma=0\) gives a stationary series).
+
+\subsubsection{Markov Chain\label{sec:methods-markov}}
+
+To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate \(\mu(t\).
+
+This Markov chain is given a constant probability of persistence which is the same for both states, such that the transition matrix is
+\begin{equation*}
+  P = \mqty[\pi & 1-\pi \\ 1-\pi & \pi]
+\end{equation*}
+This is first used to generate a sequence of states \(S(t)\).
+The value \( \mu(t) \) depends only on the \(S(t)\) and the time itself:
+\begin{equation*}
+  \mu(t) = \begin{cases}
+    \mu_{0,1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
+    \mu_{0,2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
+  \end{cases}
+\end{equation*}
+For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
+We also assume \( \mu_{0, 1} \gg \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
+
+\subsection{Estimating Future Streamflow Sequences\label{sec:estimation}}
+
+This paper does not propose new models or suggest a general ``best'' model; instead it aims to describe the total uncertainty in estimates using easily methods which have been well studied in the literature.
+
+In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 1000\footnote{Check this parameter} sequences of future streamflow are generated using Monte Carlo simulation.
+Three models are assessed for estimating future streamflow and are described in detail below.
+
+\subsubsection{Stationary LN2 Model\label{sec:method-stationary}}
+
+We first consider fitting a stationary model to the observed flood record, following classical assumptions of \iid sequences and stationarity.
+We use the log-normal model (\cref{eq:lognormal}) for stationary flood distribution; while this model is less flexible than three-parameter models such as the GEV or LP3 \citep{Vogel1996}, it is easy to understand and interpret.
+Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
+
+For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations (denoted \(Q_{\text{hist}}\) for clarity).
+Then, the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
+Weak (uninformative) priors are chosen; in the context of a synthetic data experiment using informative priors seems inappropriate but other studies\footnote{cite} have used scaling information and regional information to inform prior distributions and reduce posterior variance.
+The full probability model is thus:
+\begin{align}
+  \log Q_\text{hist} & \sim \normal \qty(\mu, \sigma) \\
+  \mu &\sim \normal \qty(0, 10) \\
+  \sigma &\sim \normal \qty(0, 2)
+\end{align}
+
+\subsubsection{Trend LN2 Model\label{sec:method-trend}}
+
+To incorporate secular change into a statistical model, many studies have used time parameterizations on certain parameters.
+We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \):
+\begin{align}
+  \log Q_\text{hist} & \sim \normal \qty(\mu_0 + \beta_\mu \qty(t - t_0), \sigma_0 + \beta_\sigma \qty(t-t_0)) \\
+  \mu_0 & \sim \normal \qty(0, 10) \\
+  \beta_\mu & \sim \normal \qty(0, 0.1) \\
+  \sigma_0 & \sim \normal \qty(0, 2) \\
+  \beta_\sigma & \sim \normal \qty(0, 0.05)
+\end{align}
+As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors.
+The influence of priors in this case is stronger than in the stationary case, and a reasonable choice of priors might more strongly favor \(\beta_mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
+However, in the context of a synthetic data exercise it seems reasonable to apply priors that are at most weakly informative.
+
+\subsubsection{Hidden Markov Model\label{sec:method-HMM}}
+
+A Hidden Markov Model (HMM) is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved (\ie{} hidden) states \(S(t)\) \citep{Rabiner1986}.
+The (unobserved) states evolve following a first-order Markov process (\(S(t)\) depends only on \(S(t-1)\)), and the observed variable (streamflow) depends only on the observed state.
+As such, HMMs are a structured prediction method which extend general mixture models to sequences of data, where position in the sequence is relevant.
+Because this structure allows the hidden state to represent hydroclimatological ``regimes'' \citep{Reinhold1982,Michelangeli1995,Merz2014}, HMMs have have been used extensively to model hydrological variables such as streamflow \citep[\ie{}][]{Bracken2016} or climate states such as ENSO.\footnote{Julian's work}
+
+We fit observed streamflow sequences \(Q_\text{hist}\) using a HMM as implemented in the python \texttt{pomegranate} package \citep{Schreiber2016}.
+The model is fit to the \(N\) observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
+This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
+To choose the number of hidden states, a log probability score is used.\footnote{Need to implement this?}
+Future floods are estimated by simulating future states from the resulting matrix and then drawing floods log-normally using the estimated parameters.
+
+\subsection{Evaluating Fitting Models}
+
+For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
+For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
+\(K\) posterior simulations of future rainfall are then drawn from the fitted model.
+We then perform a standard decomposition of uncertainty:
+\begin{equation}
+  \mathbb{E} \qty[ \qty( \hat{p}_T - p_T )^2 ] = \qty( \mathbb{E} \qty[ \hat{p}_T ] - p_T )^2 + \mathbb{E} \qty[ \qty( \hat{p}_T - \mathbb{E} \qty[ \hat{p}_T ] )^2]
+\end{equation}
+where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
+
+\subsection{Code Availability}
+
+All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
+To carry implement the flood sequence generating and estimating functions, a new package called \texttt{floodsampling} was developed and is available at \url{github.com/jdossgollin/floodsampling}.
+The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.
+
+\section{Results}
+
+We begin by showing the types of synthetic streamflow sequence that are by the two models discussed \cref{sec:methods-generating}.
+We then examine how different statistical models for estimating future flood sequences perform under stationary conditions and when a secular trend is introduced.
+
+\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
+
+The models used to generate synthetic flood sequences used simplistic dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
+
+To verify that these synthetic streamflow sequences sufficiently represent these behaviors, we plot several sequences.
+The ENSO model used a long run of the Cane-Zebiak model \citep{Zebiak1987,Ramesh2017} to determine \(\mu(t)\); a \SI{2500}{year} sub-set of the annualized time series is shown in \cref{fig:enso-ts}.
+As described in the literature,\footnote{citations, and maybe some paleo stuff as well} this model shows quasi-periodic oscillations in the \SIrange{3}{7}{year} band but with substantial low-frequency modulation of extreme activity.
+The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
+
+\Cref{fig:stationary-sequences} shows 50 streamflow sequences generated with \(M=50\) and \(N=100\).
+The ENSO sequences were generated with \(\gamma=0, \beta_\mu=0.5, \mu_0=6\), which gives a strong (but still noisy) relationship between the NINO3 index \(x(t)\)) and the annual-maximum streamflow \(Q(t)\).
+The Markov sequences were generated with \(\mu_1 = 6.55, \mu_2=5.7, \gamma_1=0, \gamma_2=0, \pi=0.9\), which gives a large difference between the two states.
+For both, the coefficient of variation \(\alpha=0.1\) and the minimum value of sigma \(\sigma_\text{min}=0.01\).
+As the time series plot shows, both generating models exhit substantial persistence in flood probability from one year to the next, though the ENSO model shows negative autocorrelation at lag 1 while the Markov model has positive autocorrelation at lag 1 (not shown).\footnote{Need to include this?}
+
+\begin{figure}[b]
+  \includegraphics[width=\textwidth]{stationary_sequences.pdf}  
+  \caption{
+    50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-nino} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
+    Both models use a stationary parameterization.
+    49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
+    The right hand panel shows the distribution over the historical and future periods for 500 sequences from the same model, visualized with a histogram and kernel density estimate.
+    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.014133333333333333} by the NINO3 sequences and \num{0.013546666666666667} by the Markov sequences.\label{fig:stationary-sequences}
+  }
+\end{figure}
+
+To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
+For the ENSO time series, this mean setting \(\gamma=0.02\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.02\).
+The resulting sequences are shown in \cref{fig:trend-sequences}.
+The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
+\begin{figure}[b]
+  \includegraphics[width=\textwidth]{trend_sequences.pdf}
+  \caption{
+    As \cref{fig:stationary-sequences} but a trend term has been added.
+    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.023306666666666667} by the NINO3 sequences and \num{0.024493333333333332} by the Markov sequences.\label{fig:trend-sequences}
+  }
+\end{figure}
+
+\subsection{Stationary Case}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{stationary_bias.pdf}
+  \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{stationary_variance.pdf}
+  \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\subsection{Trend Case}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{trend_bias.pdf}
+  \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{trend_variance.pdf}
+  \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+\end{figure}
+
+\section{Case Study}
+
+We still need to find a good case study; it can be a highly simplified one as long as there is a long streamflow record available and building levees is an option that has been proposed.
+If there's not an obvious candidate, we can use one of our models above.
+We can make some other strong assumptions
+\begin{itemize}
+  \item Risk exposure constant in time, which allows us to focus on the estimation rather than comprehensive risk manageement
+  \item Constant future discounting rate
+  \item Flood damages are linearly related to streamflow over some amount:
+  \begin{equation}
+    \mathcal{D}(t) = k \qty[ Q(t) - Q_T ]
+  \end{equation}
+\end{itemize}
+Then our two options are to build a levee, which we can assume to have a capital cost plus operation and maintenance cost which is proportional to the capital cost, which increases \(Q_T\) by some fixed amount, or to buy insurance.
+For the insurance, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
+Although \(Q_T\) is in general a decision variable, we treat it as fixed.
+Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that \(Q(t) \geq Q_T\) for \(t \in \vb{t}\), which is equivalent to \(p_T(t)\).
+For a particular year, the fair price for the insurance premium is \(\expectation \qty(p_T) C\).
+However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger \(Q_T\) is specified in the contract and thus not uncertain).
+This risk premium is proportional to the variance of \(p_T\), or more generally to its uncertainty distribution.
+The insurance premium for a particular year is thus
+\begin{equation} \label{eq:premium}
+  Z(t) = C \qty[\expectation \qty(p_T(t)) + R \variance \qty(p_T(t))]
+\end{equation}
+where \(R \geq 0\) is a coefficient describing the cost of tue uncertainty in the estimate of \(p_T(t)\).
+
+Then at each point in time, calculate:
+\begin{itemize}
+  \item Expected cost plus damage per year for the levee using perfect knowledge of the next 101 years
+  \item Expected cost plus damage per year for the levee using MCMC projections of the next 101 years from all 3 models
+  \item Expected cost plus damage per year for a 5-year insurance contract using perfect knowledge of the next 5 years
+  \item Expected cost plus damage per year for a 5-year insurance contract using MCMC projections of the next 5 years from all 3 models
+\end{itemize}
+
+Can add appropriate plot or plots.
+
+\section{Discussion}
+
+For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
+While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
+
+From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
+How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length \(M\).
+What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
+
+How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
+
+\section{Summary}
+
+The main point you are hitting is that the very nature of climate has long memory and some people point this out, many model it, and most future projections don't show this.
+All corrections (even if we like them) address the marginal distribution.
+So while the debate you refer to is interesting to a degree, the point we are making is that in terms of risk analysis it is not even tangentially relevant to reality and what we bring up -- our focus -- is strictly the risk based decision process and we are agnostic to the details of the models, other than the long memory features and spatial teleconnections associated with them that are seen as marks of the system behavior.
+We actually demonstrate that the potential loss in a 2100 scenario has to be catastrophic and really out of the range of any historical variability for it to be relevant.
+If we accept this proposition then we have a strong case for the mitigation of climate change but a much weaker case for adaptation, since if we are not able to mitigate that scenario then building a giant wall to control the 2100 scenario today will be economically and socially infeasible anyway.
+So the next obvious question is in which situation would waiting and updating the decision make sense and this can be set up and explored or just discussed.
+
+% -----------------------------------------------------------------------------
+% END HERE
+% -----------------------------------------------------------------------------
+\appendix
+\printbibliography{}
+
+\section{Thanks}
+
+The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{1000000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh2017}.
+Kyle Johnson of Columbia University for his suggestions for improving reliability and documentation of the \texttt{floodsampling} package.
+
+\section{Background Information and Conceptual Framing}
 
@@
 
-\section{Methods}\label{sec:methods}
-
-We use an idealized model to generate sequences of annual-maximum floods, fit these sequences in a probabilistic framework to three flood models that incorporate time in different ways, and evalaluate the performance of these models for multiple observational record lengths \(N\) and project planning periods \(M\).
-
-\subsection{Generating Synthetic Flood Sequences\label{sec:methods-generating}}
-
-We generate synthetic flood sequences from two different simple conceptual models that represent chaotic dynamics and long memory.
-
-For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution:
-\begin{equation} \label{eq:lognormal}
-  \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
-\end{equation}
-where the parameters \( \mu(t), \sigma(t) \) are allowed to vary in time following the two approaches outlined below.
-For simplicitiy we assume a constant coefficient of variation:
-\begin{equation}
-  \sigma(t) = \begin{cases} \alpha \mu(t) & \qqtext{if} \alpha \mu(t) \geq \sigma_\text{min} \\ \sigma_\text{min} & \qqtext{else} \end{cases}
-\end{equation}
-where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
-The two models are summarized in \cref{tab:model-generating}.
-
-Though both approaches used to generate sequences \( \mu(t) \) useh highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
-
-In \cref{sec:sequence-realistic} we show sequences of streamflow generated by both models for some choices of parameters and discuss the processes which they can reasonably represent.
-
-\begin{table}[bht]
-  \begin{center}
-    \begin{tabular}{L{1.0in} L{3.0in} L{1.5in}}
-      \toprule
-        Model Name & Description & Relevant Section \\
-      \midrule
-        NINO3 & \(\mu(t)\) proportional to synthetic NINO3 time series & \cref{sec:methods-enso} \\
-        Markov & \(\mu(t)\) depends only on sequence of latent states generated by Markov chain & \cref{sec:methods-markov} \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \caption{Summary of models used for generating synthetic flood sequences\label{tab:model-generating}}
-\end{table}
-
-\subsubsection{Synthetic ENSO Model\label{sec:methods-enso}}
-
-ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a quasi-oscillatory manner with variability on the order of \SIrange{4}{6}{year} modulated by low-frequency oscillation \citep[see][for a comprehensive review]{Sarachik2010}.
-A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
-A \SI{1000}{year} subset of this data is shown in \cref{fig:enso-ts}.
-To create an annual time series, we average the October-December months to create a time series \( x(t) \) where \( t \) is taken in annual steps \(1, \ldots, \num{100000} \).
-
-First a sequence of length \(M+N\) is randomly chosen from the \SI{100000}{year} ENSO sequence.
-Next, \(\mu\) is calculated following by
-\begin{equation}
-  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0)
-\end{equation}
-where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
-
-\subsubsection{Markov Chain\label{sec:methods-markov}}
-
-To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate \(\mu(t\).
-
-This Markov chain is given a constant probability of persistence which is the same for both states, such that the transition matrix is
-\begin{equation*}
-  P = \mqty[\pi & 1-\pi \\ 1-\pi & \pi]
-\end{equation*}
-This is first used to generate a sequence of states \(S(t)\).
-The value \( \mu(t) \) depends only on the \(S(t)\) and the time itself:
-\begin{equation*}
-  \mu(t) = \begin{cases}
-    \mu_{0,1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
-    \mu_{0,2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
-  \end{cases}
-\end{equation*}
-For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
-We also assume \( \mu_{0, 1} \gg \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
-
-\subsection{Estimating Future Streamflow Sequences\label{sec:estimation}}
-
-We do not, in general, know the true distribution of annual-maximum floods, and as discussed in \cref{sec:ffa} a plethora of approaches have been suggested in the literature.
-This paper does not propose new models or suggest a general ``best'' model; instead it aims to describe the total uncertainty in estimates using easily methods which have been well studied in the literature.
-
-In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 1000\footnote{Check this parameter} sequences of future streamflow are generated using Monte Carlo simulation.
-These estimates are then used to calculate \( \hat{p_T} \), and the estimates are evaluated as described in \cref{sec:methods-uncertainty}.
-Three models are assessed for estimating future streamflow; they are summarized in \cref{tab:model-fitting} and described in greater detail below.
-
-\begin{table}[bht]
-  \begin{center}
-    \begin{tabular}{L{1.25in} L{3.0in} L{1.25in}}
-      \toprule
-        Model Name & Description & Relevant Section \\
-      \midrule
-        \texttt{ln2\_stationary} & \( \log Q \sim N \qty(\mu, \sigma) \) & \cref{sec:method-stationary} \\
-        \texttt{ln2\_trend} & \(\log Q \sim N \qty(\mu_0 + \beta_\mu \qty[t - t_0], \sigma_0 + \beta_\sigma \qty[t - t_0] ) \) & \cref{sec:method-trend} \\
-        \texttt{HMM} & Hidden Markov Model & \cref{sec:method-HMM} \\
-      \bottomrule
-    \end{tabular}
-  \end{center}
-  \caption{Summary of models used for fitting\label{tab:model-fitting}}
-\end{table}
-
-\subsubsection{Stationary Model\label{sec:method-stationary}}
-
-We do not, in general, know the true distribution of annual-maximum floods, and there has been substantial discussion in the literature as to the choice of distribution to use for stationary flood frequency analysis.
-We consider the log-normal model (\cref{eq:lognormal}) for stationary flood distribution; while this model is less flexible than three-parameter models such as the GEV or LP3 \citep{Vogel1996}, it is computationally convenient and easy to understand.
-More importantly, since we know the true distribution of these synthetically generated flood, using the log-normal for both generating and fitting supports interpreting results as a lower bound on total uncertainty.
-In the stationary context, we assume \(\mu(t) = \mu\) and \(\sigma(t)=\sigma\).
-
-For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations.
-Then, once the appropriate time parameterization is selected (\cref{tab:model-fitting}), the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
-Weak (uninformative) priors are chosen; in the context of a synthetic data experiment using informative priors seems inappropriate but other studies\footnote{cite} have used scaling information and regional information to inform prior distributions and reduce posterior variance.
-
-\subsubsection{Trend Model\label{sec:method-trend}}
-
-Many studies\footnote{cite} have used time parameterizations for fitting.
-We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \):
-\begin{align}
-  \mu(t) = \mu_0 + \beta_\mu \qty(t - t_0) \\
-  \sigma(t) = \sigma_0 + \beta_\sigma \qty(t - t_0)
-\end{align}
-
-As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors.
-The influence of priors in this case is stronger than in the stationary case, and a reasonable choice of priors might strongly favor \(\beta_mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
-
-\subsubsection{Hidden Markov Model\label{sec:method-HMM}}
-
-A Hidden Markov Model (HMM) is a statistical Markov model in which the system being modeled is assumed to be a Markov process with unobserved (\ie{} hidden) states \citep{Rabiner1986}.
-The (unobserved) states evolve following a first-order Markov process, and the observed variable (\ie{} streamflow) depends only on the observed state.
-As such, HMMs are a structured prediction method which extend general mixture models to sequences of data, where position in the sequence is relevant.
-Because this structure allows the hidden state to represent hydroclimatological ``regimes'' \citep{Reinhold1982,Michelangeli1995,Merz2014}, HMMs have have been used extensively to model hydrological variables such as streamflow \citep{Bracken2016}.
-HMMs have also been used to model the evolution of climate states such as ENSO.\footnote{Julian's work}
-
-To supplement the distributional models described in \cref{sec:estimation}, we also fit observed streamflow sequences using a HMM as implemented in the python \texttt{pomegranate} package \citep{Schreiber2016}.
-The model is fit to the \(N\) observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
-To choose the number of hidden states, a log probability score is used.
-Future floods are estimated by simulating future states from the resulting matrix and then drawing floods log-normally using the estimated parameters.
-
-\texttt{How much more do we need to say about HMM?}
-
-\subsection{Evaluating Fitting Models}
-
-For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
-For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
-\(K\) posterior simulations of future rainfall are then drawn from the fitted model.
-We then perform a standard decomposition of uncertainty:
-\begin{equation}
-  \mathbb{E} \qty[ \qty( \hat{p}_T - p_T )^2 ] = \qty( \mathbb{E} \qty[ \hat{p}_T ] - p_T )^2 + \mathbb{E} \qty[ \qty( \hat{p}_T - \mathbb{E} \qty[ \hat{p}_T ] )^2]
-\end{equation}
-where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
-
-\subsection{Reproducability}
-
-All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
-To carry implement the flood sequence generating and estimating functions, a new package called \texttt{floodsampling} was developed and is available at \url{github.com/jdossgollin/floodsampling}.
-The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.
-
-\section{Results}
-
-We begin by showing the types of synthetic streamflow sequence that are by the two models discussed \cref{sec:methods-generating}.
-We then examine how different statistical models for estimating future flood sequences perform under stationary conditions and when a secular trend is introduced.
-
-\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
-
-The models used to generate synthetic flood sequences used simplistic dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
-
-To verify that these synthetic streamflow sequences sufficiently represent these behaviors, we plot several sequences.
-The ENSO model used a long run of the Cane-Zebiak model \citep{Zebiak1987,Ramesh2017} to determine \(\mu(t)\); a \SI{2500}{year} sub-set of the annualized time series is shown in \cref{fig:enso-ts}.
-As described in the literature,\footnote{citations, and maybe some paleo stuff as well} this model shows quasi-periodic oscillations in the \SIrange{3}{7}{year} band but with substantial low-frequency modulation of extreme activity.
-The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
-
-\Cref{fig:stationary-sequences} shows 50 streamflow sequences generated with \(M=50\) and \(N=100\).
-The ENSO sequences were generated with \(\gamma=0, \beta_\mu=0.5, \mu_0=6\), which gives a strong (but still noisy) relationship between the NINO3 index \(x(t)\)) and the annual-maximum streamflow \(Q(t)\).
-The Markov sequences were generated with \(\mu_1 = 6.55, \mu_2=5.7, \gamma_1=0, \gamma_2=0, \pi=0.9\), which gives a large difference between the two states.
-For both, the coefficient of variation \(\alpha=0.1\) and the minimum value of sigma \(\sigma_\text{min}=0.01\).
-As the time series plot shows, both generating models exhit substantial persistence in flood probability from one year to the next, though the ENSO model shows negative autocorrelation at lag 1 while the Markov model has positive autocorrelation at lag 1 (not shown).\footnote{Need to include this?}
-
-\begin{figure}[b]
-  \includegraphics[width=\textwidth]{../../figs/stationary_sequences.pdf}  
-  \caption{
-    50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-enso} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
-    Both models use a stationary parameterization.
-    49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
-    The right hand panel shows the distribution over the historical and future periods for 500 sequences from the same model, visualized with a histogram and kernel density estimate.
-    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.014133333333333333} by the NINO3 sequences and \num{0.013546666666666667} by the Markov sequences.\label{fig:stationary-sequences}
-  }
-\end{figure}
-
-To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
-For the ENSO time series, this mean setting \(\gamma=0.02\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.02\).
-The resulting sequences are shown in \cref{fig:trend-sequences}.
-The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
-\begin{figure}[b]
-  \includegraphics[width=\textwidth]{../../figs/trend_sequences.pdf}
-  \caption{
-    As \cref{fig:stationary-sequences} but a trend term has been added.
-    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.023306666666666667} by the NINO3 sequences and \num{0.024493333333333332} by the Markov sequences.\label{fig:trend-sequences}
-  }
-\end{figure}
-
-\subsection{Stationary Case}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/stationary_bias.pdf}
-  \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/stationary_variance.pdf}
-  \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
-\end{figure}
-
-\subsection{Trend Case}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/trend_bias.pdf}
-  \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
-\end{figure}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{../../figs/trend_variance.pdf}
-  \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
-\end{figure}
-
-\section{Case Study}
-
-We still need to find a good case study; it can be a highly simplified one as long as there is a long streamflow record available and building levees is an option that has been proposed.
-If there's not an obvious candidate, we can use one of our models above.
-We can make some other strong assumptions
-\begin{itemize}
-  \item Risk exposure constant in time, which allows us to focus on the estimation rather than comprehensive risk manageement
-  \item Constant future discounting rate
-  \item Flood damages are linearly related to streamflow over some amount:
-  \begin{equation}
-    \mathcal{D}(t) = k \qty[ Q(t) - Q_T ]
-  \end{equation}
-\end{itemize}
-Then our two options are to build a levee, which we can assume to have a capital cost plus operation and maintenance cost which is proportional to the capital cost, which increases \(Q_T\) by some fixed amount, or to buy insurance.
-For the insurance, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
-Although \(Q_T\) is in general a decision variable, we treat it as fixed.
-Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that \(Q(t) \geq Q_T\) for \(t \in \vb{t}\), which is equivalent to \(p_T(t)\).
-For a particular year, the fair price for the insurance premium is \(\expectation \qty(p_T) C\).
-However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger \(Q_T\) is specified in the contract and thus not uncertain).
-This risk premium is proportional to the variance of \(p_T\), or more generally to its uncertainty distribution.
-The insurance premium for a particular year is thus
-\begin{equation} \label{eq:premium}
-  Z(t) = C \qty[\expectation \qty(p_T(t)) + R \variance \qty(p_T(t))]
-\end{equation}
-where \(R \geq 0\) is a coefficient describing the cost of tue uncertainty in the estimate of \(p_T(t)\).
-
-Then at each point in time, calculate:
-\begin{itemize}
-  \item Expected cost plus damage per year for the levee using perfect knowledge of the next 101 years
-  \item Expected cost plus damage per year for the levee using MCMC projections of the next 101 years from all 3 models
-  \item Expected cost plus damage per year for a 5-year insurance contract using perfect knowledge of the next 5 years
-  \item Expected cost plus damage per year for a 5-year insurance contract using MCMC projections of the next 5 years from all 3 models
-\end{itemize}
-
-Can add appropriate plot or plots.
-
-\section{Discussion}
-
-For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
-While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
-
-From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
-How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length \(M\).
-WHat is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
-
-How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
-
-\section{Summary}
-
-The main point you are hitting is that the very nature of climate has long memory and some people point this out, many model it, and most future projections don't show this.
-All corrections (even if we like them) address the marginal distribution.
-So while the debate you refer to is interesting to a degree, the point we are making is that in terms of risk analysis it is not even tangentially relevant to reality and what we bring up -- our focus -- is strictly the risk based decision process and we are agnostic to the details of the models, other than the long memory features and spatial teleconnections associated with them that are seen as marks of the system behavior.
-We actually demonstrate that the potential loss in a 2100 scenario has to be catastrophic and really out of the range of any historical variability for it to be relevant.
-If we accept this proposition then we have a strong case for the mitigation of climate change but a much weaker case for adaptation, since if we are not able to mitigate that scenario then building a giant wall to control the 2100 scenario today will be economically and socially infeasible anyway.
-So the next obvious question is in which situation would waiting and updating the decision make sense and this can be set up and explored or just discussed.
-
-% -----------------------------------------------------------------------------
-% END HERE
-% -----------------------------------------------------------------------------
-\appendix
-\printbibliography{}
-
 \section{Notation}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{../../figs/enso.pdf}
+  \includegraphics[width=\textwidth]{enso.pdf}
   \caption{
-    A \SI{2500}{year} sub-set of the synthetic time series described in \cref{sec:methods-enso}.\label{fig:enso-ts}
+    A \SI{2500}{year} sub-set of the synthetic time series described in \cref{sec:methods-nino}.\label{fig:enso-ts}
   }

### 2018-02-21 run.sh

@@ -1,9 +1,0 @@
-python src/plot_stationary_sequences.py
-python src/plot_trend_sequences.py
-python src/plot_enso.py
-
-python src/calc_stationary.py
-python src/calc_trend.py
-
-python src/plot_stationary.py
-python src/plot_trend.py

### 2018-02-21 slurm.sh

@@ -1,0 +1,17 @@
+#!/bin/sh
+#
+# Run the code on a slurm server
+#
+# Replace <ACCOUNT> with your account name before submitting.
+#
+#SBATCH --account=cwc           # The account name for the job.
+#SBATCH --job-name=JDGMN01      # The job name.
+#SBATCH -N 1                    # The number of nodes to use
+#SBATCH --exclusive
+#SBATCH --time=10:00:00         # The time the job will take to run.
+
+module load anaconda/3-4.4.0
+source activate test_env
+make simulate                   # run the make command specified in the Makefile
+
+# End of script

### 2018-02-21 src/plot_enso.py

@@ -1,14 +1,14 @@
+import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
-import numpy as np
 
 enso = pd.read_csv('~/Documents/GitHub/floodsampling/floodsampling/data/ramesh2017.csv', index_col='year')
 
 # Plot Time Series
 enso.loc[10000:12500].plot(
     figsize=(9, 5),
-    linewidth=0.5,
     c='blue',
+    linewidth=0.5,
     legend=False,
     grid=True
 )

### 2018-02-21 src/calc_trend.py
ready for HPC

@@ -6,18 +6,14 @@ import seaborn as sns
 import pandas as pd
 import matplotlib.pyplot as plt
 
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 
 # Define Parameters
 N_try = np.array([10, 20, 30, 40, 50, 75, 100, 150])
 M_try = np.array([3, 5, 10, 20, 30, 50, 75])
-n_seq = 100 # PITIFULLY FEW!
-n_sim = 1000 # PITIFULLY SMALL!
+n_seq = 2000
+n_sim = 2000
 threshold = 5000
 figsize = (9, 6)
 

### 2018-02-21 src/plot_enso.py

@@ -2,7 +2,10 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 
-enso = pd.read_csv('~/Documents/GitHub/floodsampling/floodsampling/data/ramesh2017.csv', index_col='year')
+from floodsampling.util import get_data_path
+
+data_fname = os.path.join(get_data_path(), 'ramesh2017.csv')
+enso = pd.read_csv(data_fname, index_col='year')
 
 # Plot Time Series
 enso.loc[10000:12500].plot(

### 2018-02-21 src/plot_example_short.py

@@ -3,10 +3,6 @@ import xarray as xr
 import pandas as pd
 import matplotlib.pyplot as plt
 
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 

### 2018-02-21 src/plot_stationary_sequences.py

@@ -3,10 +3,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-import os
-import sys
-sys.path.insert(0, os.path.abspath('/Users/james/Documents/GitHub/floodsampling'))
-
 from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
 
 time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}

### 2018-02-21 writeup/main.tex
put floodsampling package in here for simplicity

@@
 To incorporate secular change into a statistical model, many studies have used time parameterizations on certain parameters.
-We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \):
+We consider a trend model where the data is log-normally distributed as \cref{eq:lognormal} with a linear time trend on both \( \mu \) and \( \sigma \).
+As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors:
 \begin{align}
@@
 \end{align}
-As for the stationary case, we fit posterior samples in a Bayesian framework using \texttt{stan} and apply weak priors.
-The influence of priors in this case is stronger than in the stationary case, and a reasonable choice of priors might more strongly favor \(\beta_mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
-However, in the context of a synthetic data exercise it seems reasonable to apply priors that are at most weakly informative.
+The influence of priors in this case is stronger than in the stationary case, and an alternative choice of priors might more strongly favor \(\beta_\mu=0\) and \(\beta_\sigma=0\), for example using the Horseshoe prior parameterization \citep{Piironen2016a}.
 
@@
 where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
+In \cref{sec:results} the bias and standard deviation (square root of variance) are averaged over all \(J\) sequences and plotted.\footnote{Check this code}
 
@@
 
-\section{Results}
+\section{Estimation Bias and Uncertainty}\label{sec:results}
 
@@
 
+\emph{Add more text once the results come through from Habanero -- they are running now!}
+
 \begin{figure}[ht]
@@
 \subsection{Trend Case}
+
+\emph{Add more text once the results come through from Habanero -- they are running now!}
 

### 2018-02-21 floodsampling/core.py

@@ -1,0 +1,167 @@
+"""Parent class for StreamflowCreator and FloodFit
+"""
+import os
+import pickle
+from hashlib import md5
+from collections import OrderedDict
+
+import numpy as np
+import xarray as xr
+
+from floodsampling.util import get_cache_path, safe_pkl_dump
+
+class BaseSequence:
+    """A parent class for streamflow sequences or monte carlo model draws
+
+    You shouldn't call this class directly.
+    It provides functionality primarily for file I/O and for repeating calculations
+    for many sequences, which are shared methods which would otherwise be duplicated.
+
+    Parameters
+    ----------
+    N : int
+        the length of the observational record, in years
+    M : int
+        the length of the future project planning period, in years
+    t0 : int, optional
+        the "zero" time (default 0)
+    n_seq : int, optional
+        the number of sequences to generate (default 3)
+
+    Attributes
+    ----------
+    time : dict
+        A dictionary with keys ['M', 'N', 't0', 'n_seq'] that describe the time
+    param : dict
+        The parameters of the model or models
+    """
+
+    def __init__(self, **kwargs):
+        self.time = {
+            'M': int(kwargs.pop('M')),
+            'N': int(kwargs.pop('N')),
+            't0': int(kwargs.pop('t0', 0)),
+            'n_seq': int(kwargs.pop('n_seq', 3))
+        }
+        self.param = {} # must be added by child class
+
+    def _get_attributes(self):
+        """Get the key parameters of the data as an ordered dictionary.
+
+        This makes it suitable for setting a file name, and also for
+        comparing the attributes of a data set read from file with the attributes
+        specified by the user.
+
+        Returns
+        -------
+        OrderedDict
+            The parameters of the model
+        """
+        attributes = self.param
+        attributes.update({
+            'M': self.time['M'],
+            'N': self.time['N'],
+            't0': self.time['t0'],
+            'n_seq': self.time['n_seq']
+        })
+        attributes = OrderedDict(attributes)
+        return attributes
+
+    def _get_filename(self):
+        """Get a file name
+
+        Uses the parameters of the model to build a dictionary of all the
+        key attributes of the model. Then converts them to a string
+        and hashes the output to a (shorter!) filename. Finally adds the path
+        to the data directory and the appropriate file suffix.
+
+        Returns
+        -------
+        str
+            A full path to the filename to which the data will be saved
+        """
+
+        attributes = self._get_attributes()
+
+        file_string = '' # initialize empty string
+        for key, val in attributes.items():
[…]

### 2018-02-21 floodsampling/fit.py

@@ -1,0 +1,450 @@
+"""Fit models to observations
+"""
+
+import os
+import warnings
+
+import pomegranate as pm
+import numpy as np
+import xarray as xr
+from tqdm import tqdm
+
+from .util import compile_model, get_data_path
+from .core import BaseSequence
+from .streamflow import StreamflowCreator
+
+class FloodFit(BaseSequence):
+    """A parent class for creating multiple synthetic streamflow sequences.
+
+    To implement your own method for creating streamflow sequences, create a child
+    class. Otherwise, you shouldn't call this class directly.
+
+    It is built on `floodfit.core.BaseSequence`, which you should also not plan
+    on calling directly.
+
+    .. note:: Time Indexing
+        The time is indexed so that the "historical record" is defined as
+        the period from time_0 - n_years + 1 to time_0, inclusive. The "future period"
+        is defined as the period from time_0 + 1 to time_0 + m_years, inclusive.
+
+    Parameters
+    ----------
+    sflow : StreamflowCreator
+        A StreamflowCreator instance
+    n_sim : int
+        Number of samples to return
+    model_name : str
+        The name of the model
+
+    Attributes
+    ----------
+    time : dict
+        A dictionary that describes the time parameters of the model.
+        Should have keys ``['M', 'M', t0', 'n_seq', 'n_sim']``.
+    param : dict
+        The parameters of _both_ the generating model and fitting model
+    sflow : StreamflowCreator
+        The StreamflowCreator object that creates the streamflow sequences
+    """
+
+    def __init__(self, sflow, n_sim, **kwargs):
+
+        model_name = kwargs.pop('model_name')
+        assert isinstance(model_name, str), 'model_name must be a string'
+        assert isinstance(sflow, StreamflowCreator), 'sflow must be a StreamflowCreator'
+
+        super().__init__(
+            N=sflow.time['N'], M=sflow.time['M'],
+            t0=sflow.time['t0'], n_seq=sflow.time['n_seq']
+        )
+        self.time.update({'n_sim': n_sim})
+        self.param.update(sflow.param)
+        self.param.update({'category': 'fit'})
+        self.param.update({'model_name': model_name})
+        self.sflow = sflow
+
+    def _calculate_one_seq(self, one_seq):
+        """Simulate a single sequence of annual maximum flood peaks
+
+        This function needs to be implemented by the child class and is not
+        implemented in this parent class.
+
+        Parameters
+        ----------
+        one_seq : xr.DataArray
[…]

### 2018-02-21 floodsampling/streamflow.py

@@ -1,0 +1,356 @@
+"""Generate synthetic streamflow sequences conditional on climate variables.
+
+Some recent literature has considered statistical-dynaimcal models for fitting
+flood time series where the flood or hydrological quantity of interest depends
+on some (fully observed) climate covariates. This module provides some classes
+for generating sequences that reverse-engineer this approach and in which the
+observed annual-maximum streamflow depends on some climate covariates, plus
+(optionally) time.
+"""
+
+import warnings
+import os
+from tqdm import tqdm
+import xarray as xr
+import numpy as np
+import pandas as pd
+import pomegranate as pm
+
+from .core import BaseSequence
+from .util import get_data_path
+
+class StreamflowCreator(BaseSequence):
+    """A parent class for creating multiple synthetic streamflow sequences.
+
+    To implement your own method for creating streamflow sequences, create a child
+    class. Otherwise, you shouldn't call this class directly.
+
+    It is built on `floodfit.core.BaseSequence`, which you should also not plan
+    on calling directly.
+
+    .. note:: Time Indexing
+        The time is indexed so that the "historical record" is defined as
+        the period from time_0 - n_years + 1 to time_0, inclusive. The "future period"
+        is defined as the period from time_0 + 1 to time_0 + m_years, inclusive.
+
+    Parameters
+    ----------
+    N : int
+        the length of the observational record, in years
+    M : int
+        the length of the future project planning period, in years
+    t0 : int, optional
+        the "zero" time (default 0)
+    model_name : str, optional
+        A name for the model used to generate the sequences (default None)
+    n_seq : int, optional
+        the number of sequences to generate (default 1)
+
+    Attributes
+    ----------
+    time : dict
+        A dictionary that describes the time parameters of the model.
+        Should have keys ``['M', 'M', t0', 'n_seq']``.
+    param : dict
+        The parameters of the model or models
+    """
+
+    def __init__(self, **kwargs):
+
+        model_name = kwargs.pop('model_name')
+        assert isinstance(model_name, str), 'model_name must be a string'
+
+        super().__init__(**kwargs)
+
+        self.param.update({'category': 'sequence'})
+        self.param.update({'model_name': model_name})
+
+    def _calculate_one_seq(self):
+        """Simulate a single sequence of annual maximum flood peaks
+
[…]

### 2018-02-21 floodsampling/tests.py

@@ -1,0 +1,55 @@
+import numpy as np
+import pytest
+
+from .streamflow import CZNINO3LN2
+from .fit import StationaryLN2Stan, TrendLN2Stan, HMM
+from .util import clear_cache
+
+clear_cache()
+
+def assert_get_sequences(sflow):
+    """Make sure we are getting data of the right shape
+    """
+    data = sflow.get_data()
+    n_seq = sflow.time['n_seq']
+    n_years = sflow.time['M'] + sflow.time['N']
+    assert data['year'].size == n_years
+    assert data['sequence'].size == n_seq
+
+def assert_get_fits(fitted):
+    """Make sure we can get fits
+    """
+    data = fitted.get_data()
+    n_seq = fitted.time['n_seq']
+    n_sim = fitted.time['n_sim']
+    n_years = fitted.time['M']
+    assert data['year'].size == n_years
+    assert data['sequence'].size == n_seq
+    assert data['sim'].size == n_sim
+
+def assert_bias(fitted):
+    """Make sure we can get fits
+    """
+    bias, _ = fitted.decompose_variance(threshold=2000)
+    assert np.abs(bias) < 1, 'bias cannot be greater than 1'
+
+def test_loop():
+    """For each possible combination, run the assertion.
+
+    The idea here is to try many different combinations of generating model,
+    fitting model, and parameters.
+    In particular, try different values of T_0
+    """
+    M_vec = [20, 50, 20, 50]
+    N_vec = [50, 20, 50, 20]
+    t0_vec = [0, 10, -10, 0]
+    all_gen_fun = [CZNINO3LN2]
+    all_fit_fun = [StationaryLN2Stan, TrendLN2Stan, HMM]
+    for (M, N, t0) in zip(M_vec, N_vec, t0_vec):
+        for gen_fun in all_gen_fun:
+            for fit_fun in all_fit_fun:
+                sflow = gen_fun(N=N, M=M, n_seq=3, t0=t0)
+                assert_get_sequences(sflow)
+                fitted = fit_fun(sflow=sflow, n_sim=10)
+                assert_get_fits(fitted)
+                assert_bias(fitted)

### 2018-02-21 floodsampling/util.py

@@ -1,0 +1,123 @@
+"""Utility functions for floodfit
+
+This module provides some functions which are useful to the entire project and
+whose utility is not limited to a single sub-module.
+"""
+from __future__ import absolute_import
+from __future__ import division
+from __future__ import print_function
+
+from pkg_resources import resource_filename
+import os
+import pickle
+from hashlib import md5
+import stat
+import pystan
+
+def compile_model(filename, model_name=''):
+    """Compile a stan model only if it hasn't already been compiled
+
+    This will automatically cache models - great if you're just running a
+    script on the command line. This code was taken nearly verbatim from
+    `Pystan` documentation.
+
+    Parameters
+    ----------
+    model_code : str
+        the model code
+    filename : str
+        the path to the .stan file
+
+    Returns
+    -------
+    pystan.StanModel
+        the compiled model object
+
+    See Also
+    --------
+    - `Pystan Documentation <http://pystan.readthedocs.io/en/latest/avoiding_recompilation.html/>`_
+    - `Aki Vehtari's Stan Utilities <https://github.com/avehtari/BDA_py_demos/blob/new_pystan_demos/utilities_and_data/stan_utility.py>`
+    """
+
+    cache_dir = get_cache_path()
+
+    with open(filename) as file:
+        model_code = file.read()
+        code_hash = md5(model_code.encode('ascii')).hexdigest()
+        cache_fn = 'cached-{}-{}.pkl'.format(model_name, code_hash)
+        cache_fn = os.path.join(cache_dir, 'stan', cache_fn)
+        try:
+            smodel = pickle.load(open(cache_fn, 'rb'))
+        except BaseException:
+            smodel = pystan.StanModel(model_code=model_code)
+            safe_pkl_dump(obj=smodel, fname=cache_fn)
+
+    return smodel
+
+def get_data_path():
+    """Get the folder where all data is saved
+
+    Returns
+    -------
+    str
+        The full path to the directory where the data is stored.
+    """
+    data_dir = resource_filename('floodsampling', 'data')
+    return data_dir
+
+def get_cache_path():
+    """Get the folder where cached simulations are saved
+
+    Returns
+    -------
+    str
+        The full path to the directory where the cache is stored.
+    """
+    data_dir = get_data_path()
+    cache_dir = os.path.join(data_dir, 'cached')
+    cache_dir = os.path.abspath(cache_dir)
+    if not os.path.isdir(cache_dir):
+        os.makedirs(cache_dir)
[…]

### 2018-02-21 src/ENSO.R

@@ -1,0 +1,22 @@
+if(!require(pacman)) install.packages('pacman')
+pacman::p_load(dplyr, readr, ggplot2, ggthemes, xts)
+
+# READ THE RAW DATA
+raw_file <- '~/Documents/GitHub/floodsampling/floodsampling/data/ramesh2017.csv'
+nino <- read_csv(raw_file)
+
+# PLOT EL NINO TIME SERIES
+nino_plot <- 
+  nino %>%
+  filter(year >= 1000 & year < 2000) %>%
+  ggplot(aes(x=year, y=nino3)) + 
+  geom_line() +
+  labs(x='Simulated Year', y='NINO3 Index')
+
+k <- kernel("daniell", m = c(40, 40, 40))
+smooth_spec <- spec.pgram(nino$nino3, kernel = k, taper = 0, plot=FALSE)
+
+par(mfrow = c(1, 2))
+plot(nino3 ~ year, data=filter(nino, year >= 1000 & year < 2000), main='NINO3 Time Series', type='line', xlab='Year', ylab='NINO3')
+plot(smooth_spec, log='no', main='Smoothed Periodogram of NINO3', ylab='Power')
+par(mfrow = c(1, 1))

### 2018-02-21 src/calc_stationary.py

@@ -12,10 +12,9 @@ from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 # Define Parameters
 N_try = np.array([10, 20, 30, 40, 50, 75, 100, 150])
 M_try = np.array([3, 5, 10, 20, 30, 50, 75])
-n_seq = 2000
-n_sim = 2000
+n_seq = 2
+n_sim = 2
 threshold = 5000
-figsize = (9, 6)
 
 # Initialize the bias and variance
 gen_names = np.array(['enso', 'markov'])

### 2018-02-21 src/calc_stationary.py
change number of runs

@@ -12,8 +12,8 @@ from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 # Define Parameters
 N_try = np.array([10, 20, 30, 40, 50, 75, 100, 150])
 M_try = np.array([3, 5, 10, 20, 30, 50, 75])
-n_seq = 2
-n_sim = 2
+n_seq = 2000
+n_sim = 2000
 threshold = 5000
 
 # Initialize the bias and variance

### 2018-02-21 slurm.sh
a few tweaks for slurm

@@ -10,8 +10,9 @@
 #SBATCH --exclusive
 #SBATCH --time=10:00:00         # The time the job will take to run.
 
-module load anaconda/3-4.4.0
-source activate test_env
+module load anaconda            # load the anaconda module
+make environment                # create the conda env (comment if not 1st time)
+source activate floodsampling   # activate the conda environment
 make simulate                   # run the make command specified in the Makefile
 
 # End of script

### 2018-02-21 src/calc_trend.py

@@ -15,7 +15,6 @@ M_try = np.array([3, 5, 10, 20, 30, 50, 75])
 n_seq = 2000
 n_sim = 2000
 threshold = 5000
-figsize = (9, 6)
 
 # Initialize the bias and variance
 gen_names = np.array(['enso', 'markov'])

### 2018-02-21 slurm.sh
dont make environment in slurm

@@ -11,7 +11,6 @@
 #SBATCH --time=10:00:00         # The time the job will take to run.
 
 module load anaconda            # load the anaconda module
-make environment                # create the conda env (comment if not 1st time)
 source activate floodsampling   # activate the conda environment
 make simulate                   # run the make command specified in the Makefile
 

### 2018-02-21 make_environment.sh
separate the make commands

@@ -1,0 +1,16 @@
+#!/bin/sh
+#
+# Run the code on a slurm server
+#
+# Replace <ACCOUNT> with your account name before submitting.
+#
+#SBATCH --account=cwc           # The account name for the job.
+#SBATCH --job-name=JDGMN01      # The job name.
+#SBATCH -N 1                    # The number of nodes to use
+#SBATCH --exclusive
+#SBATCH --time=20:00         # The time the job will take to run.
+
+module load anaconda            # load the anaconda module
+make environment                # create the conda env (comment if not 1st time)
+
+# End of script

### 2018-02-21 make_environment.sh
use updated version of anaconda

@@ -8,9 +8,9 @@
 #SBATCH --job-name=JDGMN01      # The job name.
 #SBATCH -N 1                    # The number of nodes to use
 #SBATCH --exclusive
-#SBATCH --time=20:00         # The time the job will take to run.
+#SBATCH --time=20:00            # The time the job will take to run.
 
-module load anaconda            # load the anaconda module
+module load anaconda/3-4.4.0    # load the anaconda module
 make environment                # create the conda env (comment if not 1st time)
 
 # End of script

### 2018-02-21 make_simulate.sh

@@ -10,7 +10,7 @@
 #SBATCH --exclusive
 #SBATCH --time=10:00:00         # The time the job will take to run.
 
-module load anaconda            # load the anaconda module
+ module load anaconda/3-4.4.0   # load the anaconda module
 source activate floodsampling   # activate the conda environment
 make simulate                   # run the make command specified in the Makefile
 

### 2018-02-22 make_fit_stationary.sh
separate data analysis into 2 components: stationary and trend

@@ -1,0 +1,17 @@
+#!/bin/sh
+#
+# Run the code on a slurm server
+#
+# Replace <ACCOUNT> with your account name before submitting.
+#
+#SBATCH --account=cwc           # The account name for the job.
+#SBATCH --job-name=JDGMN01      # The job name.
+#SBATCH -N 1                    # The number of nodes to use
+#SBATCH --exclusive
+#SBATCH --time=20:00:00         # The time the job will take to run.
+
+ module load anaconda/3-4.4.0   # load the anaconda module
+source activate floodsampling   # activate the conda environment
+make data/stationary.nc                   # run the make command specified in the Makefile
+
+# End of script

### 2018-02-26 src/floodsampling/core.py
use netcdf4

@@ -35,7 +35,7 @@ class BaseSequence:
         The parameters of the model or models
     """
 
-    def __init__(self, **kwargs):
+    def __init__(self, category, model_name, **kwargs):
         self.time = {
             'M': int(kwargs.pop('M')),
             'N': int(kwargs.pop('N')),
@@ -43,6 +43,8 @@ class BaseSequence:
             'n_seq': int(kwargs.pop('n_seq', 3))
         }
         self.param = {} # must be added by child class
+        self.category = category
+        self.model_name = model_name
 
     def _get_attributes(self):
         """Get the key parameters of the data as an ordered dictionary.
@@ -82,12 +84,12 @@ class BaseSequence:
 
         attributes = self._get_attributes()
 
-        file_string = '' # initialize empty string
+        file_string = 'floodsampling' # initialize empty string
         for key, val in attributes.items():
-            file_string += 'simulated_{}{}'.format(key, val)
+            file_string += '{}{}'.format(key, val)
 
         file_string = md5(file_string.encode('ascii')).hexdigest()
-        file_string += '.pkl'
+        file_string += '.nc'
 
         file_dir = os.path.join(get_cache_path(), self.model_name)
         file_dir = os.path.abspath(file_dir)
@@ -99,8 +101,7 @@ class BaseSequence:
     def to_file(self, data):
         """Save the model sequences to file
 
-        Uses the `util.safe_pkl_dump()` function to create the desired parent directory
-        and delete existing file with same name, if necessary, and then dump to file.
+        Create the desired parent directory and delete existing file with same name, if necessary, and then dump to file.
 
         Parameters
         ----------
@@ -111,7 +112,7 @@ class BaseSequence:
 
         data.attrs = self._get_attributes() # force the attrs object of the xarray object
         fname = self._get_filename()
-        safe_pkl_dump(data, fname=fname)
+        data.to_netcdf(path=fname, mode='w', format='NETCDF4')
 
     def from_file(self):
         """Get data from file
@@ -131,8 +132,7 @@ class BaseSequence:
         """
         try:
             fname = self._get_filename()
-            with open(fname, "rb") as f:
-                data = pickle.load(f)
+            data = xr.open_dataarray(fname)    
             attrs_desired = self._get_attributes()
             attr_observed = data.attrs
             success = attrs_desired == attr_observed

### 2018-02-26 src/floodsampling/core.py
use categories for better organizing

@@ -91,7 +91,7 @@ class BaseSequence:
         file_string = md5(file_string.encode('ascii')).hexdigest()
         file_string += '.nc'
 
-        file_dir = os.path.join(get_cache_path(), self.model_name)
+        file_dir = os.path.join(get_cache_path(), self.category, self.model_name)
         file_dir = os.path.abspath(file_dir)
         if not os.path.isdir(file_dir):
             os.makedirs(file_dir)

### 2018-02-26 src/floodsampling/fit.py

@@ -54,15 +54,13 @@ class FloodFit(BaseSequence):
         assert isinstance(sflow, StreamflowCreator), 'sflow must be a StreamflowCreator'
 
         super().__init__(
+            category='fit', model_name=model_name,
             N=sflow.time['N'], M=sflow.time['M'],
             t0=sflow.time['t0'], n_seq=sflow.time['n_seq']
         )
         self.time.update({'n_sim': n_sim})
         self.param.update(sflow.param)
-        self.param.update({'category': 'fit'})
         self.sflow = sflow
-
-        self.model_name = model_name
 
     def _calculate_one_seq(self, one_seq):
         """Simulate a single sequence of annual maximum flood peaks

### 2018-02-26 src/floodsampling/streamflow.py

@@ -59,11 +59,7 @@ class StreamflowCreator(BaseSequence):
 
         model_name = kwargs.pop('model_name')
         assert isinstance(model_name, str), 'model_name must be a string'
-
-        super().__init__(**kwargs)
-
-        self.param.update({'category': 'sequence'})
-        self.model_name = model_name
+        super().__init__(category='sequence', model_name=model_name, **kwargs)
 
     def _calculate_one_seq(self):
         """Simulate a single sequence of annual maximum flood peaks

### 2018-02-26 src/calc_stationary.py
should be working fine, but use only a few sequences for testing purposes

@@ -1,137 +1,0 @@
-"""Get simulations of stationary data and save to file
-"""
-import argparse
-import itertools
-import numpy as np
-import xarray as xr
-import seaborn as sns
-import pandas as pd
-
-from joblib import Parallel, delayed
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-from floodsampling.util import compile_model
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--outfile", help="the filename of the data to save")
-parser.add_argument("--n_jobs", type=int, help="number of jobs to run in parallel")
-
-def get_gen_fun(fun_name, N, M, t0, n_seq):
-    """Get a generating function as a function of the fun_name 0 or 1
-    """
-    if fun_name == 'NINO3':
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0,
-            coeff_var=0.1, sigma_min=0.01,
-            model_name='CZNINO3LN2_Stationary'
-        )
-    elif fun_name == 'Markov':
-        gen_fun = TwoStateSymmetricMarkovLN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01,
-            model_name='TwoStateSymmetricMarkovLN2_Stationary'
-        )
-    else:
-        raise ValueError('Invalid Index')
-    return gen_fun
-
-def get_fit_fun(fun_name, gen_fun, n_sim):
-    """Get a fit function as a function of the fun_name 0 or 1 or 2
-    """
-    if fun_name == 'LN2 Stationary':
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000, model_name='StationaryLN2Stan_Stationary')
-    elif fun_name == 'LN2 Trend':
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000, model_name='TrendLN2Stan_Stationary')
-    elif fun_name == 'HMM':
[…]

### 2018-02-26 src/calc_trend.py

@@ -1,137 +1,0 @@
-"""Get simulations of stationary data and save to file
-"""
-import argparse
-import itertools
-import numpy as np
-import xarray as xr
-import seaborn as sns
-import pandas as pd
-
-from joblib import Parallel, delayed
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-from floodsampling.util import compile_model
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--outfile", help="the filename of the data to save")
-parser.add_argument("--n_jobs", type=int, help="number of jobs to run in parallel")
-
-def get_gen_fun(fun_name, N, M, t0, n_seq):
-    """Get a generating function as a function of the fun_name 0 or 1
-    """
-    if fun_name == 'NINO3':
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0.015,
-            coeff_var=0.1, sigma_min=0.01,
-            model_name='CZNINO3LN2_Trend'
-        )
-    elif fun_name == 'Markov':
-        gen_fun = TwoStateSymmetricMarkovLN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_1=6.75, mu_0 = 6, gamma_1=0.015, gamma_2=0, pi=0.9,
-            coeff_var = 0.1, sigma_min = 0.01,
-            model_name='TwoStateSymmetricMarkovLN2_Trend'
-        )
-    else:
-        raise ValueError('Invalid Index')
-    return gen_fun
-
-def get_fit_fun(fun_name, gen_fun, n_sim):
-    """Get a fit function as a function of the fun_name 0 or 1 or 2
-    """
-    if fun_name == 'LN2 Stationary':
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000, model_name='StationaryLN2Stan_Trend')
-    elif fun_name == 'LN2 Trend':
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000, model_name='TrendLN2Stan_Trend')
-    elif fun_name == 'HMM':
[…]

### 2018-02-26 src/floodsampling/core.py

@@ -112,6 +112,11 @@ class BaseSequence:
 
         data.attrs = self._get_attributes() # force the attrs object of the xarray object
         fname = self._get_filename()
+        par_dir = os.path.dirname(fname)
+        if not os.path.isdir(par_dir):
+            os.makedirs(par_dir)
+        if os.path.isfile(fname):
+            os.remove(fname)
         data.to_netcdf(path=fname, mode='w', format='NETCDF4')
 
     def from_file(self):

### 2018-02-26 src/get_bias_variance.py

@@ -1,0 +1,212 @@
+"""Get simulations of stationary data and save to file
+
+This is done in parallel, but in an intelligent way.
+First, the stan models are compiled -- this is not done in parallel because it should
+only be done once.
+Second, the synthetic streamflow sequences are generated for the longest M available.
+This is done in parallel.
+Finally, the statistical fitting models are used to generate replicated sequences for
+the longest M available; this is also done in parallel.
+For many values of M* to try, the first M* years of "true" and "fit" data are selected and
+the bias and variance computed.
+"""
+import argparse
+import itertools
+from collections import OrderedDict
+import numpy as np
+import xarray as xr
+import seaborn as sns
+import pandas as pd
+
+from joblib import Parallel, delayed
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
+from floodsampling.util import compile_model
+
+parser = argparse.ArgumentParser()
+parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--n_jobs", type=int, help="number of jobs to run in parallel")
+parser.add_argument("--gamma", type=float, help="the coefficient to put on time")
+
+def get_gen_fun(fun_name, **kwargs):
+    """Get a generating function as a function of the fun_name 0 or 1
+    """
+    N = kwargs.pop('N')
+    M = kwargs.pop('M')
+    n_seq = kwargs.pop('n_seq')
+    t0 = kwargs.pop('t0')
+    gamma = kwargs.pop('gamma')
+    
+    mu0 = kwargs.pop('mu0')
+    beta_mu = kwargs.pop('beta_mu')
+    coeff_var = kwargs.pop('coeff_var')
+    sigma_min = kwargs.pop('sigma_min')
+
+    mu_1 = kwargs.pop('mu_1')
+    mu_2 = kwargs.pop('mu_2')
+    pi = kwargs.pop('pi')
+
+    if fun_name == 'NINO3':
+        gen_fun = CZNINO3LN2(
+            N=N, M=M, t0=t0, n_seq=n_seq,
+            mu_0=mu0, beta_mu=beta_mu, gamma=gamma,
+            coeff_var=coeff_var, sigma_min=sigma_min,
+        )
[…]

### 2018-02-26 make_simulate.sh
bug squash for testing

@@ -6,9 +6,9 @@
 #
 #SBATCH --account=cwc           # The account name for the job.
 #SBATCH --job-name=STATIONARY   # The job name.
-#SBATCH -N 4                    # The number of nodes to use
+#SBATCH -N 1                    # The number of nodes to use
 #SBATCH --exclusive
-#SBATCH --time=3:59:00          # The time the job will take to run.
+#SBATCH --time=10:00          # The time the job will take to run.
 
 module load anaconda/3-4.4.0    # load the anaconda module
 source activate floodsampling   # activate the conda environment

### 2018-02-26 src/floodsampling/core.py

@@ -115,8 +115,6 @@ class BaseSequence:
         par_dir = os.path.dirname(fname)
         if not os.path.isdir(par_dir):
             os.makedirs(par_dir)
-        if os.path.isfile(fname):
-            os.remove(fname)
         data.to_netcdf(path=fname, mode='w', format='NETCDF4')
 
     def from_file(self):

### 2018-02-26 src/get_bias_variance.py
next step of testing: try parallel execution

@@ -153,7 +153,7 @@ def main():
     n_jobs = args.n_jobs
     parameters = {
         't0': 0,
-        'n_seq': 2,
+        'n_seq': 3,
         'n_sim': 5,
         'gamma': args.gamma,
         'mu0': 6,

### 2018-02-26 src/floodsampling/core.py
load data all the way when reading

@@ -135,7 +135,7 @@ class BaseSequence:
         """
         try:
             fname = self._get_filename()
-            data = xr.open_dataarray(fname)    
+            data = xr.open_dataarray(fname).load()
             attrs_desired = self._get_attributes()
             attr_observed = data.attrs
             success = attrs_desired == attr_observed

### 2018-02-26 src/floodsampling/core.py
go back to using pickle instead of netcdf for storage

@@ -89,7 +89,7 @@ class BaseSequence:
             file_string += '{}{}'.format(key, val)
 
         file_string = md5(file_string.encode('ascii')).hexdigest()
-        file_string += '.nc'
+        file_string += '.pkl'
 
         file_dir = os.path.join(get_cache_path(), self.category, self.model_name)
         file_dir = os.path.abspath(file_dir)
@@ -112,10 +112,7 @@ class BaseSequence:
 
         data.attrs = self._get_attributes() # force the attrs object of the xarray object
         fname = self._get_filename()
-        par_dir = os.path.dirname(fname)
-        if not os.path.isdir(par_dir):
-            os.makedirs(par_dir)
-        data.to_netcdf(path=fname, mode='w', format='NETCDF4')
+        safe_pkl_dump(data, fname=fname)
 
     def from_file(self):
         """Get data from file
@@ -135,7 +132,8 @@ class BaseSequence:
         """
         try:
             fname = self._get_filename()
-            data = xr.open_dataarray(fname).load()
+            with open(fname, 'rb') as f:
+                data = pickle.load(f)
             attrs_desired = self._get_attributes()
             attr_observed = data.attrs
             success = attrs_desired == attr_observed

### 2018-02-26 make_simulate.sh
use more cores and time to run with 1000 seq and 2000 sim

@@ -8,7 +8,7 @@
 #SBATCH --job-name=STATIONARY   # The job name.
 #SBATCH -N 1                    # The number of nodes to use
 #SBATCH --exclusive
-#SBATCH --time=10:00          # The time the job will take to run.
+#SBATCH --time=3:59:00          # The time the job will take to run.
 
 module load anaconda/3-4.4.0    # load the anaconda module
 source activate floodsampling   # activate the conda environment

### 2018-02-26 src/get_bias_variance.py

@@ -153,8 +153,8 @@ def main():
     n_jobs = args.n_jobs
     parameters = {
         't0': 0,
-        'n_seq': 3,
-        'n_sim': 5,
+        'n_seq': 1000,
+        'n_sim': 2000,
         'gamma': args.gamma,
         'mu0': 6,
         'beta_mu': 0.5,

### 2018-02-26 launch-notebook.sh
Make plots ready for paper draft

@@ -1,14 +1,0 @@
-#!/bin/bash
-
-#SBATCH --account=cwc
-#SBATCH -J notebook
-#SBATCH --time=6:00:00
-
-
-# Setup Environment
-module load anaconda
-source activate floodsampling
-
-export XDG_RUNTIME_DIR=""
-
-jupyter notebook --no-browser --ip "*" --notebook-dir $HOME

### 2018-02-26 src/plot_bias.py

@@ -1,0 +1,44 @@
+import xarray as xr
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+import argparse
+
+parser = argparse.ArgumentParser()
+parser = argparse.ArgumentParser()
+parser.add_argument("--infile", help="the filename of the saved bias/variance data")
+parser.add_argument("--outfile", help="the filename of the figure to save")
+
+def main():
+    args = parser.parse_args()
+
+    fit = xr.open_dataset(args.infile)
+    fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
+
+    fig, axes = plt.subplots(
+        nrows=fit['Generating Function'].size, 
+        ncols=fit['Fitting Function'].size, 
+        sharex=True, sharey=True, figsize=(10, 6)
+    )
+    for col,f in enumerate(fit['Fitting Function'].values):
+        for row,g in enumerate(fit['Generating Function'].values):
+            sub = fit_df.loc[np.logical_and(fit_df['Generating Function'] == g, fit_df['Fitting Function'] == f)].reset_index().pivot('M', 'N', 'bias')
+            ax=axes[row, col]
+            sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
+            if row == 0:
+                ax.set_title(f)
+            if col  == (fit['Fitting Function'].size - 1):
+                ax.set_ylabel(g)
+                ax.yaxis.set_label_position('right')
+
+    fig.tight_layout()
+    fig.subplots_adjust(right=0.85)
+    cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
+    sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'Expected Bias'})
+
+    fig.savefig(args.outfile)
+
+if __name__ == '__main__':
+    main()

### 2018-02-26 src/plot_example.py

@@ -9,8 +9,9 @@ from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
 
 parser = argparse.ArgumentParser()
 parser.add_argument("--outfile", help="the filename of the data to save")
+parser.add_argument("--N", type=int, help="the filename of the data to save")
 
-def get_gen_fun(index):
+def get_gen_fun(index, N, M, t0, n_seq):
     """Get a generating function as a function of the index 0 or 1
     """
     if index == 0:
@@ -29,7 +30,7 @@ def get_gen_fun(index):
         raise ValueError('Invalid Index')
     return gen_fun
 
-def get_fit_fun(index, gen_fun):
+def get_fit_fun(index, gen_fun, n_sim):
     """Get a fit function as a function of the index 0 or 1 or 2
     """
     if index == 0:
@@ -43,11 +44,12 @@ def get_fit_fun(index, gen_fun):
     return fit_fun
 
 def main():
+    
+    args = parser.parse_args()
 
     M = 150
-    N = 150
     n_seq = 1
-    n_sim = 100
+    n_sim = 500
 
     # Initialize the bias and variance
     gen_names = np.array(['NINO3 Stationary', 'NINO3 Trend'])
@@ -55,23 +57,25 @@ def main():
 
     fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
     for g, gen_fun_name in enumerate(gen_names):
-        gen_fun = get_gen_fun(g)
+        gen_fun = get_gen_fun(g, N=args.N, M=M, t0=0, n_seq=n_seq)
         for f, fit_fun_name in enumerate(fit_names):
-            fit_fun = get_fit_fun(f, gen_fun=gen_fun)
+            fit_fun = get_fit_fun(f, gen_fun=gen_fun, n_sim=n_sim)
             gen_dat = gen_fun.get_data('all')
             fit_dat = fit_fun.get_data('future')
             ax = axes[f, g]
             for i in np.arange(n_sim):
-                ax.plot(fit_dat['year'], fit_dat.sel(sim=i).values.ravel(), c='gray', linewidth=0.5)
+                ax.plot(fit_dat['year'], fit_dat.sel(sim=i).values.ravel(), c='gray', linewidth=0.1, alpha=0.1)
             ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue', linewidth=1)
             ax.semilogy()
             ax.grid()
[…]

### 2018-02-26 src/plot_example_short.py

@@ -1,80 +1,0 @@
-import argparse
-import numpy as np
-import xarray as xr
-import pandas as pd
-import matplotlib.pyplot as plt
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-from floodsampling.fit import TrendLN2Stan, StationaryLN2Stan, HMM
-
-parser = argparse.ArgumentParser()
-parser.add_argument("--outfile", help="the filename of the data to save")
-
-def get_gen_fun(index):
-    """Get a generating function as a function of the index 0 or 1
-    """
-    if index == 0:
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0,
-            coeff_var=0.1, sigma_min=0.01
-        )
-    elif index == 1:
-        gen_fun = CZNINO3LN2(
-            N=N, M=M, t0=0, n_seq=n_seq,
-            mu_0=6, beta_mu=0.5, gamma=0.01,
-            coeff_var=0.1, sigma_min=0.01
-        )
-    else:
-        raise ValueError('Invalid Index')
-    return gen_fun
-
-def get_fit_fun(index, gen_fun):
-    """Get a fit function as a function of the index 0 or 1 or 2
-    """
-    if index == 0:
-        fit_fun = StationaryLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1000)
-    elif index == 1:
-        fit_fun = TrendLN2Stan(sflow=gen_fun, n_sim=n_sim, chains=1, warmup=1001)
-    elif index == 2:
-        fit_fun = HMM(sflow=gen_fun, n_sim=n_sim, n_components=2, n_init=50)
-    else:
-        raise ValueError('Invalid Index')
-    return fit_fun
-
-def main():
-
-    M = 150
-    N = 30
-    n_seq = 1
-    n_sim = 100
-
-    # Initialize the bias and variance
-    gen_names = np.array(['NINO3 Stationary', 'NINO3 Trend'])
-    fit_names = np.array(['LN2 Stationary', 'LN2 Trend', 'HMM'])
-
-    fig, axes = plt.subplots(nrows=3, ncols=2, figsize=(12, 7), sharex=True, sharey=True)
-    for g, gen_fun_name in enumerate(gen_names):
-        gen_fun = get_gen_fun(g)
-        for f, fit_fun_name in enumerate(fit_names):
[…]

### 2018-02-26 src/plot_sequences.py

@@ -1,0 +1,65 @@
+"""Stationary
+"""
+import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+
+from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
+
+parser = argparse.ArgumentParser()
+parser.add_argument("--gamma", type=float, help="the trand parameter")
+parser.add_argument("--outfile", help="the filename of the data to save")
+
+def main():
+
+    args = parser.parse_args()
+
+    time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
+    n_seq_plot = 50 # how many sequences to plot
+    threshold = 2500
+
+    enso_gen = CZNINO3LN2(
+        N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+        mu_0=6, beta_mu=0.5, gamma=args.gamma,
+        coeff_var=0.1, sigma_min=0.01, 
+        n_seq=time_param['n_seq'],
+    )
+    markov_gen = TwoStateSymmetricMarkovLN2(
+        N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
+        mu_1=6.55, mu_0=5.7, gamma_1=args.gamma, gamma_2=0, pi=0.9,
+        coeff_var = 0.1, sigma_min = 0.01,
+        n_seq=time_param['n_seq'],
+    )
+    enso_seq = enso_gen.get_data()
+    markov_seq = markov_gen.get_data()
+
+    # Setup the figure
+    fig, axes = plt.subplots(nrows=2, ncols=1, sharey=True, figsize=(12, 5), sharex='col')
+
+    ax = axes[0]
+    for i in np.arange(n_seq_plot-1):
+        enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
+    enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
+    ax.axvline(0, c='black', linewidth=0.6)
+    ax.set_xlabel('')
+    ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
+
+    ax = axes[1]
+    for i in np.arange(n_seq_plot-1):
+        markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
[…]

### 2018-02-26 src/plot_stationary.py

@@ -1,50 +1,0 @@
-import xarray as xr
-import pandas as pd
-import numpy as np
-import seaborn as sns
-import matplotlib.pyplot as plt
-
-fit = xr.open_dataset('data/stationary.nc')
-fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
-
-# PLOT THE BIAS
-fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
-for col,f in enumerate(fit['fit_fun'].values):
-    for row,g in enumerate(fit['gen_fun'].values):
-        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
-        ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
-        if row == 0:
-            ax.set_title(f)
-        if col  == (fit['fit_fun'].size - 1):
-            ax.set_ylabel(g)
-            ax.yaxis.set_label_position('right')
-
-fig.tight_layout()
-fig.subplots_adjust(right=0.85)
-cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'bias'})
-
-fig.savefig('figs/stationary_bias.pdf')
-
-
-
-# PLOT THE VARIANCE
-fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
-for col,f in enumerate(fit['fit_fun'].values):
-    for row,g in enumerate(fit['gen_fun'].values):
-        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
-        ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.15, cmap='viridis', cbar=False, linewidths=0.1)
-        if row == 0:
-            ax.set_title(f)
-        if col  == (fit['fit_fun'].size - 1):
-            ax.set_ylabel(g)
[…]

### 2018-02-26 src/plot_stationary_sequences.py

@@ -1,58 +1,0 @@
-"""Stationary
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-
-time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
-n_seq_plot = 50 # how many sequences to plot
-threshold = 2500
-
-enso_gen = CZNINO3LN2(
-    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_0=6, beta_mu=0.5, gamma=0,
-    coeff_var=0.1, sigma_min=0.01, 
-    n_seq=time_param['n_seq'],
-)
-markov_gen = TwoStateSymmetricMarkovLN2(
-    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_1=6.55, mu_0=5.7, gamma_1=0, gamma_2=0, pi=0.9,
-    coeff_var = 0.1, sigma_min = 0.01,
-    n_seq=time_param['n_seq'],
-)
-enso_seq = enso_gen.get_data()
-markov_seq = markov_gen.get_data()
-
-# Setup the figure
-fig, axes = plt.subplots(nrows=2, ncols=1, sharey=True, figsize=(12, 5), sharex='col')
-
-ax = axes[0]
-for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
-enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
-ax.axvline(0, c='black', linewidth=0.6)
-ax.set_xlabel('')
-ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
-
-ax = axes[1]
-for i in np.arange(n_seq_plot-1):
-    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
-markov_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
-ax.axvline(0, c='black', linewidth=0.6)
-ax.set_title('{} Markov Chain-Based Sequences'.format(n_seq_plot))
-
-
-for ax in axes.flat:
-    ax.axhline(threshold, c='black', linewidth=0.6)
-    ax.semilogy()
[…]

### 2018-02-26 src/plot_trend.py

@@ -1,51 +1,0 @@
-import xarray as xr
-import pandas as pd
-import numpy as np
-import seaborn as sns
-import matplotlib.pyplot as plt
-
-fit = xr.open_dataset('data/trend.nc')
-fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
-
-# PLOT THE BIAS
-fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
-for col,f in enumerate(fit['fit_fun'].values):
-    for row,g in enumerate(fit['gen_fun'].values):
-        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'bias')
-        ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
-        if row == 0:
-            ax.set_title(f)
-        if col  == (fit['fit_fun'].size - 1):
-            ax.set_ylabel(g)
-            ax.yaxis.set_label_position('right')
-
-fig.tight_layout()
-fig.subplots_adjust(right=0.85)
-cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'Bias'})
-
-fig.savefig('figs/trend_bias.pdf')
-
-
-
-# PLOT THE VARIANCE
-fig, axes = plt.subplots(nrows=fit['gen_fun'].size, ncols=fit['fit_fun'].size, sharex=True, sharey=True, figsize=(10, 6))
-for col,f in enumerate(fit['fit_fun'].values):
-    for row,g in enumerate(fit['gen_fun'].values):
-        sub = fit.sel(gen_fun = g, fit_fun = f).to_dataframe().reset_index().pivot('M', 'N', 'variance')
-        ax=axes[row, col]
-        sns.heatmap(sub, ax=ax, vmin=0, vmax=0.15, cmap='viridis', cbar=False, linewidths=0.1)
-        if row == 0:
-            ax.set_title(f)
-        if col  == (fit['fit_fun'].size - 1):
-            ax.set_ylabel(g)
[…]

### 2018-02-26 src/plot_trend_sequences.py

@@ -1,57 +1,0 @@
-"""Stationary
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-
-from floodsampling.streamflow import CZNINO3LN2, TwoStateSymmetricMarkovLN2
-
-time_param = {'N': 100, 'M': 50, 't0': 0, 'n_seq': 500}
-n_seq_plot = 50 # how many sequences to plot
-threshold = 2500
-
-enso_gen = CZNINO3LN2(
-    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mu_0=6, beta_mu=0.5, gamma=0.01,
-    coeff_var=0.1, sigma_min=0.01, 
-    n_seq=time_param['n_seq'],
-)
-markov_gen = TwoStateSymmetricMarkovLN2(
-    N=time_param['N'], M=time_param['M'], t0=time_param['t0'], 
-    mmu_1=6.55, mu_0=5.7, gamma_1=0.01, gamma_2=0, pi=0.9,
-    coeff_var = 0.1, sigma_min = 0.01,
-    n_seq=time_param['n_seq'],
-)
-enso_seq = enso_gen.get_data()
-markov_seq = markov_gen.get_data()
-
-# Setup the figure
-fig, axes = plt.subplots(nrows=2, ncols=1, sharey=True, figsize=(12, 5), sharex='col')
-
-ax = axes[0]
-for i in np.arange(n_seq_plot-1):
-    enso_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
-enso_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
-ax.axvline(0, c='black', linewidth=0.6)
-ax.set_xlabel('')
-ax.set_title('{} Synthetic ENSO-Based Sequences'.format(n_seq_plot))
-
-ax = axes[1]
-for i in np.arange(n_seq_plot-1):
-    markov_seq.sel(sequence=i).plot(ax=ax, linewidth=0.05, c='gray', alpha=0.5)
-markov_seq.sel(sequence=n_seq_plot-1).plot(ax=ax, linewidth=1, c='blue')
-ax.axvline(0, c='black', linewidth=0.6)
-ax.set_title('{} Markov Chain-Based Sequences'.format(n_seq_plot))
-
-for ax in axes.flat:
-    ax.axhline(threshold, c='black', linewidth=0.6)
-    ax.semilogy()
[…]

### 2018-02-26 src/plot_variance.py

@@ -1,0 +1,47 @@
+import xarray as xr
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+import argparse
+
+parser = argparse.ArgumentParser()
+parser = argparse.ArgumentParser()
+parser.add_argument("--infile", help="the filename of the saved bias/variance data")
+parser.add_argument("--outfile", help="the filename of the figure to save")
+
+def main():
+    args = parser.parse_args()
+
+    fit = xr.open_dataset(args.infile)
+    fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
+
+    vmin=-5
+    vmax=-1.25
+
+    fig, axes = plt.subplots(
+        nrows=fit['Generating Function'].size, 
+        ncols=fit['Fitting Function'].size, 
+        sharex=True, sharey=True, figsize=(10, 6)
+    )
+    for col,f in enumerate(fit['Fitting Function'].values):
+        for row,g in enumerate(fit['Generating Function'].values):
+            sub = fit_df.loc[np.logical_and(fit_df['Generating Function'] == g, fit_df['Fitting Function'] == f)].reset_index().pivot('M', 'N', 'variance')
+            ax=axes[row, col]
+            sns.heatmap(np.log(sub), ax=ax, vmin=vmin, vmax=vmax, cmap='viridis', cbar=False, linewidths=0.1)
+            if row == 0:
+                ax.set_title(f)
+            if col  == (fit['Fitting Function'].size - 1):
+                ax.set_ylabel(g)
+                ax.yaxis.set_label_position('right')
+
+    fig.tight_layout()
+    fig.subplots_adjust(right=0.85)
+    cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
+    sns.heatmap(np.log(sub), vmin=vmin, vmax=vmax, ax=ax, cmap='viridis', linewidths=0.1, cbar_kws={'label': 'Log of Expected Standard Deviation'}, cbar_ax=cax)
+
+    fig.savefig(args.outfile)
+    
+if __name__ == '__main__':
+    main()

### 2018-03-01 src/plot_bias.py
Adjust plots

@@ -3,6 +3,7 @@ import pandas as pd
 import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
+import colorcet as cc
 
 import argparse
 
@@ -17,6 +18,10 @@ def main():
     fit = xr.open_dataset(args.infile)
     fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
 
+    vmin = -0.02
+    vmax = 0.02
+    cmap = cc.cm['gwv']
+
     fig, axes = plt.subplots(
         nrows=fit['Generating Function'].size, 
         ncols=fit['Fitting Function'].size, 
@@ -26,19 +31,21 @@ def main():
         for row,g in enumerate(fit['Generating Function'].values):
             sub = fit_df.loc[np.logical_and(fit_df['Generating Function'] == g, fit_df['Fitting Function'] == f)].reset_index().pivot('M', 'N', 'bias')
             ax=axes[row, col]
-            sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar=False, linewidths=0.1)
+            C = sns.heatmap(sub, ax=ax, vmin=vmin, vmax=vmax, cmap=cmap, cbar=False, linewidths=0.1)
             if row == 0:
-                ax.set_title(f)
+                ax.set_title(f)            
             if col  == (fit['Fitting Function'].size - 1):
                 ax.set_ylabel(g)
                 ax.yaxis.set_label_position('right')
-
+    
     fig.tight_layout()
     fig.subplots_adjust(right=0.85)
     cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-    sns.heatmap(sub, ax=ax, vmin=-0.025, vmax=0.025, cmap='PuOr', cbar_ax=cax, linewidths=0.1, cbar_kws={'label': 'Expected Bias'})
+    sns.heatmap(sub, vmin=vmin, vmax=vmax, ax=ax, cmap=cmap, linewidths=0.1, cbar_kws={'label': 'Expected Bias'}, cbar_ax=cax)
+    ax.set_ylabel(g)
+    ax.yaxis.set_label_position('right')
 
     fig.savefig(args.outfile)
-
+    
 if __name__ == '__main__':
     main()

### 2018-03-01 src/plot_enso.py

@@ -5,7 +5,9 @@ import os
 
 import numpy as np
 import xarray as xr
+import pandas as pd
 import matplotlib.pyplot as plt
+from scipy import signal
 
 from floodsampling.util import get_data_path
 
@@ -18,15 +20,27 @@ def main():
     data_fname = os.path.join(get_data_path(), 'ramesh2017.nc')
     enso = xr.open_dataarray(data_fname)
 
-    # Plot Time Series
+    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(12, 4), gridspec_kw={'width_ratios': [2, 1]})
     enso.sel(year=slice(1000, 2000)).plot(
-        figsize=(10, 4.5),
         c='blue',
-        linewidth=0.5
+        linewidth=0.5,
+        ax=axes[0]
     )
-    plt.grid(True)
-    plt.xlabel('Year')
-    plt.ylabel('Annualized NINO3 Index')
+    axes[0].grid(True)
+    axes[0].set_xlabel('Year')
+    axes[0].set_ylabel('Annualized NINO3 Index')
+
+    # get the spectrum
+    x = enso.values
+    fs = 1
+    f, Pxx_den = signal.periodogram(x, fs)
+    periodogram = pd.Series(Pxx_den, index=f)
+    smoothed = periodogram.rolling(1000, win_type='hamming').mean()
+
+    smoothed.plot(ax=axes[1], label='smoothed')
+    axes[1].set_xlabel('frequency [Year$^{-1}$]')
+    axes[1].set_ylabel('Power Spectral Density')
+    axes[1].grid()
 
     plt.savefig(args.outfile, bbox_inches='tight')
 

### 2018-03-01 src/plot_example.py

@@ -49,7 +49,7 @@ def main():
 
     M = 150
     n_seq = 1
-    n_sim = 500
+    n_sim = 5000
 
     # Initialize the bias and variance
     gen_names = np.array(['NINO3 Stationary', 'NINO3 Trend'])
@@ -63,8 +63,19 @@ def main():
             gen_dat = gen_fun.get_data('all')
             fit_dat = fit_fun.get_data('future')
             ax = axes[f, g]
-            for i in np.arange(n_sim):
-                ax.plot(fit_dat['year'], fit_dat.sel(sim=i).values.ravel(), c='gray', linewidth=0.1, alpha=0.1)
+            year = fit_dat['year']
+
+            q25 = fit_dat.quantile(0.25, dim='sim').values.ravel()
+            q75 = fit_dat.quantile(0.75, dim='sim').values.ravel()
+            q90 = fit_dat.quantile(0.90, dim='sim').values.ravel()
+            q10 = fit_dat.quantile(0.10, dim='sim').values.ravel()
+            q995 = fit_dat.quantile(0.995, dim='sim').values.ravel()
+            q005 = fit_dat.quantile(0.005, dim='sim').values.ravel()
+            
+            ax.fill_between(year, q005, q995, color='gray', alpha=0.5)
+            ax.fill_between(year, q10, q90, color='gray', alpha=0.5)
+            ax.fill_between(year, q25, q75, color='gray', alpha=0.5)
+            
             ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue', linewidth=1)
             ax.semilogy()
             ax.grid()

### 2018-03-01 src/plot_variance.py

@@ -3,6 +3,7 @@ import pandas as pd
 import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
+import colorcet as cc
 
 import argparse
 
@@ -17,8 +18,9 @@ def main():
     fit = xr.open_dataset(args.infile)
     fit_df = fit.to_dataframe().reset_index().set_index(['M', 'N'])
 
-    vmin=-5
-    vmax=-1.25
+    vmin = -5
+    vmax= - 1
+    cmap = cc.cm['fire']
 
     fig, axes = plt.subplots(
         nrows=fit['Generating Function'].size, 
@@ -29,7 +31,7 @@ def main():
         for row,g in enumerate(fit['Generating Function'].values):
             sub = fit_df.loc[np.logical_and(fit_df['Generating Function'] == g, fit_df['Fitting Function'] == f)].reset_index().pivot('M', 'N', 'variance')
             ax=axes[row, col]
-            sns.heatmap(np.log(sub), ax=ax, vmin=vmin, vmax=vmax, cmap='viridis', cbar=False, linewidths=0.1)
+            sns.heatmap(np.log(sub), ax=ax, vmin=vmin, vmax=vmax, cmap=cmap, cbar=False, linewidths=0.1)
             if row == 0:
                 ax.set_title(f)
             if col  == (fit['Fitting Function'].size - 1):
@@ -39,8 +41,10 @@ def main():
     fig.tight_layout()
     fig.subplots_adjust(right=0.85)
     cax = fig.add_axes([0.9, 0.1, 0.02, 0.8])
-    sns.heatmap(np.log(sub), vmin=vmin, vmax=vmax, ax=ax, cmap='viridis', linewidths=0.1, cbar_kws={'label': 'Log of Expected Standard Deviation'}, cbar_ax=cax)
-
+    sns.heatmap(np.log(sub), vmin=vmin, vmax=vmax, ax=ax, cmap=cmap, linewidths=0.1, cbar_kws={'label': 'Log of Expected Standard Deviation'}, cbar_ax=cax)
+    ax.set_ylabel(g)
+    ax.yaxis.set_label_position('right')
+    
     fig.savefig(args.outfile)
     
 if __name__ == '__main__':

### 2018-03-01 writeup/main.tex
should be working fine, but use only a few sequences for testing purposes
Continue working on writing the intro

@@
 
+% -----------------------------------------------------------------------------
+% INTRODUCTION
+% -----------------------------------------------------------------------------
+
 \section{Introduction}
 
-The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies \citep[\ie{} Bulletin 17-B for flood frequency analysis;][]{IACWD1982}.
-In this framework event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
+The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies \citep[\ie{} Bulletin 17-B for flood frequency analysis;][]{IACWD1982} in which event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
 Under the assumption that the historical or simulated record of events constitutes an independent and identically distributed process, and using an appropriate statistical inference model, an estimate of this relationship and often its uncertainty based on a finite \(N\) year record is developed.
@@
 Over time it has become clear that even small nonstationarity (temporal changes in dynamics of the underlying hydroclimatalogical process) can lead to significant trends in the estimated \(T\)-year return periods or probability of exceedance \( p_T \) \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.
-This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last \(N\) years of the data.
-Each model will necessarily give different estimates, particularly as the model is forecast further into the future.
+This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last \(N\) years of the data.\footnote{Vogel}
+Each model will necessarily give different estimates, which may fit the historical record with comparable accuracy but diverge at an increasing rate as the model is forecast further into the future.
 Despite the wide suite of available models, the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied.
@@
 The uncertainties associated with this are not purely informational; both nature and future human actions will determine the outcomes, and neither is obviously predictable.
-The approach taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions and then use climate models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
-
-Despite continuing improvements in physical understanding and computational capacity, there is cause for concern with dependence on using physics-based numerical models to project future climate risk.
-Use of a ``model chain'' typically encompasses:
+The approach taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions and then use a sequence of physics-based numerical models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
+Such a= ``model chain'' typically encompasses:
 \begin{enumerate*}[label= (\roman*) ]
@@
 \end{enumerate*}
-Given poor performance of flood projections relative to \nth{20} century data, ``bias correction methods'' whose applicability under extrapolation does not inspire confidence are typically applied \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+Bias correction methods are typically applied at each step of the chain, and while this typically improves the model chain's representation of the historical record the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
 A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
 Would it be over- or under- designed?
-Would the high degree of bias and uncertainty in its estimation lead to suboptimal investment strategies in risk mitigation?
+Would the high degree of bias and uncertainty in its estimation lead to sub-optimal investment strategies in risk mitigation?
 These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.
@@
 Though the specific examples target flood risk, this method and framework may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricanes, and wind.
+
+% -----------------------------------------------------------------------------
+% METHODS
+% -----------------------------------------------------------------------------
 
@@
 Though both approaches used to generate sequences \( \mu(t) \) use highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
-In \cref{sec:sequence-realistic} we show sequences of streamflow generated by both models for some choices of parameters and discuss the processes which they can reasonably represent.
 
@@
 To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate \(\mu(t\).
-
 This Markov chain is given a constant probability of persistence which is the same for both states, such that the transition matrix is
@@
 For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
-We also assume \( \mu_{0, 1} \gg \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
+We also assume \( \mu_{0, 1} > \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
 
@@
 
-In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 1000\footnote{Check this parameter} sequences of future streamflow are generated using Monte Carlo simulation.
+In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 2000 sequences of future streamflow are generated using Monte Carlo simulation.
 Three models are assessed for estimating future streamflow and are described in detail below.
@@
 We first consider fitting a stationary model to the observed flood record, following classical assumptions of \iid sequences and stationarity.
-We use the log-normal model (\cref{eq:lognormal}) for stationary flood distribution; while this model is less flexible than three-parameter models such as the GEV or LP3 \citep{Vogel1996}, it is easy to understand and interpret.
-Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
 
+We use the log-normal model (\cref{eq:lognormal}) for its simplicity, interpretability, and good performance relative to other 2-parameter models \citep{Vogel1996}.
 For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations (denoted \(Q_{\text{hist}}\) for clarity).
@@
 where the first term is the mean squared error, the second term is the bias squared, and the third term is the variance.
-In \cref{sec:results} the bias and standard deviation (square root of variance) are averaged over all \(J\) sequences and plotted.\footnote{Check this code}
 
@@
 All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
-To carry implement the flood sequence generating and estimating functions, a new package called \texttt{floodsampling} was developed and is available at \url{github.com/jdossgollin/floodsampling}.
 The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.
+
+% -----------------------------------------------------------------------------
+% ESTIMATION BIAS AND UNCERTAINTY
+% -----------------------------------------------------------------------------
 
@@
 
-The models used to generate synthetic flood sequences used simplistic dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
+The models used to generate synthetic flood sequences used simple dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
 
@@
 The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
+Analysis of the power density spectrum shows a strong peak around 3-6 years.
+\begin{figure}
+  \includegraphics[width=\textwidth]{enso.pdf}
+  \caption{
+    (L): A \SI{1000}{year} sub-set of the synthetic time series described in \cref{sec:methods-nino}.
+    (R): The power spectral density of the annualized time series. A moving window was applied to the density for smoothing.\label{fig:enso-ts}
+  }
+\end{figure}
 
@@
 For both, the coefficient of variation \(\alpha=0.1\) and the minimum value of sigma \(\sigma_\text{min}=0.01\).
-As the time series plot shows, both generating models exhit substantial persistence in flood probability from one year to the next, though the ENSO model shows negative autocorrelation at lag 1 while the Markov model has positive autocorrelation at lag 1 (not shown).\footnote{Need to include this?}
-
 \begin{figure}[b]
-  \includegraphics[width=\textwidth]{stationary_sequences.pdf}  
+  \includegraphics[width=\textwidth]{sequences_stationary.pdf}  
   \caption{
     50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-nino} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
-    Both models use a stationary parameterization.
+    Both models use a stationary parameterization \(\gamma=0\).
     49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
-    The right hand panel shows the distribution over the historical and future periods for 500 sequences from the same model, visualized with a histogram and kernel density estimate.
-    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.014133333333333333} by the NINO3 sequences and \num{0.013546666666666667} by the Markov sequences.\label{fig:stationary-sequences}
+    Note that $y$-axis has a log scale.\label{fig:stationary-sequences}
   }
@@
 To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
-For the ENSO time series, this mean setting \(\gamma=0.02\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.02\).
+For the ENSO time series, this mean setting \(\gamma=0.015\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.015\).
 The resulting sequences are shown in \cref{fig:trend-sequences}.
 The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
-\begin{figure}[b]
-  \includegraphics[width=\textwidth]{trend_sequences.pdf}
+\begin{figure}[b]d
+  \includegraphics[width=\textwidth]{sequences_trend.pdf}
   \caption{
-    As \cref{fig:stationary-sequences} but a trend term has been added.
-    Horizontal line shows a threshold of \num{2500}, which is exceeded with probability \num{0.023306666666666667} by the NINO3 sequences and \num{0.024493333333333332} by the Markov sequences.\label{fig:trend-sequences}
+    As \cref{fig:stationary-sequences} but a trend term has been added by setting \(\gamma=0.015\).\label{fig:trend-sequences}
   }
@@
 
-\subsection{Stationary Case}
+This choice of parameterizing secular change as a linear trend on the expected value of log streamflow was made to aid interpretability, and not because of any particular physical reasoning.
+Indeed, for large \(N\) this model suggests implausibly dry conditions and implausibly wet conditions for very large \(M\).
+However, given widespread use of trend-based models in the flood frequency analysis literature it is informative to consider the ``best-case'' scenario in which the model is correctly specified and there is a true, linear trend in the data.
 
-\emph{Add more text once the results come through from Habanero -- they are running now!}
+\subsection{Stationary Case}\label{sec:results-stationary}
 
+We next consider the estimation bias and variance as a function of \(M\) and \(N\) when the true generative process is stationary.
+
+To do this we generate 1000 sequences of streamflow for each combination of \(M\), \(N\), and generative process.
+We then fit each estimation model to each sequence and draw 2000 estimates of future streamflow from the fitted model.
+These are then used to determine \(\hat{p}_T\), allowing calculation of the expected bias and variance for each combination of \(M\), \(N\), generative model, and estimating model.
+The parameters for the NINO3 model are given in \cref{tab:nino-stationary} and for the Markov model in \cref{tab:nino-stationary}.
+We use a threshold $Q_T = \SI{5000}{units}$.
+
+\begin{table}[ht]
+  \begin{tabular}{L{1in}L{3.75in}L{0.75in}}
+    \toprule
+    Paramater & Description & Value \\
+    \midrule
+    $\mu_0$ & Expected value of $\log Q$ when $x=0$ & 6 \\
+    $\beta_\mu$ & Coefficient of $x$ on $\mu$ & 0.5 \\
+    $\alpha$ & Coefficient variation & 0.1 \\
+    $\sigma_\text{min}$ & Lowest allowable value of $\sigma$ & 0.01 \\
+    $\gamma$ & Time trend of $\mu$ & 0 \\
+    \bottomrule
+  \end{tabular}
+  \caption{Parameters of NINO3 model under stationary case. $x(t)$ denotes the modeled NINO3 time series.}\label{tab:nino-stationary}
+\end{table}
+\begin{table}[ht]
+  \begin{tabular}{L{1in}L{3.75in}L{0.75in}}
+    \toprule
+    Parameter & Description & Value \\
+    \midrule
+    $\mu_1$ & Expected value of $\log Q$ for state 1 & 6.75 \\
+    $\mu_2$ & Expected value of $\log Q$ for state 2 & 6 \\
+    $\gamma_1$ & Time trend of $\mu_1$ & 0 \\
+    $\gamma_2$ & Time trend of $\mu_2$ & 0 \\
+    $\alpha$ & Coefficient variation & 0.1 \\
+    $\sigma_\text{min}$ & Lowest allowable value of $\sigma$ & 0.01 \\
+    \bottomrule
+  \end{tabular}
+  \caption{Parameters of Markov model under stationary case}\label{tab:markov-stationary}
+\end{table}
+
+\Cref{fig:stationary-bias} shows that in general, bias is quite similar across generative models but varies substantially between the models used for estimation.
+The HMM and LN2 Stationary models have no time term, and so their bias is independent of \(M\).
+In general both of these models have very low bias, though the LN2 Stationary model tends to over-estimate the probability of flooding for very short \(N\), unsurprising since the true distribution of streamflow is bimodal for both generating functions.
+The LN2 trend model registers a non-trivial probability of a positive time trend, particularly for short \(N\), leading to large bias in estimation.
+For very long \(M\) the bias is higher, as estimated probability of exceeding the threshold is exceeded.
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{stationary_bias.pdf}
-  \caption{The estimation bias for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \includegraphics[width=\textwidth]{bias_stationary.pdf}
+  \caption{The estimation bias for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
 \end{figure}
 
+Analysis of the variance, in these estimates, shown in \cref{fig:stationary-variance} as the log standard deviation, is similar.
+As suggested in the bias plots, the HMM and LN2 Stationary models have very low variance of their estimates.
+The LN2 Trend model has much higher variance in estimates, particularly for short \(N\) or short \(M\).
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{stationary_variance.pdf}
-  \caption{The estimation variance for 1000* sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \includegraphics[width=\textwidth]{variance_stationary.pdf}
+  \caption{The estimation variance for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-variance}}
 \end{figure}
@@
 
-\emph{Add more text once the results come through from Habanero -- they are running now!}
+We next repeat the previous example but introduce a time trend in the data.
+Thus the parameters of the NINO3 model are as \cref{tab:nino-stationary} except that $\gamma=0.015$ and the parameters of the Markov model are as \cref{tab:markov-stationary} except that $\gamma_1 = 0.015$.
 
+For short \(M\), particularly when \(N\) is not too big, the stationary models (HMM and LN2 Stationary) perform well for both generative functions.
+As \(M\) increases, they begin to substantially underestimate the probability of future flooding.
+As for the stationary case, when \(N\) is short the LN2 Stationary model tends to over-estimate $\hat{p}_T$ because it sees variation caused by the NINO3 time series or the Markov state switching and assigns it to internal variability, leading to unrealistically fat tails in the estimate.
+The trend model, by contrast, over-estimates $p_T$ except when $N$ is very large, mainly because the posterior uncertainty in the magnitude of the trend means that the model believes that a trend larger than observed is not unlikely.
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{trend_bias.pdf}
-  \caption{The estimation bias for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \includegraphics[width=\textwidth]{bias_trend.pdf}
+  \caption{The estimation bias for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
 \end{figure}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{trend_variance.pdf}
-  \caption{The estimation variance for 1000* sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \includegraphics[width=\textwidth]{variance_trend.pdf}
+  \caption{The estimation variance for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
 \end{figure}
 
-\section{Case Study}
+% -----------------------------------------------------------------------------
+% SPATIALLY CORRELATED RISK
+% -----------------------------------------------------------------------------
 
-We still need to find a good case study; it can be a highly simplified one as long as there is a long streamflow record available and building levees is an option that has been proposed.
-If there's not an obvious candidate, we can use one of our models above.
-We can make some other strong assumptions
+\section{Spatially Correlated Risk}
+
+The setup:
+\begin{enumerate}
+  \item Take one of the models -- say the NINO3 model -- and calculate \(\mu(t)\)
+  \item For each of \(J\) locations, sample \(Q(t)\)
+  \item Take threshold to be ``true'' 50-year flood
+  \item Plot expected damages as a function of time
+\end{enumerate}
+Some alternate ideas:
 \begin{itemize}
-  \item Risk exposure constant in time, which allows us to focus on the estimation rather than comprehensive risk manageement
-  \item Constant future discounting rate
-  \item Flood damages are linearly related to streamflow over some amount:
-  \begin{equation}
-    \mathcal{D}(t) = k \qty[ Q(t) - Q_T ]
-  \end{equation}
-\end{itemize}
-Then our two options are to build a levee, which we can assume to have a capital cost plus operation and maintenance cost which is proportional to the capital cost, which increases \(Q_T\) by some fixed amount, or to buy insurance.
-For the insurance, we consider a contract in which an insurer agrees to pay the entire coverage cost \(C\) every time that the trigger threshold \(Q_T\) is exceeded.
-Although \(Q_T\) is in general a decision variable, we treat it as fixed.
-Then, using the posterior distribution of the parameters of the chosen statistical model (from \cref{tab:model-fitting}) we estimate the probability that \(Q(t) \geq Q_T\) for \(t \in \vb{t}\), which is equivalent to \(p_T(t)\).
-For a particular year, the fair price for the insurance premium is \(\expectation \qty(p_T) C\).
-However, there is typically also a risk premium associated with the policy that prices the uncertainty in this estimate (since the trigger \(Q_T\) is specified in the contract and thus not uncertain).
-This risk premium is proportional to the variance of \(p_T\), or more generally to its uncertainty distribution.
-The insurance premium for a particular year is thus
-\begin{equation} \label{eq:premium}
-  Z(t) = C \qty[\expectation \qty(p_T(t)) + R \variance \qty(p_T(t))]
-\end{equation}
-where \(R \geq 0\) is a coefficient describing the cost of tue uncertainty in the estimate of \(p_T(t)\).
-
-Then at each point in time, calculate:
-\begin{itemize}
-  \item Expected cost plus damage per year for the levee using perfect knowledge of the next 101 years
-  \item Expected cost plus damage per year for the levee using MCMC projections of the next 101 years from all 3 models
-  \item Expected cost plus damage per year for a 5-year insurance contract using perfect knowledge of the next 5 years
-  \item Expected cost plus damage per year for a 5-year insurance contract using MCMC projections of the next 5 years from all 3 models
+  \item Represent different cities with different \(\beta\)?
+  \item OK to do \(C \propto \qty(Q - Q_T)\)
 \end{itemize}
 
-Can add appropriate plot or plots.
+% -----------------------------------------------------------------------------
+% DISCUSSION
+% -----------------------------------------------------------------------------
 
@@
 How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
+
+Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{example_long.pdf}
+  \caption{
+    A single sequence of streamflow generated by the NINO3 function with (L) no trend and (R) \(\gamma=0.01\). 
+    The sequence is fit to each fit to three different fitting models (rows).
+    Confidence intervals shown are 50\%, 80\%, and 99\%.
+    Generated with \(M=N=150\).\label{fig:example-long}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{example_short.pdf}
+  \caption{
+    As \cref{fig:example-short} except \(N=50\).\label{fig:example-short}
+  }
+\end{figure}
+
+% -----------------------------------------------------------------------------
+% SUMMARY
+% -----------------------------------------------------------------------------
 
@@
 \appendix
-\printbibliography{}
 
@@
 Kyle Johnson of Columbia University for his suggestions for improving reliability and documentation of the \texttt{floodsampling} package.
-
-\section{Background Information and Conceptual Framing}
-
-\subsection{Physical Reasoning}
-
-In order to make probabilistic predictions of future flooding in a particular location, it is helpful to begin by considering the specific physical mechanisms which may lead to flooding in particular location.
-
-Floods may be associated with a single rainfall event or be composed of a sequence of multiple, distinct events over a week or a season, such that there is pre-conditioning of the land surface soil moisture and natural and man-made stores that leads to a high flooding potential.
-In large river basins, Lagrangian tracking studies \citep{Gimeno2010} indicate that river flooding often requires the large-scale transport and convergence of moisture  which are driven by specific large-scale climate mechåanisms \citep{Dacre2014,Gimeno2014,Lu2013,Dirmeyer2010}.\footnote{Farnham et al}
-For example, major river floods in the Ohio River Basin (USA) in the 20th century exhibit remarkably consistent continental-scale circulation features in the weeks preceding the flood peak \citep{Nakamura2012,Robertson2015}.
-Other case studies of  major floods in the Balkans \citep{Stadtherr2016}, Central Europe \citep{Grams2014}, and Paraguay\footnote{Add my work}, and theoretical analyses \citep{Hoskins2015,Screen2014,Coumou2014}, support the hypothesis that persistent, intense rainfall in a particular place is associated with specific and potentially identifiable climate mechanisms.
-
-At the seasonal and sub-seasonal scales, surface temperature boundary conditions are well known to significantly change the mean, variance, persistence, recurrence and spatial structure of atmospheric circulation patterns, as well as moisture availability.
-On longer time scales, changes to the physical processes that govern hydroclimate systems, known generically as nonstationarity \citep{Milly2008} can occur due to climate change, river modification, and land use change \citep{Merz2014}.
-Thermodynamic scaling arguments predict an intensified hydrologic cycle under global warming due to the Clausius-Clapeyron scaling of the atmosphere's moisture-holding capacity with temperature \citep[see][]{Muller2011,OGorman2015}.
-However, these scaling arguments do not in general hold over land \citep{Byrne2015} and typically describe short-duration rainfall maxima rather than persistent and intense rainfall sequences.
-As compared with thermodynamic changes there is less certainty as to how large-scale dynamical circulations may respond to warming signals \citep{Shaw2016,Barnes2015}, as the dynamics governing these mechanisms are highly non-linear \citep{Palmer2013} and in general not well resolved by climate models\footnote{Farnham 2018}.
-
-\subsection{Nonstationary Flood Frequency Analysis\label{sec:ffa}}
-
-Classical methods for design of infrastructure first specify a return period \(T\) (\ie{} 100 years), and then design a project that protects against this \(T\)-year event.
-The mathematical problem of interest then becomes estimation of the probability distribution \( f(p_T) \), where \( p_T \) is binary event that the threshold \( Q_T \) is exceeded by the annual-maxiumum flood series \( Q(t) \):
-\begin{equation}
-  p_T(t) \equiv \probability \qty[ Q(t) \geq Q_T] \qqtext{for} t = \qty[t_0 + 1, \ldots, t_0 + M]
-\end{equation}
-where \( t_0 \) is the first year forecast and \( M \) is the project operation period.
-Averaging over the \(M\)-year period, we can write for convenience
-\begin{equation}
-  p_T = \frac{1}{M} \sum_{t=1}^{M} p_T(t)
-\end{equation}
-Although in a stationary setting an unbiased estimator of \( P \qty[ Q(t) \geq Q_T] \) converges asymtotically to \(1 - \frac{1}{T}\), the asymmetric uncertainty distribution of the estimate has also prompted the consideration of the potential for over- or under-design for floods if the conditional mean of the estimated distribution of \( Q_T \) is used for design \citep{Stedinger1997}\footnote{Manu has mentioned a Vogel198x citation but I haven't successfully found it.}.
-
-Under classical assumptions of stationarity, the parameters describing the hydroclimate system are assumed independent of time.
-Even when this assumption is satisfied, low-frequency variability can lead to a high probability of surprise, particularly when the length of the record used for estimation \(N\) is short \citep{Jain2001,Matalas2012}.
-However, widespread acceptance of nonstationarity as a principle has not led to easily generalizable methodologies for nonstationary flood frequency analysis.
-
-One approach is to extend the purely statistical models currently used in the United States \citep{IACWD1982}.
-Like the stationary case, there is no consensus as to the proper statistical distribution to use.
-Although Bulletin 17-B \citep{IACWD1982} mandates the use of the log-Pearson type III (LP3) model of annual-maximum floods in the United States, annual-maximum floods are also modeled three-parameter distributions such as the generalized extreme value distribution and two-parameter models such as the lognormal (LN2) \citep{Vogel1996}.
-Many studies have extended stationary statistical models to include a time trend on one or more parameters \citep{Obeysekera2014,Vogel2011,Serinaldi2015,Strupczewski2001}, an approach which is conceptually and computationally simple but that lacks physical justification.
-Rather than conditioning parameters on time, some recent studies have incorporated a climate index or set of climate indices \(X\) as predictor variables for streamflow \citep{Delgado2014,Silva2016,Sun2014,Griffis2007}\footnote{Farnham et al}.
-Other studies have used Markov or mixture models \citep{Waylen1986,Sveinsson2005,Griffis2007}, Bayesian approaches for combining information \citep{Lima2016,Bracken2017}, and other more sophisticated statistical models.
-
-An alternative approach is to use numerical models to represent the physical processes at all scales.
-Because models are not able to simultaneously represent scales ranging from global-scale climate phenomena to river hydrology, a ``model chain'' approach is used.
-A typical model chain might encompass:
-\begin{enumerate*}[label= (\roman*) ]
-  \item emission scenario;
-  \item general circulation model (GCM);
-  \item downscaling;
-  \item hydrological catchment model; and
-  \item flood frequency analysis.
-\end{enumerate*}
-Typically bias correction is applied at several or steps of the procedure, which can lead to results which are difficult to interpret in a probabilistic context \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.
-
-\subsection{Decision Approaches}
-
-Despite the abundance of approaches for estimating future flood hazard, each of which may give different estimates of future flood risk, the ultimate goal of flood frequency analysis is to inform risk management decisions.
-
-Risk-based design methods \citep[RBDM; see][]{Rosner2014} recognize that the appropriate level of design may itself be a decision variable.
-\citet{Lall1987} considered the uncertainty in the frequency with which floods may exceed the design level for different values of \(p\) (equivalent to different values of \(T\)), and also for different sample sizes for estimation \(N\) and different project operation periods \(M\).
-Some prominent methodologies for risk-based design under nonstationary in the water resources literature include robust decision making (RDM)\footnote{(Lempert et al.
-2006; Lempert and Groves 2010; Matrosov et al.
-2013)}, scenario-neutral planning\footnote{(Prudhomme et al.
-2010)}, infogap analysis\footnote{(Ben-Haim 2006; Korteling et al.
-2013)}, and decision scaling\footnote{(Brown et al.
-2011, 2012; Moody and Brown 2013; Turner et al.
-2014)}.
-
-Alternatively, sequential decision models \citep[see][]{Russell2003,Howard1960} allow decision-makers to optimize not only what action to take, but also when to take it.
-Sequential decision models have been used in the hydrological literature for sea wall optimization \citep{Lickley2014}.\footnote{need some citations for sequential decisions}
-Given that, low-frequency variability or system memory may lead to less uncertain probabilistic forecasts at short times than long times, models which can capture this uncertainty may be particularly advantageous for sequential models.
-
-There is thus \emph{a critical theoretical need for decision frameworks which can compare projects of different operation period} using an arbitrary choice of model of future flood behavior.
-In the following sections we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary environment.
-Understanding the full uncertainty associated with thresholds of interest is relevant to many other climate adaptation problems, including infrastructure design for flood management, because we are faced with the task of estimating the infrastructure's probability of failure \(p_T\) for the next \(M\) years, as well as the uncertainty and bias in this estimate.
-This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
-We proceed as follows \(\ldots{}\)
+The authors thank Columbia ITS for support running on the cluster Habanero.
 
@@
 
-\section{Supplemental Figures}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{enso.pdf}
-  \caption{
-    A \SI{2500}{year} sub-set of the synthetic time series described in \cref{sec:methods-nino}.\label{fig:enso-ts}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{example_long.pdf}
-  \caption{
-    A single sequence of streamflow generated by the NINO3 function with (L) no trend and (R) \(\gamma=0.01\). 
-    The sequence is fit to each f`it to three different fitting models (rows). 
-    Generated with \(M=N=150\).\label{fig:example-long}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{example_short.pdf}
-  \caption{
-    As \cref{fig:example-short} except \(N=30\).\label{fig:example-short}
-  }
-\end{figure}
+\printbibliography{}
 

### 2018-03-02 writeup/main.tex
working on results and intro

@@
 
-The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies \citep[\ie{} Bulletin 17-B for flood frequency analysis;][]{IACWD1982} in which event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
-Under the assumption that the historical or simulated record of events constitutes an independent and identically distributed process, and using an appropriate statistical inference model, an estimate of this relationship and often its uncertainty based on a finite \(N\) year record is developed.
-This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation, typically using an expected net present value criteria applied to the costs and residual losses expected from the mitigation instrument.
+In 2011 five successive cyclones struck Thailand, causing massive damage to public and private infrastructure and crippling global supply chains.\footnote{See Masa Haraguchi's work}
+Later in June 2013 a long rainy period was followed by three summertime cyclones tracking over the same region, leading to massive flooding in central Europe \citep{Grams2014}.
+At the same time a multi-year drought in S\~ao Paulo threatened to dry out one of the world's largest cities, and in 2017 the Atlantic Hurricane season led to over \usd{100 billion} in reinsurance loss.\footnote{\href{http://www.artemis.bm/blog/2018/02/27/berkshire-hathaway-can-easily-weather-a-400bn-mega-catastrophe-buffett/}{Click for blog post on \url{artemis.bm}}}
+Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems are overwhelmed by these multi-year droughts or multi-hurricane hits.
 
-Over time it has become clear that even small nonstationarity (temporal changes in dynamics of the underlying hydroclimatalogical process) can lead to significant trends in the estimated \(T\)-year return periods or probability of exceedance \( p_T \) \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.
-This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last \(N\) years of the data.\footnote{Vogel}
+Indeed, the scientific literature has recognized that even small nonstationarity (temporal changes in dynamics of the underlying hydroclimatalogical process) can lead to significant trends in the estimated $T$-year return periods or probability of exceedance $p_T$ \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.
+This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last $N$ years of the data.\footnote{Vogel}
 Each model will necessarily give different estimates, which may fit the historical record with comparable accuracy but diverge at an increasing rate as the model is forecast further into the future.
-Despite the wide suite of available models, the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied.
+Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied}.\footnote{All of these citations are for floods, can we look at other types of extremes?
+Even droughts would be helpful.}
 
@@
 The approach taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions and then use a sequence of physics-based numerical models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
-Such a= ``model chain'' typically encompasses:
+Such a ``model chain'' typically encompasses:
 \begin{enumerate*}[label= (\roman*) ]
@@
   \item general circulation model (GCM);
-  \item downscaling;
-  \item hydrological catchment model; and
-  \item flood frequency analysis.
+  \item downscaling; and
+  \item high-resolution model of the study region.
 \end{enumerate*}
-Bias correction methods are typically applied at each step of the chain, and while this typically improves the model chain's representation of the historical record the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+Because of errors inherent to discretization and parameterization of the laws of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
 A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
 Would it be over- or under- designed?
-Would the high degree of bias and uncertainty in its estimation lead to sub-optimal investment strategies in risk mitigation?
+Would the bias and uncertainty in its estimation lead to sub-optimal investment strategies in risk mitigation?
 These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.
@@
 
-To generate \(\mu(t)\), a sequence of length \(M+N\) is randomly chosen from the \SI{100000}{year} NINO3 sequence.
+To generate $\mu(t)$, a sequence of length $M+N$ is randomly chosen from the \SI{100000}{year} NINO3 sequence.
 Next, \(\mu\) is calculated following by
@@
   \includegraphics[width=\textwidth]{bias_trend.pdf}
-  \caption{The estimation bias for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \caption{The estimation bias for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:trend-bias}}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{variance_trend.pdf}
-  \caption{The estimation variance for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \caption{The estimation variance for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:trend-variance}}
 \end{figure}
-
-% -----------------------------------------------------------------------------
-% SPATIALLY CORRELATED RISK
-% -----------------------------------------------------------------------------
-
-\section{Spatially Correlated Risk}
-
-The setup:
-\begin{enumerate}
-  \item Take one of the models -- say the NINO3 model -- and calculate \(\mu(t)\)
-  \item For each of \(J\) locations, sample \(Q(t)\)
-  \item Take threshold to be ``true'' 50-year flood
-  \item Plot expected damages as a function of time
-\end{enumerate}
-Some alternate ideas:
-\begin{itemize}
-  \item Represent different cities with different \(\beta\)?
-  \item OK to do \(C \propto \qty(Q - Q_T)\)
-\end{itemize}
 
@@
 \section{Discussion}
+
+% PHYSICAL PLAUSABILITY OF TRENDS 
+% SPATIAL CASE
+% THIS IS RELEVANT FOR OTHER NATURAL HAZARDS -- DISCUSS
+% IMPLICATIONS FOR INSURANCE / DESIGN
 
@@
 
+Implications
+\begin{itemize}
+  \item Blind use of trend models leads to sub-optimal models \emph{even when there is a trend} for many combination of \(M,N\)
+  \item Bias-variance tradeoff of model complexity
+  \item Need to consider the structure of the underlying trend, $N$, and $M$ when choosing a model
+  \item Should consider multiple models, \eg by ``stacking'' \citep{Yao2017}
+\end{itemize}
+
 % -----------------------------------------------------------------------------
@@
 The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{1000000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh2017}.
-Kyle Johnson of Columbia University for his suggestions for improving reliability and documentation of the \texttt{floodsampling} package.
-The authors thank Columbia ITS for support running on the cluster Habanero.
+The authors thank Columbia ITS for support running simulations on the cluster.
+James thanks NSF.
+Dave thanks grant.
 

### 2018-03-02 src/plot_enso.py

@@ -20,11 +20,11 @@ def main():
     data_fname = os.path.join(get_data_path(), 'ramesh2017.nc')
     enso = xr.open_dataarray(data_fname)
 
-    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(12, 4), gridspec_kw={'width_ratios': [2, 1]})
+    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(14, 4), gridspec_kw={'width_ratios': [2, 1]})
     enso.sel(year=slice(1000, 2000)).plot(
         c='blue',
         linewidth=0.5,
-        ax=axes[0]
+        ax=axes[0],
     )
     axes[0].grid(True)
     axes[0].set_xlabel('Year')
@@ -35,7 +35,7 @@ def main():
     fs = 1
     f, Pxx_den = signal.periodogram(x, fs)
     periodogram = pd.Series(Pxx_den, index=f)
-    smoothed = periodogram.rolling(1000, win_type='hamming').mean()
+    smoothed = periodogram.rolling(100, win_type='hamming').mean()
 
     smoothed.plot(ax=axes[1], label='smoothed')
     axes[1].set_xlabel('frequency [Year$^{-1}$]')

### 2018-03-02 src/plot_example.py

@@ -65,16 +65,16 @@ def main():
             ax = axes[f, g]
             year = fit_dat['year']
 
-            q25 = fit_dat.quantile(0.25, dim='sim').values.ravel()
-            q75 = fit_dat.quantile(0.75, dim='sim').values.ravel()
-            q90 = fit_dat.quantile(0.90, dim='sim').values.ravel()
-            q10 = fit_dat.quantile(0.10, dim='sim').values.ravel()
-            q995 = fit_dat.quantile(0.995, dim='sim').values.ravel()
-            q005 = fit_dat.quantile(0.005, dim='sim').values.ravel()
+            c50l = fit_dat.quantile(0.25, dim='sim').values.ravel()
+            c50u = fit_dat.quantile(0.75, dim='sim').values.ravel()
+            c80u = fit_dat.quantile(0.90, dim='sim').values.ravel()
+            c80l = fit_dat.quantile(0.10, dim='sim').values.ravel()
+            c99u = fit_dat.quantile(0.995, dim='sim').values.ravel()
+            c99l = fit_dat.quantile(0.005, dim='sim').values.ravel()
             
-            ax.fill_between(year, q005, q995, color='gray', alpha=0.5)
-            ax.fill_between(year, q10, q90, color='gray', alpha=0.5)
-            ax.fill_between(year, q25, q75, color='gray', alpha=0.5)
+            ax.fill_between(year, c99l, c99u, color='gray', alpha=0.33, label='99% CI')
+            ax.fill_between(year, c80l, c80u, color='gray', alpha=0.33, label='80% CI')
+            ax.fill_between(year, c50l, c50u, color='gray', alpha=0.33, label='50% CI')
             
             ax.plot(gen_dat['year'], gen_dat.values.ravel(), c='blue', linewidth=1)
             ax.semilogy()

### 2018-03-02 writeup/main.tex
version to send to dave

@@
 
-\title{Nonstationary, Quasi-Periodic Risk of Climate Extremes, and Decision Analysis}
+\title{Adapting to Quasi-Periodic and Secular Changes of Climate Risk}
 \author[1,2]{James Doss-Gollin}
@@
   Despite broad recognition of nonstationarity's importance, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
-  Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period \(M\), but not for analysis of sequential decision strategies.
+  Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies.
   In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
@@
 
-In 2011 five successive cyclones struck Thailand, causing massive damage to public and private infrastructure and crippling global supply chains.\footnote{See Masa Haraguchi's work}
-Later in June 2013 a long rainy period was followed by three summertime cyclones tracking over the same region, leading to massive flooding in central Europe \citep{Grams2014}.
-At the same time a multi-year drought in S\~ao Paulo threatened to dry out one of the world's largest cities, and in 2017 the Atlantic Hurricane season led to over \usd{100 billion} in reinsurance loss.\footnote{\href{http://www.artemis.bm/blog/2018/02/27/berkshire-hathaway-can-easily-weather-a-400bn-mega-catastrophe-buffett/}{Click for blog post on \url{artemis.bm}}}
-Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems are overwhelmed by these multi-year droughts or multi-hurricane hits.
+In 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
+In June 2013, a rainy spring was followed by three summertime cyclones following highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
+At the same time a multi-year drought in S\~ao Paolo threatened to dry out one of the world's largest cities \citep{Seth2015}, and more recently the 2017 Atlantic Hurricane season led to approximately \usd{100 billion} in insured losses.\footnote{\url{http://www.swissre.com/media/news_releases/nr20171220_sigma_estimates.html}}
+Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems may be overwhelmed by events which violate these assumptions.\footnote{Do we need to add something about climate mechanisms such as wave resonance, blocking, \etc{} that can explain some of this?
+Or just cite \citet{Hannachi2017} to say the atmosphere has regimes?}
 
-Indeed, the scientific literature has recognized that even small nonstationarity (temporal changes in dynamics of the underlying hydroclimatalogical process) can lead to significant trends in the estimated $T$-year return periods or probability of exceedance $p_T$ \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.
-This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last $N$ years of the data.\footnote{Vogel}
+Indeed, the scientific literature has recognized that even small nonstationarity (temporal change in dynamics of the underlying physical process) can lead to significant trends in the estimated $T$-year return periods or probability of exceedance $p_T$ \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.\footnote{All of these citations are for floods, can we look at other types of extremes?
+Even droughts would be helpful.}
+This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last $N$ years of the data.\footnote{Rich Vogel citation?}
 Each model will necessarily give different estimates, which may fit the historical record with comparable accuracy but diverge at an increasing rate as the model is forecast further into the future.
-Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied}.\footnote{All of these citations are for floods, can we look at other types of extremes?
-Even droughts would be helpful.}
+Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied}.
 
-There are two kinds of nonstationairities of concern.
+There are two kinds of nonstationarities of concern.
 The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Merz2014}.
@@
 \end{enumerate*}
-Because of errors inherent to discretization and parameterization of the laws of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+Because of inherent errors due to, \eg{} the discretization and parameterization of the equations of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
 A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
@@
 Would the bias and uncertainty in its estimation lead to sub-optimal investment strategies in risk mitigation?
-These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.
+These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.\footnote{Perhaps these sentences belong elsewhere?}
 
-If a positive discount rate is used for the future, as is typical in public and private sector planning models,\footnote{CITE} the investment returns on major public and private investment may diminish substantially far into the future.
-This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
-Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a paticular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
+In addition to secular change, quasi-periodic inter-annual to multi-decadal modes of climate variability \citep[see][]{Hannachi2017} to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed events \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015}.
+For example, even when no secular trend exists the using past $N$ years to represent the next $N$ leads large surprise in estimates of floods relative to when the time series is fully shuffled to remove temporal clustering, even for very long ($N=\SI{200}{years}$) sequences \citep{Jain2001}.
 
-Consideration of these shorter time horizons brings up the concern associated with the second type of nonstationarity in climate: the quasi-periodic inter-annual, decadal and multidecadal modes of climate variability that have extensive spatial teleconnections and appear to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed draws of events that is central to the classical analysis \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015,Hannachi2017}.
-Using a bootstrap approach with a long, stationary ENSO data series, \citet{Jain2001} demonstrated that the using the past \(K\) years to represent the next \(K\) years led to large surprise relative to when the data was fully randomized and the temporal structure removed.
-Even for a very long sequence (\ie{} \(K=200\) years), the past was found to be a misleading representation of the future.
-Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies,\footnote{Bonnafous et al (2017a,b)} who face spatial as well as temporal clustering.
-
-Clearly, in a nonstationary world the probability of exceedance is temporally variable over the project life.
-In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period \(M\) and the length of the historical record \(N\).
+When secular and quasi-periodic variability are both considered, it becomes clear the probability of exceedance is temporally variable over the life of a climate adaptation project.
+In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period $M$ as well as the length of the historical record $N$.
 This project planning period can be interpreted as the design life of a physical structure, or as the contract length of a financial instrument.
-Our aim is not to identify the ``best'' model, but rather to build some theory as to how the underlying nature of the natural variability and nonstationarity and the choice of model used to model future flood risk map onto estimation bias and variance in the context of many different \(M\) and \(N\).
-This estimation bias and variance may map directly onto the price of an insurance contract\footnote{cite} or the possibility of over/under-design of a flood control structure.
+Our aim is not to identify the ``best'' model, but rather to develop theory as to how the underlying nature of the natural variability and nonstationarity and the choice of model used to model future flood risk map onto estimation bias and variance in the context of many different $M$ and $N$.
+This estimation bias and variance may map directly onto the price of an insurance contract, or the possibility of over/under-design of a large structure like a dam or seawall used to protect against climate hazards.
 Though the specific examples target flood risk, this method and framework may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricanes, and wind.
@@
 
-We use a highly idealized modeling framework to generate sequences of annual-maximum floods, fit these sequences in a probabilistic framework to three flood models that incorporate time in different ways, and evalaluate the performance of these models for multiple observational record lengths \(N\) and project planning periods \(M\).
+We use a statistical framework with highly idealized models to generate sequences of annual-maximum floods, fit these sequences to probabilistic models what incorporate time in different ways, and evaluate the performance of these models for multiple observational record lengths $N$ and project planning periods $M$.
 
@@
 We generate synthetic flood sequences from two different simple conceptual models that represent chaotic dynamics and long memory.
-For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution, conditional on an expected value \(\mu(t)\) which depends on individual model:
+For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution, conditional on an expected value $\mu(t)$ which depends on individual model:
 \begin{equation} \label{eq:lognormal}
@@
 \end{equation}
-where \(\sigma_\text{min}\) is a small constant that ensures non-trivial noise in the system.
+where $\sigma_\text{min}$ is a small constant that ensures non-trivial noise in the system.
 
@@
 \begin{enumerate}
-  \item Choose \(t_0\), \(M\), and \(N\) so that the time is \(\vb{t}=t_0-N+1, \ldots, t_0+M\).
-  \item Use one of the models described below to simulate \(\mu(t)\)
-  \item Use \cref{eq:calc-sigma} to calculate \(\sigma(t)\)
-  \item For each time \(t \in \vb{t}\), draw \(Q(t)\) following \cref{eq:lognormal}.
+  \item Choose $t_0$, $M$, and $N$ so that the time is $\vb{t}=t_0-N+1, \ldots, t_0+M$.
+  \item Use one of the models described below to simulate $\mu(t)$
+  \item Use \cref{eq:calc-sigma} to calculate $\sigma(t)$
+  \item For each time $t \in \vb{t}$, draw $Q(t)$ following \cref{eq:lognormal}.
 \end{enumerate}
-Though both approaches used to generate sequences \( \mu(t) \) use highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of \(\qty[M + N]\) years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
+Though both approaches used to generate sequences $\mu(t)$ use highly simplified models neglecting many important mechanisms, they enable us to generate many synthetic sequences of $\qty[M + N]$ years which exhibit chaotic but organized quasi-periodic behavior (regime behavior or low-frequency variability) plus a trend term, thereby capturing some essential features of hydrological variability.
 
@@
 
-ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a quasi-oscillatory manner with variability on the order of 3 to 7 years with further low-frequency modulation \citep{Sarachik2010}.
+ENSO is the dominant mode of tropical climate variability, impacts flood risk around the world, and varies in a chaotic, quasi-oscillatory manner with variability on the order of 3 to 7 years with further low-frequency modulation \citep{Sarachik2010}.
 A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
-A subset of this data is shown in \cref{fig:enso-ts}.
 To create an annual time series, we average the October-December values of the NINO3 index for each year.
+A \SI{1000}{year} subset of this data is plotted in \cref{fig:enso-ts} along with the average spectral power of the series.
 
 To generate $\mu(t)$, a sequence of length $M+N$ is randomly chosen from the \SI{100000}{year} NINO3 sequence.
-Next, \(\mu\) is calculated following by
+Next, $\mu(t)$ is calculated deterministically from the NINO3 time series $x(t)$ by
 \begin{equation}
@@
 \end{equation}
-where parameter \(\gamma\) controls the magnitude of the secular trend (\(\gamma=0\) gives a stationary series).
+where $\beta$ controls the sensitivity to the NINO3 variation and $\gamma$ controls the magnitude of the secular trend ($\gamma=0$ gives a stationary series).
 
@@
 
-To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate \(\mu(t\).
+To create sequences of floods with strong memory and regime-like behavior, we use a 2-state Markov chain to sample a sequence of states that are used to generate $\mu(t)$.
 This Markov chain is given a constant probability of persistence which is the same for both states, such that the transition matrix is
@@
 \end{equation*}
-This is first used to generate a sequence of states \(S(t)\).
-The value \( \mu(t) \) depends only on the \(S(t)\) and the time itself:
+This is first used to generate a sequence of states $S(t)$.
+The value $\mu(t)$ depends only on the $S(t)$ and the time itself:
 \begin{equation*}
@@
 \end{equation*}
-For simplicity, we assume that the coefficient of variation \( \alpha \) is the same for both states.
-We also assume \( \mu_{0, 1} > \mu_{0, 2} \) so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
+For simplicity, we assume that the coefficient of variation $\alpha$ is the same for both states.
+We also assume $\mu_{0, 1} > \mu_{0, 2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
 
@@
 
-In general, each historical \(N\)-year sequence of streamflow is fit to a statistical model, from which 2000 sequences of future streamflow are generated using Monte Carlo simulation.
+In general, each historical $N$-year sequence of streamflow is fit to a statistical model, from which 2000 sequences of future streamflow are generated using Monte Carlo simulation.
 Three models are assessed for estimating future streamflow and are described in detail below.
@@
 We use the log-normal model (\cref{eq:lognormal}) for its simplicity, interpretability, and good performance relative to other 2-parameter models \citep{Vogel1996}.
-For each simulation, the first \(N\) years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations (denoted \(Q_{\text{hist}}\) for clarity).
+For each simulation, the first $N$ years \(\qty(t=[t_0-N+1, \ldots, t_0])\) are treated as observations (denoted \(Q_{\text{hist}}\) for clarity).
 Then, the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
@@
 We fit observed streamflow sequences \(Q_\text{hist}\) using a HMM as implemented in the python \texttt{pomegranate} package \citep{Schreiber2016}.
-The model is fit to the \(N\) observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
+The model is fit to the $N$ observed years of data using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
 This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
@@
 
-For a given choice of \(M\), \(N\), generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
-For each of the \(J\) sequences, the fitting model is then fit on the \(N\)-year historical record.
+For a given choice of $M$, $N$, generating model, and fitting model, \(J\) sequences of streamflow of length \(M+N\) years each are generated.
+For each of the \(J\) sequences, the fitting model is then fit on the $N$-year historical record.
 \(K\) posterior simulations of future rainfall are then drawn from the fitted model.
@@
 All computation was carried out in the \texttt{python} environment, making particular use of the \texttt{xarray}, \texttt{numpy}, \texttt{pandas}, and \texttt{matplotlib} libraries \citep{Hoyer2017,vanderWalt2011,McKinney2010,Hunter2007}.
-The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.
+The codes used to generate the figures and text of this paper are available at \url{github.com/jdossgollin/MNpaper}.\footnote{It's currently a private repository}
 
@@
 
-We begin by showing the types of synthetic streamflow sequence that are by the two models discussed \cref{sec:methods-generating}.
-We then examine how different statistical models for estimating future flood sequences perform under stationary conditions and when a secular trend is introduced.
-
-\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
-
-The models used to generate synthetic flood sequences used simple dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
-
-To verify that these synthetic streamflow sequences sufficiently represent these behaviors, we plot several sequences.
-The ENSO model used a long run of the Cane-Zebiak model \citep{Zebiak1987,Ramesh2017} to determine \(\mu(t)\); a \SI{2500}{year} sub-set of the annualized time series is shown in \cref{fig:enso-ts}.
-As described in the literature,\footnote{citations, and maybe some paleo stuff as well} this model shows quasi-periodic oscillations in the \SIrange{3}{7}{year} band but with substantial low-frequency modulation of extreme activity.
-The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
-Analysis of the power density spectrum shows a strong peak around 3-6 years.
-\begin{figure}
-  \includegraphics[width=\textwidth]{enso.pdf}
-  \caption{
-    (L): A \SI{1000}{year} sub-set of the synthetic time series described in \cref{sec:methods-nino}.
-    (R): The power spectral density of the annualized time series. A moving window was applied to the density for smoothing.\label{fig:enso-ts}
-  }
-\end{figure}
-
-\Cref{fig:stationary-sequences} shows 50 streamflow sequences generated with \(M=50\) and \(N=100\).
-The ENSO sequences were generated with \(\gamma=0, \beta_\mu=0.5, \mu_0=6\), which gives a strong (but still noisy) relationship between the NINO3 index \(x(t)\)) and the annual-maximum streamflow \(Q(t)\).
-The Markov sequences were generated with \(\mu_1 = 6.55, \mu_2=5.7, \gamma_1=0, \gamma_2=0, \pi=0.9\), which gives a large difference between the two states.
-For both, the coefficient of variation \(\alpha=0.1\) and the minimum value of sigma \(\sigma_\text{min}=0.01\).
-\begin{figure}[b]
-  \includegraphics[width=\textwidth]{sequences_stationary.pdf}  
-  \caption{
-    50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-nino} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
-    Both models use a stationary parameterization \(\gamma=0\).
-    49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
-    Note that $y$-axis has a log scale.\label{fig:stationary-sequences}
-  }
-\end{figure}
-
-To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
-For the ENSO time series, this mean setting \(\gamma=0.015\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.015\).
-The resulting sequences are shown in \cref{fig:trend-sequences}.
-The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
-\begin{figure}[b]d
-  \includegraphics[width=\textwidth]{sequences_trend.pdf}
-  \caption{
-    As \cref{fig:stationary-sequences} but a trend term has been added by setting \(\gamma=0.015\).\label{fig:trend-sequences}
-  }
-\end{figure}
-
-This choice of parameterizing secular change as a linear trend on the expected value of log streamflow was made to aid interpretability, and not because of any particular physical reasoning.
-Indeed, for large \(N\) this model suggests implausibly dry conditions and implausibly wet conditions for very large \(M\).
-However, given widespread use of trend-based models in the flood frequency analysis literature it is informative to consider the ``best-case'' scenario in which the model is correctly specified and there is a true, linear trend in the data.
+We begin by considering how different statistical models for estimating future flood sequences perform under stationary conditions and when a secular trend is introduced.
 
@@
 
-We next consider the estimation bias and variance as a function of \(M\) and \(N\) when the true generative process is stationary.
+We first consider the estimation bias and variance as a function of $M$ and $N$ when the true generative process is stationary.
 
-To do this we generate 1000 sequences of streamflow for each combination of \(M\), \(N\), and generative process.
-We then fit each estimation model to each sequence and draw 2000 estimates of future streamflow from the fitted model.
-These are then used to determine \(\hat{p}_T\), allowing calculation of the expected bias and variance for each combination of \(M\), \(N\), generative model, and estimating model.
+To do this we generate 1000 sequences of streamflow for each combination of $M$, $N$, and generative process.
 The parameters for the NINO3 model are given in \cref{tab:nino-stationary} and for the Markov model in \cref{tab:nino-stationary}.
 We use a threshold $Q_T = \SI{5000}{units}$.
+We then fit each estimation model to each sequence and draw 2000 estimates of future streamflow from the fitted model.
+These are then used to determine $\hat{p}_T$, allowing calculation of the expected bias and variance in its estimation for each combination of $M$, $N$, generative model, and estimating model.
 
@@
 \Cref{fig:stationary-bias} shows that in general, bias is quite similar across generative models but varies substantially between the models used for estimation.
-The HMM and LN2 Stationary models have no time term, and so their bias is independent of \(M\).
-In general both of these models have very low bias, though the LN2 Stationary model tends to over-estimate the probability of flooding for very short \(N\), unsurprising since the true distribution of streamflow is bimodal for both generating functions.
-The LN2 trend model registers a non-trivial probability of a positive time trend, particularly for short \(N\), leading to large bias in estimation.
-For very long \(M\) the bias is higher, as estimated probability of exceeding the threshold is exceeded.
+The HMM and LN2 Stationary models have no time term, and so their bias is independent of $M$.
+In general both of these models have very low bias, though the LN2 Stationary model tends to over-estimate the probability of flooding for very short $N$, unsurprising since the true distribution of streamflow is bimodal for both generating functions.
+The LN2 trend model registers a non-trivial probability of a positive time trend, particularly for short $N$, leading to large bias in estimation.
+For very long $M$ the bias is higher, as estimated probability of exceeding the threshold is exceeded.
 \begin{figure}[ht]
   \includegraphics[width=\textwidth]{bias_stationary.pdf}
-  \caption{The estimation bias for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-bias}}
+  \caption{The estimation bias for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of $M$ and $N$.\label{fig:stationary-bias}}
 \end{figure}
@@
 As suggested in the bias plots, the HMM and LN2 Stationary models have very low variance of their estimates.
-The LN2 Trend model has much higher variance in estimates, particularly for short \(N\) or short \(M\).
+The LN2 Trend model has much higher variance in estimates, particularly for short $N$ or short $M$.
 \begin{figure}[ht]
   \includegraphics[width=\textwidth]{variance_stationary.pdf}
-  \caption{The estimation variance for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:stationary-variance}}
+  \caption{The estimation variance for 1000 sequences generated by each of two stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of $M$ and $N$.\label{fig:stationary-variance}}
 \end{figure}
@@
 
-For short \(M\), particularly when \(N\) is not too big, the stationary models (HMM and LN2 Stationary) perform well for both generative functions.
-As \(M\) increases, they begin to substantially underestimate the probability of future flooding.
-As for the stationary case, when \(N\) is short the LN2 Stationary model tends to over-estimate $\hat{p}_T$ because it sees variation caused by the NINO3 time series or the Markov state switching and assigns it to internal variability, leading to unrealistically fat tails in the estimate.
+For short $M$, particularly when $N$ is not too big, the stationary models (HMM and LN2 Stationary) perform well for both generative functions.
+As $M$ increases, they begin to substantially underestimate the probability of future flooding.
+As for the stationary case, when $N$ is short the LN2 Stationary model tends to over-estimate $\hat{p}_T$ because it sees variation caused by the NINO3 time series or the Markov state switching and assigns it to internal variability, leading to unrealistically fat tails in the estimate.
 The trend model, by contrast, over-estimates $p_T$ except when $N$ is very large, mainly because the posterior uncertainty in the magnitude of the trend means that the model believes that a trend larger than observed is not unlikely.
@@
   \includegraphics[width=\textwidth]{bias_trend.pdf}
-  \caption{The estimation bias for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:trend-bias}}
+  \caption{The estimation bias for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of $M$ and $N$.\label{fig:trend-bias}}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{variance_trend.pdf}
-  \caption{The estimation variance for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of \(M\) and \(N\).\label{fig:trend-variance}}
+  \caption{The estimation variance for 1000 sequences generated by each of two non-stationary generating functions (row) and fit by each of three fitting functions (columns) as a function of $M$ and $N$.\label{fig:trend-variance}}
 \end{figure}
-
-% -----------------------------------------------------------------------------
-% DISCUSSION
-% -----------------------------------------------------------------------------
-
-\section{Discussion}
-
-% PHYSICAL PLAUSABILITY OF TRENDS 
-% SPATIAL CASE
-% THIS IS RELEVANT FOR OTHER NATURAL HAZARDS -- DISCUSS
-% IMPLICATIONS FOR INSURANCE / DESIGN
-
-For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
-While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
-
-From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
-How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length \(M\).
-What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
-
-How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
-
-Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
 
@@
   \caption{
-    A single sequence of streamflow generated by the NINO3 function with (L) no trend and (R) \(\gamma=0.01\). 
-    The sequence is fit to each fit to three different fitting models (rows).
-    Confidence intervals shown are 50\%, 80\%, and 99\%.
-    Generated with \(M=N=150\).\label{fig:example-long}
+    A single sequence of streamflow generated by the NINO3 function with (L) $\gamma=0$ and (R) $\gamma=0.015$. 
+    The sequence is fit to each fit to three different fitting models (rows); the grey shading indicates the posterior 50\%, 80\%, and 99\% confidence intervals of $\hat{P}_T(t)$.
+    Here $M=N=150$.\label{fig:example-long}
   }
@@
 \end{figure}
+
+% -----------------------------------------------------------------------------
+% CASE STUDY
+% -----------------------------------------------------------------------------
+
+\section{Case Study}
+
+I would go back to a case study and look at a 500-year streamflow sample (could be generated synthetically if we can't get a good real one).
+Fix $N=100$ (say) and choose between $M=5$ or $M=50$. 
+Estimate this model with 
+\begin{enumerate}[label=(\roman*)]
+  \item HMM;
+  \item LN2 Stationary;
+  \item LN2 Trend; and
+  \item Perfect Information (\ie{} knowing the future exactly)
+\end{enumerate}
+Then at each time, calculate the insurance price for each model from \cref{eq:insurance-price} (the perfect information model will have a risk premium of zero) and summarize the over or under-pricing of the risk premium.
+Then explain that one would be unlikely to buy a 50-year insurance contract but that it's helpful to compare the two models using equal criteria.
+We could extend to include NPV analysis if desired.
+
+\emph{Please comment: should this be part of our paper?}
+
+% -----------------------------------------------------------------------------
+% DISCUSSION
+% -----------------------------------------------------------------------------
+
+\section{Discussion}
+
+Infrastructure in the developed world is aging \citep{Ho2017}, leading to questions of whether and how to replace it.
+In the developed world new infrastructure is being built.\footnote{citations}
+Given a positive future discount rate and uncertainty associated with future urbanization, economic growth, and climate change, the expected net benefits of this infrastructure are highly uncertain.
+At the same time the private sector has accepted an increasing share of global risk \citep{WorldEconomicForum2016}, which it has mitigated primarily through financial instruments.
+How should different contracts be priced? When does it make sense to build large infrastructure?
+These questions cannot be answered with existing methodologies.
+
+The assessment of sampling bias and variance associated with estimating the occurrence probability of an extreme event, $\hat{p}_T$, maps directly into this insurance contract, where a typical price for a $M$-year contract is given by
+\begin{equation}\label{eq:insurance-price}
+  \frac{\mathbb{C}}{M} = \lambda \qty(\mathbb{E} \qty[p_T] + c \mathbb{V} \qty[p_T])
+\end{equation}
+where $\mathbb{E}$ denotes the expectation operator, $\mathbb{V}$ the variance operator, $\lambda$ the profit margin, and $c$ the risk premium.
+Bias in $\hat{P}_T$ leads to inefficient pricing, since if the price is too low, the insurer is exposed to failure and if it is too high the buyer will be less inclined to purchase insurance.
+
+We illustrated sampling uncertainty and inter-model disagreement using several simple models for annual-maximum floods.
+In more realistic applications, analysis of flood frequency can be improved by combining information across space \citep{Lima2016,Merz2008a}, combining model forecasts, \etc{}.
+Analysis of other weather and climate hazards may use different distributions, and the time structure of the underlying climate process may be different.
+
+For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
+While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
+Recent efforts\footnote{Prad's work} to identify regions where risk of a particular event is negatively correlated in space.
+
+Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies,\footnote{Bonnafous et al (2017a,b)} who face spatial as well as temporal clustering.
+
+How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
+
+Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
+
+From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
+How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length $M$.
+What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
+If a positive discount rate is used for the future, as is typical in public and private sector planning models,\footnote{CITE} the investment returns on major public and private investment may diminish substantially far into the future.
+This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
+Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
 
@@
 
-Implications
+To really drive this paper home, we ought to end with a few recommendations for practitioners estimating and adapting to future climate risk such as:
 \begin{itemize}
-  \item Blind use of trend models leads to sub-optimal models \emph{even when there is a trend} for many combination of \(M,N\)
-  \item Bias-variance tradeoff of model complexity
-  \item Need to consider the structure of the underlying trend, $N$, and $M$ when choosing a model
-  \item Should consider multiple models, \eg by ``stacking'' \citep{Yao2017}
+  \item For some combinations of $M$ and $N$, use of trend models leads to sub-optimal decisions \emph{even when there is a truely linear trend and the model is correctly specified}
+  \item Consider the structure of the underlying trend, $N$, and $M$ when choosing a model
+  \item Consider the bias-variance tradeoff of model complexity
 \end{itemize}
+We could even offer some recommendations as to when to choose ``insurance'' projects versus ``infrastructure'' projects, but this may just needlessly antagonize readers.
 
@@
 
+\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
+
+The models used to generate synthetic flood sequences used simple dynamics, but nevertheless were designed to incorporate regime-like and quasi-period variability at multiple timescales.
+
+To verify that these synthetic streamflow sequences sufficiently represent these behaviors, we plot several sequences.
+The ENSO model used a long run of the Cane-Zebiak model \citep{Zebiak1987,Ramesh2017} to determine \(\mu(t)\); a \SI{2500}{year} sub-set of the annualized time series is shown in \cref{fig:enso-ts}.
+As described in the literature,\footnote{citations, and maybe some paleo stuff as well} this model shows quasi-periodic oscillations in the \SIrange{3}{7}{year} band but with substantial low-frequency modulation of extreme activity.
+The index is also asymmetrically distributed, with a longer tail on the positive (El Ni\~{n}o) side than the negative (La Ni\~{n}a) phase, which is consistent with observations of ENSO.
+Analysis of the power density spectrum shows a strong peak around 3-6 years.
+\begin{figure}
+  \includegraphics[width=\textwidth]{enso.pdf}
+  \caption{
+    (L): A \SI{1000}{year} sub-set of the synthetic time series described in \cref{sec:methods-nino}.
+    (R): The power spectral density of the annualized time series. A moving window was applied to the density for smoothing.\label{fig:enso-ts}
+  }
+\end{figure}
+
+\Cref{fig:stationary-sequences} shows 50 streamflow sequences generated with $M=50$ and $N=100$.
+The ENSO sequences were generated with the parameters shown in \cref{tab:nino-stationary} Markov sequences were generated with the parameters shown in \cref{tab:markov-stationary}.
+\begin{figure}[b]
+  \includegraphics[width=\textwidth]{sequences_stationary.pdf}  
+  \caption{
+    50 streamflow sequences from (Top) the ENSO-based model described in \cref{sec:methods-nino} and (Bottom) the Markov chain modle described in \cref{sec:methods-markov}.
+    Both models use a stationary parameterization $\gamma=0$.
+    49 of the 50 sequences are plotted in gray, and one (randomly chosen) is plotted in blue for emphasis.
+    Note that $y$-axis has a log scale.\label{fig:stationary-sequences}
+  }
+\end{figure}
+
+To incorporate secular variability into our model, we repeat the exercise but add a trend term to the data.
+For the ENSO time series, this mean setting \(\gamma=0.015\) while for the Markov case this meant a trend term on the wet state only, so that \(\gamma_1=0.015\).
+The resulting sequences are shown in \cref{fig:trend-sequences}.
+The trend has an extremely large impact on the probability of an extremely high flood, but the underlying time series structure is generally preserved.
+\begin{figure}[b]d
+  \includegraphics[width=\textwidth]{sequences_trend.pdf}
+  \caption{
+    As \cref{fig:stationary-sequences} but a trend term has been added by setting \(\gamma=0.015\).\label{fig:trend-sequences}
+  }
+\end{figure}
+
+This choice of parameterizing secular change as a linear trend on the expected value of log streamflow was made to aid interpretability, and not because of any particular physical reasoning.
+Indeed, for large $N$ this model suggests implausibly dry conditions and implausibly wet conditions for very large $M$.
+However, given widespread use of trend-based models in the flood frequency analysis literature it is informative to consider the ``best-case'' scenario in which the model is correctly specified and there is a true, linear trend in the data.
+
 \section{Notation}
@@
   \item[\( x(t) \)] the annualized time series of the NINO3 index
-  \item[\( p_T \)] the true probability that \( Q(t) \geq Q_T \) in the future \(M\)-year period
-  \item[\( \hat{p}_T \)] the estimated probability that \( Q(t) \geq Q_T \) in the future \(M\)-year period
+  \item[\( p_T \)] the true probability that \( Q(t) \geq Q_T \) in the future $M$-year period
+  \item[\( \hat{p}_T \)] the estimated probability that \( Q(t) \geq Q_T \) in the future $M$-year period
 \end{description}

### 2018-03-02 src/floodsampling/fit.py
use lognormal distribution in stan to account for jacobian transformation

@@ -279,7 +279,6 @@ class StationaryLN2Stan(FloodFit):
             coords={'sim': sim, 'year': years},
             dims=['sim', 'year']
         )
-        samples = np.exp(samples)
 
         return samples
 
@@ -383,7 +382,6 @@ class TrendLN2Stan(FloodFit):
             coords={'sim': sim, 'year': years},
             dims=['sim', 'year']
         )
-        samples = np.exp(samples)
         return samples
 
 class HMM(FloodFit):

### 2018-03-03 writeup/main.tex
Dave updated intro

@@
 \begin{abstract}
-  Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep[``nonstationarity'';][]{Milly2008,Merz2014,Hodgkins2017}.
+  Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep[``nonstationarity'';][]{Milly2008,Merz2014,Hodgkins2017}\footnote{Some people don't like citations in the abstract unless the paper is a rebuttal to another -- in which case citing the original is probably good.
+My impression is that our citations int he abstract are not highly necessary so I may take them out.
+Up to you.}.
   This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
@@
   Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies.
-  In this paper we illustrate some basic considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary adaptation.
+  In this paper we illustrate some considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary setting.
   We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show \(\ldots{}\).
@@
 Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems may be overwhelmed by events which violate these assumptions.\footnote{Do we need to add something about climate mechanisms such as wave resonance, blocking, \etc{} that can explain some of this?
-Or just cite \citet{Hannachi2017} to say the atmosphere has regimes?}
+Or just cite \citet{Hannachi2017} to say the atmosphere has regimes?} \footnote{As for $\uparrow$, I would avoid the impulse to get too into the dynamics in this paper (i.e.
+wave resonance, etc).
+However, I do think that this paragraph could be a bit more direct.
+There is ample evidence that climate extremes are not IID $\rightarrow$ for example, Thailand, etc $\rightarrow$ thus risk analyses need to consider this...
+Basically my critique is that three sentences about recent disasters without the context of why we particularly care about these things may be a bit of a misdirection for the reader}.
 
@@
 Each model will necessarily give different estimates, which may fit the historical record with comparable accuracy but diverge at an increasing rate as the model is forecast further into the future.
-Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance like instruments has been much less studied}.
+Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance instruments has been much less studied}.
 
@@
 \end{enumerate*}
-Because of inherent errors due to, \eg{} the discretization and parameterization of the equations of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence \citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
-A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
+Because of inherent errors due to, \eg{} the discretization and parameterization of the equations of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence\footnote{Is "inspire confidence" too informal?...
+I'm not sure}.
+\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+A critical question in this regard is how to assess whether any structure\footnote{or other risk management instrument such as an insurance product?} designed using these projections may actually perform in the future relative to the proposed level of risk protection.
 Would it be over- or under- designed?
@@
 In addition to secular change, quasi-periodic inter-annual to multi-decadal modes of climate variability \citep[see][]{Hannachi2017} to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed events \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015}.
-For example, even when no secular trend exists the using past $N$ years to represent the next $N$ leads large surprise in estimates of floods relative to when the time series is fully shuffled to remove temporal clustering, even for very long ($N=\SI{200}{years}$) sequences \citep{Jain2001}.
+For example, even when no secular trend exists, using the past $N$ years to project the next $N$ leads to different estimates of floods relative to when the time series is fully shuffled to remove temporal clustering, even for very long ($N=\SI{200}{years}$) sequences \citep{Jain2001}.
 
@@
 Though the specific examples target flood risk, this method and framework may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricanes, and wind.
+\footnote{Overall comment on the introduction:  We may consider hinting at some of the results a bit more, otherwise the introduction is fairly wide open and I'm not sure the reader gets a good sense for what she is in for...
+but then again, perhaps that's OK}.
 
@@
 
-We use a statistical framework with highly idealized models to generate sequences of annual-maximum floods, fit these sequences to probabilistic models what incorporate time in different ways, and evaluate the performance of these models for multiple observational record lengths $N$ and project planning periods $M$.
+We use a statistical framework with highly idealized models to generate sequences of annual-maximum floods, fit these sequences to probabilistic models that incorporate time in different ways, and evaluate the performance of these models for multiple observational record lengths $N$ and project planning periods $M$.
 

### 2018-03-03 writeup/main.tex
Dave updated methods

@@
 We generate synthetic flood sequences from two different simple conceptual models that represent chaotic dynamics and long memory.
-For both cases, the expected value of the annual maximum flood flood follows a log-normal distribution, conditional on an expected value $\mu(t)$ which depends on individual model:
+For both cases, the expected value of the annual maximum flood follows a log-normal distribution, conditional on an expected value $\mu(t)$, which in turn depends on individual models described in \cref{sec:methods-nino,sec:methods-markov}:
 \begin{equation} \label{eq:lognormal}
   \log Q(t) \big| \mu(t),\sigma(t) \sim \normal \qty(\mu(t), \sigma(t)).
-\end{equation}.
-For simplicitiy we assume a constant coefficient of variation:
+\end{equation}
+For simplicity we assume a constant coefficient of variation:
 \begin{equation}\label{eq:calc-sigma}
@@
 A \SI{100000}{year} integration of the Cane-Zebiak model \citep{Zebiak1987} was used to produce a monthly NINO3.4 index with stationary forcing as described in \citet{Ramesh2017}.
-To create an annual time series, we average the October-December values of the NINO3 index for each year.
+To create an annual time series, we average the October-December values of the NINO3 index for each year\footnote{So is this nino3.4 or 3?}.
 A \SI{1000}{year} subset of this data is plotted in \cref{fig:enso-ts} along with the average spectral power of the series.
@@
 
-This paper does not propose new models or suggest a general ``best'' model; instead it aims to describe the total uncertainty in estimates using easily methods which have been well studied in the literature.
+This paper does not propose new models or suggest a general ``best'' model; instead it aims to describe the total uncertainty in estimates using methods that have been well studied in the literature.
 
@@
 Then, the parameters are fit in a Bayesian framework to fully represent the posterior uncertainty as implemented in the \texttt{stan} probabilistic programming language \citep{Carpenter2016}.
-Weak (uninformative) priors are chosen; in the context of a synthetic data experiment using informative priors seems inappropriate but other studies\footnote{cite} have used scaling information and regional information to inform prior distributions and reduce posterior variance.
+Weak (uninformative) priors are chosen; in the context of a synthetic data experiment using informative priors seems inappropriate but other studies\footnote{cite} have used scaling information and regional information to inform prior distributions and reduce posterior variance\footnote{Prob you can drop this aside -- might just serve as a distraction to the main point here}.
 The full probability model is thus:

### 2018-03-03 writeup/main.tex
Dave updated results, discussion, and summary

@@
 To do this we generate 1000 sequences of streamflow for each combination of $M$, $N$, and generative process.
-The parameters for the NINO3 model are given in \cref{tab:nino-stationary} and for the Markov model in \cref{tab:nino-stationary}.
+The parameters for the NINO3 and Markov models are given in \cref{tab:nino-stationary}.
 We use a threshold $Q_T = \SI{5000}{units}$.
@@
 
-\Cref{fig:stationary-bias} shows that in general, bias is quite similar across generative models but varies substantially between the models used for estimation.
+\Cref{fig:stationary-bias}\footnote{It's more intuitive for me if the M axis is increasing upwards, but perhaps there is a reason for your orientation that I am not considering} shows that in general, bias is quite similar across generative models but varies substantially between the models used for estimation.
 The HMM and LN2 Stationary models have no time term, and so their bias is independent of $M$.
@@
 
-Analysis of the variance, in these estimates, shown in \cref{fig:stationary-variance} as the log standard deviation, is similar.
+The variance estimates show similar patterns to those of the bias and are shown in \cref{fig:stationary-variance} as the log standard deviation.
 As suggested in the bias plots, the HMM and LN2 Stationary models have very low variance of their estimates.
-The LN2 Trend model has much higher variance in estimates, particularly for short $N$ or short $M$.
+The LN2 Trend model has much higher variance in estimates, particularly for short $N$ and long $M$.
 \begin{figure}[ht]
@@
 
-\emph{Please comment: should this be part of our paper?}
+\emph{Please comment: should this be part of our paper?}\footnote{I vote yes to include this, since I think it nails the "so-what?" part}
 
@@
 
-\section{Discussion}
+\section{Discussion}\footnote{Not doing much editing from here on yet, since I think it will depend a bit on a more finalized version of the above sections.
+The beginning of the discussion does read as introductory material to me -- likely a case study in between sections 3 and 5 will help ease the reader from the discussion of bias and variance for N and M (without much context) in section 3 into the very applied considerations discussed at the beginning of section 5.
+Still I do think that the discussion could use some re-focusing.
+Also, adding some context in the results section will help to set up a productive discussion I think.}
 

### 2018-03-03 src/floodsampling/core.py
import pickle to avoid error

@@ -3,6 +3,7 @@
 import os
 from hashlib import md5
 from collections import OrderedDict
+import pickle
 
 import numpy as np
 import xarray as xr

### 2018-03-26 writeup/main.tex
Slow progress on writeup

@@
 \author[1,2]{James Doss-Gollin}
+\author[1,2]{David Farnham}
 \author[1,2]{Upmanu Lall}
-\author[1,2]{David Farnham}
 \affil[1]{Columbia Water Center, Columbia University}
@@
 
+\input{header/glossaries.tex}
 \input{header/biblatex.tex}
@@
 \begin{abstract}
-  Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep[``nonstationarity'';][]{Milly2008,Merz2014,Hodgkins2017}\footnote{Some people don't like citations in the abstract unless the paper is a rebuttal to another -- in which case citing the original is probably good.
-My impression is that our citations int he abstract are not highly necessary so I may take them out.
-Up to you.}.
-  This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
-  Despite broad recognition of nonstationarity's importance, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
-  Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies.
-  In this paper we illustrate some considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary setting.
-  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show \(\ldots{}\).
   This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
@@
 
+\section{Introduction}\label{sec:introduction}
+
+The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of climate extremes for use in risk analysis and mitigation studies. 
+In this framework, event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
+Under the assumption that the historical or simulated record of events constitutes an stationary, \gls{iid} process, and using an appropriate statistical inference model, an estimate of this relationship and often its uncertainty based on a finite $N$-year record is developed.
+This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation, typically using expected net present value analyses \citep[\ie{}][]{Rosner2014}, applied to the costs and residual losses expected from the mitigation instrument.
+The uncertainty associated with the model may be used to develop a risk premium for insurance, or for assessing the potential over- or under- design for infrastructure.
+
+Over time, it has become clear that even small fluctuations in the dynamics of the hydroclimate system, typically called nonstationarity, can lead to clustering of extreme events  and strong time dependence of a public or private actor's risk profile \citep{Bonnafous2017a,Bonnafous2017b}.
+For example, in 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
+Similarly, in June 2013 three summertime cyclones followed highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
+More recently the 2017 Atlantic hurricane season led to over \usd{130 billion} in insured losses \citep{SwissRe2017}, despite negligible losses in the preceding $N$ years.
+
+In recent years a plethora of efforts have aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Waylen1986,Sveinsson2005,Griffis2007}, or using only the last $N$ years of the data.
+Despite such efforts, how one should use these nonstationarity estimates for the design of risk mitigation infrastructure, or the pricing of insurance like instruments has been much less studied.
+
+Two forms of nonstationarity are of concern.
+The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Milly2008,Merz2014}.
+The uncertainties associated with this are not purely informational, at least in the sense that they can be inferred from a historical sample or model.
+Both nature and human actions over the course of this century will determine the magnitude and direction of secular change, and neither is fully predictable.
+The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
+Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
+While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+
+% This is a strange paragraph,.
+A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
+Would it be over- or under- designed?
+Would the high degree of bias and uncertainty in its estimation lead to suboptimal investment strategies in risk mitigation?
+These questions go well beyond what can be reliably answered at this time.
+This kind of epistemic uncertainty clearly grows as we look out into the distant future, and the importance of investment decisions that may engage us in considerations of designs that go far out into the future, e.g., 2100AD, diminish if any positive discount rate is used for the future.
+While this consideration offsets the impact of the growing uncertainty for longer project life, it also suggests that it may be more effective to understand the consequences for analysis that may focus on shorter lived mitigation alternatives, ones that may last for 1, 5, 10, 20,  30 or 50 years.
+Sequences of such projects may better reflect climate adaptation goals in any case, especially if we believe that major climate changes may only happen beyond these horizons, or if we are optimistic and believe that humans may avert much of the potential catastrophe by the end of the longer of these horizons.
+
+Focusing on these shorter time horizons brings up the concern associated with the second type of nonstationarity in climate: the quasi-periodic inter-annual, decadal and multidecadal modes of climate variability \citep[see][for a review]{Hannachi2017} that influence the spatiotemporal expression of climate extremes in ways that violate the assumption of \gls{iid} draws of events that is central to classical analyses \citep{Matalas2012}.
+For example, \citet{Jain2001} considered how \gls{enso}, an example of one such process, impacts flood risk when there is zero secular change.
+Using a \SI{1000}{year} synthetic \gls{enso} time series, the largest event from a block of $N$ years was sampled from the historical record as the target design level.
+Next, the number of times this magnitude was exceeded in the subsequent block of $N$ years was then recorded.
+A corresponding randomized experiment was also conducted where the $N$ years in each step were selected at random, consistent with an \gls{iid} process hypothesis.
+Crucially, the bootstrap frequency distribution associated with event magnitude exceedance process for the structured blocks was dramatically fatter tailed than the one for the randomized blocks, demonstrating that extreme events are significantly clustered in time relative to an \gls{iid} process.
+This finding also held when a long sequence of historical streamflow observations was used in place of synthetic data.
+
+Given the \gls{iid} assumptions in her analysis of climate extremes, a typical risk analyst uses annualized cost-benefit analyses over the projected life of a structural risk mitigation project, and the expected loss in a given year from the exceedance of the nominal design level is computed as the potential loss* probability of exceedance.
+Clearly, in a nonstationary world, the probability of exceedance could be considered to be temporally variable over the project life.
+This would however, need to be a prediction of that probability ex ante, using appropriate modeling tools applied to the historical data or to climate model based forecasts.
+Faced with nonstationarity, some analysts propose using only the most recent subset of the data, e.g., the last 20 or 30 years to build the event magnitude – probability of exceedance relationship.
+This can be thought of as a “persistence” model, which may be “better than nothing”, and may or may not include a trend in the parameters of the probability distribution.
+How many years should one retain from the recent past?
+Too few and the estimation uncertainty dominates.
+Too many and the bias may dominate.
+Of course, if the underlying climate is quasi-periodic then this choice can be arbitrarily poor, since the last batch of K years may be of exactly the opposite phase of the oscillation as what comes next.
+In summary, we need to carefully consider the number of years of historical data, N, and the number of future years M over which the instrument of interest will be evaluated, as well as efforts at using the data from the N year period to predict what will happen in the M years.
+The risk estimation from the N years needs to explicitly inform a risk prediction over the M years, and depending on the approach and the form of the underlying nonstationarity, we can expect both bias and uncertainty associated with the prediction, and we need temporally varying probabilities to address the inevitable risk clustering.
+This is then the statement of the risk mitigation design problem that we need to consider for nonstationary climate.
+This paper is the first to formally develop these ideas in the context of a decision analytic framework.
+
+We focus on structured experiments with synthetic and actual data on extremes to illustrate how one could approach the analyses and what the implications are of varying N and M on the ability to do effective inference and on the ability to judge operational performance.
+We consider idealized cases for the structural and financial instruments to convey the ideas, and defer more complex sequential decision making tasks to a subsequent paper.
+The idea that one needs to consider the joint uncertainty in the estimation and occurrence of climate extremes for the N and M periods for stationary conditions was originally considered by Lall (198x).
+That analysis corresponds to what one would expect under i.i.d.
+conditions, and would hence provide a baseline to compare the uncertainty distribution associated with a portfolio whose elements were designed using N years of data with an application to M future years at a common design probability of exceedance for the elements.
+
 \section{Introduction}
@@
 In June 2013, a rainy spring was followed by three summertime cyclones following highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
-At the same time a multi-year drought in S\~ao Paolo threatened to dry out one of the world's largest cities \citep{Seth2015}, and more recently the 2017 Atlantic Hurricane season led to approximately \usd{100 billion} in insured losses.\footnote{\url{http://www.swissre.com/media/news_releases/nr20171220_sigma_estimates.html}}
-Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems may be overwhelmed by events which violate these assumptions.\footnote{Do we need to add something about climate mechanisms such as wave resonance, blocking, \etc{} that can explain some of this?
-Or just cite \citet{Hannachi2017} to say the atmosphere has regimes?} \footnote{As for $\uparrow$, I would avoid the impulse to get too into the dynamics in this paper (i.e.
-wave resonance, etc).
-However, I do think that this paragraph could be a bit more direct.
-There is ample evidence that climate extremes are not IID $\rightarrow$ for example, Thailand, etc $\rightarrow$ thus risk analyses need to consider this...
-Basically my critique is that three sentences about recent disasters without the context of why we particularly care about these things may be a bit of a misdirection for the reader}.
+At the same time a multi-year drought in S\~ao Paolo threatened to dry out one of the world's largest cities \citep{Seth2015}, and more recently the 2017 Atlantic Hurricane season led to over \usd{130 billion} in insured losses \citep{SwissRe2017}.
+These particular events suggest that core assumptions of classical analysis of climate extremes for use in risk analysis and mitigation studies, particularly that the historical or simulated record of events constitutes an independent and identically distributed process, are inconsistent with observations.
 
@@
 
-In addition to secular change, quasi-periodic inter-annual to multi-decadal modes of climate variability \citep[see][]{Hannachi2017} to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed events \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015}.
-For example, even when no secular trend exists, using the past $N$ years to project the next $N$ leads to different estimates of floods relative to when the time series is fully shuffled to remove temporal clustering, even for very long ($N=\SI{200}{years}$) sequences \citep{Jain2001}.
+In addition to secular change, quasi-periodic inter-annual to multi-decadal modes of climate variability \citep[see][]{Hannachi2017} to influence the spatiotemporal expression of climate extremes in ways that violate the assumption of independent, identically distributed events \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015}.
 
@@
 
-We first consider fitting a stationary model to the observed flood record, following classical assumptions of \iid sequences and stationarity.
+We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences and stationarity.
 
@@
 
-Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies,\footnote{Bonnafous et al (2017a,b)} who face spatial as well as temporal clustering.
+Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies, who face spatial as well as temporal clustering \citep{Bonnafous2017a,Bonnafous2017b}.
 
@@
 % -----------------------------------------------------------------------------
+
 \appendix

### 2018-03-29 writeup/main.tex
Updates to text, primarily case study and intro

--- 2018-03-03 writeup/main.tex
@@
 
-\title{Adapting to Quasi-Periodic and Secular Changes of Climate Risk}
+\title{Adapting to Quasi-Periodic and Secular Variation of Climate Risk}
 \author[1,2]{James Doss-Gollin}
+\author[1,2]{David Farnham}
 \author[1,2]{Upmanu Lall}
-\author[1,2]{David Farnham}
 \affil[1]{Columbia Water Center, Columbia University}
@@
 \RaggedRight{}
+
 \begin{abstract}
-  Hydroclimatic systems vary on many timescales, including regime-like variability and secular variability  \citep[``nonstationarity'';][]{Milly2008,Merz2014,Hodgkins2017}\footnote{Some people don't like citations in the abstract unless the paper is a rebuttal to another -- in which case citing the original is probably good.
-My impression is that our citations int he abstract are not highly necessary so I may take them out.
-Up to you.}.
-  This multi-scale variability greatly complicates the estimation of future flood sequences, leading to ``low confidence in projections of changes in fluvial floods'' \citep{IPCC2012}.
-  Despite broad recognition of nonstationarity's importance, there has been little theoretical development of ideas for the design and operation of infrastructure that simultaneously consider the potential for predictability of secular change and regime behavior or low-frequency variability.
-  Consequently, existing theories of flood risk assessment are well-suited to analysis of flood adaptation strategies with fixed project operation period $M$, but not for analysis of sequential decision strategies.
-  In this paper we illustrate some considerations of uncertainty and risk analysis associated with the sequential decisions that may need to be made for climate risk adaptation in a nonstationary setting.
-  We simulate synthetic data and fit the resulting simulations in a Bayesian context using stationary and non-stationary models to show \(\ldots{}\).
-  This is a first conceptual step to the development of a full sequential decision analysis approach under nonstationarity.
+  Recent disasters highlight the need for improved adaptation to extreme floods and other climate hazards.
+  Climate risk mitigation has traditionally been achieved through large-scale infrastructure projects with long project lifespans.
+  However, political and economic constraints limit the ability of central governments to fund such projects, leading to an increased role for the private sector and local governments \citep{WorldEconomicForum2016}.
+  These actors typically operate on much shorter planning horizons for which the quasi-periodic variations of hydroclimate systems may be of interest.
+  We use a series of structured experiments to explore how estimated and true climate risk vary as a function of the project planning period and the available data record.
+  Strong differences in the expected value of flood risk, and the associated uncertainty, emerge between long planning periods dominated by secular trend and short planning periods dominated by quasi-periodic variability.
+  Successful climate adaptation strategies should consider decisions on multiple timescales.
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
-In 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
-In June 2013, a rainy spring was followed by three summertime cyclones following highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
-At the same time a multi-year drought in S\~ao Paolo threatened to dry out one of the world's largest cities \citep{Seth2015}, and more recently the 2017 Atlantic Hurricane season led to approximately \usd{100 billion} in insured losses.\footnote{\url{http://www.swissre.com/media/news_releases/nr20171220_sigma_estimates.html}}
-Since classical analysis of climate extremes for use in risk analysis and mitigation studies assumes that  the historical or simulated record of events constitutes an independent and identically distributed process, adaptation systems may be overwhelmed by events which violate these assumptions.\footnote{Do we need to add something about climate mechanisms such as wave resonance, blocking, \etc{} that can explain some of this?
-Or just cite \citet{Hannachi2017} to say the atmosphere has regimes?} \footnote{As for $\uparrow$, I would avoid the impulse to get too into the dynamics in this paper (i.e.
-wave resonance, etc).
-However, I do think that this paragraph could be a bit more direct.
-There is ample evidence that climate extremes are not IID $\rightarrow$ for example, Thailand, etc $\rightarrow$ thus risk analyses need to consider this...
-Basically my critique is that three sentences about recent disasters without the context of why we particularly care about these things may be a bit of a misdirection for the reader}.
+The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of weather and climate hazard.
+In this framework, event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
+Under the assumption that the historical or simulated record of events constitutes an stationary, \gls{iid} process, and using an appropriate statistical inference model, an estimate of this relationship and its uncertainty based on a finite $N$-year record is developed.
+This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation applied to the costs and residual losses expected from the mitigation instrument.
+The uncertainty associated with the model may be used to develop a risk premium for insurance, or for assessing the potential over- or under- design of infrastructure.
 
-Indeed, the scientific literature has recognized that even small nonstationarity (temporal change in dynamics of the underlying physical process) can lead to significant trends in the estimated $T$-year return periods or probability of exceedance $p_T$ \citep{Milly2008,Vogel2011,Salas2014,Merz2014,Obeysekera2016}.\footnote{All of these citations are for floods, can we look at other types of extremes?
-Even droughts would be helpful.}
-This has led to a plethora of efforts aimed at identifying the nature of the nonstationarity and modeling it, \eg{} by introducing time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Delgado2014,Griffis2007,Silva2016,Sun2014,Hall2014}, latent variable or state-based modeling approaches \citep{Griffis2007,Sveinsson2005,Waylen1986}, or using only the last $N$ years of the data.\footnote{Rich Vogel citation?}
+Over time it has become clear that even small fluctuations in the dynamics of the hydroclimate system, known as nonstationarities \citep{Milly2008}, can lead to clustering of extreme events  and strong time dependence of a public or private actor's risk profile \citep{Bonnafous2017a,Bonnafous2017b}.
+For example, in 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
+Similarly, in June 2013 three summertime cyclones followed highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
+More recently the 2017 Atlantic hurricane season led to over \usd{130 billion} in insured losses \citep{SwissRe2017}, despite negligible losses in the preceding several years.
+Each of these events was highly improbable under \gls{iid} assumptions, yet the clustering of such events is well-documented in the literature \citep{Hurst1951,Matalas2012,Merz2014} and is consistent with our understanding of the nonlinear dynamics of the climate system \citep{Hannachi2017}.
+
+Nonstationarity behavior that leads to violation of traditional assumptions takes two primary forms.
+The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Merz2014}.
+The uncertainties associated with this are not purely informational, at least in the sense that they can be inferred from a historical sample or model.
+Both nature and human actions over the course of this century will determine the magnitude and direction of secular change, and neither is fully predictable.
+The uncertainty associated with secular change clearly grows as one looks out into the distant future, and the importance of investment decisions that may engage us in considerations of designs that go far out into the future, \eg{} 2100, diminish if any positive discount rate is used for the future.
+While this consideration offsets the impact of the growing uncertainty for longer project life, it also suggests that it may be more effective to understand the consequences for analysis that may focus on shorter lived adaptation strategies which may last for 1, 5, 10, 20,  30 or 50 years.
+
+Focusing on these shorter time horizons brings up the concern associated with the second type of nonstationarity in climate: the quasi-periodic inter-annual, decadal and multidecadal modes of climate variability that influence the spatiotemporal expression of climate extremes in ways that violate the standard \gls{iid} assumption \citep{Matalas2012,Hannachi2017}.
+For example, analysis of a long synthetic data set demonstrated that using the past $N$ years to estimate return periods for the next $N$ years leads to a high degree of surprise even when secular change is exactly zero \citep{Jain2001}.
+The finding that floods at a particular location are clustered in time relative to what could be expected under an \gls{iid} process was replicated for an observed long streamflow record \citep{Jain2001} and is consistent with analysis of a \SI{1500}{year} streamflow reconstructions \citep{Swierczynski2012}.
+Analysis of annual-maximum rainfall events at sites around the world has also demonstrated a high degree of spatial clustering \citep{Bonnafous2017a}.
+
+Broad recognition of nonstationarity has led to a plethora of efforts aimed at identifying the nature of the nonstationary and modeling it, though no single approach has proven broadly successful.
+Such efforts have included time trends on statistical parameters \citep{Strupczewski2001,Vogel2011,Obeysekera2014,Serinaldi2015}, climate covariates \citep{Griffis2007,Sun2014,Hall2014,Delgado2014,Silva2016}, latent variable or state-based modeling approaches \citep{Waylen1986,Sveinsson2005,Griffis2007}, or using only the last $N$ years of the data.\footnote{Rich Vogel citation?}
 Each model will necessarily give different estimates, which may fit the historical record with comparable accuracy but diverge at an increasing rate as the model is forecast further into the future.
-Despite the wide suite of available models, \emph{the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance instruments has been much less studied}.
+Despite the wide suite of available models, the question of how one should use these nonstationarity estimates for the design of risk mitigation infrastructure or the pricing of insurance instruments has been much less studied.
 
-There are two kinds of nonstationarities of concern.
-The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Merz2014}.
-The uncertainties associated with this are not purely informational; both nature and future human actions will determine the outcomes, and neither is obviously predictable.
-The approach taken by many, including the IPCC, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions and then use a sequence of physics-based numerical models with these scenarios to develop probabilistic statements as to the potential changes in the extremes.
-Such a ``model chain'' typically encompasses:
-\begin{enumerate*}[label= (\roman*) ]
-  \item emission scenario;
-  \item general circulation model (GCM);
-  \item downscaling; and
-  \item high-resolution model of the study region.
-\end{enumerate*}
-Because of inherent errors due to, \eg{} the discretization and parameterization of the equations of motion, bias correction methods are typically applied at each step of the chain; while this typically improves the model chain's representation of the historical record, the extrapolation of these bias corrections far into the future does not inspire confidence\footnote{Is "inspire confidence" too informal?...
-I'm not sure}.
-\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
-A critical question in this regard is how to assess whether any structure\footnote{or other risk management instrument such as an insurance product?} designed using these projections may actually perform in the future relative to the proposed level of risk protection.
-Would it be over- or under- designed?
-Would the bias and uncertainty in its estimation lead to sub-optimal investment strategies in risk mitigation?
-These questions go far beyond what can be reliably answered at this time, and this type of epistemic uncertainty clearly grows as we look out into the distant future.\footnote{Perhaps these sentences belong elsewhere?}
+A critical question in this regard is how to assess whether any structure designed using these projections may actually perform in the future relative to the proposed level of risk protection.
+These questions go well beyond what can be reliably answered at this time.
+Clearly, in a nonstationary world, the probability of exceedance could be considered to be temporally variable over the project life.
+This would however, need to be a prediction of that probability ex ante, using appropriate modeling tools applied to the historical data or to climate model based forecasts.
+The risk estimation from the historical record must explicitly inform a risk prediction over the future planning period, and depending on the approach and the form of the underlying nonstationarity, we can expect both bias and uncertainty associated with the prediction, and we need temporally varying probabilities to address the inevitable risk clustering.
+This is then the statement of the risk mitigation design problem that we need to consider for nonstationary climate.
 
-In addition to secular change, quasi-periodic inter-annual to multi-decadal modes of climate variability \citep[see][]{Hannachi2017} to influence the spatio-temporal expression of climate extremes in ways that violate the assumption of independent, identically distributed events \citep[see][]{Merz2014,Serinaldi2015,Hoskins2015}.
-For example, even when no secular trend exists, using the past $N$ years to project the next $N$ leads to different estimates of floods relative to when the time series is fully shuffled to remove temporal clustering, even for very long ($N=\SI{200}{years}$) sequences \citep{Jain2001}.
-
-When secular and quasi-periodic variability are both considered, it becomes clear the probability of exceedance is temporally variable over the life of a climate adaptation project.
 In this paper we provide some initial insight into the bias and uncertainty associated with flood frequency estimation for projects as a function of the project planning period $M$ as well as the length of the historical record $N$.
-This project planning period can be interpreted as the design life of a physical structure, or as the contract length of a financial instrument.
-Our aim is not to identify the ``best'' model, but rather to develop theory as to how the underlying nature of the natural variability and nonstationarity and the choice of model used to model future flood risk map onto estimation bias and variance in the context of many different $M$ and $N$.
-This estimation bias and variance may map directly onto the price of an insurance contract, or the possibility of over/under-design of a large structure like a dam or seawall used to protect against climate hazards.
+The project planning period can be interpreted as the design life of a physical structure, or as the contract length of a financial instrument.
+We focus on structured experiments with synthetic data on extremes to illustrate how one could approach the analyses and what the implications are of varying $N$ and $M$ on the ability to do effective inference and on the ability to judge operational performance.
 Though the specific examples target flood risk, this method and framework may be used for understanding the implications of uncertain projections of the future for climate hazards including drought, hurricanes, and wind.
@@
 Next, $\mu(t)$ is calculated deterministically from the NINO3 time series $x(t)$ by
-\begin{equation}
+\begin{equation}\label{eq:nino3}
   \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0)
@@
 
-\section{Case Study}
+\section{Spatial Risk}
 
-I would go back to a case study and look at a 500-year streamflow sample (could be generated synthetically if we can't get a good real one).
-Fix $N=100$ (say) and choose between $M=5$ or $M=50$. 
-Estimate this model with 
-\begin{enumerate}[label=(\roman*)]
-  \item HMM;
-  \item LN2 Stationary;
-  \item LN2 Trend; and
-  \item Perfect Information (\ie{} knowing the future exactly)
-\end{enumerate}
-Then at each time, calculate the insurance price for each model from \cref{eq:insurance-price} (the perfect information model will have a risk premium of zero) and summarize the over or under-pricing of the risk premium.
-Then explain that one would be unlikely to buy a 50-year insurance contract but that it's helpful to compare the two models using equal criteria.
-We could extend to include NPV analysis if desired.
+Many actors, particularly those in the private sector such as reinsurance vendors, may be exposed to concurrent risk in many different locations.
+It is consequently of interest to examine how their risk portfolio risk may vary as a function of time.
 
-\emph{Please comment: should this be part of our paper?}
+We consider that the annual-maximum streamflow in each location depends only on the NINO 3.4 index as shown in \cref{eq:nino3} ($\gamma=0$ to enforce zero secular change).
+However, we allow the dependance of streamflow at each location on the NINO 3.4 index to be a random variable, such that $\beta \sim \mathcal{N} \qty(0, 1)$.
+This assumption represents observations that during El Ni\~no events drought occurs in some parts of the world and floods in others \citep{Ropelewski1987,Ward2014}.
+
+\Cref{fig:spatial-risk} shows the proportion of locations that experience 100-year floods in each year, empirically defined from a simulated \SI{2000}{year} record at that location.
+The orange line shows that if the \gls{iid} assumption is valid, the proportion of sites experiencing a 100-year flood each year is always between 0.005 and 0.015.
+However, if the data is simulated from the log-normal distribution conditional on the NINO 3.4 series, as few as zero sites or as many as 20\% of all sites may experience floods in a particular year.
+In other words, even though the portfolio has managed its risk by choosing locations with strongly positive, strongly negative, and near-zero dependence on \gls{enso}, the distribution of portfolio risk has a very fat tail, in agreement with previous analyses \citep{Bonnafous2017a}.
+\begin{figure}
+  \includegraphics[width=\textwidth]{case_study.pdf}
+  \caption{
+    Proportion of sites experiencing a 100-year flood return in each year.
+    (a): A 250-year series of data simulated with the NINO 3.4 model (blue) and using an \gls{iid} model (orange). 
+    (b): The probability distribution function of data simulated with the NINO 3.4 model and using an \gls{iid} model.
+    This distribution is estimated using a \SI{2000}{year} record, of which the first \SI{250}{year} is shown in (a).\label{fig:spatial-risk}
+  }
+\end{figure}
 
@@
 Infrastructure in the developed world is aging \citep{Ho2017}, leading to questions of whether and how to replace it.
-In the developed world new infrastructure is being built.\footnote{citations}
+In the developed world new infrastructure is being built.
 Given a positive future discount rate and uncertainty associated with future urbanization, economic growth, and climate change, the expected net benefits of this infrastructure are highly uncertain.
@@
 Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
+
+The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
+Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
+While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+
+The idea that one needs to consider the joint uncertainty in the estimation and occurrence of climate extremes for the $N$ and $M$ periods for stationary conditions was originally considered by \citet{Lall1987}.
+That analysis corresponds to what one would expect under \gls{iid} conditions, and would hence provide a baseline to compare the uncertainty distribution associated with a portfolio whose elements were designed using $N$ years of data with an application to $M$ future years at a common design probability of exceedance for the elements.
 

### 2018-03-29 src/case_study.py

@@ -1,6 +1,8 @@
 import xarray as xr
 import numpy as np
 import matplotlib.pyplot as plt
+import seaborn as sns
+import string
 
 import argparse
 
@@ -29,13 +31,43 @@ def main():
         coords = {'location': np.arange(1, N_sites + 1), 'time': np.arange(1, N_years + 1)},
         dims = ['time', 'location']
     )
-    threshold = sflow.quantile(0.99).values
+    threshold = sflow.quantile(0.99, dim='time')
+    exceedances = (sflow > threshold).mean(dim='location')
 
-    (sflow > threshold).mean(dim='location').plot(figsize=(12, 6))
-    plt.xlabel('Year')
-    plt.ylabel('Proportion of {} Sites Experiencing Flood'.format(N_sites))
-    plt.grid()
-    plt.savefig(args.outfile, bbox_inches='tight')
+    exceedances_iid = np.random.binomial(n=N_sites, p=0.01, size=N_years) / N_sites
+    exceedances_iid = xr.DataArray(exceedances_iid, coords=exceedances.coords, dims=['time'])
+    print(exceedances_iid.max())
+    print(exceedances_iid.min())
+
+    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(14, 5), gridspec_kw={'width_ratios': [3.5, 1]}, sharey=True)
+    
+    ax = axes[0]
+    sub_ex = exceedances.sel(time=slice(0, 250))
+    sub_iid = exceedances_iid.sel(time=slice(0, 250))
+    ax.plot(sub_ex['time'], sub_ex.values, label='Simulated')
+    ax.plot(sub_iid['time'], sub_iid.values, label='IID')
+    ax.set_xlabel('Year')
+    ax.set_ylabel('Proportion of Sites Experiencing 100-Year Flood'.format(N_sites))
+    ax.grid()
+    ax.axhline(0.01, c='gray')
+    ax.legend()
+
+    ax = axes[1]
+    sns.distplot(exceedances, vertical=True, label='Simulated')
+    sns.distplot(exceedances_iid, vertical=True, label='IID')
+    ax.set_xticks([])
+    ax.axhline(0.01, c='gray')
+    ax.grid()
+    ax.legend()
+
+    letters = string.ascii_lowercase
+    for i, ax in enumerate(axes.flat):
+        label = '({})'.format(letters[i])
[…]

### 2018-03-29 writeup/main.tex
working on discussion

@@
 
-\title{Adapting to Quasi-Periodic and Secular Variation of Climate Risk}
+\title{Planning for Quasi-Periodic and Secular Variability of Climate Risk}
 \author[1,2]{James Doss-Gollin}
@@
 
-\section{Spatial Risk}
+\section{Portfolio Risk Across Space}
 
@@
 
+Although the flood sequences shown in \cref{fig:example-long,fig:example-short,fig:stationary-sequences,fig:trend-sequences} exhibit the regime-like, quasi-periodic behavior observed in analysis of observational streamflow records, the assumption that the deterministic component of flood risk varies linearly with time and \gls{enso} is a simplification.
+Reality, of course, is more complicated.
+The probability of flooding in a particular location is modulated by multiple physical mechanisms, and dependence on a particular driver is rarely linear.
+Further, the linear time dependance assumed is merely a placeholder for physical drivers, such as water storage capacity of the catchment's soil or the regional average specific humidity, that experience secular change.
+
+More complex strategies also exist for estimating the risk of flood, or another natural hazard, from an observational record.
+For example, analysis of flood frequency can be improved by combining information across space \citep{Merz2008a,Lima2016}, combining predictions from many models, or by through use of frequency- or wavelet-based methods \citep{Kwon2007}.
+The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
+Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
+While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
+
+From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every $M$ years that incrementally add (or not) protection.
+In this context one could ask how the uncertainty (and hence risk) for these decisions depends on the design level, the project duration $M$, and the uncertainty associated with the estimation of $f(\hat{p}_T)$.
+This study has considered only uncertainty associated with changing flood hazard, yet the large uncertainty associated with flood risk hazard and possible adaptation technologies would necessarily be relevant.
+
+What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
+This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
+Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
+
+The results presented in \cref{sec:results} illustrate sampling uncertainty and inter-model disagreement using several simple models for annual-maximum floods.
+In reality,
+
 Infrastructure in the developed world is aging \citep{Ho2017}, leading to questions of whether and how to replace it.
@@
 
-The assessment of sampling bias and variance associated with estimating the occurrence probability of an extreme event, $\hat{p}_T$, maps directly into this insurance contract, where a typical price for a $M$-year contract is given by
-\begin{equation}\label{eq:insurance-price}
-  \frac{\mathbb{C}}{M} = \lambda \qty(\mathbb{E} \qty[p_T] + c \mathbb{V} \qty[p_T])
-\end{equation}
-where $\mathbb{E}$ denotes the expectation operator, $\mathbb{V}$ the variance operator, $\lambda$ the profit margin, and $c$ the risk premium.
-Bias in $\hat{P}_T$ leads to inefficient pricing, since if the price is too low, the insurer is exposed to failure and if it is too high the buyer will be less inclined to purchase insurance.
-
-We illustrated sampling uncertainty and inter-model disagreement using several simple models for annual-maximum floods.
-In more realistic applications, analysis of flood frequency can be improved by combining information across space \citep{Lima2016,Merz2008a}, combining model forecasts, \etc{}.
-Analysis of other weather and climate hazards may use different distributions, and the time structure of the underlying climate process may be different.
-
-For the type of extreme flood events we are interested in, the nominal return period at some location flooded during the event will typically be larger than 10 years.
-While this is a relatively rare event at a given location, across a geography as large as the United States, one can expect to see such events rather frequently, emphasizing the need to understand and predict the risk of such events over a season or few years to facilitate preparation and mitigation strategies.
-Recent efforts\footnote{Prad's work} to identify regions where risk of a particular event is negatively correlated in space.
-
-Such clustering in the occurrence of extreme rainfall and drought is manifest in the potential exposure of mining companies, who face spatial as well as temporal clustering \citep{Bonnafous2017a,Bonnafous2017b}.
-
 How could  financial risk mitigation instruments be used in conjunction with structural design to identify a robust, adaptive climate risk mitigation strategy with sequential decisions that are informed by updated estimates of risk.
@@
 Since we know the true distribution of the flood sequences (we have generated them ourselves), using the log-normal for both generating and fitting enables interpretation of results as a lower bound on total uncertainty.
-
-From a decision making perspective, one could either design and build a project that gives us protection through the year 2100 considering both types of nonstationarities, or consider a sequence of projects every M years that incrementally add (or not) protection.
-How does the uncertainty and hence risk for these decisions manifest depending on the design level p, the duration M, and the uncertainty associated with the estimation of the \( f(Qp(t)) \) given a methodology used for prediction -- \eg{} the model chain using gcms and such and models fitted to a historical data length $M$.
-What is the implication for choices of p and M under different models for stationarity and uncertainty using some simple loss and cost  functions.
-If a positive discount rate is used for the future, as is typical in public and private sector planning models,\footnote{CITE} the investment returns on major public and private investment may diminish substantially far into the future.
-This consideration suggests that it may be more effective to understand the consequences for analysis focusing on shorter lived mitigation alternatives that may last for 1, 5, 10, 20, 30, or 50 years.
-Thus, sequences of such smaller projects may in some cases better reflect climate adaptation goals, particularly given the possibilities that major climate change impacts in a particular place may occur far into the future; that climate mitigation efforts may be successful; or that the needs of human society may change drastically over these horizons.
-
-The approach being taken by many, including the \gls{ipcc}, is to consider discrete scenarios for what humans may do with respect to greenhouse gas emissions, and then attempt to use climate models to estimate the probability of extreme events conditional on occurrence of a particular scenario.
-Because of inherent errors due to the discretization and parameterization of the equations of motion, as well as uncertainty about the underlying physical parameters, bias correction methods \citep[\eg quantile-quantile mapping;][]{Rajczak2016} are typically applied to the final outputs.
-While this typically improves the model chain's representation of the historical record, this approach assumes a stationary relationship between modeled and observed rainfall at a point-by-point scale, an assumption that is unlikely to hold under the kinds of spatial and dynamical shifts anticipated under increased greenhouse gas scenarios such as expansion of the tropics and altered equator-pole temperature contrasts.\citep{Dankers2009,Ott2013,Merz2014,Dittes2017}.\footnote{Farnham et al 2018!}
 

### 2018-03-29 writeup/misc.tex
reorganize writeup
working on results and intro
working on discussion

--- 2018-02-11 writeup/Outline/misc.tex
@@
 
+Some stuff on decisions and insurance which may be relevant for the introduction?
 \begin{enumerate}
@@
 \end{enumerate}
+
+The assessment of sampling bias and variance associated with estimating the occurrence probability of an extreme event, $\hat{p}_T$, maps directly into this insurance contract, where a typical price for a $M$-year contract is given by
+\begin{equation}\label{eq:insurance-price}
+  \frac{\mathbb{C}}{M} = \lambda \qty(\mathbb{E} \qty[p_T] + c \mathbb{V} \qty[p_T])
+\end{equation}
+where $\mathbb{E}$ denotes the expectation operator, $\mathbb{V}$ the variance operator, $\lambda$ the profit margin, and $c$ the risk premium.
+Bias in $\hat{P}_T$ leads to inefficient pricing, since if the price is too low, the insurer is exposed to failure and if it is too high the buyer will be less inclined to purchase insurance.
 

### 2018-03-29 src/case_study.py

@@ -36,10 +36,10 @@ def main():
 
     exceedances_iid = np.random.binomial(n=N_sites, p=0.01, size=N_years) / N_sites
     exceedances_iid = xr.DataArray(exceedances_iid, coords=exceedances.coords, dims=['time'])
-    print(exceedances_iid.max())
-    print(exceedances_iid.min())
+    print(exceedances_iid.max().values)
+    print(exceedances_iid.min().values)
 
-    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(14, 5), gridspec_kw={'width_ratios': [3.5, 1]}, sharey=True)
+    fig, axes = plt.subplots(nrows=1, ncols=2, figsize=(12, 4), gridspec_kw={'width_ratios': [4.5, 1]}, sharey=True)
     
     ax = axes[0]
     sub_ex = exceedances.sel(time=slice(0, 250))
@@ -50,7 +50,6 @@ def main():
     ax.set_ylabel('Proportion of Sites Experiencing 100-Year Flood'.format(N_sites))
     ax.grid()
     ax.axhline(0.01, c='gray')
-    ax.legend()
 
     ax = axes[1]
     sns.distplot(exceedances, vertical=True, label='Simulated')

### 2018-04-04 writeup/main.tex
Continue writing

@@
 
-The \nth{20} century saw the evolution of a nearly universal approach to the statistical analysis of weather and climate hazard.
-In this framework, event risk is assessed as a relationship between the event magnitude and its annual probability of exceedance.
-Under the assumption that the historical or simulated record of events constitutes an stationary, \gls{iid} process, and using an appropriate statistical inference model, an estimate of this relationship and its uncertainty based on a finite $N$-year record is developed.
+Hydroclimate systems exhibit organized low-frequency and regime-like structure which cause the reliability of physical structures and financial instruments used for risk management to vary across timescales \citep{Jain2001,Merz2014,Hannachi2017}.
+This phenomenon, termed nonstationarity \citep{Milly2008} violates the core assumptions made in classical statistical analysis of weather and climate hazard, in which an $N$-year historical record is assumed to represent \gls{iid} draws from a stationary distribution in order to estimate a relationship between event magnitude and its probability of occurrence.
 This is then used as a measure for decision analysis for infrastructure, flood zoning and insurance instruments for event risk mitigation applied to the costs and residual losses expected from the mitigation instrument.
@@
 
-Over time it has become clear that even small fluctuations in the dynamics of the hydroclimate system, known as nonstationarities \citep{Milly2008}, can lead to clustering of extreme events  and strong time dependence of a public or private actor's risk profile \citep{Bonnafous2017a,Bonnafous2017b}.
-For example, in 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
-Similarly, in June 2013 three summertime cyclones followed highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
-More recently the 2017 Atlantic hurricane season led to over \usd{130 billion} in insured losses \citep{SwissRe2017}, despite negligible losses in the preceding several years.
-Each of these events was highly improbable under \gls{iid} assumptions, yet the clustering of such events is well-documented in the literature \citep{Hurst1951,Matalas2012,Merz2014} and is consistent with our understanding of the nonlinear dynamics of the climate system \citep{Hannachi2017}.
-
-Nonstationarity behavior that leads to violation of traditional assumptions takes two primary forms.
+The nonstationarity behavior which violates these assumptions takes two primary forms.
 The first, secular change, may be related to anthropogenic climate change, river modification, or land use change \citep{Merz2014}.
@@
 This is then the statement of the risk mitigation design problem that we need to consider for nonstationary climate.
+
+Given that these risk mitigation instruments are typically evaluated over a $M$-year period and based on an $N$-year historical record, the potential of a regime shift can lead to a dramatic over- or under-design of the designed instrument.
+If a physical regime associated with hydroclimate variability has duration $D$ which is of order $N$, then it is unlikely that 
+Now signal that it is important to then understand how the linkage of regimes of duration D influences risk, and what implications it has for detection given N and mis-classification given M
+
+% Perceptions about 100 year flood?
+For example, in 2011 four typhoons struck northern Thailand \citep{Gale2013}, causing severe damage to public and private infrastructure and crippling global supply chains \citep{Haraguchi2015}.
+Similarly, in June 2013 three summertime cyclones followed highly similar tracks, leading to widespread flooding in central Europe \citep{Grams2014}.
+More recently the 2017 Atlantic hurricane season led to over \usd{130 billion} in insured losses \citep{SwissRe2017}, despite negligible losses in the preceding several years.
+Each of these events was highly improbable under \gls{iid} assumptions, yet the clustering of such events is well-documented in the literature \citep{Hurst1951,Matalas2012,Merz2014} and is consistent with our understanding of the nonlinear dynamics of the climate system \citep{Hannachi2017}.
+This behavior is generically known as nonstationarity \citep{Milly2008}, and has been shown to lead to clustering of extreme events and strong time dependence of a public or private actor's risk profile \citep{Bonnafous2017a,Bonnafous2017b}.
 
@@
 
-\subsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
+\section{Supplemental Results}
+
+\subsubsection{Are Generated Sequences Realistic?\label{sec:sequence-realistic}}
 

### 2018-04-04 src/plot_enso.py

@@ -35,8 +35,8 @@ def main():
     fs = 1
     f, Pxx_den = signal.periodogram(x, fs)
     periodogram = pd.Series(Pxx_den, index=f)
-    smoothed = periodogram.rolling(100, win_type='hamming').mean()
-
+    smoothed = periodogram.rolling(100, win_type='hamming').mean().dropna()
+    
     smoothed.plot(ax=axes[1], label='smoothed')
     axes[1].set_xlabel('frequency [Year$^{-1}$]')
     axes[1].set_ylabel('Power Spectral Density')

### 2018-09-04 report-earths-future/figs/flowchart.tex
Upload the paper draft as submitted to earth's future

\documentclass{standalone}

\usepackage{tikz}
\usetikzlibrary{
  shapes,
  arrows,
  shapes.geometric,
  positioning,
}

% Define block styles
\tikzstyle{parameter} = [
rectangle,
draw,
fill=red!20,
text width=2.75cm,
text badly centered,
rounded corners
]

\tikzstyle{calculated} = [
rectangle,
draw,
text width=2.75cm,
text badly centered,
rounded corners
]

\tikzstyle{result} = [
rectangle,
draw,
fill=blue!20,
text width=2.75cm,
text badly centered,
rounded corners
]

\tikzstyle{arrow} = [draw, -latex']
\tikzstyle{noarrow} = [draw, dashed]

% set consistent font size
\tikzset{every picture/.style={font issue=\footnotesize},
         font issue/.style={execute at begin picture={#1\selectfont}}
         }

\begin{document}

\begin{tikzpicture}[node distance = 0.7cm, auto, font=\sffamily]
  % Place nodes
  \node [parameter] (genmod) {Generating Model};
  \node [parameter, above=of genmod] (N) {$N$};
  \node [parameter, below=of genmod] (M) {$M$};

  \node [calculated, right=of genmod] (seq) {Synthetic Flood Sequence};
  \node [calculated, right=of N] (historical) {First $N$ Years};
  \node [calculated, right=of M] (future) {Next $M$ Years};

  \node [result, right=of historical] (pThat) {Estimated Exceedances per Year};
  \node [result, right= of future] (pT) {Observed Exceedances per Year};

  \node [parameter, right=of pThat] (fitting) {Estimating Model};
  \node [parameter, right=of seq] (threshold) {Flood Design Level};

  % Draw edges
  \path [arrow] (N) -- (seq);
  \path [arrow] (M) -- (seq);
  \path [arrow] (genmod) -- (seq);

  \path [arrow] (seq) -- (historical);
  \path [arrow] (seq) -- (future);

  \path [arrow] (future) -- (pT);
  \path [arrow] (historical) -- (pThat);
  \path [arrow] (fitting) -- (pThat);

  \path [arrow] (threshold) -- (pThat);
  \path [arrow] (threshold) -- (pT);

\end{tikzpicture}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-09-04 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Upload the paper draft as submitted to earth's future

\documentclass[
  draft,
  linenumbers
]{agujournal2018}
\journalname{Earth's Future}

% Macros and newcommands
\usepackage{xspace}
\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\ie}{i.e.\@\xspace}
\newcommand{\normal}{\mathcal{N}}
\makeatletter
\newcommand*{\etc}{%
    \@ifnextchar{.}%
        {etc}%
        {etc.\@\xspace}%
}
\makeatother

% Package calls with options
\usepackage[letterpaper, margin=1in]{geometry} % page geometry
\usepackage[american]{babel} % language setting
\usepackage[inline]{enumitem} % in-line lists

% Other package calls
\usepackage{
  amssymb, amsmath, % for math symbols
  booktabs, % for nice tables
  graphicx, % for figures
  physics, % for physics notation
  siunitx, % for SI notation
  apacite, % AGU style citations apparently
}

% Fixed-width columns
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% Some package settings
\sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
\graphicspath{{./figs/}{../fig/}} % folders of figures
\allowdisplaybreaks{} % let the align environment span multiple pages

% Use the glossaries package
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{acc}{ACC}{anthropogenic climate change}
\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
\newacronym{cba}{CBA}{cost-benefit analysis}
\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
\newacronym{ffa}{FFA}{flood frequency analysis}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{hmm}{HMM}{hidden Markov model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
\newacronym{lbda}{LBDA}{living blended drought analysis}
\newacronym{lfv}{LFV}{low-frequency climate variability}
\newacronym{nao}{NAO}{North Atlantic Oscillation}
\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
\newacronym{s2d}{S2D}{seasonal to decadal}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}

% These package calls need to come last
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

% -----------------------------------------------------------------------------
% ABSTRACT
% -----------------------------------------------------------------------------

\begin{document}

% Title and Authors
\title{Robust Adaptation to Multi-Scale Climate Variability}
\authors{James Doss-Gollin\affil{1,2}, David J. Farnham\affil{1,2}, Scott Steinschneider\affil{3}, Upmanu Lall\affil{1,2}}
\correspondingauthor{James Doss-Gollin}{james.doss-gollin@columbia.edu}
\affiliation{1}{Department of Earth and Environmental Engineering, Columbia University}
\affiliation{2}{Columbia Water Center, Columbia University}
\affiliation{3}{Department of Biological and Environmental Engineering, Cornell University}

% Key Points
\begin{keypoints}
  \item Climate risk varies in time over a project's finite planning period
  \item The dominant driver of climate risk depends on the length of the planning period 
  \item Potentially predictable cyclical climate variability dominates near-term climate risk
  \item Successful climate adaptation requires prediction of short- and long-term climate variability over the planning period
\end{keypoints}

% Abstract
\begin{abstract}
  The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records highlight the role of \acrlong{lfv} in modulating climate extremes at interannual to multidecadal time scales, \acrlong{acc} is expected to alter their occurrence at longer time scales.
  In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
  These experiments consider both quasi-periodic and secular change for the underlying risk, as well as statistical models for estimating this risk from an $N$-year historical record.
  The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$ and the future discount rate for investment.
  Shorter design lives are preferred where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation. 
\end{abstract}

% -----------------------------------------------------------------------------
% INTRO
% -----------------------------------------------------------------------------

\section{Introduction}\label{sec:introduction}

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
This climate risk, which is  defined as the product of exposure (the damage that a particular event may cause) and hazard (the probability that this event occurs), is rapidly increasing, at least at the global scale: population growth and urbanization drive increases in exposure \citep{Jongman:2012cr} while \gls{acc} affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Merz:2014gf, Shaw:2016bo}.

Public or private sector investments in climate adaptation require not only the design of each potential structural (\eg, a levee) or non-structural (\eg, a catastrophe bond) risk mitigation instrument, but also selecting between instruments with vastly different operational planning periods.
This project planning period describes the nominal economic or physical lifespan of the structure or contract and may vary from one year for  a financial contract to 100 years or longer for a permanent structure, as illustrated in \cref{tab:real-world-M}.
Complex, real-world climate adaptation strategies often consist of many risk mitigation instruments \citep[\eg,][]{CityofNewYork:2013uh}, each of which has its own planning period or design life $M$.

\noindent\begin{table}
  {\footnotesize
    \begin{tabular}{L{1in}L{2.5in}L{0.3in}L{1.7in}}
      \toprule
      Location & Description & $M$ & Reference \\
      \midrule
      Iowa River & Purchase options for inundation of downstream agricultural lands to allow higher release flows from the flood control reservoir & 1 & \citet{Spence:2016ca} \\
      New York City & Catastrophe bond for protection against storm surge caused by named storms and earthquakes & 3 &  \\
      County of Santa Barbara, California & Emergency improvements to portions of the Santa Maria Levee to reduce risk of levee failure & 5 & \citet{USACE:2007ta} \\
      Iowa River & Raise levees by 6 feet & 30 & \citet{Spence:2016ca} \\
      Dallas, TX & Evacuation of Rockefeller Boulevard & 50 & \citet{USACE:2014vn} \\
      Central California & Tulare Lake storage and floodwater protection project & 100 & \citet{GEI:gIaEZ-gS} \\
      \bottomrule
    \end{tabular}
    \caption{
      Several examples of real-world risk mitigation instruments and the associated project planning period ($M$).
    }\label{tab:real-world-M}
  }
\end{table}

Assessing the utility of a particular risk mitigation instrument involves projecting the climate hazard against which the instrument protects over the $M$-year planning period.
However, an often neglected point is that the sources of predictability differ between projects with long and short planning periods.
As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a 100 year planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
By contrast, a three-year insurance contract may be well informed by a skillful forecast of \gls{lfv}.
Although \cref{fig:conceptual-sketch} shows a highly stylized depiction of \gls{lfv}, analyses of historical and paleo records, like those shown in \cref{fig:observed-lfv}, highlight that subseasonal to multidecadal modes of climate variability such as \gls{enso}, the \gls{pdo}, and the \gls{amo} strongly modulate the occurrence of climate extremes around the world \citep{Ropelewski:1987do, Cook:2010bz, Swierczynski:2012km, Hodgkins:2017hw, Woollings:2018ea}.

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{conceptual-sketch.pdf}
  \caption{
    A stylized illustration of (a) irreducible and (b) estimation uncertainty.
    (a): Irreducible uncertainty cannot be resolved with better models or data and is dominated in the short term by chaotic behavior of the climate, and in the long term by the uncertainty in future \acrlong{acc}.
    (b): The length of a historical record limits the potential to identify different climate signals.
    The blue line shows an idealized climate signal and the black line shows observations, which are scattered stochastically around the signal line.
    The green shading shows the true range within which observations will occur 95\% of the time, while the gray shading the 95\% confidence interval as estimated with a linear trend model.
    }\label{fig:conceptual-sketch}
\end{figure}

Since different physical processes control climate risk at different timescales, successful integration of probabilistic climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.
As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk, which may in turn lead to high premiums of an insurance policy \citep{Kunreuther:1996kp}.
The consequences of this uncertainty also interact with those of any estimation bias as illustrated in \cref{fig:conceptual-bias-variance}.
An instrument designed based on overestimated variance and positive bias will be over-designed, either causing the risk manager to avoid the investmenet, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
  \caption{
    Consequences of model bias or incorrect model representation of uncertainty.
    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive for the user.
    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
  }\label{fig:conceptual-bias-variance}
\end{figure}

In this paper we present a set of stylized experiments to assess how well one can identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy based on these projections.
We consider different temporal structures for the underlying risk which encompass quasi-periodic, regime-like, and secular change, as well as simple statistical models for estimating this risk from an $N$-year historical record.
The relative importance of estimating the short- or long-term risk associated with these extremes depends on the design life $M$, but the potential to understand and predict these different types of variability depends on the informational uncertainty in the $N$-year historical record.
Though we use floods as an example, the framework also applies to other forms of climate extremes.

% -----------------------------------------------------------------------------
% METHODS
% -----------------------------------------------------------------------------

\section{Methods}\label{sec:methods}

We use a statistical framework to
\begin{enumerate*}[label= (\roman*)]
  \item generate synthetic sequences of annual-maximum floods;
  \item fit these sequences to probabilistic models that parameterize time in different ways; and
  \item evaluate the performance of these models as a function of $M$ and $N$ (\cref{fig:methods-summary})
\end{enumerate*}.

As shown in \cref{fig:methods-summary}, we first specify $N$, $M$, the flood threshold (\eg, the design level of a structure or the level at which an insurance contract is triggered), the method for generating synthetic streamflow sequences, and the method for estimating flood risk given an $N$-year historical record.
The methods for generating synthetic streamflow sequences are discussed in \cref{sec:methods-generating} and for estimating flood risk in \cref{sec:methods-estimating}.
For each scenario of climate risk structure, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
We repeat the simulations $J = 2000$ times for each combination of experiment parameters to estimate the bias and variance of the statistical estimate (\cref{sec:methods-evaluating}).
The scenarios considered are summarized in \cref{sec:methods-experiments}.
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{flowchart.pdf}
  \caption{
    Framework for assessing flood risk from synthetic streamflow sequences.
    Red boxes indicate parameters that are chosen for each experiment, white boxes calculated quantities, and blue boxes the final values used to evalute the model fit.
  }\label{fig:methods-summary}
\end{figure}

\subsection{Sampling Climate Risk}\label{sec:methods-generating}

We consider two approaches for modeling the influence of quasi-periodic variability and secular climate change on flood risk.
For each, we take the annual-maximum sequences to follow a log-normal distribution, conditional on a location parameter which may vary in time:
\begin{equation} \label{eq:lognormal}
  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow, $\sigma(t) = \xi \mu(t)$, and apply a lower threshold on the standard deviation $\sigma(t) \geq \sigma_\text{min} > 0$.
The procedure to generate a single sequence of streamflow is thus:
\begin{enumerate}
  \item choose $t_0$, $M$, and $N$ so that the time is $\vb{t}=t_0-N+1, \ldots, t_0+M$
  \item use one of the models described below to simulate $\mu(t)$
  \item calculate $\sigma(t) = \xi \mu(t)$
  \item for each time $t \in \vb{t}$, draw $Q(t)$ stochastically following \cref{eq:lognormal}.
\end{enumerate}
We consider two approaches for simulating $\mu(t)$.

The first approach considers $\mu(t)$ as a function of a variable which explicitly represents quasi-periodic climate variability.
We use an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} plus a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We also allow for the inclusion of secular trends.
We represent \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
To generate $\mu(t)$, a sequence of length $M+N$ is randomly chosen from the \SI{20000}{year} NINO3 sequence.
Next, $\mu(t)$ is calculated from the NINO3 time series, denoted as $x(t)$, as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $\beta$ controls the sensitivity to the NINO3 variation and $\gamma$ controls the magnitude of the secular trend.

The second model to create sequences of annual-maxima is a two-state Markov chain model to explicitly enforce persistence and regime structure in $\mu(t)$.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.

We use the NINO3 and Markov Chain models described above to create three scenarios:
\begin{enumerate*}[label= (\roman*)]
  \item \gls{lfv} only;
  \item secular change only; and
  \item \gls{lfv} plus secular change.
\end{enumerate*}
\gls{lfv} is added to the NINO3 model through $\beta$, which increases the dependence on a climate signal which itself exhibits \gls{lfv}, and to the Markov chain model by making the two states more distinct (increasing $\mu_1 - \mu_2$, or $\gamma_1 - \gamma_2$).
Secular change is applied by increasing $\gamma$, which gives the time dependence.
The parameter values used to create these three scenarios are listed in \cref{sec:supp-methods}.

\subsection{Projecting Climate Risk over the Future $M$ years}\label{sec:methods-estimating}

We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid, and note that the key conclusions hold for more sophisticated \gls{ffa} methods using statistical or dynamical models.

For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations, denoted $Q_{\text{hist}}$.
Once a statistical model is fit to $Q_{\text{hist}}$, $K=\num{2000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.

We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors for regularization.
The full model is given in \cref{eq:ln2-stationary}.

Next, we expand this stationary model to incorporate secular change.
Many studies have done this by regressing certain parameters of the model on time \citep{Merz:2014gf, Salas:2018ge}.
We consider an extension of the stationary LN2 model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:ln2-trend}.

Finally, we consider a model which explicitly represents low-frequency variability.
A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie{} hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
We fit  streamflow sequences $Q_\text{hist}$ simulated using the scenarios in Section 2.1, using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

\subsection{Evaluating Fitting Models}\label{sec:methods-evaluating}

For a given choice of $M$, $N$, generating model, and fitting model, we generate a streamflow sequence of length $M+N$ years.
The fitting model is applied to the synthetic streamflow sequence's $N$-year ``historical record'', and $K=2000$ posterior simulations of future flows are then drawn from the fitted model.

The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
We then compute the variance of these $K$ estimates of $\hat{p}_T$.
We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.

Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we repeat this analysis for $J=1000$ different streamflow sequences.
These are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may be different between the $J$ sequences.
In \cref{sec:results} we report the expected bias and variance by averaging over the $J$ sequences.

\subsection{Computation}\label{sec:methods-computation}

Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, Schreiber:2017tg, Jones:2001uv, Hoyer:2017hs}.
Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke}.
The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk}\footnote{Dear reviewers: this GitHub repository is currently not publicly available, but codes are provided in the online supplemental information. The repository will be made publicly available if and when this work is published.} or in the online supplemental material of this paper.

% -----------------------------------------------------------------------------
% RESULTS
% -----------------------------------------------------------------------------

\section{Results}\label{sec:results}

We assess estimation bias and variance for three cases.
First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
These three cases are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
Projections with each of the three fitting models are also shown.
In the remainder of this section we present a more systmematic analysis of each of these three cases.
\begin{figure}
  \includegraphics[width=\textwidth]{Example-NINO3-M100-N50.pdf}
  \caption{
    An illustration of the estimation procedure.
    A single streamflow sequence with $N=50$ and $M=100$ is shown for each of the three cases (secular only, \gls{lfv} only, and secular plus \gls{lfv}) considered.
    The blue line shows the observed sequence.
    The gray shading indicates the 50\% and 95\% confidence intervals using each of the three fitting methods discussed.
    The horizontal black line indicates the flood threshold.
  }\label{fig:example-fit}
\end{figure}

\subsection{Secular Change Only}

In the idealized case where only secular change exists, accurate climate predictions need to either use a long record to identify and model this trend, or to ignore the trend and predict only a few years ahead.
This is shown in \cref{fig:secular-nino3-bias-variance}, which depicts the estimation bias and variance for each of the three estimation models for many combinations of $M$ and $N$.

The LN2 linear trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
The variance of these estimates is also large.
This again highlights the difficulty of identifying the magnitude of secular change, which strongly controls climate risk far into the future, from a single noisy time series.
By contrast, the stationary LN2 model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[\ie,][]{Muller:2014fc}.
However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace any bias reductions.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{secular-only-nino3-bias-variance.pdf}
  \caption{
    Expected estimation bias and variance for sequences generated from the NINO3 model with secular change only (no \gls{lfv}).
    Sequences were fit to each of three statistical models (colulmns) for different $N$ and $M$ ($x$ and $y$ axis, repsectively).
    Top row shows estimation bias and bottom row shows log standard deviation of estimates.
    Note the uneven spacing of the $x$ and $y$ axes.
  }\label{fig:secular-nino3-bias-variance}
\end{figure}

If the analyst could know \emph{a priori} that secular change is present in a time series, and if $M$ is long, then the use of a model that captures this secular trend is essential.
Here the LN2 linear trend model has the advantage of being correctly specified (both the generating and fitting processes assume a log-normal distribution conditional on a linear time trend), which is generally not the case in the real world \citep{Serinaldi:2015bq, Montanari:2014hl}.
As a result, in real-world settings longer $N$ may be required to identify trends whose exact form is not known.
Here, $N$ implicitly refers to the uncertainty in trend estimation and may represent an ``equivalent sample size'' if \gls{ipcc}-like projections are used.
Alternatively, if $M$ is small then it may be reasonable to use a stationary estimate, since the bias will be small and the variance substantially lower.

\subsection{Low-Frequency Variability Only}

Next, in the idealized case where  \gls{lfv} is present but there is no secular change, the identification of nonexistent trends may lead to gross over-estimation of true risk through an increase in the variance of the estimated risk.

Results for this scenario are shown in \cref{fig:lfv-nino3-bias-variance} for the NINO3 model and \cref{fig:lfv-markov-bias-variance} for the two-state Markov model.
As expected, the stationary model (right column) performs well overall, with low bias and low variance.
The \gls{hmm} (left column) actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the true multimodal distribution that emerges from dependence on the NINO3 index which exhibits several regimes.
By contrast, the linear trend model (middle column) performs poorly for low $N$ and high $M$ because a positive probability is assigned to the existence of a positive trend.

Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability of the \gls{lfv}, denoted $T_{\text{LFV}}$.
As shown in \cref{fig:enso-ts}, the NINO3 time series is most active in the \SIrange{3}{6}{year} band.
If $T_{\text{LFV}} \ll M$, then a stationary assumption may provide reasonable estimates, and fewer observations may be required (shorter $N$).
In the real world, however, many hydroclimate time series vary at multidecadal and longer frequencies.
In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic frequencies $T_{\text{LFV}}$ may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
This in turn requires a longer observational record $N$ in order to identify and predict these different signals.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-only-nino3-bias-variance.pdf}
  \caption{
    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and \gls{lfv} from the NINO3 model.
  }\label{fig:lfv-nino3-bias-variance}
\end{figure}

\subsection{Low-Frequency Variability and Secular Change}

In the final and most realistic case, where both \gls{lfv} and secular change are present, stationary models perform well for short $M$ while for long $M$ the trend must be identified from a long record and modeled explicitly.

Simulation results are presented for the NINO3 model in \cref{fig:lfv-secular-nino3-bias-variance} and for the two-state Markov chain model in \cref{fig:lfv-secular-markov-bias-variance}.
Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the relative importance of secular change and \gls{lfv} depends on $M$.
When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly (\ie, through the linear trend model).
Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-secular-nino3-bias-variance.pdf}
  \caption{
    As \cref{fig:secular-nino3-bias-variance} but for sequences generated from the NINO3 model with both \gls{lfv} and secular change.
  }\label{fig:lfv-secular-nino3-bias-variance}
\end{figure}

% -----------------------------------------------------------------------------
% DISCUSSION
% -----------------------------------------------------------------------------

\clearpage
\section{Discussion}\label{sec:discussion-conclusions}

Evaluating and implementing investments for climate risk mitigation demands making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
The success of this prediction will depend on the identifiability of different signals from an $N$-year sample, the time scales of \gls{lfv} relative to the project life of the instrument, and the uncertainty associated with the estimation.
In this paper we took a synthetic data approach to explore the implications of varying $M$ and $N$ in stylized scenarios which represent important features of real-world hydroclimate systems.

The stationary, linear trend, and \gls{hmm} models considered for estimation present an idealized alternative to more complex, physically based models for future risk projection.
They were chosen to represent the ability of simple models to capture two features of interest: \gls{lfv} and secular trends.
The models were assessed on their performance over a complete future period of length $M$ years; as discussed in \cref{sec:introduction} the bias and variance computed together represent the likelihood for over- or under-design of the instrument in a given climate scenario (see \cref{fig:conceptual-bias-variance}).
If more complex methods, such as analysis of \gls{ipcc} scenarios, are used, the sample size $N$ used in our experiments may be straightforwardly interpreted as a surrogate for the total informational uncertainty in the analysis.

The climate adaptation decisions which this analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate.
The application of a positive discount rate, as is typical in the public and private sectors, emphasizes the importance of predicting near-term risk.
Even ignoring this, however, if the uncertainty associated with the performance of the instrument increases into the future (as shown in \cref{fig:conceptual-sketch}) then adaptation strategies with a short $M$ are preferred.
This is consistent with recent emphasis on the importance of making decisions that are robust to a wide range of possible future climates \citep[\ie,][]{Haasnoot:2013im, Poff:2015jn}.

For these short-$M$ projects, cyclical climate variability is dominant over the project planning period \citep{Hodgkins:2017hw, Jain:2001hz}.
Analysis of the former using the wavelet transformation \citep{Torrence:1998jp} shows dominant spectral peaks at 8, 20, and 64 years, and of the latter at 2 and 15 years.
This is illustrated in \cref{sec:observed-lfv}, which shows the cyclical variability of two real-world hydroclimate systems: a 500 year drought reconstruction from the \gls{lbda} \citep{Cook:2010bz} and a \SI{107}{year} record of naturalized daily river stage on the American River at the Folsom Dam.
The long periods and high amplitudes of the oscillations present in these time series highlight the importance of adapting not only to secular change but also, simultaneously, to \gls{lfv}.
Interestingly, our results show that even though the two-state \gls{hmm} is an imperfect analog for the \gls{lfv} represented in the \gls{enso} process, it performs quite well for short $M$.
This highlights that even imperfect estimates of \gls{lfv} can support suffessful climate adaptation.

For projects with longer $M$, it becomes necessary to consider the role of secular change.
As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.
We find that when making risk projections that consider secular change, mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases.
This is particularly important given that in real-world settings the true form of the trend is not known and thus ``nonstationary'' projections which assume a particular type of trend are likely to perform poorly \citep{Montanari:2014hl, Serinaldi:2015bq}.

Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, as shown in \cref{fig:observed-lfv}.
Identifying and modeling these different physical signals requires a large amount of information, represented here by the length of the $N$-year historical record.
In places where long records are not available, however, information from nearby locations, reanalysis products, or \acrshort{gcm} runs may increase the amount of available information \citep{Merz:2008eh}, effectively increasing $N$.

Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future.
For most large actors, a portfolio of both large $M$ and small $M$ projects will likely be necessary, none of which precludes the need for mitigation of global and local climate change and the development and execution of vulnerability reduction strategies.

% -----------------------------------------------------------------------------
% END MATTER
% -----------------------------------------------------------------------------

\acknowledgments

The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{100000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh:2016hf}.
The authors thank John High of the United States Army Corps of Engineers for providing the naturalized daily streamflows at the Folsom Dam.
JDG thanks the NSF GRFP program for support (grant DGE 16-44869: ``Understanding \& Predicting Climate Drivers of Extreme, Mid-latitude River Floods'').
DJF thanks SERDP program for support (grant 2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate'').

\bibliography{library}

% -----------------------------------------------------------------------------
% SUPPLEMENTAL INFORMATION -- CURRENTLY INCLUDED IN THIS FILE FOR CONVENIENCE
% -----------------------------------------------------------------------------

\clearpage
\appendix

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\section{Supplemental Methods}\label{sec:supp-methods}

In this section we provide further equations and parameter values used to generate and fit synthetic streamflow sequences.

\subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}

As described in \cref{sec:methods-generating}, we use a synthetic NINO3 time series, as described in \citep{Ramesh:2016hf}, to represent low-frequency climate variability.
\Cref{fig:enso-ts} shows the wavelet spectrum of this time serties, calculated using WaveletComp software \citep{Roesch:wlBQQoIs}.
The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.
There are, however, modes of \gls{lfv} between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
\begin{figure}
  \includegraphics[width=\textwidth]{enso_wavelet.png}
  \caption{
    Wavelet analysis of the synthetic annual NINO3 time series described in \cref{sec:methods-generating}.
    (L): wavelet power spectrum.
    (R): global (average) power spectrum.
    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
  }\label{fig:enso-ts}
\end{figure}

\subsection{Projecting Climate Risk over the Future $M$ Years}

As described in \cref{sec:methods}, we implement the stationary and trend lognormal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
The full model, including priors, for the stationary lognormal model is given by \cref{eq:ln2-stationary}:
\begin{align}\label{eq:ln2-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
    \mu &\sim \normal \qty(0, 10) \\
    \sigma &\sim \normal^+ \qty(0, 2)
  \end{split}
\end{align}
where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
The full model, including priors, for the trend lognormal model is given by \cref{eq:ln2-trend}:
\begin{align}\label{eq:ln2-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
  \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
  \mu_0 & \sim \normal \qty(0, 10) \\
  \beta_\mu & \sim \normal \qty(0, 0.1) \\
  \log \xi & \sim \normal \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

\subsection{Experiment Design}\label{sec:methods-experiments}

In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
To explore the temporal structure of risk, we generate $J=2000$ sequences of streamflow, using both the NINO3 and Markov models, for each of several combinations of $M$ and $N$.
Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.
Both the NINO3 model and the Markov model are used; the parameters are shown in \cref{tab:nino-stationary} and \cref{tab:markov-stationary}, respectively.

\begin{table}[ht]
  \centering
  \begin{tabular}{llll}
    \toprule
    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
    \midrule
    $\mu_0$             & 6     & 6.5   & 6 \\
    $\gamma$            & 0     & 0.015 & 0.015\\
    $\beta$             & 0.5   & 0     & 0.5\\
    $\xi$               & 0.1   & 0.1   & 0.1\\
    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
    Threshold           & 3000  & 3000  & 3000 \\
    \bottomrule
  \end{tabular}
  \caption{
    Parameters of the NINO3 model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
  }\label{tab:nino-stationary}
\end{table}
\begin{table}[ht]
  \centering
  \begin{tabular}{llll}
    \toprule
    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
    \midrule
    $\mu_1$             & 6.75  & 6.5   & 6.75\\
    $\mu_2$             & 6     & 6.5   & 6\\
    $\gamma_1$          & 0     & 0.015 & 0.015\\
    $\gamma_2$          & 0     & 0.015 & 0\\
    $\xi$               & 0.1   & 0.1   & 0.1\\
    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
    Threshold           & 3000  & 3000  & 3000 \\
    \bottomrule
  \end{tabular}
  \caption{
    Parameters of the two-state Markov chain model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
  }\label{tab:markov-stationary}
\end{table}

\section{Supplemental Results}

\subsection{Observed Low-Frequency Hydroclimate Variability}\label{sec:observed-lfv}

\Acrfull{lfv} of the climate is present in a wide range of real-world time series \citep{Hodgkins:2017hw, Swierczynski:2012km, Cook:2010bz}.
This variability is illustrated in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet \citep{Torrence:1998jp} average spectrum for both.
Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
This clearly illustrates the strong observed multidecadal variability.
For further discussion of the \gls{lbda} time series we refer the reader to \citet{Cook:2010bz}.
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observed-lfv.pdf}
  \caption{
    Hydroclimate time series vary on many time scales.
    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
    High numbers indicate more severe drought.
    A 20-year running mean is also shown in blue.
    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
    (d) wavelet spectrum, like (c), for the American River data.
  }\label{fig:observed-lfv}
\end{figure}

\subsection{Results for the Markov Model}

In \cref{sec:results} all results given use the NINO3 model to generate synthetic flood sequences.
In this section equivalent results are given for sequences generated with the two-state Markov chain model described in \cref{sec:methods}.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
  \caption{
    \Gls{lfv} only: as \cref{fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
  }\label{fig:lfv-markov-bias-variance}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
  \caption{
    \Gls{lfv} plus secular change: as \cref{fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
  }\label{fig:lfv-secular-markov-bias-variance}
\end{figure}

\end{document}

### 2018-10-30 poster-nexus-conference/blocks/at-a-glance.tex
put together a poster for the NEXUS conference

\begin{block}{At a Glance}
    Assessing the utility of a particular risk mitigation instrument involves projecting the climate hazard against which the instrument protects over the $M$-year planning period.
    \begin{itemize}
        \item Climate risk varies in time over a project’s finite planning period, $M$, which may be 1-5 years for a financial risk mitigation instrument and 30-100 years for a structural instrument.
        \item Cyclical climate variability (anthropogenic climate change) dominates near-term (long-term) climate risk
        \item Successful climate adaptation requires prediction of short- and long-term climate variability over the finite planning period $M$
    \end{itemize}
    An often neglected point is that the sources of predictability differ between projects with long and short planning periods.
    \emph{We present a set of stylized experiments to assess how well one can identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy based on these projections.}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/climate-change.tex
put together a poster for the NEXUS conference

\begin{block}{Secular Change}
  Risk, defined as the product of \emph{hazard} and \emph{exposure}, is changing (``nonstationary''):
  \begin{itemize}
    \item \gls{acc} affects the intensity, location, and frequency of hydroclimate extremes \cite{Held:2006jo, OGorman:2009hj}
    \item changes in land use, river channels, and water use affect local hydrologic cycles \cite{Merz:2014gf}
    \item urbanization and development drive increasing exposure to floods \cite{Jongman:2012cr} and hurricanes \cite{Peduzzi:2012iq}
  \end{itemize}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/conclusions.tex
put together a poster for the NEXUS conference

\begin{block}{Key Findings}
    \begin{itemize}
        \item If risk uncertainty increases into future (i.e., \cref{fig:conceptual-sketch}) then adaptation strategies with short $M$ are preferred \cite{Haasnoot:2013im, Zeff:2016hx}
        \item Risk profile of short-$M$ projects dominated by \acrlong{lfv} \cite{Jain:2001hz, Hodgkins:2017hw}
        \item Even though the \gls{hmm} is an imperfect analog for \gls{enso}, it performs well for short $M$
        \item When $M$ is long, trends must be estimated explicitly (but this requires more information -- longer $N$)
        \item Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future
    \end{itemize}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/experiment-design.tex
put together a poster for the NEXUS conference

--- 2018-09-04 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
@@
-\documentclass[
-  draft,
-  linenumbers
-]{agujournal2018}
-\journalname{Earth's Future}
-
-% Macros and newcommands
-\usepackage{xspace}
-\newcommand{\eg}{e.g.\@\xspace}
-\newcommand{\ie}{i.e.\@\xspace}
-\newcommand{\normal}{\mathcal{N}}
-\makeatletter
-\newcommand*{\etc}{%
-    \@ifnextchar{.}%
-        {etc}%
-        {etc.\@\xspace}%
-}
-\makeatother
-
-% Package calls with options
-\usepackage[letterpaper, margin=1in]{geometry} % page geometry
-\usepackage[american]{babel} % language setting
-\usepackage[inline]{enumitem} % in-line lists
-
-% Other package calls
-\usepackage{
-  amssymb, amsmath, % for math symbols
-  booktabs, % for nice tables
-  graphicx, % for figures
-  physics, % for physics notation
-  siunitx, % for SI notation
-  apacite, % AGU style citations apparently
-}
-
-% Fixed-width columns
-\usepackage{array}
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-% Some package settings
-\sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
-\graphicspath{{./figs/}{../fig/}} % folders of figures
-\allowdisplaybreaks{} % let the align environment span multiple pages
-
-% Use the glossaries package
-\usepackage[acronym]{glossaries}
-\makeglossaries
-\newacronym{acc}{ACC}{anthropogenic climate change}
-\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
-\newacronym{cba}{CBA}{cost-benefit analysis}
-\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
-\newacronym{ffa}{FFA}{flood frequency analysis}
-\newacronym{gcm}{GCM}{general circulation model}
-\newacronym{hmm}{HMM}{hidden Markov model}
-\newacronym{iid}{IID}{independent and identically distributed}
-\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
-\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
-\newacronym{lbda}{LBDA}{living blended drought analysis}
-\newacronym{lfv}{LFV}{low-frequency climate variability}
-\newacronym{nao}{NAO}{North Atlantic Oscillation}
-\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
-\newacronym{s2d}{S2D}{seasonal to decadal}
-\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
-
-% These package calls need to come last
-\usepackage[hidelinks]{hyperref}
-\usepackage{cleveref}
-
-% -----------------------------------------------------------------------------
-% ABSTRACT
-% -----------------------------------------------------------------------------
-
-\begin{document}
-
-% Title and Authors
-\title{Robust Adaptation to Multi-Scale Climate Variability}
-\authors{James Doss-Gollin\affil{1,2}, David J. Farnham\affil{1,2}, Scott Steinschneider\affil{3}, Upmanu Lall\affil{1,2}}
-\correspondingauthor{James Doss-Gollin}{james.doss-gollin@columbia.edu}
-\affiliation{1}{Department of Earth and Environmental Engineering, Columbia University}
-\affiliation{2}{Columbia Water Center, Columbia University}
-\affiliation{3}{Department of Biological and Environmental Engineering, Cornell University}
-
-% Key Points
-\begin{keypoints}
-  \item Climate risk varies in time over a project's finite planning period
-  \item The dominant driver of climate risk depends on the length of the planning period 
-  \item Potentially predictable cyclical climate variability dominates near-term climate risk
-  \item Successful climate adaptation requires prediction of short- and long-term climate variability over the planning period
-\end{keypoints}
-
-% Abstract
-\begin{abstract}
-  The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
-  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records highlight the role of \acrlong{lfv} in modulating climate extremes at interannual to multidecadal time scales, \acrlong{acc} is expected to alter their occurrence at longer time scales.
-  In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
-  These experiments consider both quasi-periodic and secular change for the underlying risk, as well as statistical models for estimating this risk from an $N$-year historical record.
-  The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
-  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$ and the future discount rate for investment.
-  Shorter design lives are preferred where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
-\end{abstract}
-
-% -----------------------------------------------------------------------------
-% INTRO
-% -----------------------------------------------------------------------------
-
-\section{Introduction}\label{sec:introduction}
-
-Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
-This climate risk, which is  defined as the product of exposure (the damage that a particular event may cause) and hazard (the probability that this event occurs), is rapidly increasing, at least at the global scale: population growth and urbanization drive increases in exposure \citep{Jongman:2012cr} while \gls{acc} affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Merz:2014gf, Shaw:2016bo}.
-
-Public or private sector investments in climate adaptation require not only the design of each potential structural (\eg, a levee) or non-structural (\eg, a catastrophe bond) risk mitigation instrument, but also selecting between instruments with vastly different operational planning periods.
-This project planning period describes the nominal economic or physical lifespan of the structure or contract and may vary from one year for  a financial contract to 100 years or longer for a permanent structure, as illustrated in \cref{tab:real-world-M}.
-Complex, real-world climate adaptation strategies often consist of many risk mitigation instruments \citep[\eg,][]{CityofNewYork:2013uh}, each of which has its own planning period or design life $M$.
-
-\noindent\begin{table}
-  {\footnotesize
-    \begin{tabular}{L{1in}L{2.5in}L{0.3in}L{1.7in}}
-      \toprule
-      Location & Description & $M$ & Reference \\
-      \midrule
-      Iowa River & Purchase options for inundation of downstream agricultural lands to allow higher release flows from the flood control reservoir & 1 & \citet{Spence:2016ca} \\
-      New York City & Catastrophe bond for protection against storm surge caused by named storms and earthquakes & 3 &  \\
-      County of Santa Barbara, California & Emergency improvements to portions of the Santa Maria Levee to reduce risk of levee failure & 5 & \citet{USACE:2007ta} \\
-      Iowa River & Raise levees by 6 feet & 30 & \citet{Spence:2016ca} \\
-      Dallas, TX & Evacuation of Rockefeller Boulevard & 50 & \citet{USACE:2014vn} \\
-      Central California & Tulare Lake storage and floodwater protection project & 100 & \citet{GEI:gIaEZ-gS} \\
-      \bottomrule
-    \end{tabular}
-    \caption{
-      Several examples of real-world risk mitigation instruments and the associated project planning period ($M$).
-    }\label{tab:real-world-M}
-  }
-\end{table}
-
-Assessing the utility of a particular risk mitigation instrument involves projecting the climate hazard against which the instrument protects over the $M$-year planning period.
-However, an often neglected point is that the sources of predictability differ between projects with long and short planning periods.
-As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a 100 year planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
-By contrast, a three-year insurance contract may be well informed by a skillful forecast of \gls{lfv}.
-Although \cref{fig:conceptual-sketch} shows a highly stylized depiction of \gls{lfv}, analyses of historical and paleo records, like those shown in \cref{fig:observed-lfv}, highlight that subseasonal to multidecadal modes of climate variability such as \gls{enso}, the \gls{pdo}, and the \gls{amo} strongly modulate the occurrence of climate extremes around the world \citep{Ropelewski:1987do, Cook:2010bz, Swierczynski:2012km, Hodgkins:2017hw, Woollings:2018ea}.
-
+\begin{block}{Experiment Design}
     \begin{figure}
         \centering
-  \includegraphics[width=0.7\textwidth]{conceptual-sketch.pdf}
-  \caption{
-    A stylized illustration of (a) irreducible and (b) estimation uncertainty.
-    (a): Irreducible uncertainty cannot be resolved with better models or data and is dominated in the short term by chaotic behavior of the climate, and in the long term by the uncertainty in future \acrlong{acc}.
-    (b): The length of a historical record limits the potential to identify different climate signals.
-    The blue line shows an idealized climate signal and the black line shows observations, which are scattered stochastically around the signal line.
-    The green shading shows the true range within which observations will occur 95\% of the time, while the gray shading the 95\% confidence interval as estimated with a linear trend model.
-    }\label{fig:conceptual-sketch}
-\end{figure}
-
-Since different physical processes control climate risk at different timescales, successful integration of probabilistic climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
-A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.
-As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk, which may in turn lead to high premiums of an insurance policy \citep{Kunreuther:1996kp}.
-The consequences of this uncertainty also interact with those of any estimation bias as illustrated in \cref{fig:conceptual-bias-variance}.
-An instrument designed based on overestimated variance and positive bias will be over-designed, either causing the risk manager to avoid the investmenet, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
-Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
-  \caption{
-    Consequences of model bias or incorrect model representation of uncertainty.
-    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive for the user.
-    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
-  }\label{fig:conceptual-bias-variance}
-\end{figure}
-
-In this paper we present a set of stylized experiments to assess how well one can identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy based on these projections.
-We consider different temporal structures for the underlying risk which encompass quasi-periodic, regime-like, and secular change, as well as simple statistical models for estimating this risk from an $N$-year historical record.
-The relative importance of estimating the short- or long-term risk associated with these extremes depends on the design life $M$, but the potential to understand and predict these different types of variability depends on the informational uncertainty in the $N$-year historical record.
-Though we use floods as an example, the framework also applies to other forms of climate extremes.
-
-% -----------------------------------------------------------------------------
-% METHODS
-% -----------------------------------------------------------------------------
-
-\section{Methods}\label{sec:methods}
-
-We use a statistical framework to
-\begin{enumerate*}[label= (\roman*)]
-  \item generate synthetic sequences of annual-maximum floods;
-  \item fit these sequences to probabilistic models that parameterize time in different ways; and
-  \item evaluate the performance of these models as a function of $M$ and $N$ (\cref{fig:methods-summary})
-\end{enumerate*}.
-
-As shown in \cref{fig:methods-summary}, we first specify $N$, $M$, the flood threshold (\eg, the design level of a structure or the level at which an insurance contract is triggered), the method for generating synthetic streamflow sequences, and the method for estimating flood risk given an $N$-year historical record.
-The methods for generating synthetic streamflow sequences are discussed in \cref{sec:methods-generating} and for estimating flood risk in \cref{sec:methods-estimating}.
-For each scenario of climate risk structure, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
-We repeat the simulations $J = 2000$ times for each combination of experiment parameters to estimate the bias and variance of the statistical estimate (\cref{sec:methods-evaluating}).
-The scenarios considered are summarized in \cref{sec:methods-experiments}.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{flowchart.pdf}
+        \includegraphics[width=\textwidth]{flowchart.pdf}
         \caption{
           Framework for assessing flood risk from synthetic streamflow sequences.
-    Red boxes indicate parameters that are chosen for each experiment, white boxes calculated quantities, and blue boxes the final values used to evalute the model fit.
         }\label{fig:methods-summary}
       \end{figure}
-
-\subsection{Sampling Climate Risk}\label{sec:methods-generating}
-
-We consider two approaches for modeling the influence of quasi-periodic variability and secular climate change on flood risk.
-For each, we take the annual-maximum sequences to follow a log-normal distribution, conditional on a location parameter which may vary in time:
-\begin{equation} \label{eq:lognormal}
-  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
-\end{equation}
-We further assume a constant coefficient of variation of the log streamflow, $\sigma(t) = \xi \mu(t)$, and apply a lower threshold on the standard deviation $\sigma(t) \geq \sigma_\text{min} > 0$.
-The procedure to generate a single sequence of streamflow is thus:
+      Stylized experiment design (\cref{fig:methods-summary}):
       \begin{enumerate}
-  \item choose $t_0$, $M$, and $N$ so that the time is $\vb{t}=t_0-N+1, \ldots, t_0+M$
-  \item use one of the models described below to simulate $\mu(t)$
-  \item calculate $\sigma(t) = \xi \mu(t)$
-  \item for each time $t \in \vb{t}$, draw $Q(t)$ stochastically following \cref{eq:lognormal}.
+        \item For each, generate many synthetic streamflow sequences embedded with one of
+        \begin{enumerate}
+          \item Secular trend only
+          \item \gls{lfv} from \gls{enso} only
+          \item both \gls{lfv} and secular trend
         \end{enumerate}
-We consider two approaches for simulating $\mu(t)$.
-
-The first approach considers $\mu(t)$ as a function of a variable which explicitly represents quasi-periodic climate variability.
-We use an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} plus a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
-We also allow for the inclusion of secular trends.
-We represent \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
-To create an annual time series, we average the October-December values of the NINO3 index for each year.
-\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
-To generate $\mu(t)$, a sequence of length $M+N$ is randomly chosen from the \SI{20000}{year} NINO3 sequence.
-Next, $\mu(t)$ is calculated from the NINO3 time series, denoted as $x(t)$, as
-\begin{equation}\label{eq:nino3}
-  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
-\end{equation}
-where $\beta$ controls the sensitivity to the NINO3 variation and $\gamma$ controls the magnitude of the secular trend.
-
-The second model to create sequences of annual-maxima is a two-state Markov chain model to explicitly enforce persistence and regime structure in $\mu(t)$.
-A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
-The transition matrix is
-\begin{equation}
-  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
-\end{equation}
-The transition matrix is first used to generate a sequence of states $S(t)$.
-The value $\mu(t)$ depends only on $S(t)$ and on time itself:
-\begin{equation}
-  \mu(t) = \begin{cases}
-    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
-    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
-  \end{cases}
-\end{equation}
-For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
-We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
-
-We use the NINO3 and Markov Chain models described above to create three scenarios:
-\begin{enumerate*}[label= (\roman*)]
-  \item \gls{lfv} only;
-  \item secular change only; and
-  \item \gls{lfv} plus secular change.
-\end{enumerate*}
-\gls{lfv} is added to the NINO3 model through $\beta$, which increases the dependence on a climate signal which itself exhibits \gls{lfv}, and to the Markov chain model by making the two states more distinct (increasing $\mu_1 - \mu_2$, or $\gamma_1 - \gamma_2$).
-Secular change is applied by increasing $\gamma$, which gives the time dependence.
-The parameter values used to create these three scenarios are listed in \cref{sec:supp-methods}.
-
-\subsection{Projecting Climate Risk over the Future $M$ years}\label{sec:methods-estimating}
-
-We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
-One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
-We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid, and note that the key conclusions hold for more sophisticated \gls{ffa} methods using statistical or dynamical models.
-
-For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations, denoted $Q_{\text{hist}}$.
-Once a statistical model is fit to $Q_{\text{hist}}$, $K=\num{2000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
-
-We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
-In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
-The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors for regularization.
-The full model is given in \cref{eq:ln2-stationary}.
-
-Next, we expand this stationary model to incorporate secular change.
-Many studies have done this by regressing certain parameters of the model on time \citep{Merz:2014gf, Salas:2018ge}.
-We consider an extension of the stationary LN2 model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:ln2-trend}.
-
-Finally, we consider a model which explicitly represents low-frequency variability.
-A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie{} hidden) states $S(t)$ \citep{Rabiner:1986jk}.
-The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
-We fit  streamflow sequences $Q_\text{hist}$ simulated using the scenarios in Section 2.1, using a \gls{hmm} with two states.
-The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
-This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
-For simplicity, we fit only a two-state \gls{hmm} to each sequence.
-Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.
-
-\subsection{Evaluating Fitting Models}\label{sec:methods-evaluating}
-
-For a given choice of $M$, $N$, generating model, and fitting model, we generate a streamflow sequence of length $M+N$ years.
-The fitting model is applied to the synthetic streamflow sequence's $N$-year ``historical record'', and $K=2000$ posterior simulations of future flows are then drawn from the fitted model.
-
-The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
-We then compute the variance of these $K$ estimates of $\hat{p}_T$.
-We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
-
-Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we repeat this analysis for $J=1000$ different streamflow sequences.
-These are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may be different between the $J$ sequences.
-In \cref{sec:results} we report the expected bias and variance by averaging over the $J$ sequences.
-
-\subsection{Computation}\label{sec:methods-computation}
-
-Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, Schreiber:2017tg, Jones:2001uv, Hoyer:2017hs}.
-Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
-Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke}.
-The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk}\footnote{Dear reviewers: this GitHub repository is currently not publicly available, but codes are provided in the online supplemental information.
-The repository will be made publicly available if and when this work is published.} or in the online supplemental material of this paper.
-
-% -----------------------------------------------------------------------------
-% RESULTS
-% -----------------------------------------------------------------------------
-
-\section{Results}\label{sec:results}
-
-We assess estimation bias and variance for three cases.
-First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
-Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
-Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
-These three cases are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
-Projections with each of the three fitting models are also shown.
-In the remainder of this section we present a more systmematic analysis of each of these three cases.
+        plus stochastic variability.
+        \item Fit using simple Bayesian model
+        \item Evaluate estimation bias and variance across all synthetic streamflow sequences
+        \item Repeat for many combinations of $N$ (length of historical record: proxy for informational uncertainty) and $M$ (project planning period)
+      \end{enumerate}
+      This approach is illustrated in \cref{fig:example-fit}
       \begin{figure}
@@
       \end{figure}
-
-\subsection{Secular Change Only}
-
-In the idealized case where only secular change exists, accurate climate predictions need to either use a long record to identify and model this trend, or to ignore the trend and predict only a few years ahead.
-This is shown in \cref{fig:secular-nino3-bias-variance}, which depicts the estimation bias and variance for each of the three estimation models for many combinations of $M$ and $N$.
-
-The LN2 linear trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
-The variance of these estimates is also large.
-This again highlights the difficulty of identifying the magnitude of secular change, which strongly controls climate risk far into the future, from a single noisy time series.
-By contrast, the stationary LN2 model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
-As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
-This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[\ie,][]{Muller:2014fc}.
-However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace any bias reductions.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{secular-only-nino3-bias-variance.pdf}
-  \caption{
-    Expected estimation bias and variance for sequences generated from the NINO3 model with secular change only (no \gls{lfv}).
-    Sequences were fit to each of three statistical models (colulmns) for different $N$ and $M$ ($x$ and $y$ axis, repsectively).
-    Top row shows estimation bias and bottom row shows log standard deviation of estimates.
-    Note the uneven spacing of the $x$ and $y$ axes.
-  }\label{fig:secular-nino3-bias-variance}
-\end{figure}
-
-If the analyst could know \emph{a priori} that secular change is present in a time series, and if $M$ is long, then the use of a model that captures this secular trend is essential.
-Here the LN2 linear trend model has the advantage of being correctly specified (both the generating and fitting processes assume a log-normal distribution conditional on a linear time trend), which is generally not the case in the real world \citep{Serinaldi:2015bq, Montanari:2014hl}.
-As a result, in real-world settings longer $N$ may be required to identify trends whose exact form is not known.
-Here, $N$ implicitly refers to the uncertainty in trend estimation and may represent an ``equivalent sample size'' if \gls{ipcc}-like projections are used.
-Alternatively, if $M$ is small then it may be reasonable to use a stationary estimate, since the bias will be small and the variance substantially lower.
-
-\subsection{Low-Frequency Variability Only}
-
-Next, in the idealized case where  \gls{lfv} is present but there is no secular change, the identification of nonexistent trends may lead to gross over-estimation of true risk through an increase in the variance of the estimated risk.
-
-Results for this scenario are shown in \cref{fig:lfv-nino3-bias-variance} for the NINO3 model and \cref{fig:lfv-markov-bias-variance} for the two-state Markov model.
-As expected, the stationary model (right column) performs well overall, with low bias and low variance.
-The \gls{hmm} (left column) actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the true multimodal distribution that emerges from dependence on the NINO3 index which exhibits several regimes.
-By contrast, the linear trend model (middle column) performs poorly for low $N$ and high $M$ because a positive probability is assigned to the existence of a positive trend.
-
-Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability of the \gls{lfv}, denoted $T_{\text{LFV}}$.
-As shown in \cref{fig:enso-ts}, the NINO3 time series is most active in the \SIrange{3}{6}{year} band.
-If $T_{\text{LFV}} \ll M$, then a stationary assumption may provide reasonable estimates, and fewer observations may be required (shorter $N$).
-In the real world, however, many hydroclimate time series vary at multidecadal and longer frequencies.
-In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic frequencies $T_{\text{LFV}}$ may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
-This in turn requires a longer observational record $N$ in order to identify and predict these different signals.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-only-nino3-bias-variance.pdf}
-  \caption{
-    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and \gls{lfv} from the NINO3 model.
-  }\label{fig:lfv-nino3-bias-variance}
-\end{figure}
-
-\subsection{Low-Frequency Variability and Secular Change}
-
-In the final and most realistic case, where both \gls{lfv} and secular change are present, stationary models perform well for short $M$ while for long $M$ the trend must be identified from a long record and modeled explicitly.
-
-Simulation results are presented for the NINO3 model in \cref{fig:lfv-secular-nino3-bias-variance} and for the two-state Markov chain model in \cref{fig:lfv-secular-markov-bias-variance}.
-Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the relative importance of secular change and \gls{lfv} depends on $M$.
-When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly (\ie, through the linear trend model).
-Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-secular-nino3-bias-variance.pdf}
-  \caption{
-    As \cref{fig:secular-nino3-bias-variance} but for sequences generated from the NINO3 model with both \gls{lfv} and secular change.
-  }\label{fig:lfv-secular-nino3-bias-variance}
-\end{figure}
-
-% -----------------------------------------------------------------------------
-% DISCUSSION
-% -----------------------------------------------------------------------------
-
-\clearpage
-\section{Discussion}\label{sec:discussion-conclusions}
-
-Evaluating and implementing investments for climate risk mitigation demands making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
-The success of this prediction will depend on the identifiability of different signals from an $N$-year sample, the time scales of \gls{lfv} relative to the project life of the instrument, and the uncertainty associated with the estimation.
-In this paper we took a synthetic data approach to explore the implications of varying $M$ and $N$ in stylized scenarios which represent important features of real-world hydroclimate systems.
-
-The stationary, linear trend, and \gls{hmm} models considered for estimation present an idealized alternative to more complex, physically based models for future risk projection.
-They were chosen to represent the ability of simple models to capture two features of interest: \gls{lfv} and secular trends.
-The models were assessed on their performance over a complete future period of length $M$ years; as discussed in \cref{sec:introduction} the bias and variance computed together represent the likelihood for over- or under-design of the instrument in a given climate scenario (see \cref{fig:conceptual-bias-variance}).
-If more complex methods, such as analysis of \gls{ipcc} scenarios, are used, the sample size $N$ used in our experiments may be straightforwardly interpreted as a surrogate for the total informational uncertainty in the analysis.
-
-The climate adaptation decisions which this analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate.
-The application of a positive discount rate, as is typical in the public and private sectors, emphasizes the importance of predicting near-term risk.
-Even ignoring this, however, if the uncertainty associated with the performance of the instrument increases into the future (as shown in \cref{fig:conceptual-sketch}) then adaptation strategies with a short $M$ are preferred.
-This is consistent with recent emphasis on the importance of making decisions that are robust to a wide range of possible future climates \citep[\ie,][]{Haasnoot:2013im, Poff:2015jn}.
-
-For these short-$M$ projects, cyclical climate variability is dominant over the project planning period \citep{Hodgkins:2017hw, Jain:2001hz}.
-Analysis of the former using the wavelet transformation \citep{Torrence:1998jp} shows dominant spectral peaks at 8, 20, and 64 years, and of the latter at 2 and 15 years.
-This is illustrated in \cref{sec:observed-lfv}, which shows the cyclical variability of two real-world hydroclimate systems: a 500 year drought reconstruction from the \gls{lbda} \citep{Cook:2010bz} and a \SI{107}{year} record of naturalized daily river stage on the American River at the Folsom Dam.
-The long periods and high amplitudes of the oscillations present in these time series highlight the importance of adapting not only to secular change but also, simultaneously, to \gls{lfv}.
-Interestingly, our results show that even though the two-state \gls{hmm} is an imperfect analog for the \gls{lfv} represented in the \gls{enso} process, it performs quite well for short $M$.
-This highlights that even imperfect estimates of \gls{lfv} can support suffessful climate adaptation.
-
-For projects with longer $M$, it becomes necessary to consider the role of secular change.
-As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.
-We find that when making risk projections that consider secular change, mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases.
-This is particularly important given that in real-world settings the true form of the trend is not known and thus ``nonstationary'' projections which assume a particular type of trend are likely to perform poorly \citep{Montanari:2014hl, Serinaldi:2015bq}.
-
-Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, as shown in \cref{fig:observed-lfv}.
-Identifying and modeling these different physical signals requires a large amount of information, represented here by the length of the $N$-year historical record.
-In places where long records are not available, however, information from nearby locations, reanalysis products, or \acrshort{gcm} runs may increase the amount of available information \citep{Merz:2008eh}, effectively increasing $N$.
-
-Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future.
-For most large actors, a portfolio of both large $M$ and small $M$ projects will likely be necessary, none of which precludes the need for mitigation of global and local climate change and the development and execution of vulnerability reduction strategies.
-
-% -----------------------------------------------------------------------------
-% END MATTER
-% -----------------------------------------------------------------------------
-
-\acknowledgments
-
-The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{100000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh:2016hf}.
-The authors thank John High of the United States Army Corps of Engineers for providing the naturalized daily streamflows at the Folsom Dam.
-JDG thanks the NSF GRFP program for support (grant DGE 16-44869: ``Understanding \& Predicting Climate Drivers of Extreme, Mid-latitude River Floods'').
-DJF thanks SERDP program for support (grant 2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate'').
-
-\bibliography{library}
-
-% -----------------------------------------------------------------------------
-% SUPPLEMENTAL INFORMATION -- CURRENTLY INCLUDED IN THIS FILE FOR CONVENIENCE
-% -----------------------------------------------------------------------------
-
-\clearpage
-\appendix
-
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\theequation}{S\arabic{equation}}
-\setcounter{equation}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
-
-\section{Supplemental Methods}\label{sec:supp-methods}
-
-In this section we provide further equations and parameter values used to generate and fit synthetic streamflow sequences.
-
-\subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}
-
-As described in \cref{sec:methods-generating}, we use a synthetic NINO3 time series, as described in \citep{Ramesh:2016hf}, to represent low-frequency climate variability.
-\Cref{fig:enso-ts} shows the wavelet spectrum of this time serties, calculated using WaveletComp software \citep{Roesch:wlBQQoIs}.
-The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.
-There are, however, modes of \gls{lfv} between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
-\begin{figure}
-  \includegraphics[width=\textwidth]{enso_wavelet.png}
-  \caption{
-    Wavelet analysis of the synthetic annual NINO3 time series described in \cref{sec:methods-generating}.
-    (L): wavelet power spectrum.
-    (R): global (average) power spectrum.
-    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
-  }\label{fig:enso-ts}
-\end{figure}
-
-\subsection{Projecting Climate Risk over the Future $M$ Years}
-
-As described in \cref{sec:methods}, we implement the stationary and trend lognormal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
-The full model, including priors, for the stationary lognormal model is given by \cref{eq:ln2-stationary}:
-\begin{align}\label{eq:ln2-stationary}
-  \begin{split}
-    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
-    \mu &\sim \normal \qty(0, 10) \\
-    \sigma &\sim \normal^+ \qty(0, 2)
-  \end{split}
-\end{align}
-where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
-The full model, including priors, for the trend lognormal model is given by \cref{eq:ln2-trend}:
-\begin{align}\label{eq:ln2-trend}
-  \begin{split}
-    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
-  \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
-  \mu_0 & \sim \normal \qty(0, 10) \\
-  \beta_\mu & \sim \normal \qty(0, 0.1) \\
-  \log \xi & \sim \normal \qty(0.1, 0.1)
-  \end{split}
-\end{align}
-where $\xi$ is an estimated coefficient of variation.
-
-\subsection{Experiment Design}\label{sec:methods-experiments}
-
-In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
-To explore the temporal structure of risk, we generate $J=2000$ sequences of streamflow, using both the NINO3 and Markov models, for each of several combinations of $M$ and $N$.
-Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.
-Both the NINO3 model and the Markov model are used; the parameters are shown in \cref{tab:nino-stationary} and \cref{tab:markov-stationary}, respectively.
-
-\begin{table}[ht]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
-    \midrule
-    $\mu_0$             & 6     & 6.5   & 6 \\
-    $\gamma$            & 0     & 0.015 & 0.015\\
-    $\beta$             & 0.5   & 0     & 0.5\\
-    $\xi$               & 0.1   & 0.1   & 0.1\\
-    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
-    Threshold           & 3000  & 3000  & 3000 \\
-    \bottomrule
-  \end{tabular}
-  \caption{
-    Parameters of the NINO3 model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
-  }\label{tab:nino-stationary}
-\end{table}
-\begin{table}[ht]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
-    \midrule
-    $\mu_1$             & 6.75  & 6.5   & 6.75\\
-    $\mu_2$             & 6     & 6.5   & 6\\
-    $\gamma_1$          & 0     & 0.015 & 0.015\\
-    $\gamma_2$          & 0     & 0.015 & 0\\
-    $\xi$               & 0.1   & 0.1   & 0.1\\
-    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
-    Threshold           & 3000  & 3000  & 3000 \\
-    \bottomrule
-  \end{tabular}
-  \caption{
-    Parameters of the two-state Markov chain model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
-  }\label{tab:markov-stationary}
-\end{table}
-
-\section{Supplemental Results}
-
-\subsection{Observed Low-Frequency Hydroclimate Variability}\label{sec:observed-lfv}
-
-\Acrfull{lfv} of the climate is present in a wide range of real-world time series \citep{Hodgkins:2017hw, Swierczynski:2012km, Cook:2010bz}.
-This variability is illustrated in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet \citep{Torrence:1998jp} average spectrum for both.
-Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
-This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
-This clearly illustrates the strong observed multidecadal variability.
-For further discussion of the \gls{lbda} time series we refer the reader to \citet{Cook:2010bz}.
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{observed-lfv.pdf}
-  \caption{
-    Hydroclimate time series vary on many time scales.
-    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
-    High numbers indicate more severe drought.
-    A 20-year running mean is also shown in blue.
-    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
-    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
-    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
-    (d) wavelet spectrum, like (c), for the American River data.
-  }\label{fig:observed-lfv}
-\end{figure}
-
-\subsection{Results for the Markov Model}
-
-In \cref{sec:results} all results given use the NINO3 model to generate synthetic flood sequences.
-In this section equivalent results are given for sequences generated with the two-state Markov chain model described in \cref{sec:methods}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
-  \caption{
-    \Gls{lfv} only: as \cref{fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-  }\label{fig:lfv-markov-bias-variance}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
-  \caption{
-    \Gls{lfv} plus secular change: as \cref{fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-  }\label{fig:lfv-secular-markov-bias-variance}
-\end{figure}
-
-\end{document}
+\end{block}

### 2018-10-30 poster-nexus-conference/blocks/observed-lfv.tex
put together a poster for the NEXUS conference

\begin{block}{Observed \acrshort{lfv}}
  Our analysis is motivated by analysis of historical and paleo records of hydroclimate systems, which often show key modes of variability on interannual to multidecadal time scales \cite{Merz:2014gf, Swierczynski:2012km}.
  The following time series and wavelet global spectra are representative of this \gls{lfv}.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{observed-lfv.pdf}
    \caption{
      Hydroclimate time series vary on many time scales.
      (a) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
      (b) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda} \cite{Cook:2010bz}.
      (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
      (d) like (c), for the \gls{lbda} data.
    }\label{fig:observed-lfv}
  \end{figure}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/systematic-experiments.tex
put together a poster for the NEXUS conference

\begin{block}{Experiment Results}
  We evaluate bias and variance as a function of $M$, $N$, and estimation method for three scenarios.
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{secular-only-nino3-bias-variance.pdf}
    \caption{
      Expected estimation bias and variance for sequences with secular change only (no \gls{lfv}).
    }\label{fig:secular-nino3-bias-variance}
  \end{figure}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{lfv-only-nino3-bias-variance.pdf}
    \caption{
      As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and \gls{lfv} from the \gls{enso} index.
    }\label{fig:lfv-nino3-bias-variance}
  \end{figure}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{lfv-secular-nino3-bias-variance.pdf}
    \caption{
      As \cref{fig:secular-nino3-bias-variance} but for sequences generated from the NINO3 model with both \gls{lfv} and secular change.
    }\label{fig:lfv-secular-nino3-bias-variance}
  \end{figure}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/uncertainty-implications.tex
put together a poster for the NEXUS conference

\begin{block}{Estimation Bias and Variance}
    The insurance premium for a financial risk mitigation instrument on event $X$ can be parameterized as
    \begin{equation*}
        P = \mathbb{E}\qty[X] + \lambda \sigma \qty[X]
    \end{equation*}
    Thus, if an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive for the user.
    Conversely, if an estimate has negative bias and underestimates uncertainty, it will be likely to fail (\cref{fig:conceptual-bias-variance}).
    A key question is thus \emph{whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.}
    \begin{figure}
        \centering
        \includegraphics[width=0.6\textwidth]{Bias-Variance-Sketch.pdf}
        \caption{
        Consequences of model bias or incorrect model representation of uncertainty.
        }\label{fig:conceptual-bias-variance}
    \end{figure}
\end{block}

### 2018-10-30 poster-nexus-conference/blocks/uncertainty.tex
put together a poster for the NEXUS conference

\begin{block}{Types of Uncertainty}
    \begin{figure}
        \centering
        \caption{
            A stylized illustration of (a) irreducible and (b) estimation uncertainty.
        }\label{fig:conceptual-sketch}
        \includegraphics[width=\textwidth]{conceptual-sketch.pdf}
    \end{figure}
    \begin{description}
        \item[Irreducible uncertainty] cannot be resolved with better models or data and is dominated in the short term by chaotic behavior of the climate, and in the long term by the uncertainty in future \acrlong{acc} (\cref{fig:conceptual-sketch}a) 
        \item[estimation uncertainty:] the length of a historical record limits the potential to identify different climate signals (\cref{fig:conceptual-sketch}b)
    \end{description}
\end{block}

### 2018-10-30 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
update the key points because only 3 are allowed
put together a poster for the NEXUS conference

@@
   \item Climate risk varies in time over a project's finite planning period
-  \item The dominant driver of climate risk depends on the length of the planning period 
-  \item Potentially predictable cyclical climate variability dominates near-term climate risk
+  \item Cyclical climate variability (anthropogenic climate change) dominates near-term (long-term) climate risk
   \item Successful climate adaptation requires prediction of short- and long-term climate variability over the planning period
@@
 As described in \cref{sec:methods-generating}, we use a synthetic NINO3 time series, as described in \citep{Ramesh:2016hf}, to represent low-frequency climate variability.
-\Cref{fig:enso-ts} shows the wavelet spectrum of this time serties, calculated using WaveletComp software \citep{Roesch:wlBQQoIs}.
+\Cref{fig:enso-ts} shows the wavelet spectrum of this time series, calculated using WaveletComp software \citep{Roesch:wlBQQoIs}.
 The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.

### 2018-11-12 poster-nexus-conference/blocks/conclusions.tex
Add poster

@@
+{   
+    \large
+    \setbeamercolor*{block body}{fg=black,bg=wcblue!30}
     \begin{block}{Key Findings}
+        Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future:
         \begin{itemize}
-        \item If risk uncertainty increases into future (i.e., \cref{fig:conceptual-sketch}) then adaptation strategies with short $M$ are preferred \cite{Haasnoot:2013im, Zeff:2016hx}
+            \item If uncertainty increases into future then adaptation strategies with short planning periods are preferred
             \item Risk profile of short-$M$ projects dominated by \acrlong{lfv} \cite{Jain:2001hz, Hodgkins:2017hw}
-        \item Even though the \gls{hmm} is an imperfect analog for \gls{enso}, it performs well for short $M$
-        \item When $M$ is long, trends must be estimated explicitly (but this requires more information -- longer $N$)
-        \item Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future
+            \item Even though the \gls{hmm} is an imperfect analog for \gls{enso}, it performs well for short planning periods
+            \item When the planning period is long, trends must be estimated explicitly (requiring more data)
         \end{itemize}
     \end{block}
+}

### 2018-11-12 poster-nexus-conference/blocks/systematic-experiments.tex
Add poster

@@
     \centering
-    \includegraphics[width=0.9\textwidth]{secular-only-nino3-bias-variance.pdf}
+    \includegraphics[width=0.95\textwidth]{secular-only-nino3-bias-variance.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{lfv-only-nino3-bias-variance.pdf}
+    \includegraphics[width=0.95\textwidth]{lfv-only-nino3-bias-variance.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{lfv-secular-nino3-bias-variance.pdf}
+    \includegraphics[width=0.95\textwidth]{lfv-secular-nino3-bias-variance.pdf}
     \caption{

### 2018-11-12 make_simulate.sh

@@ -11,7 +11,7 @@
 #SBATCH --time=3:59:00          # The time the job will take to run.
 
 module load anaconda/3-4.4.0    # load the anaconda module
-source activate floodsampling   # activate the conda environment
+source activate gcm-lfv   # activate the conda environment
 make simulate		        # run the make command specified in the Makefile
 
 # End of script

### 2018-12-04 slides-agu/presentation.tex
first draft of slides for AGU

\documentclass[
  10pt,     % set default generic font size
  handout   % ignores all the \pause commands
]{beamer}
\usetheme[block=fill]{metropolis}

% -----------------------------------------------------------------------------
% Title and Author Information -- Need to Edit
% -----------------------------------------------------------------------------

\title{H52F-05H: Robust Adaptation to\\Multi-Scale Climate Variability}
\subtitle{Toward Better Water Planning and Management in an Uncertain World I}
\date{19 October 2018}
\author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
\institute{$^1$: Columbia University Department of Earth and Environmental Engineering\\$^2$: Carnegie Institution for Science\\$^3$: Department of Biological and Environmental Engineering, Cornell University}
\titlegraphic{\hfill\includegraphics[height=1.25cm]{SeasCrown_blue.png}}

% -----------------------------------------------------------------------------
% Package Configuration -- Don't Necessarily Need to Edit
% -----------------------------------------------------------------------------

% Packages with Options
\usepackage[english]{babel}

% Package List
\usepackage{
  array,                              % for custom table widths
  appendixnumberbeamer,               % don't count appendix slides in progress bar
	booktabs,                           % for better (alternative?) tables
  natbib,                             % references!
	physics,                            % for better notation
  siunitx,                            % for SI notation
}

% cool fonts
\usepackage{fontspec}
\usepackage{fontawesome5}

% Use TIKZ
\input{tikz_setup.tex}

% figures
\usepackage{graphicx}
\graphicspath{{../fig/}} % can add more

% Fixed-width columns
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% Change the captions
\setbeamerfont{caption}{size=\scriptsize}

% macros
\usepackage{xspace}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\makeatletter
\newcommand*{\etc}{%
    \@ifnextchar{.}%
        {etc}%
        {etc.\@\xspace}%
}
\makeatother

% Biblatex Setup using a file called library.bib
\setbeamertemplate{bibliography item}[text] % don't print the symbols

% this has to come last
\usepackage{cleveref}

% -----------------------------------------------------------------------------
% BEGIN DOCUMENT HERE
% -----------------------------------------------------------------------------

\begin{document}

% TITLE PAGE
\maketitle

\begin{frame}{Thanks to...}
  \begin{itemize}
    \item Co-authors
    \begin{itemize}
      \item Upmanu Lall
      \item David J. Farnham
      \item Scott Steinschneider
    \end{itemize}
    \item Funders
    \begin{itemize}
      \item NSF GRPF
      \item Columbia University Fu Foundation SEAS
      \item DoD SERDP
    \end{itemize}
    \item Conveners
  \end{itemize}
\end{frame}

\begin{frame}{NYC After Sandy \citep{CityofNewYork:2013uh}}
  \begin{columns}%[T]
    \begin{column}{0.5\textwidth}
      \begin{figure}
        \includegraphics[width=\columnwidth]{three-barriers.png}
        \caption{Possible construction of barriers for storm surge protection in New York}
      \end{figure}
      Structural ``Mega-Project'':
      \begin{itemize}
        \item Expensive: $\approx$ \$ 25 billion
        \item Environmental, social impacts
      \end{itemize}
    \end{column}
    \pause
    \begin{column}{0.5\textwidth}
      Alternative:
      \begin{itemize}
        \item Lots of local-scale projects
        \item Resilience boosting and vulnerability reduction
        \item Ecological restoration, financial instruments, zoning changes, structures where needed
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Planning Periods}
  Projects are designed for finite ``planning periods'' $M$ over which climate risk must be estimated.
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{itemize}
        \item Success of ``mega-project'' depends on climate conditions $M \geq \SI{50}{years}$ in the future
        \item Success of flexible, non-structural strategy (\ie financial instrument) depends on climate conditions $M \leq \SI{5}{years}$ in the future
      \end{itemize}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{figure}
        \includegraphics[width=\columnwidth]{Bias-Variance-Sketch.pdf}
        \caption{Consequences of over- or under-estimating climate risk and associated uncertainty}
      \end{figure}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Drivers of Future Climate Risk}
  The physical drivers of hazard depend on the projection horizon ($M$).
  However, our ability to identify these drivers depends on information available (\eg, the length of an $N$-year observational record).
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.55\textheight,keepaspectratio=true]{conceptual-sketch.pdf}
    \caption{Schematic description of (a) intrinsic and (b) estimation uncertainty.}
  \end{figure}
\end{frame}

\begin{frame}{Real-World Climate Signals}
  Real-world hydroclimate systems vary on many time and space scales
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{observed-lfv.pdf}
    \caption{
      (a) \SI{500}{year} reconstruction of summer rainfall over Arizona from LBDA \citep{Cook:2010bz}.
      (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
      (c),(d): wavelet global (average) spectra.
    }\label{fig:observed-lfv}
  \end{figure}
\end{frame}

\begin{frame}{Research Objective}
  How well can one identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy given limited information (for which an $N$-year record is a proxy).
  In particular, we consider:
  \begin{itemize}
    \item the specific roles of $M$ and $N$;
    \item the choice of estimation strategy; and
    \item the form of the underlying climate signal.
  \end{itemize}
\end{frame}

\section{Stylized Experiments}

\begin{frame}{Research Plan}
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{flowchart.pdf}
    \caption{Flow chart of methods used.}
  \end{figure}
  \pause
  \emph{N.B.}: models used for generating sequences and estimating risk chosen for \emph{interpretability}, not general \emph{validity} in applications.
  Results are similar with other models.
\end{frame}

\begin{frame}{Example Sequences}
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{Example-NINO3-M100-N50.pdf}
    \caption{
      Synthetic flood sequences (blue) and estimated 50\%, 95\% intervals (gray).
      Three scenarios are used for generating sequences (columns) and three models are used for estimating risk (rows).
      Here $N=50$ and $M=100$.
    }
  \end{figure}
\end{frame}

\begin{frame}{Stationarity: LFV Only}
  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
  Explicitly modeling LFV improves estimates.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
    \caption{
      Expected estimation bias and variance for sequences with LFV but zero trend.
    }\label{fig:lfv-nino3-bias-variance}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationarity: Secular Change Only}
  Stationary models under-estimate risk, particularly for long $M$.
  For short $M$ and short $N$, the uncertainty by estimating a trend is not worth the bias reduction.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
    \caption{
      Expected estimation bias and variance for sequences with secular change only (no LFV).
    }\label{fig:secular-nino3-bias-variance}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationarity: Secular Change + LFV}
  Long $M$ requires trend estimation, but without sufficient data (long $N$) trend estimates are strongly biased.

  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
    \caption{
      As \cref{fig:secular-nino3-bias-variance} but for sequences generated with both LFV and secular change.
    }\label{fig:lfv-secular-nino3-bias-variance}
  \end{figure}
\end{frame}

\section{Discussion}

\begin{frame}{Idealized Experiments $\iff$ Real World}
  The idealizations and ``toy'' models used here are analogues for more complex real-world situations.
  \begin{table}
    \begin{tabular}{L{0.425\textwidth}L{0.525\textwidth}}
      \toprule
      Analysis & Real World \\\midrule
      $N$-year record & Total informational uncertainty of an estimate \\\midrule
      Statistical models of increasing complexity and \# parameters & Statistical and dynamical model chains of increasing complexity and \# parameters \\\midrule
      LFV comes from ENSO only & LFV comes from many signals \\\midrule
      Linear trends & Secular changes of unknown form \\\midrule
      Trends and LFV Independent & Trends and LFV modes interact\\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

\begin{frame}{Key Findings}
  Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future:
  \begin{itemize}
      \item As uncertainty increases, adaptation strategies with short planning periods are preferred
      \item Risk profile of short-$M$ projects dominated by LFV \citep{Jain:2001hz, Hodgkins:2017hw}
      \item When the planning period is long, trends must be estimated explicitly -- requiring more data
  \end{itemize}
\end{frame}

\begin{frame}{Some Potential Implications}
  \begin{table}
    \begin{tabular}{L{0.425\textwidth}L{0.525\textwidth}}
      \toprule
      Question & Analysis \\\midrule
      Linear trend models? & No, that was just for demonstration -- avoid un-physical extrapolation \\\midrule
      Use only last 30 years to address nonstationarity? & No, this increases variance; use paleo records, observations, and climate models for risk estimation \\\midrule
      Build large infrastructure to manage a changing climate? & Sometimes, but consider alternatives, including postponing construction to increase $N$\\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

% -----------------------------------------------------------------------------
% QUESTIONS, BIBLIOGRAPHY
% -----------------------------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
  \renewcommand*{\bibfont}{\footnotesize}
  \renewcommand{\bibsection}{}
  \nocite{DossGollin:TjTkb07T}
	\bibliographystyle{agufull08}
  \bibliography{library}
\end{frame}

\begin{frame}[standout]
  \alert{Thanks for your attention!}\\\vspace{1.5cm}
  \begin{tabular}{rl}
    \faIcon[regular]{twitter},\faIcon[regular]{github} & \url{@jdossgollin} \\
    \faIcon[regular]{envelope} & \url{james.doss-gollin@columbia.edu}\\
    \faIcon[regular]{paperclip} & \url{www.jamesdossgollin.me}
  \end{tabular}
\end{frame}

% -----------------------------------------------------------------------------
% BACKUP SLIDES
% -----------------------------------------------------------------------------

\appendix
\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\theequation}{A\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{table}{0}
\section{Supplemental Figures}

\begin{frame}
  We can add more figures here
\end{frame}

\end{document}

### 2018-12-04 slides-agu/presentation.tex
Update slides and add figure

@@
   10pt,     % set default generic font size
-  handout   % ignores all the \pause commands
+  %handout   % ignores all the \pause commands
 ]{beamer}
@@
 \author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
-\institute{$^1$: Columbia University Department of Earth and Environmental Engineering\\$^2$: Carnegie Institution for Science\\$^3$: Department of Biological and Environmental Engineering, Cornell University}
+\institute{
+  $^1$Columbia University Department of Earth and Environmental Engineering\\
+  $^2$Carnegie Institution for Science\\
+  $^3$Department of Biological and Environmental Engineering, Cornell University}
 \titlegraphic{\hfill\includegraphics[height=1.25cm]{SeasCrown_blue.png}}
@@
 
-% Use TIKZ
-\input{tikz_setup.tex}
-
 % figures
@@
 \makeatother
+\newcommand{\usd}[1]{\SI[round-precision=2,round-mode=places,round-integer-to-decimal]{#1}[\$]{}}
+\newcommand{\normal}{\mathcal{N}}
 
@@
         \includegraphics[width=\columnwidth]{three-barriers.png}
-        \caption{Possible construction of barriers for storm surge protection in New York}
       \end{figure}
-      Structural ``Mega-Project'':
       \begin{itemize}
-        \item Expensive: $\approx$ \$ 25 billion
         \item Environmental, social impacts
+        \item Permanent and inflexible: $\mathcal{O} \qty(\$25 \text{billion})$ is ``locked-in''
+        \item Potential for over- or under-design
       \end{itemize}
@@
     \begin{column}{0.5\textwidth}
-      Alternative:
+      \begin{figure}
+        \includegraphics[width=\columnwidth]{flood-concept.png}
+      \end{figure}
       \begin{itemize}
-        \item Lots of local-scale projects
-        \item Resilience boosting and vulnerability reduction
-        \item Ecological restoration, financial instruments, zoning changes, structures where needed
+        \item Local resilience and vulnerability reduction
+        \item Portfolio of long- and short-term investments
+        \item Update portfolio regularly
       \end{itemize}
@@
 \setcounter{table}{0}
-\section{Supplemental Figures}
 
-\begin{frame}
-  We can add more figures here
+\section{Generating Synthetic Streamflow Sequences}
+
+\begin{frame}{Equations for Synthetic Streamflow Generation}
+  First
+  \begin{equation} \label{eq:lognormal}
+    \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
+  \end{equation}
+  Where $\sigma(t) = \xi \mu(t)$, with $\sigma(t) \geq \sigma_\text{min} > 0$.
+  Then,
+  \begin{equation}\label{eq:nino3}
+    \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
+  \end{equation}
+  and where $x(t)$ is NINO3.4 index from realistic ENSO model \citep{Zebiak:1987cl,Ramesh:2016hf}
+\end{frame}
+
+\begin{frame}{Spectrum of LFV Used}
+  \begin{figure}
+    \includegraphics[width=\textwidth,height=0.6\textheight,keepaspectratio=true]{enso_wavelet}
+    \caption{Wavelet spectrum of (sub-set of) ENSO model used to embed synthetic streamflow sequences with low-frequency variability}
+  \end{figure}
+\end{frame}
+
+\section{Climate Risk Estimation}
+
+\begin{frame}{Stationary LN2 Model}
+  Treat the $N$ historical observations as IID draws from stationary distribution
+  \begin{align}\label{eq:ln2-stationary}
+    \begin{split}
+      \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
+      \mu &\sim \normal \qty(7, 1.5) \\
+      \sigma &\sim \normal^+ \qty(1, 1)
+    \end{split}
+  \end{align}
+  where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
+  Fit in Bayesian framework using stan \citep{Carpenter:2017ke}.
+\end{frame}
+
+\begin{frame}{Trend LN2 Model}
+  Treat the $N$ historical observations as IID draws from log-normal distribution with linear trend
+  \begin{align}\label{eq:ln2-trend}
+    \begin{split}
+      \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
+    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
+    \mu_0 & \sim \normal \qty(7, 1.5) \\
+    \beta_\mu & \sim \normal \qty(0, 0.1) \\
+    \log \xi & \sim \normal \qty(0.1, 0.1)
+    \end{split}
+  \end{align}
+  where $\xi$ is an estimated coefficient of variation.
+  Also fit in stan.
+\end{frame}
+
+\begin{frame}{Hidden Markov Model}
+  Two-state HMM \citep[see][]{Rabiner:1986jk} implemented using pomegranate python package \citep{Schreiber:2017tg}.
+  See package documentation for reference.
 \end{frame}

### 2018-12-04 slides-agu/presentation.tex
add a shameless plug

@@
 
-\begin{frame}{Some Potential Implications}
-  \begin{table}
-    \begin{tabular}{L{0.425\textwidth}L{0.525\textwidth}}
-      \toprule
-      Question & Analysis \\\midrule
-      Linear trend models? & No, that was just for demonstration -- avoid un-physical extrapolation \\\midrule
-      Use only last 30 years to address nonstationarity? & No, this increases variance; use paleo records, observations, and climate models for risk estimation \\\midrule
-      Build large infrastructure to manage a changing climate? & Sometimes, but consider alternatives, including postponing construction to increase $N$\\
-      \bottomrule
-    \end{tabular}
-  \end{table}
+\begin{frame}{A Shameless Pitch!}
+  I'm working to quantify these principles.
+  Specifically:
+  \begin{itemize}
+    \item Statistical models for integrating paleo records, observations, and GCMs towards climate risk projection
+    \item Financial valuation of the ``option'' to build a particular project
+    \item Optimizing portfolios of long- and short-term investments
+  \end{itemize}
+  If you have a working model of your system and are making long-/short-term investment choices I'd \alert{love to talk} further!
 \end{frame}

### 2018-12-08 slides-agu/presentation.tex
Revise the talk after practicing it a bit

@@
 \subtitle{Toward Better Water Planning and Management in an Uncertain World I}
-\date{19 October 2018}
+\date{14 December 2018}
 \author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
@@
   siunitx,                            % for SI notation
+  appendixnumberbeamer,               % don't number slides / sections in appendix
 }
@@
 \newcommand{\normal}{\mathcal{N}}
+\newcommand*{\ditto}{---''---}
 
@@
 
+% Use the glossaries package
+\usepackage[acronym]{glossaries}
+\makeglossaries
+\newacronym{acc}{ACC}{anthropogenic climate change}
+\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
+\newacronym{cba}{CBA}{Cost-Benefit Analysis}
+\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
+\newacronym{ffa}{FFA}{flood frequency analysis}
+\newacronym{gcm}{GCM}{general circulation model}
+\newacronym{hmm}{HMM}{hidden Markov model}
+\newacronym{iid}{IID}{independent and identically distributed}
+\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
+\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
+\newacronym{lbda}{LBDA}{living blended drought analysis}
+\newacronym{lfv}{LFV}{low-frequency climate variability}
+\newacronym{nao}{NAO}{North Atlantic Oscillation}
+\newacronym{npv}{NPV}{Net Present Value}
+\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
+\newacronym{s2d}{S2D}{seasonal to decadal}
+\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
+
 % this has to come last
 \usepackage{cleveref}
+\usepackage{hyperref}
 
@@
     \end{itemize}
-    \item Conveners
+    \item Conveners, colleagues, collaborators
   \end{itemize}
@@
 
+\section{Setup}
+
 \begin{frame}{NYC After Sandy \citep{CityofNewYork:2013uh}}
-  \begin{columns}%[T]
+  \begin{columns}[T]
+    \pause
     \begin{column}{0.5\textwidth}
@@
         \item Environmental, social impacts
-        \item Permanent and inflexible: $\mathcal{O} \qty(\$25 \text{billion})$ is ``locked-in''
-        \item Potential for over- or under-design
+        \item Permanent: $\approx \$25 \text{billion}$ ``locked-in''
       \end{itemize}
@@
       \begin{itemize}
-        \item Local resilience and vulnerability reduction
-        \item Portfolio of long- and short-term investments
-        \item Update portfolio regularly
+        \item Targeted local resilience $\Uparrow$ and vulnerability $\Downarrow$
+        \item Portfolio of investments
+        \item Update regularly
+      \end{itemize}
+    \end{column}
+  \end{columns}
+  \pause
+  \begin{alertblock}{How should NYC allocate resources, over time,}
+    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
+  \end{alertblock}
+\end{frame}
+
+\begin{frame}{Idea 1: Risk Estimates over Finite Future Periods}
+  \begin{alertblock}{Typical Approach:}
+    \gls{cba} over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
+  \end{alertblock}
+  \begin{columns}
+    \pause
+    \begin{column}{0.55\textwidth}
+      Thus, evaluation of project depends on climate conditions over the next $M$ years:
+      \begin{itemize}
+        \item For ``mega-project'', $M \geq \SI{50}{years}$
+        \item For small, flexible project, $M \leq \SI{5}{years}$
+      \end{itemize}
+      \pause
+      So, misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
+    \end{column}
+    \pause
+    \begin{column}{0.4\textwidth}
+      \includegraphics[width=\textwidth]{Bias-Variance-Sketch}
+      {\small Estimation bias and variance drive project success!}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Idea 2: Physical Drivers of Risk Depend on $M$}
+  The physical drivers of hazard depend on the projection horizon ($M$).
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-a.png}\\
+  \end{figure}
+  \pause
+  However, our ability to identify these drivers depends on information available (\eg, the length of an $N$-year observational record).
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-b.png}
+  \end{figure}
+  \emph{(informational uncertainty $\Leftarrow$ clouds, sea ice, subsurface, \etc)}
+\end{frame}
+
+\begin{frame}{Research Objective}
+  \begin{alertblock}{How well can one identify and predict risk}
+     associated with cyclical and secular climate signals over a finite design life ($M$),  given limited information (\eg, an $N$-year record).
+  \end{alertblock}
+  \pause
+  In particular, we consider:
+  \begin{itemize}
+    \item the specific roles of $M$ and $N$;
+    \pause\item the choice of estimation strategy; and
+    \pause\item the form of the underlying climate signal.
+  \end{itemize}
+  \pause
+  Use simplest possible representation of key processes.
+\end{frame}
+
+\section{Stylized Experiments}
+
+\begin{frame}{Breathless Overview of Methods}
+  \begin{columns}[T]
+    \begin{column}{0.475\textwidth}
+      We will consider:
+      \begin{itemize}
+        \item 3 scenarios:
+        \begin{itemize}
+          \item \gls{lfv} only
+          \item linear trend only
+          \item trend + \gls{lfv}
+        \end{itemize}
+        \item 3 fitting models:
+        \begin{itemize}
+          \item stationary
+          \item trend
+          \item \gls{hmm}
+        \end{itemize}
+        \item Many values of $\qty{M, N}$
+      \end{itemize}
+    \end{column}
+    \pause
+    \begin{column}{0.475\textwidth}
+      \begin{itemize}
+        \item \gls{lfv} from \gls{enso} model \citep{Ramesh:2016hf, Zebiak:1987cl}
+        \item For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance (recall: these map to bias of \gls{cba})
       \end{itemize}
@@
 
-\begin{frame}{Planning Periods}
-  Projects are designed for finite ``planning periods'' $M$ over which climate risk must be estimated.
-  \begin{columns}
-    \begin{column}{0.6\textwidth}
-      \begin{itemize}
-        \item Success of ``mega-project'' depends on climate conditions $M \geq \SI{50}{years}$ in the future
-        \item Success of flexible, non-structural strategy (\ie financial instrument) depends on climate conditions $M \leq \SI{5}{years}$ in the future
-      \end{itemize}
-    \end{column}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[width=\columnwidth]{Bias-Variance-Sketch.pdf}
-        \caption{Consequences of over- or under-estimating climate risk and associated uncertainty}
-      \end{figure}
-    \end{column}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Drivers of Future Climate Risk}
-  The physical drivers of hazard depend on the projection horizon ($M$).
-  However, our ability to identify these drivers depends on information available (\eg, the length of an $N$-year observational record).
+\begin{frame}{Stationary Scenario (LFV Only)}
+  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
+  Explicitly modeling \gls{lfv} improves estimates, particularly for long $N$.
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth,height=0.55\textheight,keepaspectratio=true]{conceptual-sketch.pdf}
-    \caption{Schematic description of (a) intrinsic and (b) estimation uncertainty.}
+    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
   \end{figure}
@@
 
-\begin{frame}{Real-World Climate Signals}
-  Real-world hydroclimate systems vary on many time and space scales
+\begin{frame}{Nonstationary Scenario I (Secular Change Only)}
+  For long $M$, it's important to estimate the trend -- but only works well if $N$ also long.
+  For short $M$ and short $N$, the uncertainty added by estimating a trend is not worth the bias reduction.
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{observed-lfv.pdf}
-    \caption{
-      (a) \SI{500}{year} reconstruction of summer rainfall over Arizona from LBDA \citep{Cook:2010bz}.
-      (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
-      (c),(d): wavelet global (average) spectra.
-    }\label{fig:observed-lfv}
+    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
   \end{figure}
@@
 
-\begin{frame}{Research Objective}
-  How well can one identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy given limited information (for which an $N$-year record is a proxy).
-  In particular, we consider:
-  \begin{itemize}
-    \item the specific roles of $M$ and $N$;
-    \item the choice of estimation strategy; and
-    \item the form of the underlying climate signal.
-  \end{itemize}
-\end{frame}
-
-\section{Stylized Experiments}
-
-\begin{frame}{Research Plan}
+\begin{frame}{Nonstationary Scenario II (Secular Change + LFV)}
+  Conclusions are similar to ``Secular Change Only'' but variance has increased -- \ie as the system becomes more complex, more data is needed to understand it.
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{flowchart.pdf}
-    \caption{Flow chart of methods used.}
-  \end{figure}
-  \pause
-  \emph{N.B.}: models used for generating sequences and estimating risk chosen for \emph{interpretability}, not general \emph{validity} in applications.
-  Results are similar with other models.
-\end{frame}
-
-\begin{frame}{Example Sequences}
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{Example-NINO3-M100-N50.pdf}
-    \caption{
-      Synthetic flood sequences (blue) and estimated 50\%, 95\% intervals (gray).
-      Three scenarios are used for generating sequences (columns) and three models are used for estimating risk (rows).
-      Here $N=50$ and $M=100$.
-    }
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Stationarity: LFV Only}
-  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
-  Explicitly modeling LFV improves estimates.
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
-    \caption{
-      Expected estimation bias and variance for sequences with LFV but zero trend.
-    }\label{fig:lfv-nino3-bias-variance}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Nonstationarity: Secular Change Only}
-  Stationary models under-estimate risk, particularly for long $M$.
-  For short $M$ and short $N$, the uncertainty by estimating a trend is not worth the bias reduction.
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
-    \caption{
-      Expected estimation bias and variance for sequences with secular change only (no LFV).
-    }\label{fig:secular-nino3-bias-variance}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Nonstationarity: Secular Change + LFV}
-  Long $M$ requires trend estimation, but without sufficient data (long $N$) trend estimates are strongly biased.
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth,height=0.65\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
-    \caption{
-      As \cref{fig:secular-nino3-bias-variance} but for sequences generated with both LFV and secular change.
-    }\label{fig:lfv-secular-nino3-bias-variance}
+    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
   \end{figure}
@@
 
-\begin{frame}{Idealized Experiments $\iff$ Real World}
-  The idealizations and ``toy'' models used here are analogues for more complex real-world situations.
-  \begin{table}
-    \begin{tabular}{L{0.425\textwidth}L{0.525\textwidth}}
-      \toprule
-      Analysis & Real World \\\midrule
-      $N$-year record & Total informational uncertainty of an estimate \\\midrule
-      Statistical models of increasing complexity and \# parameters & Statistical and dynamical model chains of increasing complexity and \# parameters \\\midrule
-      LFV comes from ENSO only & LFV comes from many signals \\\midrule
-      Linear trends & Secular changes of unknown form \\\midrule
-      Trends and LFV Independent & Trends and LFV modes interact\\
-      \bottomrule
-    \end{tabular}
-  \end{table}
-\end{frame}
-
 \begin{frame}{Key Findings}
-  Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future:
+  \begin{alertblock}{How should NYC allocate resources, over time,}
+    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
+  \end{alertblock}
   \begin{itemize}
+    \pause
+    \item When planning period is short, \gls{lfv} dominates \citep{Hodgkins:2017hw} $\Rightarrow$ uncertainty small
+    \pause
+    \item When $M$ long, future changes must be estimated using complex models which require more information $\Rightarrow$ more uncertainty
+    \pause
     \item As uncertainty increases, adaptation strategies with short planning periods are preferred
-      \item Risk profile of short-$M$ projects dominated by LFV \citep{Jain:2001hz, Hodgkins:2017hw}
-      \item When the planning period is long, trends must be estimated explicitly -- requiring more data
   \end{itemize}
@@
 \begin{frame}{A Shameless Pitch!}
-  I'm working to quantify these principles.
+  I'm working on turning these concepts for project \emph{evaluation} into models for \emph{decision-making}.
   Specifically:
+  \pause
   \begin{itemize}
-    \item Statistical models for integrating paleo records, observations, and GCMs towards climate risk projection
-    \item Financial valuation of the ``option'' to build a particular project
+    \item Integrating paleo records, observations, and \glspl{gcm} for climate risk projection
+    \pause
+    \item Valuing the ``option'' to build a particular project
+    \pause
     \item Optimizing portfolios of long- and short-term investments
   \end{itemize}
+  \pause
   If you have a working model of your system and are making long-/short-term investment choices I'd \alert{love to talk} further!
@@
 \begin{frame}[allowframebreaks]{References}
-  \renewcommand*{\bibfont}{\footnotesize}
+  \renewcommand*{\bibfont}{\scriptsize}
   \renewcommand{\bibsection}{}
   \nocite{DossGollin:TjTkb07T}
-	\bibliographystyle{agufull08}
+	\bibliographystyle{agu}
   \bibliography{library}
@@
   \begin{tabular}{rl}
-    \faIcon[regular]{twitter},\faIcon[regular]{github} & \url{@jdossgollin} \\
-    \faIcon[regular]{envelope} & \url{james.doss-gollin@columbia.edu}\\
+    \faIcon[regular]{twitter},\faIcon[regular]{github} & \href{https://twitter.com/jdossgollin}{@jdossgollin} \\
+    \faIcon[regular]{envelope} & \href{mailto:james.doss-gollin@columbia.edu}{james.doss-gollin@columbia.edu}\\
     \faIcon[regular]{paperclip} & \url{www.jamesdossgollin.me}
@@
 
+\section{Relating our Analysis to Observations}
+
+\begin{frame}{Idealized Experiments $\iff$ Real World}
+  The idealized models used here are analogs:
+  \begin{table}
+    \centering
+    \begin{tabular}{L{0.425\textwidth}L{0.525\textwidth}}
+      \toprule
+      Analysis & Real World \\\midrule
+      $N$-year record & Total informational uncertainty of an estimate \\\midrule
+      Statistical models of increasing complexity and \# parameters & Statistical and dynamical model chains of increasing complexity and \# parameters \\\midrule
+      Linear trends & Secular changes of unknown form \\\midrule
+      \gls{lfv} from \gls{enso} & \gls{lfv} from many sources \\\midrule
+      \gls{lfv} and trend additive & \gls{lfv} and trend interact \\
+      \bottomrule
+    \end{tabular}
+  \end{table}
+\end{frame}
+
+\begin{frame}{Real-World Climate Signals}
+  Real-world hydroclimate systems vary on many time and space scales
+  \begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{observed-lfv.pdf}
+    \caption{
+      (a) \SI{500}{year} reconstruction of summer rainfall over Arizona from LBDA \citep{Cook:2010bz}.
+      (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
+      (c),(d): wavelet global (average) spectra.
+    }\label{fig:observed-lfv}
+  \end{figure}
+\end{frame}
+
 \section{Generating Synthetic Streamflow Sequences}
@@
   \end{equation}
-  and where $x(t)$ is NINO3.4 index from realistic ENSO model \citep{Zebiak:1987cl,Ramesh:2016hf}
+  and where $x(t)$ is NINO3.4 index from realistic \gls{enso} model \citep{Zebiak:1987cl,Ramesh:2016hf}
 \end{frame}
 
-\begin{frame}{Spectrum of LFV Used}
+\begin{frame}{Spectrum of \gls{lfv} Used}
   \begin{figure}
     \includegraphics[width=\textwidth,height=0.6\textheight,keepaspectratio=true]{enso_wavelet}
-    \caption{Wavelet spectrum of (sub-set of) ENSO model used to embed synthetic streamflow sequences with low-frequency variability}
+    \caption{
+      Wavelet spectrum of (sub-set of) \gls{enso} model used to embed synthetic streamflow sequences with low-frequency variability.
+      \gls{enso} data from \citet{Ramesh:2016hf}.
+    }
   \end{figure}
@@
 \begin{frame}{Stationary LN2 Model}
-  Treat the $N$ historical observations as IID draws from stationary distribution
+  Treat the $N$ historical observations as \gls{iid} draws from stationary distribution
   \begin{align}\label{eq:ln2-stationary}
@@
 \begin{frame}{Trend LN2 Model}
-  Treat the $N$ historical observations as IID draws from log-normal distribution with linear trend
+  Treat the $N$ historical observations as \gls{iid} draws from log-normal distribution with linear trend
   \begin{align}\label{eq:ln2-trend}
@@
 \begin{frame}{Hidden Markov Model}
-  Two-state HMM \citep[see][]{Rabiner:1986jk} implemented using pomegranate python package \citep{Schreiber:2017tg}.
+  Two-state \gls{hmm} \citep[see][]{Rabiner:1986jk} implemented using pomegranate python package \citep{Schreiber:2017tg}.
   See package documentation for reference.

### 2018-12-10 slides-agu/presentation.tex
Update slides -- new look

@@
   10pt,     % set default generic font size
-  %handout   % ignores all the \pause commands
+  handout   % ignores all the \pause commands
 ]{beamer}
@@
 \titlegraphic{\hfill\includegraphics[height=1.25cm]{SeasCrown_blue.png}}
+
+% Add footer for AGU
+\setbeamertemplate{footline}[text line]{%
+  \parbox{0.8\linewidth}{
+    \vspace*{-8pt}James Doss-Gollin (\url{james.doss-gollin@columbia.edu})
+  }
+  \hfill%
+  \parbox{0.15\linewidth}{
+    \vspace*{-8pt}\raggedleft\insertpagenumber
+  }
+}
 
@@
 
-\section{Setup}
-
-\begin{frame}{NYC After Sandy \citep{CityofNewYork:2013uh}}
+\begin{frame}{Motivating Example}
   \begin{columns}[T]
@@
       \begin{figure}
-        \includegraphics[width=\columnwidth]{three-barriers.png}
+        \centering
+        \caption{
+          Permanent structural solution would pose environmental and social impacts and lock in $\approx \$25 \text{billion}$;
+          a portfolio of smaller projects may allow a more flexible strategy.
+        }
+        \includegraphics[width=0.85\columnwidth]{three-barriers.png}\\
+        \includegraphics[width=0.85\columnwidth]{flood-concept.png}
       \end{figure}
-      \begin{itemize}
-        \item Environmental, social impacts
-        \item Permanent: $\approx \$25 \text{billion}$ ``locked-in''
-      \end{itemize}
     \end{column}
@@
     \begin{column}{0.5\textwidth}
-      \begin{figure}
-        \includegraphics[width=\columnwidth]{flood-concept.png}
-      \end{figure}
       \begin{itemize}
-        \item Targeted local resilience $\Uparrow$ and vulnerability $\Downarrow$
-        \item Portfolio of investments
-        \item Update regularly
+        \item ``Because present decisions impact a system's ability to adapt to future needs, flexibility in activating, delaying, and replacing engineering projects should be considered in least‐cost water supply intervention scheduling,'' \citep{erfani:2018}.
+        \item How to build a portfolio of risk management investments, over time, given uncertain future climate/society/technology?
       \end{itemize}
@@
   \end{columns}
-  \pause
-  \begin{alertblock}{How should NYC allocate resources, over time,}
-    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
-  \end{alertblock}
 \end{frame}
+
+\section{Underlying Ideas}
 
@@
   \begin{alertblock}{Typical Approach:}
-    \gls{cba} over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
+    \gls{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
   \end{alertblock}
@@
     \pause
-    \begin{column}{0.55\textwidth}
+    \begin{column}{0.5\textwidth}
       Thus, evaluation of project depends on climate conditions over the next $M$ years:
@@
       \end{itemize}
-      \pause
-      So, misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
     \end{column}
     \pause
-    \begin{column}{0.4\textwidth}
-      \includegraphics[width=\textwidth]{Bias-Variance-Sketch}
-      {\small Estimation bias and variance drive project success!}
+    \begin{column}{0.5\textwidth}
+      Misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
+      \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio=true]{Bias-Variance-Sketch}
     \end{column}
@@
   \pause
-  However, our ability to identify these drivers depends on information available (\eg, the length of an $N$-year observational record).
+  However, our ability to identify these mechanisms depends on information available (\eg, the length of an $N$-year observational record).
   \begin{figure}
@@
   \end{figure}
-  \emph{(informational uncertainty $\Leftarrow$ clouds, sea ice, subsurface, \etc)}
 \end{frame}
+
+\section{Stylized Experiments}
 
@@
     \item the specific roles of $M$ and $N$;
-    \pause\item the choice of estimation strategy; and
-    \pause\item the form of the underlying climate signal.
+    \pause
+    \item estimation strategy (few processes/parameters $\Rightarrow$ many processes/parameters)
+    \pause
+    \item the form of the underlying climate signal.
   \end{itemize}
@@
 \end{frame}
-
-\section{Stylized Experiments}
 
@@
     \begin{column}{0.475\textwidth}
-      \begin{itemize}
-        \item \gls{lfv} from \gls{enso} model \citep{Ramesh:2016hf, Zebiak:1987cl}
-        \item For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance (recall: these map to bias of \gls{cba})
-      \end{itemize}
+      \begin{figure}
+        \caption{Over- or under-estimating uncertainty leads to biased estimates of $p(X > X^*)$ even if estimate of $X$ is unbiased}
+        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-LN2.jpg}
+      \end{figure}
+      For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance of $p(X > X^*)$
     \end{column}
@@
     \centering
-    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
+    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
   \end{figure}
@@
     \centering
-    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
+    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
   \end{figure}
@@
     \centering
-    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
+    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
   \end{figure}
@@
 
-\begin{frame}{Key Findings}
-  \begin{alertblock}{How should NYC allocate resources, over time,}
-    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
-  \end{alertblock}
+\begin{frame}{Summary}
+  Starting point:
+  \begin{enumerate}
+    \item Investment evaluation depends on climate condition over finite planning period
+    \item Physical drivers of risk depend on planning period
+  \end{enumerate}
+  \pause
+  As a consequence:
   \begin{itemize}
+    \item \emph{Ability} to identify and predict different climate signals depends on information available (\eg, $N$)
+    \item \emph{Importance} of identifying and predicting different climate signals depends on degree of extrapolation desired (\ie, $M$)
+  \end{itemize}
   \pause
-    \item When planning period is short, \gls{lfv} dominates \citep{Hodgkins:2017hw} $\Rightarrow$ uncertainty small
-    \pause
-    \item When $M$ long, future changes must be estimated using complex models which require more information $\Rightarrow$ more uncertainty
-    \pause
-    \item As uncertainty increases, adaptation strategies with short planning periods are preferred
-  \end{itemize}
+  In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
 \end{frame}
@@
   Specifically:
-  \pause
   \begin{itemize}
     \item Integrating paleo records, observations, and \glspl{gcm} for climate risk projection
-    \pause
     \item Valuing the ``option'' to build a particular project
-    \pause
     \item Optimizing portfolios of long- and short-term investments
   \end{itemize}
-  \pause
-  If you have a working model of your system and are making long-/short-term investment choices I'd \alert{love to talk} further!
+  If you have a working model of your system and are making long/short-term investment choices I'd \alert{love to talk} further!
 \end{frame}
@@
 
+\begin{frame}{Example Sequences and Fits}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{Example-NINO3-M100-N50.pdf}
+    \caption{Example of sequences generated with $M=100$ and $N=50$}
+  \end{figure}
+\end{frame}
+
 \begin{frame}{Equations for Synthetic Streamflow Generation}

### 2018-12-11 slides-agu/presentation.tex
Update draft of slides for AGU

@@
 
-\begin{frame}{Thanks to...}
-  \begin{itemize}
-    \item Co-authors
-    \begin{itemize}
-      \item Upmanu Lall
-      \item David J. Farnham
-      \item Scott Steinschneider
-    \end{itemize}
-    \item Funders
-    \begin{itemize}
-      \item NSF GRPF
-      \item Columbia University Fu Foundation SEAS
-      \item DoD SERDP
-    \end{itemize}
-    \item Conveners, colleagues, collaborators
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Motivating Example}
+\begin{frame}{Motivating Example \citep{CityofNewYork:2013uh}}
   \begin{columns}[T]
-    \pause
     \begin{column}{0.5\textwidth}
       \begin{figure}
-        \centering
+        \includegraphics[width=\columnwidth]{three-barriers.png}
         \caption{
-          Permanent structural solution would pose environmental and social impacts and lock in $\approx \$25 \text{billion}$;
-          a portfolio of smaller projects may allow a more flexible strategy.
+          Large, permanent project would lock in $\approx \$25 \text{billion}$ for fixed protection.
         }
-        \includegraphics[width=0.85\columnwidth]{three-barriers.png}\\
-        \includegraphics[width=0.85\columnwidth]{flood-concept.png}
       \end{figure}
     \end{column}
-    \pause
     \begin{column}{0.5\textwidth}
-      \begin{itemize}
-        \item ``Because present decisions impact a system's ability to adapt to future needs, flexibility in activating, delaying, and replacing engineering projects should be considered in least‐cost water supply intervention scheduling,'' \citep{erfani:2018}.
-        \item How to build a portfolio of risk management investments, over time, given uncertain future climate/society/technology?
-      \end{itemize}
+      \begin{figure}
+        \includegraphics[width=\columnwidth]{flood-concept.png}
+        \caption{
+          A sequenced portfolio of smaller investments may be more flexible and adaptive.
+        }
+      \end{figure}
     \end{column}
   \end{columns}
+  \begin{alertblock}{Scheduling and sequencing:}
+    \alert{If} and \alert{when} to build the permanent project?
+  \end{alertblock}
 \end{frame}
@@
   \end{alertblock}
-  \begin{columns}
   \pause
-    \begin{column}{0.5\textwidth}
-      Thus, evaluation of project depends on climate conditions over the next $M$ years:
+  Project should be evaluated on climate conditions over this finite planning period:
   \begin{itemize}
@@
   \end{itemize}
-    \end{column}
-    \pause
-    \begin{column}{0.5\textwidth}
-      Misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
-      \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio=true]{Bias-Variance-Sketch}
-    \end{column}
-  \end{columns}
 \end{frame}
 
-\begin{frame}{Idea 2: Physical Drivers of Risk Depend on $M$}
-  The physical drivers of hazard depend on the projection horizon ($M$).
+\begin{frame}{Idea 2: Hydroclimate Systems Vary on Many Scales}
+  \begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{observed-lfv.pdf}
+    \caption{
+      (a) \SI{500}{year} reconstruction of summer rainfall over Arizona from LBDA \citep{Cook:2010bz}.
+      (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
+      (c),(d): wavelet global (average) spectra.
+    }\label{fig:observed-lfv}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Idea 3: Physical Drivers of Risk Depend on $M$}
+  The physical drivers of hazard depend on the projection horizon ($M$),
   \begin{figure}
@@
   \pause
-  However, our ability to identify these mechanisms depends on information available (\eg, the length of an $N$-year observational record).
+  but our ability to identify these mechanisms depends on information available (\eg, the length of an $N$-year observational record).
   \begin{figure}
@@
   \pause
-  Use simplest possible representation of key processes.
+  Use parsimonious representation of key processes.
 \end{frame}
@@
 \begin{frame}{Summary}
-  Starting point:
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      Assertions:
       \begin{enumerate}
         \item Investment evaluation depends on climate condition over finite planning period
+        \item Physical hydroclimate systems vary on many scales
         \item Physical drivers of risk depend on planning period
       \end{enumerate}    
-  \pause
-  As a consequence:
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      Implications:
       \begin{itemize}
@@
         \item \emph{Importance} of identifying and predicting different climate signals depends on degree of extrapolation desired (\ie, $M$)
+        \item In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
       \end{itemize}    
-  \pause
-  In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
+    \end{column}
+  \end{columns}
 \end{frame}
@@
   \end{table}
-\end{frame}
-
-\begin{frame}{Real-World Climate Signals}
-  Real-world hydroclimate systems vary on many time and space scales
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{observed-lfv.pdf}
-    \caption{
-      (a) \SI{500}{year} reconstruction of summer rainfall over Arizona from LBDA \citep{Cook:2010bz}.
-      (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
-      (c),(d): wavelet global (average) spectra.
-    }\label{fig:observed-lfv}
-  \end{figure}
 \end{frame}

### 2018-12-12 slides-agu/presentation.tex
Try to sketch out some text

@@
   siunitx,                            % for SI notation
-  appendixnumberbeamer,               % don't number slides / sections in appendix
 }
@@
 
-\begin{frame}{Motivating Example \citep{CityofNewYork:2013uh}}
-  \begin{columns}[T]
-    \begin{column}{0.5\textwidth}
+\begin{frame}{Motivating Example}
+  What to do after Sandy? \citep{CityofNewYork:2013uh}
   \begin{figure}
-        \includegraphics[width=\columnwidth]{three-barriers.png}
+    \centering
+    \includegraphics[height=0.3\textheight]{three-barriers.png}~
+    \includegraphics[height=0.3\textheight]{flood-concept.png}
     \caption{
-          Large, permanent project would lock in $\approx \$25 \text{billion}$ for fixed protection.
+      A large, permanent project would lock in $\approx \$25 \text{billion}$ for fixed protection; a portfolio of smaller investments may be more flexible and adaptive.
     }
   \end{figure}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      \begin{figure}
-        \includegraphics[width=\columnwidth]{flood-concept.png}
-        \caption{
-          A sequenced portfolio of smaller investments may be more flexible and adaptive.
-        }
-      \end{figure}
-    \end{column}
-  \end{columns}
+  \pause
   \begin{alertblock}{Scheduling and sequencing:}
-    \alert{If} and \alert{when} to build the permanent project?
+    When (and how) to build a flexible portfolio vs mega-project
   \end{alertblock}
@@
 
-\section{Underlying Ideas}
+\begin{frame}{Table of contents}
+  \setbeamertemplate{section in toc}[sections numbered]
+  \tableofcontents[hideallsubsections]
+\end{frame}
+
+\section{Hypotheses}
 
@@
 \begin{frame}{Idea 2: Hydroclimate Systems Vary on Many Scales}
+  Inter-annual to multi-decadal variability very important in near future (\ie, for small $M$) \citep{Hodgkins:2017hw}
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{observed-lfv.pdf}
+    \includegraphics[width=\textwidth,height=0.55\textheight,keepaspectratio=true]{observed-lfv.pdf}
     \caption{
@@
   \begin{alertblock}{How well can one identify and predict risk}
-     associated with cyclical and secular climate signals over a finite design life ($M$),  given limited information (\eg, an $N$-year record).
+     associated with cyclical and secular climate signals over a finite planning period ($M$),  given limited information (\eg, an $N$-year record).
   \end{alertblock}
   \pause
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
       In particular, we consider:
@@
       \end{itemize}
-  \pause
-  Use parsimonious representation of key processes.
-\end{frame}
-
-\begin{frame}{Breathless Overview of Methods}
-  \begin{columns}[T]
-    \begin{column}{0.475\textwidth}
-      We will consider:
-      \begin{itemize}
-        \item 3 scenarios:
-        \begin{itemize}
-          \item \gls{lfv} only
-          \item linear trend only
-          \item trend + \gls{lfv}
-        \end{itemize}
-        \item 3 fitting models:
-        \begin{itemize}
-          \item stationary
-          \item trend
-          \item \gls{hmm}
-        \end{itemize}
-        \item Many values of $\qty{M, N}$
-      \end{itemize}
     \end{column}
     \pause
-    \begin{column}{0.475\textwidth}
+    \begin{column}{0.5\textwidth}
       \begin{figure}
-        \caption{Over- or under-estimating uncertainty leads to biased estimates of $p(X > X^*)$ even if estimate of $X$ is unbiased}
         \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-LN2.jpg}
+        \caption{
+          What is the bias and spread of $p(X > X^*)$?
+          Here, $M=100$ and $N=50$.
+          Subsequent results are tested systematically over $J=1000$ sequences.
+        }
       \end{figure}
-      For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance of $p(X > X^*)$
     \end{column}
@@
 \begin{frame}{Summary}
-  \begin{columns}
+  \begin{columns}[T]
     \begin{column}{0.5\textwidth}
@@
     \end{column}
+    \pause
     \begin{column}{0.5\textwidth}
@@
       \begin{itemize}
+        \pause
         \item \emph{Ability} to identify and predict different climate signals depends on information available (\eg, $N$)
+        \pause
         \item \emph{Importance} of identifying and predicting different climate signals depends on degree of extrapolation desired (\ie, $M$)
+        \pause
         \item In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
@@
   \end{columns}
-\end{frame}
-
-\begin{frame}{A Shameless Pitch!}
-  I'm working on turning these concepts for project \emph{evaluation} into models for \emph{decision-making}.
-  Specifically:
-  \begin{itemize}
-    \item Integrating paleo records, observations, and \glspl{gcm} for climate risk projection
-    \item Valuing the ``option'' to build a particular project
-    \item Optimizing portfolios of long- and short-term investments
-  \end{itemize}
-  If you have a working model of your system and are making long/short-term investment choices I'd \alert{love to talk} further!
 \end{frame}
@@
 \begin{frame}[standout]
-  \alert{Thanks for your attention!}\\\vspace{1.5cm}
+  \alert{Thanks for your attention!}\\
+  \vspace{1.5cm}
+  Interested in making these ideas more concrete?
+  I'd love to collaborate!\\
+  \vspace{1.5cm}
   \begin{tabular}{rl}

### 2018-12-12 slides-agu/presentation.tex
update slides to be a bit less verbose

@@
   \begin{alertblock}{Typical Approach:}
-    \gls{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
+    \acrfull{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years.
   \end{alertblock}
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-LN2.jpg}
+        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-Stationary-Trend.pdf}
         \caption{
@@
 \begin{frame}{Nonstationary Scenario II (Secular Change + LFV)}
-  Conclusions are similar to ``Secular Change Only'' but variance has increased -- \ie as the system becomes more complex, more data is needed to understand it.
+  As the system becomes more complex, more data is needed to understand it.
   \begin{figure}
@@
       Assertions:
-      \begin{enumerate}
+      \begin{itemize}
         \item Investment evaluation depends on climate condition over finite planning period
@@
         \item Physical drivers of risk depend on planning period
-      \end{enumerate}    
+      \end{itemize}    
     \end{column}

### 2018-12-13 slides-agu/presentation.tex
potentially a final version of slides?

@@
   \parbox{0.15\linewidth}{
-    \vspace*{-8pt}\raggedleft\insertpagenumber
+    \vspace*{-8pt}\raggedleft\insertframenumber
   }
@@
 \begin{frame}{Research Objective}
-  \begin{alertblock}{How well can one identify and predict risk}
-     associated with cyclical and secular climate signals over a finite planning period ($M$),  given limited information (\eg, an $N$-year record).
+  \begin{alertblock}{How to identify \& predict}
+    cyclical and secular climate signals over a finite planning period ($M$),  given limited information (\eg, an $N$-year record).
  \end{alertblock}
@@
         \item the specific roles of $M$ and $N$;
+        \item estimation strategy
+        \item structure of climate signal
+      \end{itemize}
       \pause
-        \item estimation strategy (few processes/parameters $\Rightarrow$ many processes/parameters)
-        \pause
-        \item the form of the underlying climate signal.
-      \end{itemize}
+      To analyze predictions, we assess the \emph{bias} and \emph{variance} of predicted flood occurrence.
     \end{column}
-    \pause
     \begin{column}{0.5\textwidth}
       \begin{figure}
-        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-Stationary-Trend.pdf}
-        \caption{
-          What is the bias and spread of $p(X > X^*)$?
-          Here, $M=100$ and $N=50$.
-          Subsequent results are tested systematically over $J=1000$ sequences.
-        }
+        \includegraphics[width=\textwidth]{Bias-Variance-Sketch.pdf}
       \end{figure}
@@
 \begin{frame}{Stationary Scenario (LFV Only)}
-  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
-  Explicitly modeling \gls{lfv} improves estimates, particularly for long $N$.
+  With limited data, the uncertainties caused by extrapolating from complex models lead to poor performance.
   \begin{figure}
@@
 \begin{frame}{Nonstationary Scenario I (Secular Change Only)}
-  For long $M$, it's important to estimate the trend -- but only works well if $N$ also long.
+  For long $M$, it's important to estimate the trend -- but this works well only if $N$ also long.
   For short $M$ and short $N$, the uncertainty added by estimating a trend is not worth the bias reduction.
@@
         \pause
-        \item \emph{Importance} of identifying and predicting different climate signals depends on degree of extrapolation desired (\ie, $M$)
+        \item \emph{Importance} of predicting different climate signals depends on extrapolation desired (\ie, planning period)
         \pause
-        \item In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
+        \item In general, low risk tolerance and/or limited information favor investments with short planning periods.
       \end{itemize}
@@
 
-\section{Relating our Analysis to Observations}
+\section{Supplemental Discussion}
 

### 2018-12-13 slides-agu/presentation.tex
Improve README text

@@
     \centering
-    \includegraphics[height=0.3\textheight]{three-barriers.png}~
+    \includegraphics[height=0.3\textheight]{three-barriers.png}
     \includegraphics[height=0.3\textheight]{flood-concept.png}
     \caption{
-      A large, permanent project would lock in $\approx \$25 \text{billion}$ for fixed protection; a portfolio of smaller investments may be more flexible and adaptive.
+      A large, permanent structure (L) would lock in $\approx \$25 \text{billion}$; a portfolio of smaller investments (R) may be more flexible and robust.
     }
@@
  \pause
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
-      In particular, we consider:
-      \begin{itemize}
-        \item the specific roles of $M$ and $N$;
-        \item estimation strategy
-        \item structure of climate signal
-      \end{itemize}
-      \pause
-      To analyze predictions, we assess the \emph{bias} and \emph{variance} of predicted flood occurrence.
-    \end{column}
-    \begin{column}{0.5\textwidth}
   \begin{figure}
-        \includegraphics[width=\textwidth]{Bias-Variance-Sketch.pdf}
+    \includegraphics[width=\textwidth]{flowchart.pdf}
   \end{figure}
-    \end{column}
-  \end{columns}
+  defining $P^* \def \mathbb{P} \qty(X > X^*)$, the insurance premium
+  \begin{equation*}
+    R = \mathbb{E}\qty[ P^*] + \lambda \mathbb{V} \qty[P^*]
+  \end{equation*}
 \end{frame}

### 2018-12-13 slides-agu/presentation.tex
Close to final draft, we hope

@@
     \centering
-    \includegraphics[height=0.3\textheight]{three-barriers.png}
-    \includegraphics[height=0.3\textheight]{flood-concept.png}
-    \caption{
-      A large, permanent structure (L) would lock in $\approx \$25 \text{billion}$; a portfolio of smaller investments (R) may be more flexible and robust.
-    }
+    \includegraphics[height=0.35\textheight]{three-barriers.png}~
+    \includegraphics[height=0.35\textheight]{dull-disasters.png}\\
+    \includegraphics[height=0.3\textheight]{flood-concept.png}~
+    \includegraphics[height=0.3\textheight]{coastal-resiliency.png}
   \end{figure}
-  \pause
-  \begin{alertblock}{Scheduling and sequencing:}
-    When (and how) to build a flexible portfolio vs mega-project
-  \end{alertblock}
-\end{frame}
-
-\begin{frame}{Table of contents}
-  \setbeamertemplate{section in toc}[sections numbered]
-  \tableofcontents[hideallsubsections]
 \end{frame}
@@
 \begin{frame}{Idea 2: Hydroclimate Systems Vary on Many Scales}
-  Inter-annual to multi-decadal variability very important in near future (\ie, for small $M$) \citep{Hodgkins:2017hw}
+  Inter-annual to multi-decadal cyclical variability key (for small $M$)
   \begin{figure}
@@
 
-\begin{frame}{Research Objective}
-  \begin{alertblock}{How to identify \& predict}
-    cyclical and secular climate signals over a finite planning period ($M$),  given limited information (\eg, an $N$-year record).
+\begin{frame}{Experiment Setup}
+  \begin{alertblock}{Research Objective}
+    How well can one identify \& predict cyclical and secular climate signals over a finite planning period ($M$),  given limited information?
  \end{alertblock}
- \pause
-  \begin{figure}
-    \includegraphics[width=\textwidth]{flowchart.pdf}
-  \end{figure}
-  defining $P^* \def \mathbb{P} \qty(X > X^*)$, the insurance premium
+  Let $P^* \def \mathbb{P} \qty(X > X^*)$.
+  Note that the insurance premium (or risk factor) is:
   \begin{equation*}
@@
   \end{equation*}
+  \pause
+  \begin{columns}
+    \begin{column}{0.5\textwidth}
+      \begin{figure}
+        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-Stationary-Trend.pdf}
+      \end{figure}
+    \end{column}
+    \begin{column}{0.5\textwidth}
+      Systematic, stylized experiments:
+      what happens as we vary $M$, $N$, climate structure, estimating model?
+    \end{column}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}{Stationary Scenario (LFV Only)}
-  With limited data, the uncertainties caused by extrapolating from complex models lead to poor performance.
+  \emph{With limited data, the uncertainties caused by extrapolating from complex models lead to poor performance.}
   \begin{figure}
@@
 \begin{frame}{Nonstationary Scenario I (Secular Change Only)}
-  For long $M$, it's important to estimate the trend -- but this works well only if $N$ also long.
-  For short $M$ and short $N$, the uncertainty added by estimating a trend is not worth the bias reduction.
+  \emph{Long planning periods need trend estimation, but this demands lots of information.
+  For short planning periods, simple models may be better.}
   \begin{figure}
@@
 \begin{frame}{Nonstationary Scenario II (Secular Change + LFV)}
-  As the system becomes more complex, more data is needed to understand it.
+  \emph{As the system becomes more complex, more data is needed to understand it.}
   \begin{figure}

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Add poster
update and send to dave for comments

@@
 \documentclass[
-  draft,
-  linenumbers
+  %draft,
+  %linenumbers
 ]{agujournal2018}
@@
 \newacronym{lbda}{LBDA}{living blended drought analysis}
-\newacronym{lfv}{LFV}{low-frequency climate variability}
+\newacronym{lfv}{LFV}{low-frequency variability}
 \newacronym{nao}{NAO}{North Atlantic Oscillation}
@@
   \item Climate risk varies in time over a project's finite planning period
-  \item Cyclical climate variability (anthropogenic climate change) dominates near-term (long-term) climate risk
-  \item Successful climate adaptation requires prediction of short- and long-term climate variability over the planning period
+  \item Importance of predicting different climate signals depends on extrapolation desired (planning period)
+  \item Ability to identify and predict different climate signals depends on information available (\eg, data)
 \end{keypoints}
@@
 Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
-This climate risk, which is  defined as the product of exposure (the damage that a particular event may cause) and hazard (the probability that this event occurs), is rapidly increasing, at least at the global scale: population growth and urbanization drive increases in exposure \citep{Jongman:2012cr} while \gls{acc} affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Merz:2014gf, Shaw:2016bo}.
+Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
+For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
+At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
+Even if mitigation efforts are successful, novel adaptation strategies are needed.
 
-Public or private sector investments in climate adaptation require not only the design of each potential structural (\eg, a levee) or non-structural (\eg, a catastrophe bond) risk mitigation instrument, but also selecting between instruments with vastly different operational planning periods.
-This project planning period describes the nominal economic or physical lifespan of the structure or contract and may vary from one year for  a financial contract to 100 years or longer for a permanent structure, as illustrated in \cref{tab:real-world-M}.
-Complex, real-world climate adaptation strategies often consist of many risk mitigation instruments \citep[\eg,][]{CityofNewYork:2013uh}, each of which has its own planning period or design life $M$.
+This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
+Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs,\footnote{cite the environmental costs of big infrastructure projects} and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have slowed their deployment.
+Instead, actors such as New York City have turned to a combination of structural, operational, and financial instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+These instruments are not typically implemented in isolation or statically.
+Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.
+
+Despite recent insights, important questions remain.
+How should a portfolio of risk mitigation instruments be optimized?
+How should one choose between permanent and transient instruments?
+Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
+In this paper we focus on how the temporal structure of climate risk, and the uncertainties associated with estimating it, influence the answers to this questions.
+We continue this section with three specific observations about climate risk which, while apparently obvious, have important and subtle implications which we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
+
+\subsection{Planning decisions are made with finite horizons}\label{sec:intro-finite}
+
+Public or private sector investments in climate adaptation require not only the design of each potential structural instrument, but also selecting between instruments with vastly different operational planning periods.
+This project planning period, which we define as being $M$ years, describes the nominal economic or physical lifespan of the structure or contract.
+Typical planning periods may vary from $M=1$ year or less for a financial contract to $M=100$ years or longer for a structural instrument, as illustrated in \cref{tab:real-world-M}.
+The planning period can also be interpreted as the finite period over which \gls{cba} is conducted when assessing the project.\footnote{we should add a citation about \gls{cba} here, perhaps a technical spec from the Army Corps?}
 
@@
     \caption{
-      Several examples of real-world risk mitigation instruments and the associated project planning period ($M$).
+      Six real-world risk mitigation instruments and the associated project planning period ($M$).
     }\label{tab:real-world-M}
@@
 
-Assessing the utility of a particular risk mitigation instrument involves projecting the climate hazard against which the instrument protects over the $M$-year planning period.
-However, an often neglected point is that the sources of predictability differ between projects with long and short planning periods.
-As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a 100 year planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
-By contrast, a three-year insurance contract may be well informed by a skillful forecast of \gls{lfv}.
-Although \cref{fig:conceptual-sketch} shows a highly stylized depiction of \gls{lfv}, analyses of historical and paleo records, like those shown in \cref{fig:observed-lfv}, highlight that subseasonal to multidecadal modes of climate variability such as \gls{enso}, the \gls{pdo}, and the \gls{amo} strongly modulate the occurrence of climate extremes around the world \citep{Ropelewski:1987do, Cook:2010bz, Swierczynski:2012km, Hodgkins:2017hw, Woollings:2018ea}.
+Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
+Each has its own planning period, which may be separate from the planning horizon of the portfolio as a whole.
+This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this long-term plan may be achieved through a series of flexible and adaptive instruments with short individual planning periods.
+The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
+
+\subsection{Climate risk varies on many scales}\label{sec:intro-lfv}
+
+Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
+Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on river floods, droughts, hurricanes, urban flooding, and many other climate hazards are well documented.\footnote{Add citations of \gls{acc} influences on climate hazards}
+Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
+In combination, these effects highlight that the past may not be an adequate representation of future climate risk \citep[termed ``nonstationarity'' by][]{Milly:2008dg}.
+
+Permanent or secular change is not, however, a prerequisite for the inadequacy of relying on historical records.
+For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \acrfull{lfv}, due, \eg, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
+We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet average spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
+Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
+This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
+A detailed analysis of these time series is beyond the scope of this paper, but we note that these findings are illustrative of other analyses of climatic \gls{lfv} \citep{Jain:2001hz,Kiem:2002kq,Cook:2010bz,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw,Cassou:2018cs}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{observed-lfv.pdf}
+  \caption{
+    Hydroclimate time series vary on many time scales.
+    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
+    High numbers indicate more severe drought.
+    A 20-year running mean is also shown in blue.
+    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
+    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
+    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
+    (d) wavelet spectrum, like (c), for the American River data.
+  }\label{fig:observed-lfv}
+\end{figure}
+
+\subsection{The dominant processes depend on the planning period}\label{sec:intro-dominant}
+
+Assessing the utility of a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
+Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
+As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
+Even with a perfect climate model, these uncertainties will be large.
+By contrast, our perfect climate model may usefully inform climate hazard over a three-year insurance contract with much less associated uncertainty.
+
+Of course, scientists are not equipped with perfect models.
+Since different physical processes control climate risk at different timescales, successful integration of climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
+A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.
+As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
 
@@
 
-Since different physical processes control climate risk at different timescales, successful integration of probabilistic climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
-A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.
-As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk, which may in turn lead to high premiums of an insurance policy \citep{Kunreuther:1996kp}.
-The consequences of this uncertainty also interact with those of any estimation bias as illustrated in \cref{fig:conceptual-bias-variance}.
-An instrument designed based on overestimated variance and positive bias will be over-designed, either causing the risk manager to avoid the investmenet, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
+% -----------------------------------------------------------------------------
+% METHODS
+% -----------------------------------------------------------------------------
+
+\section{Methods}\label{sec:methods}
+
+We consider a set of stylized experiments to assess how well one can identify and predict risk associated with cyclical and secular climate signals for the $M$-year planning period and the probability of over- or under-design of a climate adaptation strategy based on these projections.
+We consider different temporal structures for the underlying risk which encompass quasi-periodic, regime-like, and secular change, as well as simple statistical models for estimating this risk from an $N$-year historical record.
+The relative importance of estimating the short- or long-term risk associated with these extremes depends on the design life $M$, but the potential to understand and predict these different types of variability depends on the informational uncertainty in the $N$-year historical record.
+Though we use floods as an example, the framework also applies to other forms of climate extremes.
+
+We consider three scenarios for climate risk, which we define by the structure of the underlying climate signal:
+\begin{enumerate*}[label= (\roman*)]
+  \item \gls{lfv} only;
+  \item \gls{lfv} plus secular change; and
+  \item secular change only.
+\end{enumerate*}
+For each scenario, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
+We repeat the simulations $J = 2000$ times for each combination of experiment parameters to estimate the bias and variance of the statistical estimate (\cref{sec:methods-evaluating}).
+
+We caution the reader that the models for sampling climate risk and for statistically projecting future risk were chosen for their intuitive interpretation, rather than their general validity \citep[see][for a thoughtful discussion of the value of simple models]{Held:2005cj}.
+We do not, in general, endorse these models for practical use but instead argue that the conclusions drawn from these simple models may be straightforwardly applied to more complex and realistic models.
+This discussion continues in \cref{sec:discussion-conclusions}.
+
+\subsection{Sampling Climate Risk}\label{sec:methods-generating}
+
+The first step is to sample climate risk by generating synthetic streamflow sequences.
+To do this, we take the annual-maximum sequences to follow a log-normal distribution, conditional on a location parameter which may vary in time:
+\begin{equation} \label{eq:log-normal}
+  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
+\end{equation}
+We further assume a constant coefficient of variation of the log streamflow,
+\begin{equation}
+  \sigma(t) = \xi \mu(t)  
+\end{equation}
+and apply a lower threshold on the standard deviation
+\begin{equation}
+  \sigma(t) \geq \sigma_\text{min} > 0.
+\end{equation}
+This flexible framing describes the \begin{enumerate*}[label= (\roman*)]
+  \item \gls{lfv} only;
+  \item \gls{lfv} plus secular change; and
+  \item secular change only
+\end{enumerate*}
+scenarios in a single equation; the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
+\begin{equation}\label{eq:nino3}
+  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
+\end{equation}
+where $x(t)$ represents a climate time series which itself exhibits \gls{lfv}.
+This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
+Thus, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
+The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.
+
+We represent the climate index $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} plus a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
+We represent \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
+To create an annual time series, we average the October-December values of the NINO3 index for each year.
+\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
+In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, \ie an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
+
+\subsection{Projecting Climate Risk over the Future $M$ years}\label{sec:methods-estimating}
+
+Once a synthetic streamflow sequence has been generated, we evaluate the identifiability and predictability of the dominant climate modes by fitting the sequence to statistical models and creating probabilistic projections of the future.
+We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
+One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
+We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
+For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations, denoted $Q_{\text{hist}}$.
+Once a statistical model is fit to $Q_{\text{hist}}$, then $K=\num{2000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
+
+We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
+In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
+The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors for regularization.
+The full model is given in \cref{eq:log-normal-stationary}.
+
+Next, we expand this stationary model to incorporate secular change.
+Many studies have done this by regressing certain parameters of the model on time \citep[see][]{Salas:2018ge}.
+We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
+
+Finally, we consider a model which explicitly represents \acrlong{lfv}.
+A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie hidden) states $S(t)$ \citep{Rabiner:1986jk}.
+The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
+We fit  streamflow sequences $Q_\text{hist}$ simulated using the scenarios in Section 2.1, using a \gls{hmm} with two states.
+The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
+This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
+For simplicity, we fit only a two-state \gls{hmm} to each sequence.
+Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.
+
+\subsection{Evaluating Fitting Models}\label{sec:methods-evaluating}
+
+Both estimation bias and estimation uncertainty affect the utility of a climate risk projection, as illustrated in \cref{fig:conceptual-bias-variance}.
+An instrument designed based on overestimated variance and positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
 Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
@@
 
-In this paper we present a set of stylized experiments to assess how well one can identify and predict risk associated with cyclical and secular climate signals for the design life ($M$ years) and the probability of over- or under-design of a climate adaptation strategy based on these projections.
-We consider different temporal structures for the underlying risk which encompass quasi-periodic, regime-like, and secular change, as well as simple statistical models for estimating this risk from an $N$-year historical record.
-The relative importance of estimating the short- or long-term risk associated with these extremes depends on the design life $M$, but the potential to understand and predict these different types of variability depends on the informational uncertainty in the $N$-year historical record.
-Though we use floods as an example, the framework also applies to other forms of climate extremes.
-
-% -----------------------------------------------------------------------------
-% METHODS
-% -----------------------------------------------------------------------------
-
-\section{Methods}\label{sec:methods}
-
-We use a statistical framework to
-\begin{enumerate*}[label= (\roman*)]
-  \item generate synthetic sequences of annual-maximum floods;
-  \item fit these sequences to probabilistic models that parameterize time in different ways; and
-  \item evaluate the performance of these models as a function of $M$ and $N$ (\cref{fig:methods-summary})
-\end{enumerate*}.
-
-As shown in \cref{fig:methods-summary}, we first specify $N$, $M$, the flood threshold (\eg, the design level of a structure or the level at which an insurance contract is triggered), the method for generating synthetic streamflow sequences, and the method for estimating flood risk given an $N$-year historical record.
-The methods for generating synthetic streamflow sequences are discussed in \cref{sec:methods-generating} and for estimating flood risk in \cref{sec:methods-estimating}.
-For each scenario of climate risk structure, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
-We repeat the simulations $J = 2000$ times for each combination of experiment parameters to estimate the bias and variance of the statistical estimate (\cref{sec:methods-evaluating}).
-The scenarios considered are summarized in \cref{sec:methods-experiments}.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{flowchart.pdf}
-  \caption{
-    Framework for assessing flood risk from synthetic streamflow sequences.
-    Red boxes indicate parameters that are chosen for each experiment, white boxes calculated quantities, and blue boxes the final values used to evalute the model fit.
-  }\label{fig:methods-summary}
-\end{figure}
-
-\subsection{Sampling Climate Risk}\label{sec:methods-generating}
-
-We consider two approaches for modeling the influence of quasi-periodic variability and secular climate change on flood risk.
-For each, we take the annual-maximum sequences to follow a log-normal distribution, conditional on a location parameter which may vary in time:
-\begin{equation} \label{eq:lognormal}
-  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
-\end{equation}
-We further assume a constant coefficient of variation of the log streamflow, $\sigma(t) = \xi \mu(t)$, and apply a lower threshold on the standard deviation $\sigma(t) \geq \sigma_\text{min} > 0$.
-The procedure to generate a single sequence of streamflow is thus:
-\begin{enumerate}
-  \item choose $t_0$, $M$, and $N$ so that the time is $\vb{t}=t_0-N+1, \ldots, t_0+M$
-  \item use one of the models described below to simulate $\mu(t)$
-  \item calculate $\sigma(t) = \xi \mu(t)$
-  \item for each time $t \in \vb{t}$, draw $Q(t)$ stochastically following \cref{eq:lognormal}.
-\end{enumerate}
-We consider two approaches for simulating $\mu(t)$.
-
-The first approach considers $\mu(t)$ as a function of a variable which explicitly represents quasi-periodic climate variability.
-We use an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} plus a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
-We also allow for the inclusion of secular trends.
-We represent \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
-To create an annual time series, we average the October-December values of the NINO3 index for each year.
-\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
-To generate $\mu(t)$, a sequence of length $M+N$ is randomly chosen from the \SI{20000}{year} NINO3 sequence.
-Next, $\mu(t)$ is calculated from the NINO3 time series, denoted as $x(t)$, as
-\begin{equation}\label{eq:nino3}
-  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
-\end{equation}
-where $\beta$ controls the sensitivity to the NINO3 variation and $\gamma$ controls the magnitude of the secular trend.
-
-The second model to create sequences of annual-maxima is a two-state Markov chain model to explicitly enforce persistence and regime structure in $\mu(t)$.
-A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
-The transition matrix is
-\begin{equation}
-  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
-\end{equation}
-The transition matrix is first used to generate a sequence of states $S(t)$.
-The value $\mu(t)$ depends only on $S(t)$ and on time itself:
-\begin{equation}
-  \mu(t) = \begin{cases}
-    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
-    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
-  \end{cases}
-\end{equation}
-For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
-We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
-
-We use the NINO3 and Markov Chain models described above to create three scenarios:
-\begin{enumerate*}[label= (\roman*)]
-  \item \gls{lfv} only;
-  \item secular change only; and
-  \item \gls{lfv} plus secular change.
-\end{enumerate*}
-\gls{lfv} is added to the NINO3 model through $\beta$, which increases the dependence on a climate signal which itself exhibits \gls{lfv}, and to the Markov chain model by making the two states more distinct (increasing $\mu_1 - \mu_2$, or $\gamma_1 - \gamma_2$).
-Secular change is applied by increasing $\gamma$, which gives the time dependence.
-The parameter values used to create these three scenarios are listed in \cref{sec:supp-methods}.
-
-\subsection{Projecting Climate Risk over the Future $M$ years}\label{sec:methods-estimating}
-
-We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
-One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
-We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid, and note that the key conclusions hold for more sophisticated \gls{ffa} methods using statistical or dynamical models.
-
-For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations, denoted $Q_{\text{hist}}$.
-Once a statistical model is fit to $Q_{\text{hist}}$, $K=\num{2000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
-
-We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
-In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
-The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors for regularization.
-The full model is given in \cref{eq:ln2-stationary}.
-
-Next, we expand this stationary model to incorporate secular change.
-Many studies have done this by regressing certain parameters of the model on time \citep{Merz:2014gf, Salas:2018ge}.
-We consider an extension of the stationary LN2 model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:ln2-trend}.
-
-Finally, we consider a model which explicitly represents low-frequency variability.
-A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie{} hidden) states $S(t)$ \citep{Rabiner:1986jk}.
-The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
-We fit  streamflow sequences $Q_\text{hist}$ simulated using the scenarios in Section 2.1, using a \gls{hmm} with two states.
-The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
-This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
-For simplicity, we fit only a two-state \gls{hmm} to each sequence.
-Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.
-
-\subsection{Evaluating Fitting Models}\label{sec:methods-evaluating}
-
-For a given choice of $M$, $N$, generating model, and fitting model, we generate a streamflow sequence of length $M+N$ years.
-The fitting model is applied to the synthetic streamflow sequence's $N$-year ``historical record'', and $K=2000$ posterior simulations of future flows are then drawn from the fitted model.
-
+We evaluate both the estimation bias and estimation uncertainty.
+For a given choice of $M$, $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=2000$ posterior simulations of future flows.
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
-We then compute the variance of these $K$ estimates of $\hat{p}_T$.
+We then compute the variance of these $K$ estimates of their variance.
 We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
-
 Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we repeat this analysis for $J=1000$ different streamflow sequences.
 These are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may be different between the $J$ sequences.
-In \cref{sec:results} we report the expected bias and variance by averaging over the $J$ sequences.
 
@@
 Projections with each of the three fitting models are also shown.
-In the remainder of this section we present a more systmematic analysis of each of these three cases.
+In the remainder of this section we present a more systematic analysis of each of these three cases.
 \begin{figure}
@@
 
-The LN2 linear trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
+The log-normal trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
 The variance of these estimates is also large.
 This again highlights the difficulty of identifying the magnitude of secular change, which strongly controls climate risk far into the future, from a single noisy time series.
-By contrast, the stationary LN2 model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
+By contrast, the stationary log-normal model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
 As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
@@
     Expected estimation bias and variance for sequences generated from the NINO3 model with secular change only (no \gls{lfv}).
-    Sequences were fit to each of three statistical models (colulmns) for different $N$ and $M$ ($x$ and $y$ axis, repsectively).
+    Sequences were fit to each of three statistical models (columns) for different $N$ and $M$ ($x$ and $y$ axis, respectively).
     Top row shows estimation bias and bottom row shows log standard deviation of estimates.
@@
 If the analyst could know \emph{a priori} that secular change is present in a time series, and if $M$ is long, then the use of a model that captures this secular trend is essential.
-Here the LN2 linear trend model has the advantage of being correctly specified (both the generating and fitting processes assume a log-normal distribution conditional on a linear time trend), which is generally not the case in the real world \citep{Serinaldi:2015bq, Montanari:2014hl}.
+Here the log-normal linear trend model has the advantage of being correctly specified (both the generating and fitting processes assume a log-normal distribution conditional on a linear time trend), which is generally not the case in the real world \citep{Montanari:2014hl,Serinaldi:2015bq}.
 As a result, in real-world settings longer $N$ may be required to identify trends whose exact form is not known.
@@
 The models were assessed on their performance over a complete future period of length $M$ years; as discussed in \cref{sec:introduction} the bias and variance computed together represent the likelihood for over- or under-design of the instrument in a given climate scenario (see \cref{fig:conceptual-bias-variance}).
-If more complex methods, such as analysis of \gls{ipcc} scenarios, are used, the sample size $N$ used in our experiments may be straightforwardly interpreted as a surrogate for the total informational uncertainty in the analysis.
+If more complex methods, such as model chains based on \glspl{gcm} and hydrologic models, are used, the sample size $N$ used in our experiments may be straightforwardly interpreted as a surrogate for the total informational uncertainty in the analysis.
 
@@
 
-For these short-$M$ projects, cyclical climate variability is dominant over the project planning period \citep{Hodgkins:2017hw, Jain:2001hz}.
-Analysis of the former using the wavelet transformation \citep{Torrence:1998jp} shows dominant spectral peaks at 8, 20, and 64 years, and of the latter at 2 and 15 years.
-This is illustrated in \cref{sec:observed-lfv}, which shows the cyclical variability of two real-world hydroclimate systems: a 500 year drought reconstruction from the \gls{lbda} \citep{Cook:2010bz} and a \SI{107}{year} record of naturalized daily river stage on the American River at the Folsom Dam.
-The long periods and high amplitudes of the oscillations present in these time series highlight the importance of adapting not only to secular change but also, simultaneously, to \gls{lfv}.
-Interestingly, our results show that even though the two-state \gls{hmm} is an imperfect analog for the \gls{lfv} represented in the \gls{enso} process, it performs quite well for short $M$.
-This highlights that even imperfect estimates of \gls{lfv} can support suffessful climate adaptation.
+For projects where $M$ is sufficiently short, cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
+The long periods and high amplitudes of the oscillations shown in \cref{fig:observed-lfv} and discussed in \cref{sec:intro-lfv} highlight the importance of adapting not only to secular change but also, simultaneously, to \gls{lfv}.
 
@@
 
-As described in \cref{sec:methods}, we implement the stationary and trend lognormal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
-The full model, including priors, for the stationary lognormal model is given by \cref{eq:ln2-stationary}:
-\begin{align}\label{eq:ln2-stationary}
+As described in \cref{sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
+The full model, including priors, for the stationary log-normal model is given by \cref{eq:log-normal-stationary}:
+\begin{align}\label{eq:log-normal-stationary}
   \begin{split}
     \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
-    \mu &\sim \normal \qty(0, 10) \\
-    \sigma &\sim \normal^+ \qty(0, 2)
+    \mu &\sim \normal \qty(7, 1.5) \\
+    \sigma &\sim \normal^+ \qty(1, 1)
   \end{split}
@@
 where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
-The full model, including priors, for the trend lognormal model is given by \cref{eq:ln2-trend}:
-\begin{align}\label{eq:ln2-trend}
+The full model, including priors, for the trend log-normal model is given by \cref{eq:log-normal-trend}:
+\begin{align}\label{eq:log-normal-trend}
   \begin{split}
@@
     \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
-  \mu_0 & \sim \normal \qty(0, 10) \\
+    \mu_0 & \sim \normal \qty(7, 1.5) \\
     \beta_\mu & \sim \normal \qty(0, 0.1) \\
@@
   \caption{
-    Parameters of the NINO3 model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
+    Parameters of \cref{eq:nino3} for the \gls{lfv} only, secular only, and secular plus \gls{lfv} scenarios.
   }\label{tab:nino-stationary}
 \end{table}
+
+\section{Sensitivity to climate generating mechanism}\label{sec:markov-generating}
+
+In the main text we sampled climate risk by considering dependence on a single climate index, as shown in \cref{eq:nino3}.
+Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
+A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
+The transition matrix is given by
+\begin{equation}
+  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
+\end{equation}
+The transition matrix is first used to generate a sequence of states $S(t)$.
+The value $\mu(t)$ depends only on $S(t)$ and on time itself:
+\begin{equation}
+  \mu(t) = \begin{cases}
+    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
+    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
+  \end{cases}
+\end{equation}
+For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
+We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
+Parameters for the three experiments are given in \cref{tab:markov-stationary}.
+Results for these experiments are shown in \cref{fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.
 \begin{table}[ht]
@@
 \end{table}
-
-\section{Supplemental Results}
-
-\subsection{Observed Low-Frequency Hydroclimate Variability}\label{sec:observed-lfv}
-
-\Acrfull{lfv} of the climate is present in a wide range of real-world time series \citep{Hodgkins:2017hw, Swierczynski:2012km, Cook:2010bz}.
-This variability is illustrated in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet \citep{Torrence:1998jp} average spectrum for both.
-Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
-This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
-This clearly illustrates the strong observed multidecadal variability.
-For further discussion of the \gls{lbda} time series we refer the reader to \citet{Cook:2010bz}.
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{observed-lfv.pdf}
-  \caption{
-    Hydroclimate time series vary on many time scales.
-    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
-    High numbers indicate more severe drought.
-    A 20-year running mean is also shown in blue.
-    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
-    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
-    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
-    (d) wavelet spectrum, like (c), for the American River data.
-  }\label{fig:observed-lfv}
-\end{figure}
-
-\subsection{Results for the Markov Model}
-
-In \cref{sec:results} all results given use the NINO3 model to generate synthetic flood sequences.
-In this section equivalent results are given for sequences generated with the two-state Markov chain model described in \cref{sec:methods}.
-
 \begin{figure}
@@
   }\label{fig:lfv-markov-bias-variance}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
+  \caption{
+    Secular change: as \cref{fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
+  }\label{fig:secular-only-markov-bias-variance}
 \end{figure}

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Update again before sending to Dave for comments

@@
 Projections with each of the three fitting models are also shown.
+This figure highlights that even where projections of average streamflow are unbiased, if the spread is too large then projection of the threshold exceedance probability may be too large.
+
 In the remainder of this section we present a more systematic analysis of each of these three cases.
@@
 
-Evaluating and implementing investments for climate risk mitigation demands making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
+Evaluating and implementing investments for climate risk mitigation involves making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
 The success of this prediction will depend on the identifiability of different signals from an $N$-year sample, the time scales of \gls{lfv} relative to the project life of the instrument, and the uncertainty associated with the estimation.
@@
 
+For projects where $M$ is sufficiently short, cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
+The long periods and high amplitudes of the oscillations shown in \cref{fig:observed-lfv} and discussed in \cref{sec:intro-lfv} highlight the \emph{importance} of identifying and predicting this variability.
+However, one's \emph{ability} to do so depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
+If sufficient information is not available, then simple models which represent fewer processes may be preferred.
+
+For projects with longer $M$, it becomes necessary to consider the role of secular change.
+As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
+As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.\footnote{Let's add a citation here from the climate literature}
+We show that mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases, an important caution for the field of ``nonstationary'' flood frequency analysis as noted by \citet{Montanari:2014hl} and \citet{Serinaldi:2015bq}.
+
 The stationary, linear trend, and \gls{hmm} models considered for estimation present an idealized alternative to more complex, physically based models for future risk projection.
-They were chosen to represent the ability of simple models to capture two features of interest: \gls{lfv} and secular trends.
-The models were assessed on their performance over a complete future period of length $M$ years; as discussed in \cref{sec:introduction} the bias and variance computed together represent the likelihood for over- or under-design of the instrument in a given climate scenario (see \cref{fig:conceptual-bias-variance}).
-If more complex methods, such as model chains based on \glspl{gcm} and hydrologic models, are used, the sample size $N$ used in our experiments may be straightforwardly interpreted as a surrogate for the total informational uncertainty in the analysis.
+For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf}.
+We suggest, then, that the sample size $N$ defined in our experiments may be straightforwardly interpreted as a measure of the total informational uncertainty in the analysis; as $N$ increases, informational uncertainty decreases.
+
+Similarly, real-world climate adaptation plans will typically include multiple instruments which may be placed in different locations and times in a sequential fashion.
+Even if the planning period of a portfolio is long, the individual instruments within the portfolio may have short planning periods.
+Since \cref{sec:results} shows that the bias and variance of climate risk projections tend to increase with $M$, the total bias and variance associated with sequencing 20 consecutive $M=\SI{5}{year}$ projects will be less than that associated with making a single $M=\SI{100}{year}$ project.
+This effect will be exacerbated by the fact that if the first $M=\SI{5}{year}$ project is based on estimates with informational uncertainty $N$, the second will have $N+5$, the third $N+10$, and so on.
 
@@
 The application of a positive discount rate, as is typical in the public and private sectors, emphasizes the importance of predicting near-term risk.
-Even ignoring this, however, if the uncertainty associated with the performance of the instrument increases into the future (as shown in \cref{fig:conceptual-sketch}) then adaptation strategies with a short $M$ are preferred.
-This is consistent with recent emphasis on the importance of making decisions that are robust to a wide range of possible future climates \citep[\ie,][]{Haasnoot:2013im, Poff:2015jn}.
+Consequently, projects with long $M$ must overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.\footnote{Should we create a new section called summary for these last two paragraphs?}
 
-For projects where $M$ is sufficiently short, cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
-The long periods and high amplitudes of the oscillations shown in \cref{fig:observed-lfv} and discussed in \cref{sec:intro-lfv} highlight the importance of adapting not only to secular change but also, simultaneously, to \gls{lfv}.
+\section{Summary}
 
-For projects with longer $M$, it becomes necessary to consider the role of secular change.
-As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.
-We find that when making risk projections that consider secular change, mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases.
-This is particularly important given that in real-world settings the true form of the trend is not known and thus ``nonstationary'' projections which assume a particular type of trend are likely to perform poorly \citep{Montanari:2014hl, Serinaldi:2015bq}.
-
+In this paper we considered how the temporal structure of the climate affects the potential for successful prediction over a finite $M$-year future period.
+We began with three premises, or observations, about the nature of climate risk:
+\begin{enumerate*}[label= (\roman*)]
+  \item that planning decisions are made with finite horizons;
+  \item that climate risk varies on many scales; and
+  \item that the dominant processes that control this risk depend on the planning period itself.
+\end{enumerate*}
 Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, as shown in \cref{fig:observed-lfv}.

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
send to Manu for comments

@@
 \documentclass[
-  %draft,
-  %linenumbers
+  draft,
+  linenumbers
 ]{agujournal2018}
@@
 This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
-Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs,\footnote{cite the environmental costs of big infrastructure projects} and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have slowed their deployment.
+Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have slowed their deployment.
 Instead, actors such as New York City have turned to a combination of structural, operational, and financial instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
@@
 Typical planning periods may vary from $M=1$ year or less for a financial contract to $M=100$ years or longer for a structural instrument, as illustrated in \cref{tab:real-world-M}.
-The planning period can also be interpreted as the finite period over which \gls{cba} is conducted when assessing the project.\footnote{we should add a citation about \gls{cba} here, perhaps a technical spec from the Army Corps?}
+The planning period can also be interpreted as the finite period over which \gls{cba} is conducted when assessing the project.
 
@@
 Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
-Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on river floods, droughts, hurricanes, urban flooding, and many other climate hazards are well documented.\footnote{Add citations of \gls{acc} influences on climate hazards}
+Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on river floods, droughts, hurricanes, urban flooding, and many other climate hazards are well documented \citep[\eg,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
 Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
@@
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
-We then compute the variance of these $K$ estimates of their variance.
+We then compute the variance of these $K$ estimates.
 We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
@@
 Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke}.
-The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk}\footnote{Dear reviewers: this GitHub repository is currently not publicly available, but codes are provided in the online supplemental information.
-The repository will be made publicly available if and when this work is published.} or in the online supplemental material of this paper.
+The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.
 
@@
 
-Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability of the \gls{lfv}, denoted $T_{\text{LFV}}$.
+Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability $T_{\text{LFV}}$ of the \gls{lfv}.
 As shown in \cref{fig:enso-ts}, the NINO3 time series is most active in the \SIrange{3}{6}{year} band.
@@
 In the real world, however, many hydroclimate time series vary at multidecadal and longer frequencies.
-In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic frequencies $T_{\text{LFV}}$ may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
+In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic frequencies $T_{\text{LFV}}`$ may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
 This in turn requires a longer observational record $N$ in order to identify and predict these different signals.
@@
 As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.\footnote{Let's add a citation here from the climate literature}
+As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.
 We show that mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases, an important caution for the field of ``nonstationary'' flood frequency analysis as noted by \citet{Montanari:2014hl} and \citet{Serinaldi:2015bq}.
@@
 
-The climate adaptation decisions which this analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate.
+The climate adaptation decisions which this analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate \citep{sodastrom:1999,powers:2003}.
 The application of a positive discount rate, as is typical in the public and private sectors, emphasizes the importance of predicting near-term risk.
-Consequently, projects with long $M$ must overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.\footnote{Should we create a new section called summary for these last two paragraphs?}
+Consequently, projects with long $M$ must overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.
 

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Add a figure showing M and N

@@
 This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
-A detailed analysis of these time series is beyond the scope of this paper, but we note that these findings are illustrative of other analyses of climatic \gls{lfv} \citep{Jain:2001hz,Kiem:2002kq,Cook:2010bz,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw,Cassou:2018cs}.
+A detailed analysis of these time series is beyond the scope of this paper, but we note that these findings are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Jain:2001hz,Kiem:2002kq,Cook:2010bz,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw,Cassou:2018cs}.
 
@@
 Even with a perfect climate model, these uncertainties will be large.
-By contrast, our perfect climate model may usefully inform climate hazard over a three-year insurance contract with much less associated uncertainty.
+By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.
 
@@
 For each scenario, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
-We repeat the simulations $J = 2000$ times for each combination of experiment parameters to estimate the bias and variance of the statistical estimate (\cref{sec:methods-evaluating}).
+We repeat the simulations $J = 2000$ times for each combination of experiment parameters to obtain estimates of the expected bias and variance for each scenario given $M$ and $N$ (\cref{sec:methods-evaluating}).
 
@@
 The first step is to sample climate risk by generating synthetic streamflow sequences.
-To do this, we take the annual-maximum sequences to follow a log-normal distribution, conditional on a location parameter which may vary in time:
+To do this, we model  annual-maximum flood peaks following a log-normal distribution, conditional on a location parameter which may vary in time:
 \begin{equation} \label{eq:log-normal}
@@
 
-We represent the climate index $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} plus a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
-We represent \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
+We represent the climate state variable $x(t)$ by an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
+We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
 To create an annual time series, we average the October-December values of the NINO3 index for each year.
@@
 We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
-For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations, denoted $Q_{\text{hist}}$.
-Once a statistical model is fit to $Q_{\text{hist}}$, then $K=\num{2000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
+For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
+Once a statistical model is fit to these ``observations'', then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
 
@@
 
-Next, we expand this stationary model to incorporate secular change.
-Many studies have done this by regressing certain parameters of the model on time \citep[see][]{Salas:2018ge}.
+Next, we modify this stationary model to incorporate secular change.
+Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
 We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
+This model gives a lower bound on total informational uncertainty because it correctly represents the form of a trend, whereas in real-world analyses the form of the trend is unknown.
 
@@
 
-Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, Schreiber:2017tg, Jones:2001uv, Hoyer:2017hs}.
+Computation was carried out in the python programming language, making particular use of the matplotlib, numpy, pandas, pomegranate, scipy, and xarray libraries for scientific computing \citep{Hunter:2007ih, vanderWalt:2011dp, McKinney:2010un, schreiber:2017, Jones:2001uv, Hoyer:2017hs}.
 Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
@@
 For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf}.
+In places where long records are not available, for example, information from nearby locations, reanalysis products, or \acrshort{gcm} runs may increase the amount of available information \citep{Merz:2008eh}, effectively increasing $N$.
 We suggest, then, that the sample size $N$ defined in our experiments may be straightforwardly interpreted as a measure of the total informational uncertainty in the analysis; as $N$ increases, informational uncertainty decreases.
@@
 Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, as shown in \cref{fig:observed-lfv}.
-Identifying and modeling these different physical signals requires a large amount of information, represented here by the length of the $N$-year historical record.
-In places where long records are not available, however, information from nearby locations, reanalysis products, or \acrshort{gcm} runs may increase the amount of available information \citep{Merz:2008eh}, effectively increasing $N$.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{M-N-Sketch.pdf}
+  \caption{
+    The importance of predicting different signals, and the identifiability and predictability of the signals, depends on the degree of informational uncertainty ($N$) and the project planning period ($M$).
+  }\label{fig:m-n-sketch}
+\end{figure}
+As shown in \cref{fig:m-n-sketch}, the importance of predicting, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
+When $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and increases in model complexity only add uncertainty.
+If $M$ is short and $N$ is long, however, the \gls{lfv} in the system may be skillfully predicted, thereby reducing the cost of such an instrument.
+If $N$ remains long but $M$ increases, then informational uncertainty may be reduced, but intrinsic uncertainty will remain, and may be large.
+Finally, if $M$ is long but $N$ is short, then the analysis will lead to the ``danger zone'' of extrapolation and a high potential for misleading estimates.
 

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
fix language in discussion of new figure

@@
 \end{figure}
-As shown in \cref{fig:m-n-sketch}, the importance of predicting, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
-When $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and increases in model complexity only add uncertainty.
-If $M$ is short and $N$ is long, however, the \gls{lfv} in the system may be skillfully predicted, thereby reducing the cost of such an instrument.
-If $N$ remains long but $M$ increases, then informational uncertainty may be reduced, but intrinsic uncertainty will remain, and may be large.
+We also showed that the importance of predicting, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
+\Cref{fig:m-n-sketch} shows that when $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and more complex models  only add uncertainty.
+As $N$ increases, however, the potential for skillful prediction of \gls{lfv} over the short planning period increases, thereby reducing the cost of a flexible or adaptive instrument.
+If $M$ is large, \ie a permanent plan is considered, then informational uncertainty may be addressed with sufficient information (large $N$), but large intrinsic uncertainty may remain.
 Finally, if $M$ is long but $N$ is short, then the analysis will lead to the ``danger zone'' of extrapolation and a high potential for misleading estimates.

### 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
clean up conclusions and acknowledgements

@@
 \end{figure}
-We also showed that the importance of predicting, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
+We also showed that the importance of predicting different climate signals, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
 \Cref{fig:m-n-sketch} shows that when $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and more complex models  only add uncertainty.
@@
 The authors thank John High of the United States Army Corps of Engineers for providing the naturalized daily streamflows at the Folsom Dam.
-JDG thanks the NSF GRFP program for support (grant DGE 16-44869: ``Understanding \& Predicting Climate Drivers of Extreme, Mid-latitude River Floods'').
-DJF thanks SERDP program for support (grant 2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate'').
+JDG thanks the NSF GRFP program (grant DGE 16-44869: ``Understanding \& Predicting Climate Drivers of Extreme, Mid-latitude River Floods'') and SERDP program (grant 2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate'') for support.
 

### 2019-01-09 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Incorporate manu comments

@@
 Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
-Each has its own planning period, which may be separate from the planning horizon of the portfolio as a whole.
+Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
 This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this long-term plan may be achieved through a series of flexible and adaptive instruments with short individual planning periods.
 The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
+
+A critical point for decision-making under uncertainty is that if precise climate information over the next few years is available, then this can significantly alter the choice between a large, long-duration project which may take over a year to construct, and a sequence of smaller, short duration projects that can be executed quickly.
+For example, if there is credible information that the risk to be mitigated will be much higher in the next few years, then a more costly project would be justified under traditional \gls{cba}.
+However, if the construction period were too long, a financial risk mitigation instrument might be needed in the immediate term to cover potential losses before the large project is constructed.
+Conversely, if the near term risk is projected to be low, then deferral of the large, expensive project may be warranted, and smaller or nonstructural alternatives may be preferred.
+If the risk changes in a cyclical manner, then these decisions become even more nuanced, but there may be an opportunity for significantly improving adaptation using a portfolio risk management approach.
+Thus, the precision with how near term and long term directions of climate risk evolve plays directly into adaptation choice.
 
@@
 
+For the sake of presentation we motivate our results through a simple flood risk example.
+The points we seek to make apply to other hydroclimatic risks, and in particular those which may be easily characterized through a time series of annual maxima or minima of the process.
 The first step is to sample climate risk by generating synthetic streamflow sequences.
@@
 \Cref{fig:m-n-sketch} shows that when $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and more complex models  only add uncertainty.
-As $N$ increases, however, the potential for skillful prediction of \gls{lfv} over the short planning period increases, thereby reducing the cost of a flexible or adaptive instrument.
+As $N$ increases, however, the potential for skillful prediction of \gls{lfv} over the short planning period increases, potentially favoring implementation of a flexible or adaptive instrument.
 If $M$ is large, \ie a permanent plan is considered, then informational uncertainty may be addressed with sufficient information (large $N$), but large intrinsic uncertainty may remain.

### 2019-01-10 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
some last revisions for collaborator review

--- 2019-01-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
@@
 This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
-Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have slowed their deployment.
+Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have recently slowed their implementation.
 Instead, actors such as New York City have turned to a combination of structural, operational, and financial instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
@@
 Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
-Each has its own planning period, which may be separate from the planning horizon of the portfolio as a whole.
-This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this long-term plan may be achieved through a series of flexible and adaptive instruments with short individual planning periods.
+Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
+This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
 The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
+
+The availability of precise climate information in the near future may significantly alter the choice between a large, long-duration instrument and a sequence of smaller, short duration instruments that can be executed quickly.
+For example, if above-average climate risk is projected over the next few years, a more costly project might be justified.
+However, in the plausible case of a long construction period for the large, permanent instrument, a financial risk mitigation instrument might be needed in the immediate term to cover potential losses before the large project is completed.
+Conversely, if the near-term risk is projected to be low, then deferral of the large, potentially expensive instrument may be warranted.
+These cases highlight how the precision of short- and long-term climate risk projections plays directly into climate adaptation.
 
@@
 This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
-A detailed analysis of these time series is beyond the scope of this paper, but we note that these findings are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Jain:2001hz,Kiem:2002kq,Cook:2010bz,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw,Cassou:2018cs}.
+A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
 
@@
 
-Assessing the utility of a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
+Evaluating a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
 Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
@@
 As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
+Although \cref{fig:conceptual-sketch} focuses on physical processes, similar conclusions would also be valid for the socioeconomic processes which drive exposure to floods and other hydroclimate hazards.
 
@@
 For each scenario, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
-We repeat the simulations $J = 2000$ times for each combination of experiment parameters to obtain estimates of the expected bias and variance for each scenario given $M$ and $N$ (\cref{sec:methods-evaluating}).
+We repeat the simulations $J = 1000$ times for each combination of experiment parameters to obtain estimates of the expected bias and variance for each scenario given $M$ and $N$ (\cref{sec:methods-evaluating}).
 
-We caution the reader that the models for sampling climate risk and for statistically projecting future risk were chosen for their intuitive interpretation, rather than their general validity \citep[see][for a thoughtful discussion of the value of simple models]{Held:2005cj}.
+We caution the reader that the models for sampling climate risk (\cref{sec:methods-generating}) and for statistically projecting future risk (\cref{sec:methods-estimating}) were chosen for their intuitive interpretation, rather than their general validity \citep[see][for a thoughtful discussion of the value of simple models]{Held:2005cj}.
 We do not, in general, endorse these models for practical use but instead argue that the conclusions drawn from these simple models may be straightforwardly applied to more complex and realistic models.
@@
 
+We illustrate our findings with a simple flood risk example.
+The conclusions drawn apply to other hydroclimate hazards, and in particular those which may be easily characterized through a time series of annual maxima or minima of the process.
 The first step is to sample climate risk by generating synthetic streamflow sequences.
-To do this, we model  annual-maximum flood peaks following a log-normal distribution, conditional on a location parameter which may vary in time:
+To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
 \begin{equation} \label{eq:log-normal}
@@
 \end{equation}
-This flexible framing describes the \begin{enumerate*}[label= (\roman*)]
+This formulation describes the \begin{enumerate*}[label= (\roman*)]
   \item \gls{lfv} only;
@@
 \end{equation}
-where $x(t)$ represents a climate time series which itself exhibits \gls{lfv}.
+where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
 This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
-Thus, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
+Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
 The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.
 
-We represent the climate state variable $x(t)$ by an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
+We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
 We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
@@
 In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
-The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors for regularization.
-The full model is given in \cref{eq:log-normal-stationary}.
+The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
+The full model, including priors, is given in \cref{eq:log-normal-stationary}.
 
@@
 We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
-This model gives a lower bound on total informational uncertainty because it correctly represents the form of a trend, whereas in real-world analyses the form of the trend is unknown.
+This model gives a lower bound on total informational uncertainty because it correctly represents the form of the trend, which is known, whereas in real-world analyses the form of the trend is unknown.
 
@@
 We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
-Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we repeat this analysis for $J=1000$ different streamflow sequences.
-These are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may be different between the $J$ sequences.
+Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we average the bias and variance across $J=1000$ different streamflow sequences to compute expected values of both.
+These sequences are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may differ between the $J$ sequences.
 
@@
 Wavelet analysis was conducted using the WaveletComp package \citep{Roesch:wlBQQoIs} in the R programming language.
-Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke}.
-The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.
+Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke} using the No U-Turn Sampler \citep{Hoffman:2011wm,Betancourt:2017vd}.
+The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.\footnote{Dear Reviewers: our codes are not presently available on GitHub, as we have chosen to keep the repository private until the article is accepted for publication, but they are available to you through the online supplemental material for this submission.}
 
@@
 The variance of these estimates is also large.
-This again highlights the difficulty of identifying the magnitude of secular change, which strongly controls climate risk far into the future, from a single noisy time series.
+This again highlights the difficulty of fitting complex models for estimating risk when informational uncertainty is large.
 By contrast, the stationary log-normal model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
@@
 This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[\ie,][]{Muller:2014fc}.
-However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace any bias reductions.
+However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace the utility of any bias reductions.
 
@@
   \caption{
-    Expected estimation bias and variance for sequences generated from the NINO3 model with secular change only (no \gls{lfv}).
+    Expected estimation bias and variance for sequences generated with secular change only (no \gls{lfv}).
     Sequences were fit to each of three statistical models (columns) for different $N$ and $M$ ($x$ and $y$ axis, respectively).
@@
 
-If the analyst could know \emph{a priori} that secular change is present in a time series, and if $M$ is long, then the use of a model that captures this secular trend is essential.
+If the analyst could know \emph{a priori} that secular change is present in a time series, and if $M$ is long, then the use of a complex model which represents the processes causing this change is required.
 Here the log-normal linear trend model has the advantage of being correctly specified (both the generating and fitting processes assume a log-normal distribution conditional on a linear time trend), which is generally not the case in the real world \citep{Montanari:2014hl,Serinaldi:2015bq}.
 As a result, in real-world settings longer $N$ may be required to identify trends whose exact form is not known.
-Here, $N$ implicitly refers to the uncertainty in trend estimation and may represent an ``equivalent sample size'' if \gls{ipcc}-like projections are used.
 Alternatively, if $M$ is small then it may be reasonable to use a stationary estimate, since the bias will be small and the variance substantially lower.
@@
 
-Next, in the idealized case where  \gls{lfv} is present but there is no secular change, the identification of nonexistent trends may lead to gross over-estimation of true risk through an increase in the variance of the estimated risk.
+We next turn to the idealized case where  \gls{lfv} is present but there is no secular change in the system.
+\Cref{fig:lfv-nino3-bias-variance} highlights that identification of nonexistent trends from limited data may lead to gross over-estimation of true risk through an increase in the variance of the estimated risk.
+As expected, the stationary log-normal model performs well overall, with low bias and low variance.
+The \gls{hmm} actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the multimodal distribution that emerges from dependence on the \gls{enso} index, which exhibits several regimes (see \cref{fig:enso-ts}).
+By contrast, the linear trend model performs poorly for low $N$ and high $M$ because a positive probability is assigned to the existence of a positive trend.
 
-Results for this scenario are shown in \cref{fig:lfv-nino3-bias-variance} for the NINO3 model and \cref{fig:lfv-markov-bias-variance} for the two-state Markov model.
-As expected, the stationary model (right column) performs well overall, with low bias and low variance.
-The \gls{hmm} (left column) actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the true multimodal distribution that emerges from dependence on the NINO3 index which exhibits several regimes.
-By contrast, the linear trend model (middle column) performs poorly for low $N$ and high $M$ because a positive probability is assigned to the existence of a positive trend.
-
-Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability $T_{\text{LFV}}$ of the \gls{lfv}.
-As shown in \cref{fig:enso-ts}, the NINO3 time series is most active in the \SIrange{3}{6}{year} band.
-If $T_{\text{LFV}} \ll M$, then a stationary assumption may provide reasonable estimates, and fewer observations may be required (shorter $N$).
+Of particular relevance to analysis of real-world data sets is the ratio of the project planning period $M$ to the characteristic periods of variability of the \gls{lfv}.
+If this period is much larger than $M$, then a stationary assumption may provide reasonable estimates, and fewer observations may be required (shorter $N$).
+As shown in \cref{fig:enso-ts}, the \gls{enso} time series is most active in the \SIrange{3}{6}{year} band.
 In the real world, however, many hydroclimate time series vary at multidecadal and longer frequencies.
-In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic frequencies $T_{\text{LFV}}`$ may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
+In this case, as illustrated in \cref{fig:conceptual-sketch}, the characteristic periods may be as large or larger than $M$, particularly if multidecadal modes such as the \gls{pdo} or \gls{amo} are involved, and the \gls{lfv} must therefore be estimated explicitly.
 This in turn requires a longer observational record $N$ in order to identify and predict these different signals.
@@
   \caption{
-    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and \gls{lfv} from the NINO3 model.
+    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with zero secular change and strong \gls{lfv}.
   }\label{fig:lfv-nino3-bias-variance}
@@
 
-Simulation results are presented for the NINO3 model in \cref{fig:lfv-secular-nino3-bias-variance} and for the two-state Markov chain model in \cref{fig:lfv-secular-markov-bias-variance}.
-Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the relative importance of secular change and \gls{lfv} depends on $M$.
-When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly (\ie, through the linear trend model).
+Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the results of \cref{fig:lfv-secular-nino3-bias-variance} highlight that the relative importance of secular change and \gls{lfv} depends on $M$.
+When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly with a more complex model (\ie, the linear trend model).
 Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
+As before, when the informational uncertainty is large (small $N$), the identifiability and predictability of the trend are  limited.
 \begin{figure}
@@
   \caption{
-    As \cref{fig:secular-nino3-bias-variance} but for sequences generated from the NINO3 model with both \gls{lfv} and secular change.
+    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with both \gls{lfv} and secular change.
   }\label{fig:lfv-secular-nino3-bias-variance}
@@
 Evaluating and implementing investments for climate risk mitigation involves making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
-The success of this prediction will depend on the identifiability of different signals from an $N$-year sample, the time scales of \gls{lfv} relative to the project life of the instrument, and the uncertainty associated with the estimation.
+The success of this prediction will depend on the identifiability of different signals from limited information, the time scales of \gls{lfv} relative to the project life of the instrument, and the degree of intrinsic uncertainty in the system.
 In this paper we took a synthetic data approach to explore the implications of varying $M$ and $N$ in stylized scenarios which represent important features of real-world hydroclimate systems.
 
-For projects where $M$ is sufficiently short, cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
-The long periods and high amplitudes of the oscillations shown in \cref{fig:observed-lfv} and discussed in \cref{sec:intro-lfv} highlight the \emph{importance} of identifying and predicting this variability.
-However, one's \emph{ability} to do so depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
-If sufficient information is not available, then simple models which represent fewer processes may be preferred.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{M-N-Sketch.pdf}
+  \caption{
+    The importance of predicting different signals, and the identifiability and predictability of the signals, depends on the degree of informational uncertainty ($N$) and the project planning period ($M$).
+  }\label{fig:m-n-sketch}
+\end{figure}
+
+For projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
+However, one's \emph{ability} to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
+In this case, the project may be in the ``potential predictability zone'' of \cref{fig:m-n-sketch}.
+If sufficient information is not available, however, then simple models which represent fewer processes may be preferred (the ``rough guess zone'').
 
 For projects with longer $M$, it becomes necessary to consider the role of secular change.
-As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, uncertainties compound and increase \citep{Dittes:2017he}.
-We show that mis-specification of a trend's mathematical form yields estimates which are increasingly unreliable as $M$ increases, an important caution for the field of ``nonstationary'' flood frequency analysis as noted by \citet{Montanari:2014hl} and \citet{Serinaldi:2015bq}.
+As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
+As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
+With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
+Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation in this ``danger zone''.
 
-The stationary, linear trend, and \gls{hmm} models considered for estimation present an idealized alternative to more complex, physically based models for future risk projection.
-For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf}.
-In places where long records are not available, for example, information from nearby locations, reanalysis products, or \acrshort{gcm} runs may increase the amount of available information \citep{Merz:2008eh}, effectively increasing $N$.
+These findings were derived using simple models for risk projection, but the principles are applicable to more complex, physically based methods.
+For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.
 We suggest, then, that the sample size $N$ defined in our experiments may be straightforwardly interpreted as a measure of the total informational uncertainty in the analysis; as $N$ increases, informational uncertainty decreases.
@@
 
-The climate adaptation decisions which this analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate \citep{sodastrom:1999,powers:2003}.
-The application of a positive discount rate, as is typical in the public and private sectors, emphasizes the importance of predicting near-term risk.
-Consequently, projects with long $M$ must overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.
+The climate adaptation decisions which our analysis can inform are typically framed as economic cost-benefit analyses which discount future cash flows at some annual rate \citep{sodastrom:1999,powers:2003}.
+The application of a positive discount rate, mandated for many public sector projects in the United States \citep{powers:2003}, further emphasizes the importance of predicting near-term risk.
+Projects with long planning periods must therefore overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.
 
@@
   \item that climate risk varies on many scales; and
-  \item that the dominant processes that control this risk depend on the planning period itself.
+  \item that the processes which dominate this risk over the planning period depend on the planning period itself.
 \end{enumerate*}
-Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, as shown in \cref{fig:observed-lfv}.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{M-N-Sketch.pdf}
-  \caption{
-    The importance of predicting different signals, and the identifiability and predictability of the signals, depends on the degree of informational uncertainty ($N$) and the project planning period ($M$).
-  }\label{fig:m-n-sketch}
-\end{figure}
-We also showed that the importance of predicting different climate signals, and potential for doing so skillfully, depend on the project planning period $M$ and the information available, $N$.
-\Cref{fig:m-n-sketch} shows that when $M$ and $N$ are short, only a ``rough guess'' using simple models is feasible, and more complex models  only add uncertainty.
-As $N$ increases, however, the potential for skillful prediction of \gls{lfv} over the short planning period increases, thereby reducing the cost of a flexible or adaptive instrument.
-If $M$ is large, \ie a permanent plan is considered, then informational uncertainty may be addressed with sufficient information (large $N$), but large intrinsic uncertainty may remain.
-Finally, if $M$ is long but $N$ is short, then the analysis will lead to the ``danger zone'' of extrapolation and a high potential for misleading estimates.
+Although the simulations presented here are neatly divided into secular change, \gls{lfv} only, and \gls{lfv} plus secular change, real-world hydroclimate time series exhibit \gls{lfv} on many timescales and several sources of (not necessarily linear) secular change, adding further informational and intrinsic uncertainties.
 
 Depending on the specific climate mechanisms that impact a particular site, and the predictability thereof, the cost and risk associated with a sequence of short-term adaptation projects may be lower than with building a single, permanent structure to prepare for a worst-case scenario far into the future.
-For most large actors, a portfolio of both large $M$ and small $M$ projects will likely be necessary, none of which precludes the need for mitigation of global and local climate change and the development and execution of vulnerability reduction strategies.
+For most large actors, a portfolio of both large $M$ and small $M$ projects will likely be necessary, none of which precludes the need for mitigation of global and local climate change and the development or the execution of vulnerability reduction strategies.
 

### 2019-01-10 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
final tweaks for submission

@@
 This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
-Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and concerns of making adaptation decisions based on potentially incorrect climate projections \citep{lempert:2007} have recently slowed their implementation.
+Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
 Instead, actors such as New York City have turned to a combination of structural, operational, and financial instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
@@
     (a): Irreducible uncertainty cannot be resolved with better models or data and is dominated in the short term by chaotic behavior of the climate, and in the long term by the uncertainty in future \acrlong{acc}.
-    (b): The length of a historical record limits the potential to identify different climate signals.
+    (b): Informational uncertainty limits the potential to identify different climate signals.
     The blue line shows an idealized climate signal and the black line shows observations, which are scattered stochastically around the signal line.
@@
 The relative importance of estimating the short- or long-term risk associated with these extremes depends on the design life $M$, but the potential to understand and predict these different types of variability depends on the informational uncertainty in the $N$-year historical record.
-Though we use floods as an example, the framework also applies to other forms of climate extremes.
+Though we illustrate our findings with a simple flood risk example, the conclusions drawn apply to other hydroclimate hazards, and in particular those typically characterized through a time series of annual maxima or minima.
 
@@
 
-We illustrate our findings with a simple flood risk example.
-The conclusions drawn apply to other hydroclimate hazards, and in particular those which may be easily characterized through a time series of annual maxima or minima of the process.
 The first step is to sample climate risk by generating synthetic streamflow sequences.
@@
 We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
-This model gives a lower bound on total informational uncertainty because it correctly represents the form of the trend, which is known, whereas in real-world analyses the form of the trend is unknown.
+This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.
 
-Finally, we consider a model which explicitly represents \acrlong{lfv}.
+We finally consider a model which explicitly represents \acrlong{lfv}.
 A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie hidden) states $S(t)$ \citep{Rabiner:1986jk}.
@@
 We evaluate both the estimation bias and estimation uncertainty.
-For a given choice of $M$, $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=2000$ posterior simulations of future flows.
+For a given choice of $M$, $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=1000$ posterior simulations of future flows.
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
@@
     The blue line shows the observed sequence.
-    The gray shading indicates the 50\% and 95\% confidence intervals using each of the three fitting methods discussed.
+    The gray shading indicates the 50\% and 95\% confidence intervals using each of the three fitting methods discussed (rows).
     The horizontal black line indicates the flood threshold.
@@
 
-As described in \cref{sec:methods-generating}, we use a synthetic NINO3 time series, as described in \citep{Ramesh:2016hf}, to represent low-frequency climate variability.
-\Cref{fig:enso-ts} shows the wavelet spectrum of this time series, calculated using WaveletComp software \citep{Roesch:wlBQQoIs}.
+As described in \cref{sec:methods-generating}, we use a synthetic \gls{enso} time series, specifically a NINO3 time series, from \citep{Ramesh:2016hf} to represent low-frequency climate variability.
+\Cref{fig:enso-ts} shows the wavelet spectrum of this time series, calculated using the WaveletComp package \citep{Roesch:wlBQQoIs}.
 The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.
@@
 In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
-To explore the temporal structure of risk, we generate $J=2000$ sequences of streamflow, using both the NINO3 and Markov models, for each of several combinations of $M$ and $N$.
+To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
 Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.
-Both the NINO3 model and the Markov model are used; the parameters are shown in \cref{tab:nino-stationary} and \cref{tab:markov-stationary}, respectively.
 

### 2019-01-10 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
add two more references

@@
 The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
-We fit  streamflow sequences $Q_\text{hist}$ simulated using the scenarios in Section 2.1, using a \gls{hmm} with two states.
+\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
+We fit  streamflow sequences using a \gls{hmm} with two states.
 The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.

### 2019-03-27 report-earths-future/reviewer-response.md
addressed all minor reviewer comments -- major comments still open

# Response to Reviewers

We thank

## Reviewer 1

### Overall comments

I enjoyed reading this paper, which confirms that uncertainty dominates when predicting climate change over an extended lead time.

The analysis is carried out through a series of synthetic experiments.
It is interesting to see what are the limits of predictability in the presence of natural variability and secular change.

I think this paper is timely in that it confirms that particular care should be taken when making long term predictions in the presence of a changing climate.
As the analysis clearly shows, it is not easy to recognize climate signals.
Natural variability may interact with climate change therefore masking inherent behaviours.

I think the paper is well written and organized. I made some comments in the attached pdf file.
My comments are essentially minor.

I would like to point out that I missed in the paper a discussion on the Hurst effect.
There is an extended literature on it. Such effect implies the presence of low frequency variability (long term cycles) which were observed in several secular time series of environmental variables.
The Hurst effect may be difficult to distinguish from anthropogenic climate change (ACC).
This effect is ignored by several climate studies. I believe this paper would be a good opportunity to discuss the interaction of the Hurst effect with ACC.
To this end, a proper discussion should be included in the manuscript.

I would like to congratulate the authors for their interesting work.

Best regards,
Alberto Montanari

### Line comments

7: It is not clear to me what this key point means

> We have updated this key point following suggestions from reviewer 2

10: I believe these key points are not much informative. I am not sure of my opinion, but my feeling it that they should reformulated to better explain what the paper discovered.

> Working on it!

34: What do you mean exactly? I don't think mitigation efforts proved to be successful.

> We (unfortunately!) agree that mitigation efforts have not yet proven successful. We have re-phrased this sentence. It now reads "Even if future mitigation efforts are successful, existing levels of atmospheric CO2 and ocean heat content mean that novel adaptation strategies are needed."

84: I believe the impact of ACC on floods and drought is far from being well documented. It is not yet clear whether droughts like the Millenium Drought and the Californian drought have been caused by ACC. California is subjected to drought since the XIX century. Please see  https://en.wikipedia.org/wiki/Droughts_in_California and references therein.

> We agree with the reviewer that the influence of ACC on hydrological extremes is unclear and complex. We have re-worded this sentence to be more clear about this point.  It now reads "Of these processes, ACC has received the most attention in the climate adaptation literature and its influence on some river floods, droughts, hurricanes, urban flooding, and many other climate hazards has been the subject of substantial investigation..."

87: I fully agree with this sentence.

> Thank you very much

103: LFV is strictly related to the Hurst effect. There is an extended literature. I think a mention to the Hurst effect would be appropriate here.

> We have added a short paragraph here which seeks to join some disparate literatures on paleoclimate, Hurst effects, and ocean-atmosphere dynamics. These all explain similar phenomena but use different perspectives.

115: I am not sure it is appropriate to make mention to a hypothetically perfect climate model. It should be made clear that this is a VERY ideal assumption. Perfect environmental models do not exist and, as hydrologists, we should be clear on that.

> We strongly agree with this assessment. In a presentation of this paper at the 2018 AGU Fall Meeting we referred to opening a bottle on a beach and using one of three wishes granted by a magical genie on a climate model to emphasize the ridiculousness of having a perfect model. In this slightly more formal setting we have re-worded the sentence to "Even in the idealized, and unrealistic, case of a perfect climate model, these uncertainties will be large."

118: This is why I believe we should not make reference to perfect models.

> See previous comment

231: I believe these details on the software should not be given here, in the body of the paper. These may be part of the user guide of the produced software

> We are deferring here to the editor. However, I feel that as a consumer of open source software I have an ethical responsibility to cite the primary packages that I use. Further, as a scientist committed to reproducibility I believe that describing software used is as important as describing the experiment design. However, making code accessible on GitHub partially alleviates this concern.

## Reviewer 2

### Overall comments

The authors have teased apart climatic risks that are considered for planning purposes across multiple time horizons.
It is an interesting exploration the interaction of various types of climatic variability, planning horizons, and length of time series, and how this interaction impacts the choice of instruments of climate risk mitigation.
As this paper uses the word robust in the title, the introduction needs to link the concept of robustness to their questions and methods.
Suggest citing literature on robust adaptation if you keep ‘robust’ in the title.

> We have updated the key point in the summary

In Sections 2.1 and 2.2 you describe the three models of climatic change, and three ways of modeling each climate model.
Is there a way to clarify the structure of your computational experiment in these two sections, including your intent of using multiple methods of generating these types of climate variability?
Please explicitly link the lognormal model with the LN2 acronym in text before using in figures.

It would be interesting to see whether there is a relationship in the results between bias and variance; e.g. can the variance inform or predict the bias in each of the three types of climate variability?

The discussion needs to reference the computational results, at minimum to justify the computational experiment as opposed to solely a conceptual exploration.

### Key points

To me, planning implies a process rather than the expected lifespan of a project.
Can you clarify which meaning you use?
If the latter, perhaps the following would be more readable: Climate risk varies over the planned lifespan of a project.
This can also be clarified in other locations of the paper, e.g. line 364 in the summary.

### Line comments

15-16: Can you restate phrase copied below, unclear terms in bold:
**the role of low-frequency variability in modulating climate extremes**

> We agree and have re-phrased this sentence. It now reads "... historical and paleo climate records emphasize interannual to multidecadal modes of variability"

23: The discount rate is given as a decision factor but this is not modeled in the paper. I suggest relegating mention of a discount rate to the introduction or discussion.

> We have removed the distracting mention of a discount rate

24: Suggest because instead of where.

> Our point here is that if variability can be identified and predicted -- which is not a given -- shorter design lives are preferred. The sentence now reads "Shorter design lives are preferred for situations where inter-annual to decadal variability can be successfully identified and predicted..."

38-39: Is this formatting of reference correct?

> Per the document "Brief Guide to AGU Style and Grammar" document available at https://publications.agu.org/brief-guide-agu-style-grammar/, "AGU’s latex template and APA style in reference management software do not include this exception. You do not need to update your manuscript with this exception; it will be applied by the production vendor during the copy-editing process."

41: perhaps a good place to insert a definition and example of operational and financial instruments? An example is given for a structural instrument and that is what people are most likely to understand already.

> This is a helpful suggestion. The sentence now reads "Rather, actors such as New York city have turned to a combination of structural (e.g., index insurance), operational (e.g., improved evacuation routes), and financial (eg, index insurance) investments for reducing vulnerability and increasing resilience to climate extremes." We note that all of these examples are contained in the City of New York report referenced. We believe that these examples are sufficient since a strict definition of different classes of instruments would substantially disrupt the structure of the paragraph.

50-51: Phrasing could be reworked; at minimum change this to these at the end of the sentence.

> The sentence now reads "In this paper we focus on the more narrow question of how the temporal structure of climate risk, and the uncertainties associated with its estimation, influence the answers to these questions."

53: comma after implications

> Added

67: perhaps a good place to define length of long-term short. Is this the interannual to multidecadal described in the abstract or the one-year term of a financial contract (line 59)?

> This is an important question, as we are certainly a bit vague here. The points are clearest in the limits of prediction at time scales less than one year or greater than a century; most real-world centuries are in between. To add some clarity, we have added another sentence with an example. This sentence reads "For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved." We hope that our wording does not imply that we have performed any analysis for New York City.

90-91: Phrasing is a bit obtuse.

> This sentence now reads "Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk."

122: portend (singular)

> Fixed

Figure 1: Inconsistent capitalization. Can you add units?

> Working on this, but going to follow up RE units of American River

147: Recommend preserving the order you present the climate scenarios throughout the paper.

> This is reasonable. We order them (i) secular change only; (ii) LFV only; (iii) LFV plus secular change. This also mirrors the results.

185: Is the word ‘consider’ necessary in these paragraphs?

> We have changed two sentences. The first now reads "In the first case we fit a stationary model to the observed flood record, following...". The second now reads "Finally, we explicitly model LFV using a hidden Markov model (HMM)."

198: be consistent with use of acronyms (i.e. just say LFV).

> Sentence changed; see previous comment

205: suggest “log-normal distribution that is conditional only on the unobserved state
variables” (or S(t) instead of state).

> Thank you for the helpful suggestion

215 and 217: Can you say ‘that’ instead of ‘designed based on’? If these words are meaningful, can you elaborate?

> We agree that the wording was vague. The sentence now reads "An instrument whose design was based on projections with overestimated variance or positive bias will be..."

219: suggest “M, N, and

> This is helpful

### 2019-03-27 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
addressed all minor reviewer comments -- major comments still open

@@
   amssymb, amsmath,   % for math symbols
+  apacite,            % AGU style citations apparently
   booktabs,           % for nice tables
@@
   siunitx,            % for SI notation
-  apacite, % AGU style citations apparently
+  trackchanges,       % to track any new text
 }
@@
 \begin{keypoints}
-  \item Climate risk varies in time over a project's finite planning period
+  \item Climate risk varies over the planned lifespan of a project
   \item Importance of predicting different climate signals depends on extrapolation desired (planning period)
@@
   The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
-  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records highlight the role of \acrlong{lfv} in modulating climate extremes at interannual to multidecadal time scales, \acrlong{acc} is expected to alter their occurrence at longer time scales.
+  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records \add{emphasize} \remove{highlight the role of low-frequency variability (LFV) in modulating climate extremes at} interannual to multidecadal \remove{time scales} \add{modes of variability}, \acrlong{acc} is expected to alter their occurrence at longer time scales.
   In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
@@
   The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
-  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$ and the future discount rate for investment.
-  Shorter design lives are preferred where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
+  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$\remove{ and the future discount rate for investment}.
+  Shorter design lives are preferred \add{for situations} where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
 \end{abstract}
@@
 At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
-Even if mitigation efforts are successful, novel adaptation strategies are needed.
+Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content mean that} novel adaptation strategies are needed.
 
@@
 Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
-Instead, actors such as New York City have turned to a combination of structural, operational, and financial instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (\eg, a stormwater barrier)}, operational\add{ (\eg, improved evacuation routes)}, and financial\add{ (\eg, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
 These instruments are not typically implemented in isolation or statically.
@@
 Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
-In this paper we focus on how the temporal structure of climate risk, and the uncertainties associated with estimating it, influence the answers to this questions.
-We continue this section with three specific observations about climate risk which, while apparently obvious, have important and subtle implications which we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
+In this paper we focus on \add{the more narrow question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
+We continue this section with three specific observations about climate risk which, while apparently obvious, have important and subtle implications\add{,} which we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
 
@@
 This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
+\add{For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved.}
 The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
@@
 Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
-Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on river floods, droughts, hurricanes, urban flooding, and many other climate hazards are well documented \citep[\eg,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
+Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[\eg,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
 Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
@@
 
-Permanent or secular change is not, however, a prerequisite for the inadequacy of relying on historical records.
+\remove{Permanent or secular change is not, however, a prerequisite for the inadequacy of relying on historical records.}
+\add{Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.}
 For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \acrfull{lfv}, due, \eg, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
@@
 
+\add{A WHOLE PARAGRAPH ABOUT HURST EFFECT}
+
 \subsection{The dominant processes depend on the planning period}\label{sec:intro-dominant}
@@
 As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
-Even with a perfect climate model, these uncertainties will be large.
+Even \remove{with} \add{in the idealized, and unrealistic, case of} a perfect climate model, these uncertainties will be large.
 By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.
@@
 Since different physical processes control climate risk at different timescales, successful integration of climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
-A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portends for risk mitigation instruments with a planning period ranging from a few years to several decades.
+A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portend\remove{s} for risk mitigation instruments with a planning period ranging from a few years to several decades.
 As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
@@
 
-We consider three scenarios for climate risk, which we define by the structure of the underlying climate signal:
+We consider three scenarios for climate risk, which we define by the structure of the underlying climate \annote{signal}{changed the order}:
 \begin{enumerate*}[label= (\roman*)]
-  \item \gls{lfv} only;
-  \item \gls{lfv} plus secular change; and
-  \item secular change only.
+  \item secular change only;
+  \item \gls{lfv} only; and
+  \item \gls{lfv} plus secular change.
 \end{enumerate*}
@@
 
-We first consider fitting a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
+\add{In the first case we} \remove{We first consider fitting}\add{fit} a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
 In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
@@
 
-We finally consider a model which explicitly represents \acrlong{lfv}.
+\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
 A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie hidden) states $S(t)$ \citep{Rabiner:1986jk}.
@@
 We fit  streamflow sequences using a \gls{hmm} with two states.
-The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution conditional only on the state.
+The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
 This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
@@
 Both estimation bias and estimation uncertainty affect the utility of a climate risk projection, as illustrated in \cref{fig:conceptual-bias-variance}.
-An instrument designed based on overestimated variance and positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
+An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
 Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
@@
 We evaluate both the estimation bias and estimation uncertainty.
-For a given choice of $M$, $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=1000$ posterior simulations of future flows.
+For a given choice of $M$, \add{and} $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=1000$ posterior simulations of future flows.
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
@@
 \begin{enumerate*}[label= (\roman*)]
-  \item that planning decisions are made with finite horizons;
+  \item that \remove{planning decisions are made with finite horizons}\add{different climate risk mitigation instruments have different planned lifespans};
   \item that climate risk varies on many scales; and

### 2019-03-27 report-earths-future/reviewer-response.md
Get feedback from Dave and Manu

@@
 
-> We have updated the key point in the summary
+> See Line 39 (verify)
 
@@
 Is there a way to clarify the structure of your computational experiment in these two sections, including your intent of using multiple methods of generating these types of climate variability?
+
+> **QUESTION**: I think we just need to modify the beginning of the results section. Is there anything we can make clearer?
+
 Please explicitly link the lognormal model with the LN2 acronym in text before using in figures.
+
+> We have done this in section 2.2
 
@@
 
+> This is certainly an interesting question.
+However, the precise structure of bias and variance as a function of $N$ and $M$ depends strongly on the specific model used to create nonstationarity and to estimate future climate risk.
+As such, we cannot expect that any answer we obtain to this question would generalize beyond the specific assumptions of our stylized computational experiment.
+For this reason we have determined not to address this question in this work.
+
 The discussion needs to reference the computational results, at minimum to justify the computational experiment as opposed to solely a conceptual exploration.
+
+> We have added some modifications to the discussion to reference our computational results more directly and clearly.
 

### 2019-03-27 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Get feedback from Dave and Manu

@@
 \begin{keypoints}
-  \item Climate risk varies over the planned lifespan of a project
-  \item Importance of predicting different climate signals depends on extrapolation desired (planning period)
-  \item Ability to identify and predict different climate signals depends on information available (\eg, data)
+  \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
+  \item 
 \end{keypoints}
@@
 This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
+\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
 Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
@@
 \add{Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.}
+\annote{One well-studied example}{This whole paragraph is added, but the track changes packages doens't allow math or references.} in the hydrological literature is the Hurst phenomenon, $\frac{R_n}{S_n} \sim n^h$, where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
+A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none satisfactorily \citep{klemes:1974}.
+Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
+If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
 For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \acrfull{lfv}, due, \eg, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
+When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
+
 We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet average spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
@@
 \end{figure}
-
-\add{A WHOLE PARAGRAPH ABOUT HURST EFFECT}
 
@@
 In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
+\add{We refer to this model as ``LN2 Stationary.''}
 The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
@@
 We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
+\add{We refer tho this model as ``LN2 Linear Trend''}
 This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.
@@
 These three cases are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
-Projections with each of the three fitting models are also shown.
+Projections with each of the three fitting models \add{described in} \cref{sec:methods-estimating} are also shown.
 This figure highlights that even where projections of average streamflow are unbiased, if the spread is too large then projection of the threshold exceedance probability may be too large.
@@
 
-For projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
+\Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} \add{show that} for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
 However, one's \emph{ability} to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
@@
 
-For projects with longer $M$, it becomes necessary to consider the role of secular change.
+For projects with longer $M$, \add{our results highlight the importance of identifying and predicting} \remove{it becomes necessary to consider the role of} secular change.
 As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
@@
 
-These findings were derived using simple models for risk projection, but the principles are applicable to more complex, physically based methods.
+These findings were derived \add{conceptually and through} \remove{using simple models} \add{idealized computational experiments} for \add{simulating and predicting climate} risk\emph{ projection}, but the principles are applicable to more complex, physically based methods.
 For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.

### 2019-03-28 report-earths-future/figs/flowchart.tex
Update presentation text
put stuff into separate supp info file using latest terrible template from agu

@@
   % Place nodes
-  \node [parameter] (genmod) {Generating Model};
+  \node [parameter] (genmod) {Scenario};
   \node [parameter, above=of genmod] (N) {$N$};
@@
 
-  \node [result, right=of historical] (pThat) {Estimated Exceedances per Year};
-  \node [result, right= of future] (pT) {Observed Exceedances per Year};
+  \node [calculated, right=of historical] (pThat) {Estimated Exceedances per Year};
+  \node [calculated, right= of future] (pT) {Observed Exceedances per Year};
 
-  \node [parameter, right=of pThat] (fitting) {Estimating Model};
+  \node [parameter, above=of pThat] (fitting) {Estimating Model};
   \node [parameter, right=of seq] (threshold) {Flood Design Level};
+
+  \node [result, right=of pThat] (bias) {bias of mean};
+  \node [result, right=of pT] (variance) {bias of variance};
 
@@
 
+  \path[arrow] (pThat) -- (bias);
+  \path[arrow] (pT) -- (bias);
+  \path[arrow] (pThat) -- (variance);
+  \path[arrow] (pT) -- (variance);
+
 \end{tikzpicture}

### 2019-03-28 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
put stuff into separate supp info file using latest terrible template from agu

@@
   linenumbers
-]{agujournal2018}
+]{agujournal2019}
 \journalname{Earth's Future}
 
-% Macros and newcommands
-\usepackage{xspace}
-\newcommand{\eg}{e.g.\@\xspace}
-\newcommand{\ie}{i.e.\@\xspace}
-\newcommand{\normal}{\mathcal{N}}
-\makeatletter
-\newcommand*{\etc}{%
-    \@ifnextchar{.}%
-        {etc}%
-        {etc.\@\xspace}%
-}
-\makeatother
-
 % Package calls with options
-\usepackage[letterpaper, margin=1in]{geometry} % page geometry
-\usepackage[american]{babel} % language setting
 \usepackage[inline]{enumitem} % in-line lists
@@
   graphicx,           % for figures
+  natbib,             % for references
   physics,            % for physics notation
@@
   trackchanges,       % to track any new text
+  xspace,             % for ie and eg
 }
@@
 
-This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
+This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
 \add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
-Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
-\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (\eg, a stormwater barrier)}, operational\add{ (\eg, improved evacuation routes)}, and financial\add{ (\eg, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
+\remove{Instead}\add{Rather}, factors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, a stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
 These instruments are not typically implemented in isolation or statically.
@@
 Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
-This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
+This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
 \add{For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved.}
@@
 Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
-Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[\eg,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
+Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
 Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
@@
 
-\remove{Permanent or secular change is not, however, a prerequisite for the inadequacy of relying on historical records.}
-\add{Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.}
-\annote{One well-studied example}{This whole paragraph is added, but the track changes packages doens't allow math or references.} in the hydrological literature is the Hurst phenomenon, $\frac{R_n}{S_n} \sim n^h$, where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
-A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none satisfactorily \citep{klemes:1974}.
+Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
+The Hurst phenomenon associated with stochastic long-memory processes has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series:
+\begin{equation}
+  \frac{R_n}{S_n} \sim n^h,
+\end{equation}
+where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
+A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none fully satisfactorily \citep{klemes:1974}.
+The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
+\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular.
+Based on this theory, regimes and quasi-periodic phenomena with increasing amplitude as their recurrence interval increases would be the norm rather than the exception.
+Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
+Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
+
 Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
 If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
-For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \acrfull{lfv}, due, \eg, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
+For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due, e.g.\@\xspace, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
 When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
@@
 \begin{equation} \label{eq:log-normal}
-  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
+  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
 \end{equation}
@@
 \Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
-In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, \ie an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
+In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, i.e.\@\xspace an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
 
@@
 \remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
-A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie hidden) states $S(t)$ \citep{Rabiner:1986jk}.
-The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
+A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
+The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
 \Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
@@
 As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
-This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[\ie,][]{Muller:2014fc}.
+This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[i.e.\@\xspace,][]{Muller:2014fc}.
 However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace the utility of any bias reductions.
@@
 Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the results of \cref{fig:lfv-secular-nino3-bias-variance} highlight that the relative importance of secular change and \gls{lfv} depends on $M$.
-When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly with a more complex model (\ie, the linear trend model).
+When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly with a more complex model (i.e.\@\xspace, the linear trend model).
 Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
@@
 As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
+As the physical mechanisms cascade from global (e.g.\@\xspace, global mean surface temperature) to regional  \citep[e.g.\@\xspace, storm track position][]{Barnes:2015gl} and local (e.g.\@\xspace, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
 With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
@@
 
-% -----------------------------------------------------------------------------
-% SUPPLEMENTAL INFORMATION -- CURRENTLY INCLUDED IN THIS FILE FOR CONVENIENCE
-% -----------------------------------------------------------------------------
-
-\clearpage
-\appendix
-
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\theequation}{S\arabic{equation}}
-\setcounter{equation}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{table}{0}
-
-\section{Supplemental Methods}\label{sec:supp-methods}
-
-In this section we provide further equations and parameter values used to generate and fit synthetic streamflow sequences.
-
-\subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}
-
-As described in \cref{sec:methods-generating}, we use a synthetic \gls{enso} time series, specifically a NINO3 time series, from \citep{Ramesh:2016hf} to represent low-frequency climate variability.
-\Cref{fig:enso-ts} shows the wavelet spectrum of this time series, calculated using the WaveletComp package \citep{Roesch:wlBQQoIs}.
-The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.
-There are, however, modes of \gls{lfv} between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
-\begin{figure}
-  \includegraphics[width=\textwidth]{enso_wavelet.png}
-  \caption{
-    Wavelet analysis of the synthetic annual NINO3 time series described in \cref{sec:methods-generating}.
-    (L): wavelet power spectrum.
-    (R): global (average) power spectrum.
-    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
-  }\label{fig:enso-ts}
-\end{figure}
-
-\subsection{Projecting Climate Risk over the Future $M$ Years}
-
-As described in \cref{sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
-The full model, including priors, for the stationary log-normal model is given by \cref{eq:log-normal-stationary}:
-\begin{align}\label{eq:log-normal-stationary}
-  \begin{split}
-    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
-    \mu &\sim \normal \qty(7, 1.5) \\
-    \sigma &\sim \normal^+ \qty(1, 1)
-  \end{split}
-\end{align}
-where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
-The full model, including priors, for the trend log-normal model is given by \cref{eq:log-normal-trend}:
-\begin{align}\label{eq:log-normal-trend}
-  \begin{split}
-    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
-    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
-    \mu_0 & \sim \normal \qty(7, 1.5) \\
-    \beta_\mu & \sim \normal \qty(0, 0.1) \\
-    \log \xi & \sim \normal \qty(0.1, 0.1)
-  \end{split}
-\end{align}
-where $\xi$ is an estimated coefficient of variation.
-
-\subsection{Experiment Design}\label{sec:methods-experiments}
-
-In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
-To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
-Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.
-
-\begin{table}[ht]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
-    \midrule
-    $\mu_0$             & 6     & 6.5   & 6 \\
-    $\gamma$            & 0     & 0.015 & 0.015\\
-    $\beta$             & 0.5   & 0     & 0.5\\
-    $\xi$               & 0.1   & 0.1   & 0.1\\
-    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
-    Threshold           & 3000  & 3000  & 3000 \\
-    \bottomrule
-  \end{tabular}
-  \caption{
-    Parameters of \cref{eq:nino3} for the \gls{lfv} only, secular only, and secular plus \gls{lfv} scenarios.
-  }\label{tab:nino-stationary}
-\end{table}
-
-\section{Sensitivity to climate generating mechanism}\label{sec:markov-generating}
-
-In the main text we sampled climate risk by considering dependence on a single climate index, as shown in \cref{eq:nino3}.
-Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
-A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
-The transition matrix is given by
-\begin{equation}
-  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
-\end{equation}
-The transition matrix is first used to generate a sequence of states $S(t)$.
-The value $\mu(t)$ depends only on $S(t)$ and on time itself:
-\begin{equation}
-  \mu(t) = \begin{cases}
-    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
-    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
-  \end{cases}
-\end{equation}
-For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
-We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
-Parameters for the three experiments are given in \cref{tab:markov-stationary}.
-Results for these experiments are shown in \cref{fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.
-\begin{table}[ht]
-  \centering
-  \begin{tabular}{llll}
-    \toprule
-    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
-    \midrule
-    $\mu_1$             & 6.75  & 6.5   & 6.75\\
-    $\mu_2$             & 6     & 6.5   & 6\\
-    $\gamma_1$          & 0     & 0.015 & 0.015\\
-    $\gamma_2$          & 0     & 0.015 & 0\\
-    $\xi$               & 0.1   & 0.1   & 0.1\\
-    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
-    Threshold           & 3000  & 3000  & 3000 \\
-    \bottomrule
-  \end{tabular}
-  \caption{
-    Parameters of the two-state Markov chain model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
-  }\label{tab:markov-stationary}
-\end{table}
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
-  \caption{
-    \Gls{lfv} only: as \cref{fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-  }\label{fig:lfv-markov-bias-variance}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
-  \caption{
-    Secular change: as \cref{fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-  }\label{fig:secular-only-markov-bias-variance}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
-  \caption{
-    \Gls{lfv} plus secular change: as \cref{fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-  }\label{fig:lfv-secular-markov-bias-variance}
-\end{figure}
-
 \end{document}

### 2019-03-28 report-earths-future/si_template_2019.tex
put stuff into separate supp info file using latest terrible template from agu

\documentclass[
      draft,
      ef,
]{agutexSI2019}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  SUPPORTING INFORMATION TEMPLATE
%
%% ------------------------------------------------------------------------ %%
%
%
%Please use this template when formatting and submitting your Supporting Information.

%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
%
%
%OVERVIEW
%
%Please note that all supporting information will be peer reviewed with your manuscript. It will not be copyedited if the paper is accepted.
%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
%The key criteria are that the data:
% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
%    including %data so the experiment can be reproducible);
% 2. are likely to be usable or used by other scientists working in the field;
% 3. are described with sufficient precision that other scientists can understand them, and
% 4. are not exe files.
%
%USING THIS TEMPLATE
%
%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
%
%All Supporting text and figures should be included in this document. Insert supporting information content into each appropriate section of the template. To add additional captions, simply copy and paste each sample as needed.

%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting. Data sets, large tables, movie files, and audio files should be uploaded separately. Include their captions in this document and list the file name with the caption. You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”

%IMPORTANT NOTE ON FIGURES AND TABLES
% Placeholders for figures and tables appear after the \end{article} command, after references.
% DO NOT USE \psfrag or \subfigure commands.
%
 \usepackage{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
 \setkeys{Gin}{draft=false}
%
% You may need to use one of these options for graphicx depending on the driver program you are using. 
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

\usepackage{siunitx}
\usepackage{apacite}
\graphicspath{{./figs/}{../fig/}} % folders of figures

% Package calls with options
\usepackage[inline]{enumitem} % in-line lists

% Other package calls
\usepackage{
  amssymb, amsmath,   % for math symbols
  booktabs,           % for nice tables
  physics,            % for physics notation
}

% Fixed-width columns
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% Author names in capital letters:
\authorrunninghead{DOSS-GOLLIN ET AL.}

% Shorter version of title entered in capital letters:
\titlerunninghead{ROBUST ADAPTATION TO MULTI-SCALE CLIMATE VARIABILITY}

%Corresponding author mailing address and e-mail address:
%\authoraddr{Corresponding author: A. B. Smith,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

%\includegraphics{agu_pubart-white_reduced.eps}

\title{Supporting Information for ``Robust Adaptation to Multi-Scale Climate Variability''}
%
% e.g., \title{Supporting Information for "Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
%
%DOI: 10.1002/%insert paper number here%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authoraddr{Corresponding author: James Doss-Gollin,
(james.doss-gollin@columbia.edu)}

\authors{James Doss-Gollin\affil{1,2}, David J. Farnham\affil{1,2}, Scott Steinschneider\affil{3}, Upmanu Lall\affil{1,2}}
\affiliation{1}{Department of Earth and Environmental Engineering, Columbia University}
\affiliation{2}{Columbia Water Center, Columbia University}
\affiliation{3}{Department of Biological and Environmental Engineering, Cornell University}

%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%

% The body of the article must start with a \begin{article} command
%
% \end{article} must follow the references section, before the figures
%  and tables.

\begin{article}

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\section{Supplemental Methods}\label{sec:supp-methods}

In this section we provide further equations and parameter values used to generate and fit synthetic streamflow sequences.

\subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}

As described in Section XXX, we use a synthetic ENSO time series, specifically a NINO3 time series, from \cite{Ramesh:2016hf} to represent low-frequency climate variability.
Figure XXX shows the wavelet spectrum of this time series, calculated using the WaveletComp package \cite{Roesch:wlBQQoIs}.
The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of ENSO although interactions with other modes of LFV found in some studies \cite{Jin:1994wq} is dampened in this model.
There are, however, modes of LFV between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.

\subsection{Projecting Climate Risk over the Future $M$ Years}

As described in {sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by {eq:log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \sigma) \\
    \mu &\sim \mathcal{N} \qty(7, 1.5) \\
    \sigma &\sim \mathcal{N}^+ \qty(1, 1)
  \end{split}
\end{align}
where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by {eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \mathcal{N} \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \mathcal{N} \qty(7, 1.5) \\
    \beta_\mu & \sim \mathcal{N} \qty(0, 0.1) \\
    \log \xi & \sim \mathcal{N} \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

\subsection{Experiment Design}\label{sec:methods-experiments}

In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.

\section{Sensitivity to climate generating mechanism}\label{sec:markov-generating}

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in {eq:nino3}.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in {tab:markov-stationary}.
Results for these experiments are shown in {fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.

\bibliography{library}

\end{article}
\clearpage

\begin{figure}
      \includegraphics[width=\textwidth]{enso_wavelet.png}
      \caption{
        Wavelet analysis of the synthetic annual NINO3 time series described in Section XXX.
        (L): wavelet power spectrum.
        (R): global (average) power spectrum.
        Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
      }\label{fig:enso-ts}
    \end{figure}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_0$             & 6     & 6.5   & 6 \\
        $\gamma$            & 0     & 0.015 & 0.015\\
        $\beta$             & 0.5   & 0     & 0.5\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of {eq:nino3} for the LFV only, secular only, and secular plus LFV scenarios.
      }\label{tab:nino-stationary}
    \end{table}

    \begin{table}[ht]
      \centering
      \begin{tabular}{llll}
        \toprule
        Parameter & LFV Only & Secular Only & Secular + LFV \\
        \midrule
        $\mu_1$             & 6.75  & 6.5   & 6.75\\
        $\mu_2$             & 6     & 6.5   & 6\\
        $\gamma_1$          & 0     & 0.015 & 0.015\\
        $\gamma_2$          & 0     & 0.015 & 0\\
        $\xi$               & 0.1   & 0.1   & 0.1\\
        $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
        Threshold           & 3000  & 3000  & 3000 \\
        \bottomrule
      \end{tabular}
      \caption{
        Parameters of the two-state Markov chain model for the LFV only, secular only, and secular plus LFV experiments.
      }\label{tab:markov-stationary}
    \end{table}
    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
      \caption{
        LFV only: as {fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
      \caption{
        Secular change: as {fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:secular-only-markov-bias-variance}
    \end{figure}

    \begin{figure}
      \centering
      \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
      \caption{
        LFV plus secular change: as {fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
      }\label{fig:lfv-secular-markov-bias-variance}
    \end{figure}
\end{document}

### 2019-03-29 report-earths-future/reviewer-response.md
incorporate scott suggestions

@@
 
-We thank
+We thank Dr. Montanari and one anonymous reviewer for their helpful critique of our paper "Robust Adaptation to Multi-Scale Climate Variability".
+Major comments focused on the key points, the Hurst phenomenon, and connections between the introduction, results, and discussion.
+We draw the attention of the reviewers to section 1.2, which has been completely re-written (the track changes package was not used because the section was completely re-organized) to include a more thorough theoretical discussion of the Hurst phenomenon, low-frequency variability, and our experimental design.
+
+We also draw the attention of the reviewers to section 2.4, which now includes a diagram and some additional text to contextualize the experiment design.
+THis section has also been completely re-written and we have also neglected to track changes in this section.
+
+We also thank the reviewers for thoughtful specific comments.
+We have incorporated nearly all of the changes as suggested; any discrepancies are discussed below.
 
@@
 Alberto Montanari
+
+> Thank you for your thoughtful and positive response.
+We share responses to your specific (line) comments below.
+We would like to point your attention to section 1.2, which has been substantially re-written and now provides a greater theoretical discussion of the Hurst phenomenon and low-frequency climate variability.
+Though this is a very deep literature and
 
@@
 
-> **QUESTION**: I think we just need to modify the beginning of the results section. Is there anything we can make clearer?
+> We have re-named section 2.4 experiment design and added a flow chart plus some clarifying text to this section.
 
@@
 
-> We have done this in section 2.2
+> Thank you for noticing this omission. We have defined this acronym in section 2.2
 
@@
 
-> We have added some modifications to the discussion to reference our computational results more directly and clearly.
+> We have added some modifications to the discussion to reference our computational results more directly and clearly. Please see underlined changes in section 4.
 
@@
 This can also be clarified in other locations of the paper, e.g. line 364 in the summary.
+
+> Thank you for your comment on the key points. Based on suggestions from both reviewers we have completely re-written the key points.
 
@@
 
-> This is helpful
+> This is helpful, we have incorporated this suggestion.

### 2019-03-29 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
incorporate scott suggestions

@@
   siunitx,            % for SI notation
-  trackchanges,       % to track any new text
   xspace,             % for ie and eg
@@
 \newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
-\newacronym{lbda}{LBDA}{living blended drought analysis}
+\newacronym{lbda}{LBDA}{Living Blended Drought Analysis}
 \newacronym{lfv}{LFV}{low-frequency variability}
@@
 At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
-Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content mean that} novel adaptation strategies are needed.
+Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content motivate the development of} novel adaptation strategies\remove{ are needed}.
 
@@
 Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
-\remove{Instead}\add{Rather}, factors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, a stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
 These instruments are not typically implemented in isolation or statically.
@@
 Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
-In this paper we focus on \add{the more narrow question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
-We continue this section with three specific observations about climate risk which, while apparently obvious, have important and subtle implications\add{,} which we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
+In this paper we focus on \add{the narrower question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
+We continue this section with three specific observations about climate risk which, while \remove{apparently}\add{seemingly} obvious, have important and subtle implications \add{that}\remove{which} we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
 
@@
 Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
-Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
+Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on some river floods, droughts, hurricanes, urban flooding, and many other climate hazards has been the subject of substantial investigation \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
 Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
@@
 Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
-The Hurst phenomenon associated with stochastic long-memory processes has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series:
-\begin{equation}
-  \frac{R_n}{S_n} \sim n^h,
+The Hurst phenomenon associated with stochastic long-memory processes and has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series.
+This phenomena can be characterized as:
+\begin{equation}\label{eq:hurst}
+  \frac{R_n}{S_n} \sim n^H,
 \end{equation}
-where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
-A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none fully satisfactorily \citep{klemes:1974}.
-The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
+where $H \in [0.5, 1]$ is the Hurst coefficient, $R_n$ is the range of the first $n$ cumulative deviations from the mean, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016,klemes:1974}.
+\Cref{eq:hurst} measures the tendency of a time series to either regress to the mean or cluster towards the extremes.
+The Hurst exponent $H$ is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
 \citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular.
-Based on this theory, regimes and quasi-periodic phenomena with increasing amplitude as their recurrence interval increases would be the norm rather than the exception.
+Based on this theory, regimes and quasi-periodic phenomena with amplitude that increases with recurrence interval would be the norm rather than the exception.
 Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
@@
 Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
-If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
-For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due, e.g.\@\xspace, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
-When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
+If these climate modes vary on decadal or longer time scales, such as the \gls{pdo} \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
+For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
+When records are short and low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
+t is also difficult to identify the dominant scales of variability \citep{newman:2016,Jin:1994wq}.
 
-We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet average spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
+We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
 Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
-This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
+This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
 A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
@@
     (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
-    High numbers indicate more severe drought.
+    \remove{High numbers}\add{Lower values} indicate more severe drought.
     A 20-year running mean is also shown in blue.
-    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
-    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
-    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
-    (d) wavelet spectrum, like (c), for the American River data.
+    (b) A \SI{100}{year} record of annual-maximum streamflow for the American River at Folsom.
+    Daily streamflow values were divided by the catchment area to yield a normalized flow in units of \si{\milli\meter\per\day}.
+    (c) The global wavelet power spectrum of the \gls{lbda} time series (a).
+    Blue (red) dots indicate frequencies which are significant at $\alpha=0.10(0.05)$ compared to white noise.
+    (d) Global wavelet power spectrum, like (c), for the American River data.
   }\label{fig:observed-lfv}
@@
 As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
-Even \remove{with} \add{in the idealized, and unrealistic, case of} a perfect climate model, these uncertainties will be large.
+Even \remove{with} \add{in the idealized and unrealistic case of} a perfect climate model, these uncertainties will be large.
 By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.
@@
 \end{equation}
-This formulation describes the \begin{enumerate*}[label= (\roman*)]
-  \item \gls{lfv} only;
-  \item \gls{lfv} plus secular change; and
-  \item secular change only
-\end{enumerate*}
-scenarios in a single equation; the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
+This formulation describes \add{all} scenarios \add{for future climate considered in this paper within} \remove{in} a single equation\remove{;}\add{.}
+\add{To add climate variability to the system,} the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
 \begin{equation}\label{eq:nino3}
@@
 Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
-The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.
+The values of all parameters used for sampling climate risk are listed in the online supporting information for each of the three scenarios considered.
 
@@
 To create an annual time series, we average the October-December values of the NINO3 index for each year.
-\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
-In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, i.e.\@\xspace an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
+Supplemental figure S1 shows a wavelet spectrum and time series plot of the resulting annual time series.
+In the online supplemental information we consider an alternative parameterization of $\mu(t)$, \remove{i.e.\@\xspace an alternative to,} which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
 
@@
 For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
-Once a statistical model is fit to these ``observations'', then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
+Once a statistical model is fit to these \remove{``}observations\remove{''}, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
 
@@
 The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
-The full model, including priors, is given in \cref{eq:log-normal-stationary}.
+The full model, including priors, is given in supplemental equation (S1).
 
@@
 Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
-We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
-\add{We refer tho this model as ``LN2 Linear Trend''}
+We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in supplemental equation (S2).
+\add{We refer to this model as ``LN2 Linear Trend''}
 This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.
@@
 \remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
-A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
+An \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
 The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
@@
 An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
-Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
+Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the \remove{user}\add{public}.
 \begin{figure}
@@
     Consequences of model bias or incorrect model representation of uncertainty.
-    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive for the user.
+    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive.
     If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
@@
 We evaluate both the estimation bias and estimation uncertainty.
-For a given choice of $M$, \add{and} $N$, generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'', and the $K=1000$ posterior simulations of future flows.
+For a given choice of $M$, $N$, \add{and} generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'' and the $K=1000$ posterior simulations of future flows.
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
@@
 
-\subsection{Computation}\label{sec:methods-computation}
+\subsection{Experiment design}\label{sec:methods-computation}
+
+\Cref{fig:flowchart} describes the experimental design.
+We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
+First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
+Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
+Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
+Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{flowchart.pdf}
+  \caption{
+    Flow chart describing experiment design.
+    Parameters are shown in red.
+    Calculated quantities are shown in white.
+    Quantities used for analysis are shown in blue.
+  }\label{fig:flowchart}
+\end{figure}
 
@@
 
-We assess estimation bias and variance for three cases.
-First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
-Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
-Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
-These three cases are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
-Projections with each of the three fitting models \add{described in} \cref{sec:methods-estimating} are also shown.
+These three \remove{cases}\add{scenarios for future climate considered} are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
+\add{We also show projected future climate risk} with each of the three \remove{fitting}\add{estimating} models \add{described in} \cref{sec:methods-estimating}\remove{are also shown}.
 This figure highlights that even where projections of average streamflow are unbiased, if the spread is too large then projection of the threshold exceedance probability may be too large.
+In the remainder of this section we present a more systematic analysis of each of these three cases.
 
-In the remainder of this section we present a more systematic analysis of each of these three cases.
 \begin{figure}
@@
 As expected, the stationary log-normal model performs well overall, with low bias and low variance.
-The \gls{hmm} actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the multimodal distribution that emerges from dependence on the \gls{enso} index, which exhibits several regimes (see \cref{fig:enso-ts}).
+The \gls{hmm} actually out-performs the stationary model, with slightly lower variance than the stationary model, because it better captures the multimodal distribution that emerges from dependence on the \gls{enso} index, which exhibits several regimes (see supplemental figure S1).
 By contrast, the linear trend model performs poorly for low $N$ and high $M$ because a positive probability is assigned to the existence of a positive trend.
@@
 If this period is much larger than $M$, then a stationary assumption may provide reasonable estimates, and fewer observations may be required (shorter $N$).
-As shown in \cref{fig:enso-ts}, the \gls{enso} time series is most active in the \SIrange{3}{6}{year} band.
+As shown in supplemental figure S1, the \gls{enso} time series is most active in the \SIrange{3}{6}{year} band.
 In the real world, however, many hydroclimate time series vary at multidecadal and longer frequencies.
@@
 With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
-Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation in this ``danger zone''.
+Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation \add{(i.e.\xspace a }``danger zone''\add{ for planning)}.
 
-These findings were derived \add{conceptually and through} \remove{using simple models} \add{idealized computational experiments} for \add{simulating and predicting climate} risk\emph{ projection}, but the principles are applicable to more complex, physically based methods.
+These findings were derived \add{conceptually and through} \remove{using simple models} \add{idealized computational experiments} for \add{simulating and predicting climate} risk\remove{ projection}, but the principles are applicable to more complex, physically based methods.
 For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.
@@
 The application of a positive discount rate, mandated for many public sector projects in the United States \citep{powers:2003}, further emphasizes the importance of predicting near-term risk.
-Projects with long planning periods must therefore overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$ rather than a sequence of $N$, $N+\Delta N, \ldots$.
+Projects with long planning periods must therefore overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$.
+\add{By contrast, the informational uncertainties for a sequence of short-term instruments are $N, N+M, N+2M, \ldots$, which may lead to improved identifiability and predictability of relevant climate signals.}\remove{rather than a sequence of $N$, $N+\Delta N, \ldots$}.
 
@@
 
+The authors thank Dr. Alberto Montanari and one anonymous reviewer for insightful comments on this manuscript.
 The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{100000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh:2016hf}.

### 2019-03-29 report-earths-future/si_template_2019.tex
incorporate scott suggestions

@@
 
-As described in Section XXX, we use a synthetic ENSO time series, specifically a NINO3 time series, from \cite{Ramesh:2016hf} to represent low-frequency climate variability.
-Figure XXX shows the wavelet spectrum of this time series, calculated using the WaveletComp package \cite{Roesch:wlBQQoIs}.
+As described in Section 2.1, we use a synthetic ENSO time series, specifically a NINO3 time series, from \cite{Ramesh:2016hf} to represent low-frequency climate variability.
+Figure S1 shows the wavelet spectrum of this time series, calculated using the WaveletComp package \cite{Roesch:wlBQQoIs}.
 The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of ENSO although interactions with other modes of LFV found in some studies \cite{Jin:1994wq} is dampened in this model.
-There are, however, modes of LFV between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
+There are, however, modes of LFV between 16 and 128 years which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
 
@@
 
-As described in {sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
-The full model, including priors, for the stationary log-normal model is given by {eq:log-normal-stationary}:
+As described in Section 2, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
+The full model, including priors, for the stationary log-normal model is given by equation \eqref{log-normal-stationary}:
 \begin{align}\label{eq:log-normal-stationary}
@@
 where $\mathcal{N}$ denotes the normal distribution and $\mathcal{N}^+$ denotes a half-normal distribution.
-The full model, including priors, for the trend log-normal model is given by {eq:log-normal-trend}:
+The full model, including priors, for the trend log-normal model is given by equation \eqref{eq:log-normal-trend}:
 \begin{align}\label{eq:log-normal-trend}
@@
 
-In the main text we sampled climate risk by considering dependence on a single climate index, as shown in {eq:nino3}.
+In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 5.
 Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
@@
 We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
-Parameters for the three experiments are given in {tab:markov-stationary}.
-Results for these experiments are shown in {fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.
+Parameters for the three experiments are given in Table \ref{tab:markov-stationary}.
+Results for these experiments are shown in Figure \ref{fig:lfv-markov-bias-variance}, \ref{fig:lfv-secular-markov-bias-variance}, and \ref{fig:secular-only-markov-bias-variance}.
 
@@
       \caption{
-        Wavelet analysis of the synthetic annual NINO3 time series described in Section XXX.
+        Wavelet analysis of the synthetic annual NINO3 time series described in Section 2.1.
         (L): wavelet power spectrum.
@@
       \caption{
-        Parameters of {eq:nino3} for the LFV only, secular only, and secular plus LFV scenarios.
+        Parameters of equation (5) for the LFV only, secular only, and secular plus LFV scenarios.
       }\label{tab:nino-stationary}
@@
     \end{table}
-    \begin{figure}
-      \centering
-      \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
-      \caption{
-        LFV only: as {fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
-      }\label{fig:lfv-markov-bias-variance}
-    \end{figure}
 
@@
       \caption{
-        Secular change: as {fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
+        Secular change: as Figure 5 for sequences generated with the two-state Markov chain model.
       }\label{fig:secular-only-markov-bias-variance}
+    \end{figure}
+
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
+      \caption{
+        LFV only: as Figure 6 for sequences generated with the two-state Markov chain model.
+      }\label{fig:lfv-markov-bias-variance}
     \end{figure}
@@
       \caption{
-        LFV plus secular change: as {fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
+        LFV plus secular change: as Figure 7 for sequences generated with the two-state Markov chain model.
       }\label{fig:lfv-secular-markov-bias-variance}

### 2019-04-04 report-earths-future/reviewer-response.md
version for final coauthor comments

--- 2019-03-27 report-earths-future/reviewer-response.md
@@
 
-We thank
+We thank Dr. Montanari and one anonymous reviewer for their helpful critique of our paper "Robust Adaptation to Multi-Scale Climate Variability".
+
+Major comments focused on three areas.
+First, the key points were confusingly worded and uninformative.
+Second, the theoretical discussion of low-frequency climate variability lacked connections to literature on the Hurst phenomenon.
+Finally, more explicit links between the introduction, results, and discussion were needed.
+We draw the attention of the reviewers to sections 1.2 and 2.4, which have been completely re-written to address these points.
+We note that we have not applied track changes to these two sections, as they have been completely re-written.
+
+We also thank the reviewers for thoughtful specific comments, discussed below.
 
@@
 
+> Thank you for your thoughtful and positive response.
+We share responses to your specific (line) comments below.
+We would like to point your attention to section 1.2, which has been substantially re-written and now provides a greater theoretical discussion of the Hurst phenomenon, low-frequency climate variability, fractals, and nonlinear dynamics.
+
 ### Line comments
@@
 
-> We have updated this key point following suggestions from reviewer 2
+> Please see revised key points
 
@@
 
-> Working on it!
+> Please see revised key points
 
@@
 
-> Thank you very much
+> Thank you very much for your important work on this topic
 
@@
 
-> We have added a short paragraph here which seeks to join some disparate literatures on paleoclimate, Hurst effects, and ocean-atmosphere dynamics.
-These all explain similar phenomena but use different perspectives.
+> We have re-written this subsection to discuss literature on paleoclimate, Hurst effects, and ocean-atmosphere dynamics.
 
@@
 
-> See previous comment
+> Please see previous comment
 
@@
 Further, as a scientist committed to reproducibility I believe that describing software used is as important as describing the experiment design.
-However, making code accessible on GitHub partially alleviates this concern.
+However, it is true that making code accessible on GitHub partially alleviates this concern.
 
@@
 
-> See Line 39 (verify)
+> Thank you for this suggestion. Please see sentence beginning L45.
 
@@
 
-> We have removed the distracting mention of a discount rate
+> We have removed the distracting mention of a discount rate from this section.
 
@@
 > This is a helpful suggestion.
-The sentence now reads "Rather, actors such as New York city have turned to a combination of structural (e.g., index insurance), operational (e.g., improved evacuation routes), and financial (eg, index insurance) investments for reducing vulnerability and increasing resilience to climate extremes." We note that all of these examples are contained in the City of New York report referenced.
+The sentence now reads "Rather, actors such as New York city have turned to a combination of structural (e.g., a stormwater barrier), operational (e.g., improved evacuation routes), and financial (eg, index insurance) investments for reducing vulnerability and increasing resilience to climate extremes." We note that all of these examples are contained in the City of New York report referenced.
 We believe that these examples are sufficient since a strict definition of different classes of instruments would substantially disrupt the structure of the paragraph.
@@
 > This is an important question, as we are certainly a bit vague here.
-The points are clearest in the limits of prediction at time scales less than one year or greater than a century; most real-world centuries are in between.
+The points are clearest in the limits of prediction at time scales less than one year or greater than a century; most real-world projects are in between.
 To add some clarity, we have added another sentence with an example.
@@
 
-> Working on this, but going to follow up RE units of American River
+> Thank you for this suggestion. Figure 1 now has units defined, uniform capitalization, and lettering (a)...(d).
 
@@
 
-> This is reasonable. We order them (i) secular change only; (ii) LFV only; (iii) LFV plus secular change. This also mirrors the results.
+> Thank you for this helpful suggestion. We now order the scenarios (i) secular change only; (ii) LFV only; (iii) LFV plus secular change. This also mirrors the results.
 

### 2019-04-04 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
version for final coauthor comments

@@
   \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
-  \item 
+  \item \add{We carry out stylized computational experiments to assess the bias and variance of future risk estimated from limited data sets}
+  \item \add{Lack of data limits the identifiability of climate signals, favoring prediction of short-term risks which require less extrapolation}
 \end{keypoints}
@@
 At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
-Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content motivate the development of} novel adaptation strategies\remove{ are needed}.
+Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of} novel adaptation strategies\remove{ are needed}.
 
@@
 \add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
-Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
+Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently \remove{slowed their implementation}\add{dampened enthusiasm}.
 \remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
@@
 Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
-The Hurst phenomenon associated with stochastic long-memory processes and has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series.
-This phenomena can be characterized as:
-\begin{equation}\label{eq:hurst}
-  \frac{R_n}{S_n} \sim n^H,
-\end{equation}
-where $H \in [0.5, 1]$ is the Hurst coefficient, $R_n$ is the range of the first $n$ cumulative deviations from the mean, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016,klemes:1974}.
-\Cref{eq:hurst} measures the tendency of a time series to either regress to the mean or cluster towards the extremes.
-The Hurst exponent $H$ is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
-\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular.
-Based on this theory, regimes and quasi-periodic phenomena with amplitude that increases with recurrence interval would be the norm rather than the exception.
-Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
-Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
+The Hurst phenomenon is very well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
+The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
+The Hurst phenomena has also been connected to low frequency quasi-periodic phenomena, especially where fractal scaling is expected.
+For example, wavelet methods have been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
+The Hurst phenomenon also provides a link between catchment hydrology and global climate dynamics \citep{bloschl:2010,montanari:2003}.
+The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
+These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013}.
 
-Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
-If these climate modes vary on decadal or longer time scales, such as the \gls{pdo} \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
-For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
-When records are short and low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
-t is also difficult to identify the dominant scales of variability \citep{newman:2016,Jin:1994wq}.
-
-We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
+External forcing from structured climate signals \citep[``teleconnections'';][]{Angstrom:1935ej} and catchment dynamics are both useful in explaining the \gls{lfv} observed in natural hydroclimate time series.
+We illustrate such \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
 Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
@@
 A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
+The key implication is that the observations, \citep{Jain:2001hz}, trends \citep{bhattacharya:1983}, and frequencies \citep{newman:2016} observed in the past are often poor predictors of future behavior.
 
@@
 We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
-First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped.
-Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time.
-Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change.
+First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped (``secular change only'').
+Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time (``\acrlong{lfv} only'').
+Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change (``\acrlong{lfv} plus secular change'').
 Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.
@@
 Bayesian models were written in the stan probabilistic programming language \citep{Carpenter:2017ke} using the No U-Turn Sampler \citep{Hoffman:2011wm,Betancourt:2017vd}.
-The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.\footnote{Dear Reviewers: our codes are not presently available on GitHub, as we have chosen to keep the repository private until the article is accepted for publication, but they are available to you through the online supplemental material for this submission.}
+The codes used to generate the figures and text of this paper are available at \url{https://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk} or in the online supplemental material of this paper.
 
@@
 \Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} \add{show that} for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
-However, one's \emph{ability} to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
+However, one's ability to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
 In this case, the project may be in the ``potential predictability zone'' of \cref{fig:m-n-sketch}.

### 2019-04-04 report-earths-future/si_template_2019.tex
version for final coauthor comments

@@
 As described in Section 2, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \cite{Carpenter:2017ke}.
-The full model, including priors, for the stationary log-normal model is given by equation \eqref{log-normal-stationary}:
+The full model, including priors, for the stationary log-normal model is given by equation \eqref{eq:log-normal-stationary}:
 \begin{align}\label{eq:log-normal-stationary}
@@
 where $\xi$ is an estimated coefficient of variation.
+The stan models used are available with other codes at \url{https://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk}.
 
@@
 To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
-Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.
+Streamflow sequences are generated using the stationary generating functions, both with and without superimposing a secular trend.
 
@@
 
-In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 5.
+In the main text we sampled climate risk by considering dependence on a single climate index, as shown in equation 4.
 Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
@@
 Parameters for the three experiments are given in Table \ref{tab:markov-stationary}.
-Results for these experiments are shown in Figure \ref{fig:lfv-markov-bias-variance}, \ref{fig:lfv-secular-markov-bias-variance}, and \ref{fig:secular-only-markov-bias-variance}.
+Results for these experiments are shown in figures \ref{fig:secular-only-markov-bias-variance}, \ref{fig:lfv-markov-bias-variance}, and \ref{fig:lfv-secular-markov-bias-variance}.
 

### 2019-04-04 src/slurm-LFV-Only.sh
use updated name of conda environment

@@ -9,7 +9,7 @@
 #SBATCH --time=3:59:00          # The time the job will take to run.
 
 module load anaconda            # load the anaconda module
-source activate STXCluster      # activate the conda environment
+source activate robust-adaptation-cyclical-risk      # activate the conda environment
 
 # run the python
 python 11-Run-LFV-Only.py

### 2019-04-04 src/slurm-LFV-Secular.sh

@@ -9,7 +9,7 @@
 #SBATCH --time=3:59:00          # The time the job will take to run.
 
 module load anaconda            # load the anaconda module
-source activate STXCluster      # activate the conda environment
+source activate robust-adaptation-cyclical-risk      # activate the conda environment
 
 # run the python
 python 13-Run-LFV-Secular.py

### 2019-04-04 src/slurm-Secular-Only.sh

@@ -9,7 +9,7 @@
 #SBATCH --time=3:59:00          # The time the job will take to run.
 
 module load anaconda            # load the anaconda module
-source activate STXCluster      # activate the conda environment
+source activate robust-adaptation-cyclical-risk      # activate the conda environment
 
 # run the python
 python 12-Run-Secular-Only.py

### 2019-04-04 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Incorporate minor edits from dave

@@
   \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
-  \item \add{We carry out stylized computational experiments to assess the bias and variance of future risk estimated from limited data sets}
+  \item \add{We carry out stylized computational experiments to assess the bias and variance of projected future risk estimated from limited data sets}
   \item \add{Lack of data limits the identifiability of climate signals, favoring prediction of short-term risks which require less extrapolation}
@@
 Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
-The Hurst phenomenon is very well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
+The Hurst phenomenon is well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
 The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
-The Hurst phenomena has also been connected to low frequency quasi-periodic phenomena, especially where fractal scaling is expected.
+The Hurst phenomenon has also been connected to low frequency quasi-periodic phenomenon, especially where fractal scaling is expected.
 For example, wavelet methods have been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
@@
 The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
-These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013}.
+These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013} and the \gls{pdo} \citep{Mantua:1997kj}.
 
@@
 
-Evaluating and implementing investments for climate risk mitigation involves making projections of climate risk, which exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
+Evaluating and implementing investments for climate risk mitigation involves making projections of climate risk, which generally exhibits both \gls{lfv} and secular trends, over the $M$-year project life of the instrument.
 The success of this prediction will depend on the identifiability of different signals from limited information, the time scales of \gls{lfv} relative to the project life of the instrument, and the degree of intrinsic uncertainty in the system.
-In this paper we took a synthetic data approach to explore the implications of varying $M$ and $N$ in stylized scenarios which represent important features of real-world hydroclimate systems.
+In this paper we took a synthetic data approach to explore the implications of varying $M$ and $N$ in stylized scenarios that represent important features of real-world hydroclimate systems.
 
@@
 As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
-As the physical mechanisms cascade from global (e.g.\@\xspace, global mean surface temperature) to regional  \citep[e.g.\@\xspace, storm track position][]{Barnes:2015gl} and local (e.g.\@\xspace, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
+As the physical mechanisms cascade from global (e.g.\@\xspace, global mean surface temperature) to regional  \citep[e.g.\@\xspace, storm track position;][]{Barnes:2015gl} and local (e.g.\@\xspace, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
 With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
-Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation \add{(i.e.\xspace a }``danger zone''\add{ for planning)}.
+Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation \add{(i.e.\xspace} a ``danger zone''\add{ for planning)}.
 
 These findings were derived \add{conceptually and through} \remove{using simple models} \add{idealized computational experiments} for \add{simulating and predicting climate} risk\remove{ projection}, but the principles are applicable to more complex, physically based methods.
-For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on \glspl{gcm} and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.
-We suggest, then, that the sample size $N$ defined in our experiments may be straightforwardly interpreted as a measure of the total informational uncertainty in the analysis; as $N$ increases, informational uncertainty decreases.
+For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on general circulation models and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.
+We suggest that the sample size $N$ defined in our experiments may be straightforwardly interpreted as a measure of the total informational uncertainty in the analysis; as $N$ increases, informational uncertainty decreases.
 
@@
 Since \cref{sec:results} shows that the bias and variance of climate risk projections tend to increase with $M$, the total bias and variance associated with sequencing 20 consecutive $M=\SI{5}{year}$ projects will be less than that associated with making a single $M=\SI{100}{year}$ project.
-This effect will be exacerbated by the fact that if the first $M=\SI{5}{year}$ project is based on estimates with informational uncertainty $N$, the second will have $N+5$, the third $N+10$, and so on.
+This effect will be compounded by the fact that if the first $M=\SI{5}{year}$ project is based on estimates with informational uncertainty $N$, the second will have $N+5$, the third $N+10$, and so on.
 
@@
 Projects with long planning periods must therefore overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$.
-\add{By contrast, the informational uncertainties for a sequence of short-term instruments are $N, N+M, N+2M, \ldots$, which may lead to improved identifiability and predictability of relevant climate signals.}\remove{rather than a sequence of $N$, $N+\Delta N, \ldots$}.
+\add{By contrast, the informational uncertainties for a sequence of short-term instruments are $N, N+M, N+2M, \ldots$, potentially yielding improved identifiability and predictability of relevant climate signals.}\remove{rather than a sequence of $N$, $N+\Delta N, \ldots$}.
 

### 2019-04-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Final version with manu suggestion

@@
 \begin{keypoints}
-  \item \add{Climate risk varies over the planned lifespan of a climate risk mitigation instrument}
-  \item \add{We carry out stylized computational experiments to assess the bias and variance of projected future risk estimated from limited data sets}
-  \item \add{Lack of data limits the identifiability of climate signals, favoring prediction of short-term risks which require less extrapolation}
+  \item \add{Quasi-periodic and secular climate signals with different identifiability and predictability control uncertainty and risk into the future}
+  \item \add{Adaptation strategies need to consider how the bias and variance in these risk predictions influences choices and decisions pathways}
+  \item \add{Stylized experiments reveal how bias and variance of climate risk estimates over a finite planning periodinfluence risk mitigation}
 \end{keypoints}
@@
 Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently \remove{slowed their implementation}\add{dampened enthusiasm}.
-\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, a catastrophe bond)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
 These instruments are not typically implemented in isolation or statically.
@@
 This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
-\add{For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved.}
+\add{For example, the optimal policy for New York City to manage uncertain hurricane risk in the 21st century might potentially be to keep areas devastated by hurricane Sandy zoned for low-impact development for the next 10 years.
+This would reduce future risk over all climate scenarios while postponing major investments until large uncertainties as to the magnitude of future sea level rise are resolved.}
 The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
@@
 We illustrate such \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
-Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
+Peaks for the American River time series are apparent at 2.3 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
 This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
@@
 
-Both estimation bias and estimation uncertainty affect the utility of a climate risk projection, as illustrated in \cref{fig:conceptual-bias-variance}.
+Both estimation bias and estimation uncertainty affect the utility of a climate risk projection.
 An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
 Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the \remove{user}\add{public}.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
-  \caption{
-    Consequences of model bias or incorrect model representation of uncertainty.
-    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive.
-    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
-  }\label{fig:conceptual-bias-variance}
-\end{figure}
 
@@
 
-\bibliography{library}
+% -----------------------------------------------------------------------------
+% REFERENCES
+% -----------------------------------------------------------------------------
+
+\begin{thebibliography}{}
+
+  \bibitem [\protect \citeauthoryear {%
+  Barnes%
+  \ \BBA {} Screen%
+  }{%
+  Barnes%
+  \ \BBA {} Screen%
+  }{%
+  {\protect \APACyear {2015}}%
+  }]{%
+  Barnes:2015gl}
+  \APACinsertmetastar {%
+  Barnes:2015gl}%
+  \begin{APACrefauthors}%
+  Barnes, E\BPBI A.%
+  \BCBT {}\ \BBA {} Screen, J\BPBI A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2015}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The Impact of {{Arctic}} Warming on the Midlatitude
+    Jet-Stream: {{Can}} It? {{Has}} It? {{Will}} It?} {The impact of {{Arctic}}
+    warming on the midlatitude jet-stream: {{Can}} it? {{Has}} it? {{Will}}
+    it?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Wiley Interdisciplinary Reviews-Climate
+    Change}{6}{3}{277-286}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/wcc.337} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Beran%
+  }{%
+  Beran%
+  }{%
+  {\protect \APACyear {1994}}%
+  }]{%
+  beran:1994}
+  \APACinsertmetastar {%
+  beran:1994}%
+  \begin{APACrefauthors}%
+  Beran, J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYear{1994}.
+  \newblock
+  \APACrefbtitle {Statistics for Long-Memory Processes} {Statistics for
+    long-memory processes}.
+  \newblock
+  \APACaddressPublisher{New York}{{Chapman \& Hall}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Betancourt%
+  }{%
+  Betancourt%
+  }{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  Betancourt:2017vd}
+  \APACinsertmetastar {%
+  Betancourt:2017vd}%
+  \begin{APACrefauthors}%
+  Betancourt, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A {{Conceptual Introduction}} to {{Hamiltonian Monte
+    Carlo}}} {A {{Conceptual Introduction}} to {{Hamiltonian Monte
+    Carlo}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{arXiv.org}{}{}{}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Bhattacharya%
+  , Gupta%
+  \BCBL {}\ \BBA {} Waymire%
+  }{%
+  Bhattacharya%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {1983}}%
+  }]{%
+  bhattacharya:1983}
+  \APACinsertmetastar {%
+  bhattacharya:1983}%
+  \begin{APACrefauthors}%
+  Bhattacharya, R\BPBI N.%
+  , Gupta, V\BPBI K.%
+  \BCBL {}\ \BBA {} Waymire, E.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1983}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Hurst}} Effect under Trends} {The {{Hurst}} effect
+    under trends}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Applied Probability}{20}{3}{649-662}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.2307/3213900} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Blöschl%
+  \ \BBA {} Montanari%
+  }{%
+  Blöschl%
+  \ \BBA {} Montanari%
+  }{%
+  {\protect \APACyear {2010}}%
+  }]{%
+  bloschl:2010}
+  \APACinsertmetastar {%
+  bloschl:2010}%
+  \begin{APACrefauthors}%
+  Blöschl, G.%
+  \BCBT {}\ \BBA {} Montanari, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2010}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Climate Change Impacts—Throwing the Dice?} {Climate
+    change impacts—throwing the dice?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrological Processes}{24}{3}{374-381}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/hyp.7574} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Borgomeo%
+  , {Mortazavi-Naeini}%
+  , Hall%
+  \BCBL {}\ \BBA {} Guillod%
+  }{%
+  Borgomeo%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2018}}%
+  }]{%
+  Borgomeo:2018hk}
+  \APACinsertmetastar {%
+  Borgomeo:2018hk}%
+  \begin{APACrefauthors}%
+  Borgomeo, E.%
+  , {Mortazavi-Naeini}, M.%
+  , Hall, J\BPBI W.%
+  \BCBL {}\ \BBA {} Guillod, B\BPBI P.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2018}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Risk, {{Robustness}} and {{Water Resources Planning
+    Under Uncertainty}}} {Risk, {{Robustness}} and {{Water Resources Planning
+    Under Uncertainty}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Earth's Future}{6}{3}{468-487}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2017EF000730} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Bracken%
+  , Rajagopalan%
+  \BCBL {}\ \BBA {} Woodhouse%
+  }{%
+  Bracken%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Bracken:2016ba}
+  \APACinsertmetastar {%
+  Bracken:2016ba}%
+  \begin{APACrefauthors}%
+  Bracken, C.%
+  , Rajagopalan, B.%
+  \BCBL {}\ \BBA {} Woodhouse, C.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A {{Bayesian}} Hierarchical Nonhomogeneous Hidden
+    {{Markov}} Model for Multisite Streamflow Reconstructions} {A {{Bayesian}}
+    hierarchical nonhomogeneous hidden {{Markov}} model for multisite streamflow
+    reconstructions}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{52}{10}{7837-7850}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2016WR018887} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Brown%
+  }{%
+  Brown%
+  }{%
+  {\protect \APACyear {2010}}%
+  }]{%
+  brown:2010}
+  \APACinsertmetastar {%
+  brown:2010}%
+  \begin{APACrefauthors}%
+  Brown, C\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2010}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{End}} of {{Reliability}}} {The {{End}} of
+    {{Reliability}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Water Resources Planning and
+    Management}{136}{2}{143-145}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1061/(ASCE)WR.1943-5452.65} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Bullmore%
+  \ \protect \BOthers {.}}{%
+  Bullmore%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2001}}%
+  }]{%
+  bullmore:2001}
+  \APACinsertmetastar {%
+  bullmore:2001}%
+  \begin{APACrefauthors}%
+  Bullmore, E.%
+  , Long, C.%
+  , Suckling, J.%
+  , Fadili, J.%
+  , Calvert, G.%
+  , Zelaya, F.%
+  \BDBL {}Brammer, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2001}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Colored Noise and Computational Inference in
+    Neurophysiological ({{fMRI}}) Time Series Analysis: {{Resampling}} Methods in
+    Time and Wavelet Domains} {Colored noise and computational inference in
+    neurophysiological ({{fMRI}}) time series analysis: {{Resampling}} methods in
+    time and wavelet domains}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Human Brain Mapping}{12}{2}{61-78}.
+  \newblock
+  \begin{APACrefDOI}
+    \doi{10.1002/1097-0193(200102)12:2<61::AID-HBM1004>3.0.CO;2-W}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Carpenter%
+  \ \protect \BOthers {.}}{%
+  Carpenter%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  Carpenter:2017ke}
+  \APACinsertmetastar {%
+  Carpenter:2017ke}%
+  \begin{APACrefauthors}%
+  Carpenter, B.%
+  , Gelman, A.%
+  , Hoffman, M\BPBI D.%
+  , Lee, D.%
+  , Goodrich, B.%
+  , Betancourt, M.%
+  \BDBL {}Riddell, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Stan: {{A Probabilistic Programming Language}}} {Stan:
+    {{A Probabilistic Programming Language}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal Of Statistical Software}{76}{1}{1-29}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.18637/jss.v076.i01} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Chamoli%
+  , Ram~Bansal%
+  \BCBL {}\ \BBA {} Dimri%
+  }{%
+  Chamoli%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2007}}%
+  }]{%
+  chamoli:2007}
+  \APACinsertmetastar {%
+  chamoli:2007}%
+  \begin{APACrefauthors}%
+  Chamoli, A.%
+  , Ram~Bansal, A.%
+  \BCBL {}\ \BBA {} Dimri, V\BPBI P.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2007}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Wavelet and Rescaled Range Approach for the {{Hurst}}
+    Coefficient for Short and Long Time Series} {Wavelet and rescaled range
+    approach for the {{Hurst}} coefficient for short and long time
+    series}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Computers \& Geosciences}{33}{1}{83-93}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.cageo.2006.05.008} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  {City of New York}%
+  }{%
+  {City of New York}%
+  }{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  CityofNewYork:2013uh}
+  \APACinsertmetastar {%
+  CityofNewYork:2013uh}%
+  \begin{APACrefauthors}%
+  {City of New York}.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  \APACrefbtitle {A {{Stronger}}, {{More Resilient New York}}} {A {{Stronger}},
+    {{More Resilient New York}}}\ \APACbVolEdTR{}{\BTR{}}.
+  \newblock
+  \APACaddressInstitution{New York}{}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Cook%
+  \ \protect \BOthers {.}}{%
+  Cook%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2010}}%
+  }]{%
+  Cook:2010bz}
+  \APACinsertmetastar {%
+  Cook:2010bz}%
+  \begin{APACrefauthors}%
+  Cook, E\BPBI R.%
+  , Seager, R.%
+  , Heim~Jr, R\BPBI R.%
+  , Vose, R\BPBI S.%
+  , Herweijer, C.%
+  \BCBL {}\ \BBA {} Woodhouse, C.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2010}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Megadroughts in {{North America}}: Placing {{IPCC}}
+    Projections of Hydroclimatic Change in a Long-Term Palaeoclimate Context}
+    {Megadroughts in {{North America}}: Placing {{IPCC}} projections of
+    hydroclimatic change in a long-term palaeoclimate context}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Quaternary Science}{25}{1}{48-61}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/jqs.1303} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Coumou%
+  \ \BBA {} Rahmstorf%
+  }{%
+  Coumou%
+  \ \BBA {} Rahmstorf%
+  }{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  Coumou:2012bc}
+  \APACinsertmetastar {%
+  Coumou:2012bc}%
+  \begin{APACrefauthors}%
+  Coumou, D.%
+  \BCBT {}\ \BBA {} Rahmstorf, S.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2012}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A Decade of Weather Extremes} {A decade of weather
+    extremes}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Nature Climate Change}{10}{7}{P100013-496}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1038/nclimate1452} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Delgado%
+  , Merz%
+  \BCBL {}\ \BBA {} Apel%
+  }{%
+  Delgado%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Delgado:2014ey}
+  \APACinsertmetastar {%
+  Delgado:2014ey}%
+  \begin{APACrefauthors}%
+  Delgado, J\BPBI M.%
+  , Merz, B.%
+  \BCBL {}\ \BBA {} Apel, H.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Projecting Flood Hazard under Climate Change: An
+    Alternative Approach to Model Chains} {Projecting flood hazard under climate
+    change: An alternative approach to model chains}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Natural Hazards and Earth System
+    Science}{14}{6}{1579-1589}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.5194/nhess-14-1579-2014} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Di~Baldassarre%
+  \ \protect \BOthers {.}}{%
+  Di~Baldassarre%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2018}}%
+  }]{%
+  baldassarre:2018}
+  \APACinsertmetastar {%
+  baldassarre:2018}%
+  \begin{APACrefauthors}%
+  Di~Baldassarre, G.%
+  , Kreibich, H.%
+  , Vorogushyn, S.%
+  , Aerts, J.%
+  , {Arnbjerg-Nielsen}, K.%
+  , Barendrecht, M.%
+  \BDBL {}Ward, P\BPBI J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2018}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Hess {{Opinions}}: {{An}} Interdisciplinary Research
+    Agenda to Explore the Unintended Consequences of Structural Flood Protection}
+    {Hess {{Opinions}}: {{An}} interdisciplinary research agenda to explore the
+    unintended consequences of structural flood protection}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrology and Earth System
+    Sciences}{22}{11}{5629-5637}.
+  \newblock
+  \begin{APACrefDOI} \doi{https://doi.org/10.5194/hess-22-5629-2018}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Dittes%
+  , Špačková%
+  , Schoppa%
+  \BCBL {}\ \BBA {} Straub%
+  }{%
+  Dittes%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  Dittes:2017he}
+  \APACinsertmetastar {%
+  Dittes:2017he}%
+  \begin{APACrefauthors}%
+  Dittes, B.%
+  , Špačková, O.%
+  , Schoppa, L.%
+  \BCBL {}\ \BBA {} Straub, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Climate Uncertainty in Flood Protection Planning}
+    {Climate uncertainty in flood protection planning}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrological and Earth System Science}{}{}{1-28}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.5194/hess-2017-576} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Dugan%
+  \ \protect \BOthers {.}}{%
+  Dugan%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2010}}%
+  }]{%
+  dugan:2010}
+  \APACinsertmetastar {%
+  dugan:2010}%
+  \begin{APACrefauthors}%
+  Dugan, P\BPBI J.%
+  , Barlow, C.%
+  , Agostinho, A\BPBI A.%
+  , Baran, E.%
+  , Cada, G\BPBI F.%
+  , Chen, D.%
+  \BDBL {}Winemiller, K\BPBI O.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2010}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Fish {{Migration}}, {{Dams}}, and {{Loss}} of
+    {{Ecosystem Services}} in the {{Mekong Basin}}} {Fish {{Migration}},
+    {{Dams}}, and {{Loss}} of {{Ecosystem Services}} in the {{Mekong
+    Basin}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{AMBIO}{39}{4}{344-348}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1007/s13280-010-0036-1} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Farnham%
+  , {Doss-Gollin}%
+  \BCBL {}\ \BBA {} Lall%
+  }{%
+  Farnham%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2018}}%
+  }]{%
+  Farnham:2018gs}
+  \APACinsertmetastar {%
+  Farnham:2018gs}%
+  \begin{APACrefauthors}%
+  Farnham, D\BPBI J.%
+  , {Doss-Gollin}, J.%
+  \BCBL {}\ \BBA {} Lall, U.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2018}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Regional {{Extreme Precipitation Events}}: {{Robust
+    Inference From Credibly Simulated GCM Variables}}} {Regional {{Extreme
+    Precipitation Events}}: {{Robust Inference From Credibly Simulated GCM
+    Variables}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{}{}{}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2017wr021318} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Feng%
+  , Willemain%
+  \BCBL {}\ \BBA {} Shang%
+  }{%
+  Feng%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2005}}%
+  }]{%
+  feng:2005}
+  \APACinsertmetastar {%
+  feng:2005}%
+  \begin{APACrefauthors}%
+  Feng, H.%
+  , Willemain, T\BPBI R.%
+  \BCBL {}\ \BBA {} Shang, N.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2005}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Wavelet-{{Based Bootstrap}} for {{Time Series
+    Analysis}}} {Wavelet-{{Based Bootstrap}} for {{Time Series
+    Analysis}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Communications in Statistics - Simulation and
+    Computation}{34}{2}{393-413}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1081/SAC-200055722} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  {GEI Consultants, Inc.}%
+  }{%
+  {GEI Consultants, Inc.}%
+  }{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  GEI:gIaEZ-gS}
+  \APACinsertmetastar {%
+  GEI:gIaEZ-gS}%
+  \begin{APACrefauthors}%
+  {GEI Consultants, Inc.}%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  \APACrefbtitle {Tulare {{Lake Storage}} and {{Floodwater Protection Feasibility
+    Report}}} {Tulare {{Lake Storage}} and {{Floodwater Protection Feasibility
+    Report}}}\ \APACbVolEdTR{}{\BTR{}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Gelman%
+  , Simpson%
+  \BCBL {}\ \BBA {} Betancourt%
+  }{%
+  Gelman%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  gelman:2017}
+  \APACinsertmetastar {%
+  gelman:2017}%
+  \begin{APACrefauthors}%
+  Gelman, A.%
+  , Simpson, D.%
+  \BCBL {}\ \BBA {} Betancourt, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Prior Can Often Only Be Understood}} in the
+    {{Context}} of the {{Likelihood}}} {The {{Prior Can Often Only Be
+    Understood}} in the {{Context}} of the {{Likelihood}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Entropy}{19}{10}{555}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.3390/e19100555} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Geweke%
+  \ \BBA {} Porter‐Hudak%
+  }{%
+  Geweke%
+  \ \BBA {} Porter‐Hudak%
+  }{%
+  {\protect \APACyear {1983}}%
+  }]{%
+  geweke:1983}
+  \APACinsertmetastar {%
+  geweke:1983}%
+  \begin{APACrefauthors}%
+  Geweke, J.%
+  \BCBT {}\ \BBA {} Porter‐Hudak, S.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1983}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Estimation}} and {{Application}} of {{Long Memory
+    Time Series Models}}} {The {{Estimation}} and {{Application}} of {{Long
+    Memory Time Series Models}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Time Series Analysis}{4}{4}{221-238}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1111/j.1467-9892.1983.tb00371.x} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Haasnoot%
+  , Kwakkel%
+  , Walker%
+  \BCBL {}\ \BBA {} {ter Maat}%
+  }{%
+  Haasnoot%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  Haasnoot:2013im}
+  \APACinsertmetastar {%
+  Haasnoot:2013im}%
+  \begin{APACrefauthors}%
+  Haasnoot, M.%
+  , Kwakkel, J\BPBI H.%
+  , Walker, W\BPBI E.%
+  \BCBL {}\ \BBA {} {ter Maat}, J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Dynamic Adaptive Policy Pathways: {{A}} Method for
+    Crafting Robust Decisions for a Deeply Uncertain World} {Dynamic adaptive
+    policy pathways: {{A}} method for crafting robust decisions for a deeply
+    uncertain world}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Global Environmental Change}{23}{2}{485-498}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.gloenvcha.2012.12.006} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Haasnoot%
+  , Schellekens%
+  , Beersma%
+  , Middelkoop%
+  \BCBL {}\ \BBA {} Kwadijk%
+  }{%
+  Haasnoot%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2015}}%
+  }]{%
+  Haasnoot:2015dh}
+  \APACinsertmetastar {%
+  Haasnoot:2015dh}%
+  \begin{APACrefauthors}%
+  Haasnoot, M.%
+  , Schellekens, J.%
+  , Beersma, J\BPBI J.%
+  , Middelkoop, H.%
+  \BCBL {}\ \BBA {} Kwadijk, J\BPBI C\BPBI J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2015}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Transient Scenarios for Robust Climate Change Adaptation
+    Illustrated for Water Management in {{The Netherlands}}} {Transient scenarios
+    for robust climate change adaptation illustrated for water management in
+    {{The Netherlands}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Environmental Research Letters}{10}{10}{105008}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1088/1748-9326/10/10/105008} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Held%
+  }{%
+  Held%
+  }{%
+  {\protect \APACyear {2005}}%
+  }]{%
+  Held:2005cj}
+  \APACinsertmetastar {%
+  Held:2005cj}%
+  \begin{APACrefauthors}%
+  Held, I\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2005}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Gap}} between {{Simulation}} and {{Understanding}}
+    in {{Climate Modeling}}} {The {{Gap}} between {{Simulation}} and
+    {{Understanding}} in {{Climate Modeling}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Bulletin of the American Meteorological
+    Society}{86}{11}{1609-1614}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/BAMS-86-11-1609} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Hodgkins%
+  \ \protect \BOthers {.}}{%
+  Hodgkins%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  Hodgkins:2017hw}
+  \APACinsertmetastar {%
+  Hodgkins:2017hw}%
+  \begin{APACrefauthors}%
+  Hodgkins, G\BPBI A.%
+  , Whitfield, P\BPBI H.%
+  , Burn, D\BPBI H.%
+  , Hannaford, J.%
+  , Renard, B.%
+  , Stahl, K.%
+  \BDBL {}Wilson, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Climate-Driven Variability in the Occurrence of Major
+    Floods across {{North America}} and {{Europe}}} {Climate-driven variability
+    in the occurrence of major floods across {{North America}} and
+    {{Europe}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Hydrology}{552}{}{704-717}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.jhydrol.2017.07.027} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Hoffman%
+  \ \BBA {} Gelman%
+  }{%
+  Hoffman%
+  \ \BBA {} Gelman%
+  }{%
+  {\protect \APACyear {2011}}%
+  }]{%
+  Hoffman:2011wm}
+  \APACinsertmetastar {%
+  Hoffman:2011wm}%
+  \begin{APACrefauthors}%
+  Hoffman, M\BPBI D.%
+  \BCBT {}\ \BBA {} Gelman, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2011}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{No}}-{{U}}-{{Turn Sampler}}: {{Adaptively Setting
+    Path Lengths}} in {{Hamiltonian Monte Carlo}}} {The {{No}}-{{U}}-{{Turn
+    Sampler}}: {{Adaptively Setting Path Lengths}} in {{Hamiltonian Monte
+    Carlo}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{arXiv.org}{}{}{}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Hoyer%
+  \ \BBA {} Hamman%
+  }{%
+  Hoyer%
+  \ \BBA {} Hamman%
+  }{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  Hoyer:2017hs}
+  \APACinsertmetastar {%
+  Hoyer:2017hs}%
+  \begin{APACrefauthors}%
+  Hoyer, S.%
+  \BCBT {}\ \BBA {} Hamman, J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Xarray: {{N}}-{{D}} Labeled {{Arrays}} and {{Datasets}}
+    in {{Python}}} {Xarray: {{N}}-{{D}} labeled {{Arrays}} and {{Datasets}} in
+    {{Python}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Open Research Software}{5}{1}{304}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.5334/jors.148} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Hunter%
+  }{%
+  Hunter%
+  }{%
+  {\protect \APACyear {2007}}%
+  }]{%
+  Hunter:2007ih}
+  \APACinsertmetastar {%
+  Hunter:2007ih}%
+  \begin{APACrefauthors}%
+  Hunter, J\BPBI D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2007}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Matplotlib: {{A 2D}} Graphics Environment} {Matplotlib:
+    {{A 2D}} graphics environment}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Computing in Science \& Engineering}{9}{3}{90-95}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1109/MCSE.2007.55} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  {IPCC}%
+  }{%
+  {IPCC}%
+  }{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  IPCC:2012wt}
+  \APACinsertmetastar {%
+  IPCC:2012wt}%
+  \begin{APACrefauthors}%
+  {IPCC}.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYear{2012}.
+  \newblock
+  \APACrefbtitle {Managing the {{Risks}} of {{Extreme Events}} and {{Disasters}}
+    to {{Advance Climate Change Adaptation}}} {Managing the {{Risks}} of
+    {{Extreme Events}} and {{Disasters}} to {{Advance Climate Change
+    Adaptation}}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Jain%
+  \ \BBA {} Lall%
+  }{%
+  Jain%
+  \ \BBA {} Lall%
+  }{%
+  {\protect \APACyear {2001}}%
+  }]{%
+  Jain:2001hz}
+  \APACinsertmetastar {%
+  Jain:2001hz}%
+  \begin{APACrefauthors}%
+  Jain, S.%
+  \BCBT {}\ \BBA {} Lall, U.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2001}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Floods in a Changing Climate: {{Does}} the Past
+    Represent the Future?} {Floods in a changing climate: {{Does}} the past
+    represent the future?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{37}{12}{3193-3205}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/2001WR000495} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Jin%
+  , Neelin%
+  \BCBL {}\ \BBA {} Ghil%
+  }{%
+  Jin%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {1994}}%
+  }]{%
+  Jin:1994wq}
+  \APACinsertmetastar {%
+  Jin:1994wq}%
+  \begin{APACrefauthors}%
+  Jin, F\BPBI F.%
+  , Neelin, J\BPBI D.%
+  \BCBL {}\ \BBA {} Ghil, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1994}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {El-{{Nino}} on the {{Devils Staircase}} - {{Annual
+    Subharmonic Steps}} to {{Chaos}}} {El-{{Nino}} on the {{Devils Staircase}} -
+    {{Annual Subharmonic Steps}} to {{Chaos}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Science}{264}{5155}{70-72}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Jones%
+  , Oliphant%
+  \BCBL {}\ \BBA {} Peterson%
+  }{%
+  Jones%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2001}}%
+  }]{%
+  Jones:2001uv}
+  \APACinsertmetastar {%
+  Jones:2001uv}%
+  \begin{APACrefauthors}%
+  Jones, E.%
+  , Oliphant, T.%
+  \BCBL {}\ \BBA {} Peterson, P.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2001}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {{{SciPy}}: {{Open Source Scientific Tools}} for
+    {{Python}}} {{{SciPy}}: {{Open Source Scientific Tools}} for
+    {{Python}}}.{\BBCQ}
+  \newblock
+
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Jongman%
+  , Ward%
+  \BCBL {}\ \BBA {} Aerts%
+  }{%
+  Jongman%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  Jongman:2012cr}
+  \APACinsertmetastar {%
+  Jongman:2012cr}%
+  \begin{APACrefauthors}%
+  Jongman, B.%
+  , Ward, P\BPBI J.%
+  \BCBL {}\ \BBA {} Aerts, J\BPBI C\BPBI J\BPBI H.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2012}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Global Exposure to River and Coastal Flooding: {{Long}}
+    Term Trends and Changes} {Global exposure to river and coastal flooding:
+    {{Long}} term trends and changes}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Global Environmental Change}{22}{4}{823-835}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.gloenvcha.2012.07.004} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Kiem%
+  , Franks%
+  \BCBL {}\ \BBA {} Kuczera%
+  }{%
+  Kiem%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2002}}%
+  }]{%
+  Kiem:2002kq}
+  \APACinsertmetastar {%
+  Kiem:2002kq}%
+  \begin{APACrefauthors}%
+  Kiem, A\BPBI S.%
+  , Franks, S\BPBI W.%
+  \BCBL {}\ \BBA {} Kuczera, G.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2002}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Multi-Decadal Variability of Flood Risk} {Multi-decadal
+    variability of flood risk}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Geophysical Research Letters}{30}{2}{3}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/2002GL015992} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Koutsoyiannis%
+  }{%
+  Koutsoyiannis%
+  }{%
+  {\protect \APACyear {2003}}%
+  }]{%
+  koutsoyiannis:2003}
+  \APACinsertmetastar {%
+  koutsoyiannis:2003}%
+  \begin{APACrefauthors}%
+  Koutsoyiannis, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2003}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Climate Change, the {{Hurst}} Phenomenon, and
+    Hydrological Statistics} {Climate change, the {{Hurst}} phenomenon, and
+    hydrological statistics}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrological Sciences Journal}{48}{1}{3-24}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1623/hysj.48.1.3.43481} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Kwon%
+  , Lall%
+  \BCBL {}\ \BBA {} Khalil%
+  }{%
+  Kwon%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2007}}%
+  }]{%
+  Kwon:2007fj}
+  \APACinsertmetastar {%
+  Kwon:2007fj}%
+  \begin{APACrefauthors}%
+  Kwon, H\BHBI H.%
+  , Lall, U.%
+  \BCBL {}\ \BBA {} Khalil, A\BPBI F.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2007}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Stochastic Simulation Model for Nonstationary Time
+    Series Using an Autoregressive Wavelet Decomposition: {{Applications}} to
+    Rainfall and Temperature} {Stochastic simulation model for nonstationary time
+    series using an autoregressive wavelet decomposition: {{Applications}} to
+    rainfall and temperature}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{43}{5}{287}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/2006WR005258} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Lempert%
+  \ \BBA {} Collins%
+  }{%
+  Lempert%
+  \ \BBA {} Collins%
+  }{%
+  {\protect \APACyear {2007}}%
+  }]{%
+  lempert:2007}
+  \APACinsertmetastar {%
+  lempert:2007}%
+  \begin{APACrefauthors}%
+  Lempert, R\BPBI J.%
+  \BCBT {}\ \BBA {} Collins, M\BPBI T.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2007}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Managing the {{Risk}} of {{Uncertain Threshold
+    Responses}}: {{Comparison}} of {{Robust}}, {{Optimum}}, and {{Precautionary
+    Approaches}}} {Managing the {{Risk}} of {{Uncertain Threshold Responses}}:
+    {{Comparison}} of {{Robust}}, {{Optimum}}, and {{Precautionary
+    Approaches}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Risk Analysis}{27}{4}{1009-1026}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1111/j.1539-6924.2007.00940.x} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Lima%
+  , Lall%
+  , Troy%
+  \BCBL {}\ \BBA {} Devineni%
+  }{%
+  Lima%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Lima:2016kd}
+  \APACinsertmetastar {%
+  Lima:2016kd}%
+  \begin{APACrefauthors}%
+  Lima, C\BPBI H\BPBI R.%
+  , Lall, U.%
+  , Troy, T.%
+  \BCBL {}\ \BBA {} Devineni, N.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A Hierarchical {{Bayesian GEV}} Model for Improving
+    Local and Regional Flood Quantile Estimates} {A hierarchical {{Bayesian GEV}}
+    model for improving local and regional flood quantile estimates}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Hydrology}{541}{}{816-823}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.jhydrol.2016.07.042} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Lovejoy%
+  }{%
+  Lovejoy%
+  }{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  lovejoy:2013}
+  \APACinsertmetastar {%
+  lovejoy:2013}%
+  \begin{APACrefauthors}%
+  Lovejoy, S.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {What {{Is Climate}}?} {What {{Is Climate}}?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{EOS, Transactions American Geophysical
+    Union}{94}{1}{1-2}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2013EO010001} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Lovejoy%
+  \ \BBA {} Schertzer%
+  }{%
+  Lovejoy%
+  \ \BBA {} Schertzer%
+  }{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  lovejoy:2012}
+  \APACinsertmetastar {%
+  lovejoy:2012}%
+  \begin{APACrefauthors}%
+  Lovejoy, S.%
+  \BCBT {}\ \BBA {} Schertzer, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2012}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Low-{{Frequency Weather}} and the {{Emergence}} of the
+    {{Climate}}} {Low-{{Frequency Weather}} and the {{Emergence}} of the
+    {{Climate}}}.{\BBCQ}
+  \newblock
+  \BIn{} \APACrefbtitle {Extreme {{Events}} and {{Natural Hazards}}: {{The
+    Complexity Perspective}}} {Extreme {{Events}} and {{Natural Hazards}}: {{The
+    Complexity Perspective}}}\ (\BPG~231-254).
+  \newblock
+  \APACaddressPublisher{}{{American Geophysical Union (AGU)}}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/2011GM001087} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Lovejoy%
+  \ \BBA {} Schertzer%
+  }{%
+  Lovejoy%
+  \ \BBA {} Schertzer%
+  }{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  lovejoy:2013a}
+  \APACinsertmetastar {%
+  lovejoy:2013a}%
+  \begin{APACrefauthors}%
+  Lovejoy, S.%
+  \BCBT {}\ \BBA {} Schertzer, D.%
+  \end{APACrefauthors}%
+  \ (\BEDS).
+  \unskip\
+  \newblock
+  \APACrefYear{2013}.
+  \newblock
+  \APACrefbtitle {The {{Weather}} and {{Climate}} {{Emergent Laws}} and
+    {{Multifractal Cascades}} /} {The {{Weather}} and {{Climate}} {{Emergent
+    Laws}} and {{Multifractal Cascades}} /}.
+  \newblock
+  \APACaddressPublisher{Cambridge :}{{Cambridge University Press,}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Mandelbrot%
+  }{%
+  Mandelbrot%
+  }{%
+  {\protect \APACyear {1985}}%
+  }]{%
+  mandelbrot:1985}
+  \APACinsertmetastar {%
+  mandelbrot:1985}%
+  \begin{APACrefauthors}%
+  Mandelbrot, B\BPBI B.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1985}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Self-{{Affine Fractals}} and {{Fractal Dimension}}}
+    {Self-{{Affine Fractals}} and {{Fractal Dimension}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Physica Scripta}{32}{4}{257--260}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1088/0031-8949/32/4/001} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Mandelbrot%
+  \ \BBA {} Wallis%
+  }{%
+  Mandelbrot%
+  \ \BBA {} Wallis%
+  }{%
+  {\protect \APACyear {1969}}%
+  }]{%
+  mandelbrot:1969}
+  \APACinsertmetastar {%
+  mandelbrot:1969}%
+  \begin{APACrefauthors}%
+  Mandelbrot, B\BPBI B.%
+  \BCBT {}\ \BBA {} Wallis, J\BPBI R.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1969}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Robustness of the Rescaled Range {{R}}/{{S}} in the
+    Measurement of Noncyclic Long Run Statistical Dependence} {Robustness of the
+    rescaled range {{R}}/{{S}} in the measurement of noncyclic long run
+    statistical dependence}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{5}{5}{967-988}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/WR005i005p00967} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Mantua%
+  , Hare%
+  , Zhang%
+  , Wallace%
+  \BCBL {}\ \BBA {} Francis%
+  }{%
+  Mantua%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {1997}}%
+  }]{%
+  Mantua:1997kj}
+  \APACinsertmetastar {%
+  Mantua:1997kj}%
+  \begin{APACrefauthors}%
+  Mantua, N\BPBI J.%
+  , Hare, S\BPBI R.%
+  , Zhang, Y.%
+  , Wallace, J\BPBI M.%
+  \BCBL {}\ \BBA {} Francis, R\BPBI C.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1997}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A {{Pacific}} Interdecadal Climate Oscillation with
+    Impacts on Salmon Production} {A {{Pacific}} interdecadal climate oscillation
+    with impacts on salmon production}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Bulletin of the American Meteorological
+    Society}{78}{6}{1069-1079}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/1520-0477(1997)078<1069:APICOW>2.0.CO;2}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Markonis%
+  \ \BBA {} Koutsoyiannis%
+  }{%
+  Markonis%
+  \ \BBA {} Koutsoyiannis%
+  }{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  markonis:2013}
+  \APACinsertmetastar {%
+  markonis:2013}%
+  \begin{APACrefauthors}%
+  Markonis, Y.%
+  \BCBT {}\ \BBA {} Koutsoyiannis, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Climatic {{Variability Over Time Scales Spanning Nine
+    Orders}} of {{Magnitude}}: {{Connecting Milankovitch Cycles}} with
+    {{Hurst}}–{{Kolmogorov Dynamics}}} {Climatic {{Variability Over Time Scales
+    Spanning Nine Orders}} of {{Magnitude}}: {{Connecting Milankovitch Cycles}}
+    with {{Hurst}}–{{Kolmogorov Dynamics}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Surveys in Geophysics}{34}{2}{181-207}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1007/s10712-012-9208-9} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Maruyama%
+  }{%
+  Maruyama%
+  }{%
+  {\protect \APACyear {2018}}%
+  }]{%
+  maruyama:2018}
+  \APACinsertmetastar {%
+  maruyama:2018}%
+  \begin{APACrefauthors}%
+  Maruyama, F.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2018}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Relation}} among the {{Solar Activity}}, the
+    {{Total Ozone}}, {{QBO}}, {{NAO}}, and {{ENSO}} by {{Wavelet}}-{{Based
+    Multifractal Analysis}}} {The {{Relation}} among the {{Solar Activity}}, the
+    {{Total Ozone}}, {{QBO}}, {{NAO}}, and {{ENSO}} by {{Wavelet}}-{{Based
+    Multifractal Analysis}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Applied Mathematics and
+    Physics}{06}{}{1301}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.4236/jamp.2018.66109} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  McKinney%
+  }{%
+  McKinney%
+  }{%
+  {\protect \APACyear {2010}}%
+  }]{%
+  McKinney:2010un}
+  \APACinsertmetastar {%
+  McKinney:2010un}%
+  \begin{APACrefauthors}%
+  McKinney, W.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2010}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Data {{Structures}} for {{Statistical Computing}} in
+    {{Python}}} {Data {{Structures}} for {{Statistical Computing}} in
+    {{Python}}}.{\BBCQ}
+  \newblock
+  \BIn{} \APACrefbtitle {Proceedings of the 9th {{Python}} in {{Science}} Ldots.}
+    {Proceedings of the 9th {{Python}} in {{Science}} ldots.}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  B.~Merz%
+  \ \protect \BOthers {.}}{%
+  B.~Merz%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Merz:2014gf}
+  \APACinsertmetastar {%
+  Merz:2014gf}%
+  \begin{APACrefauthors}%
+  Merz, B.%
+  , Aerts, J\BPBI C\BPBI J\BPBI H.%
+  , {Arnbjerg-Nielsen}, K.%
+  , Baldi, M.%
+  , Becker, A.%
+  , Bichet, A.%
+  \BDBL {}Nied, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Floods and Climate: Emerging Perspectives for Flood Risk
+    Assessment and Management} {Floods and climate: Emerging perspectives for
+    flood risk assessment and management}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Natural Hazards and Earth System
+    Science}{14}{7}{1921-1942}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.5194/nhess-14-1921-2014} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  R.~Merz%
+  \ \BBA {} Blöschl%
+  }{%
+  R.~Merz%
+  \ \BBA {} Blöschl%
+  }{%
+  {\protect \APACyear {2008}}%
+  }]{%
+  Merz:2008eh}
+  \APACinsertmetastar {%
+  Merz:2008eh}%
+  \begin{APACrefauthors}%
+  Merz, R.%
+  \BCBT {}\ \BBA {} Blöschl, G.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2008}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Flood Frequency Hydrology: 1. {{Temporal}}, Spatial, and
+    Causal Expansion of Information} {Flood frequency hydrology: 1. {{Temporal}},
+    spatial, and causal expansion of information}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{44}{8}{3}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1029/2007WR006744} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Milly%
+  \ \protect \BOthers {.}}{%
+  Milly%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2008}}%
+  }]{%
+  Milly:2008dg}
+  \APACinsertmetastar {%
+  Milly:2008dg}%
+  \begin{APACrefauthors}%
+  Milly, P\BPBI C\BPBI D.%
+  , Betancourt, J.%
+  , Falkenmark, M.%
+  , Hirsch, R\BPBI M.%
+  , Kundzewicz, Z\BPBI W.%
+  , Lettenmaier, D\BPBI P.%
+  \BCBL {}\ \BBA {} Stouffer, R\BPBI J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2008}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Stationarity {{Is Dead}}: {{Whither Water Management}}?}
+    {Stationarity {{Is Dead}}: {{Whither Water Management}}?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Science}{319}{5863}{573-574}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1126/science.1151915} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Montanari%
+  }{%
+  Montanari%
+  }{%
+  {\protect \APACyear {2003}}%
+  }]{%
+  montanari:2003}
+  \APACinsertmetastar {%
+  montanari:2003}%
+  \begin{APACrefauthors}%
+  Montanari, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2003}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Long-Range Dependence in Hydrology} {Long-range
+    dependence in hydrology}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Theory and applications of long-range
+    dependence}{}{}{461-472}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Montanari%
+  \ \BBA {} Koutsoyiannis%
+  }{%
+  Montanari%
+  \ \BBA {} Koutsoyiannis%
+  }{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Montanari:2014hl}
+  \APACinsertmetastar {%
+  Montanari:2014hl}%
+  \begin{APACrefauthors}%
+  Montanari, A.%
+  \BCBT {}\ \BBA {} Koutsoyiannis, D.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Modeling and Mitigating Natural Hazards:
+    {{Stationarity}} Is Immortal!} {Modeling and mitigating natural hazards:
+    {{Stationarity}} is immortal!}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{50}{12}{9748-9756}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2014WR016092} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Müller%
+  , Wild%
+  , Driesse%
+  \BCBL {}\ \BBA {} Behrens%
+  }{%
+  Müller%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Muller:2014fc}
+  \APACinsertmetastar {%
+  Muller:2014fc}%
+  \begin{APACrefauthors}%
+  Müller, B.%
+  , Wild, M.%
+  , Driesse, A.%
+  \BCBL {}\ \BBA {} Behrens, K.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Rethinking Solar Resource Assessments in the Context of
+    Global Dimming and Brightening} {Rethinking solar resource assessments in the
+    context of global dimming and brightening}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Solar Energy}{99}{}{272-282}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.solener.2013.11.013} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Newman%
+  \ \protect \BOthers {.}}{%
+  Newman%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  newman:2016}
+  \APACinsertmetastar {%
+  newman:2016}%
+  \begin{APACrefauthors}%
+  Newman, M.%
+  , Alexander, M\BPBI A.%
+  , Ault, T\BPBI R.%
+  , Cobb, K\BPBI M.%
+  , Deser, C.%
+  , Di~Lorenzo, E.%
+  \BDBL {}Smith, C\BPBI A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Pacific Decadal Oscillation}}, {{Revisited}}} {The
+    {{Pacific Decadal Oscillation}}, {{Revisited}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Climate}{29}{12}{4399-4427}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/JCLI-D-15-0508.1} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  O'Gorman%
+  \ \BBA {} Schneider%
+  }{%
+  O'Gorman%
+  \ \BBA {} Schneider%
+  }{%
+  {\protect \APACyear {2009}}%
+  }]{%
+  OGorman:2009hj}
+  \APACinsertmetastar {%
+  OGorman:2009hj}%
+  \begin{APACrefauthors}%
+  O'Gorman, P\BPBI A.%
+  \BCBT {}\ \BBA {} Schneider, T.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2009}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The Physical Basis for Increases in Precipitation
+    Extremes in Simulations of 21st-Century Climate Change} {The physical basis
+    for increases in precipitation extremes in simulations of 21st-century
+    climate change}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Proceedings of the National Academy of Sciences of the
+    United States of America}{106}{35}{14773-14777}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1073/pnas.0907610106} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  O’Connell%
+  \ \protect \BOthers {.}}{%
+  O’Connell%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  oconnell:2016}
+  \APACinsertmetastar {%
+  oconnell:2016}%
+  \begin{APACrefauthors}%
+  O’Connell, P\BPBI E.%
+  , Koutsoyiannis, D.%
+  , Lins, H\BPBI F.%
+  , Markonis, Y.%
+  , Montanari, A.%
+  \BCBL {}\ \BBA {} Cohn, T.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The Scientific Legacy of {{Harold Edwin Hurst}}
+    (1880–1978)} {The scientific legacy of {{Harold Edwin Hurst}}
+    (1880–1978)}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrological Sciences Journal}{61}{9}{1571-1590}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1080/02626667.2015.1125998} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Palmer%
+  }{%
+  Palmer%
+  }{%
+  {\protect \APACyear {1993}}%
+  }]{%
+  Palmer:1993ef}
+  \APACinsertmetastar {%
+  Palmer:1993ef}%
+  \begin{APACrefauthors}%
+  Palmer, T\BPBI N.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1993}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A Nonlinear Dynamical Perspective on Climate Change} {A
+    nonlinear dynamical perspective on climate change}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Weather}{48}{10}{314-326}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/j.1477-8696.1993.tb05802.x} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Papakonstantinou%
+  , Maier%
+  , Kaldany%
+  , Belelieu%
+  \BCBL {}\ \BBA {} Drexler%
+  }{%
+  Papakonstantinou%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Papakonstantinou:2016ve}
+  \APACinsertmetastar {%
+  Papakonstantinou:2016ve}%
+  \begin{APACrefauthors}%
+  Papakonstantinou, V.%
+  , Maier, T.%
+  , Kaldany, R\BPBI R.%
+  , Belelieu, A.%
+  \BCBL {}\ \BBA {} Drexler, M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  \APACrefbtitle {Risk {{Mitigation Instruments}} in {{Infrastructure}}} {Risk
+    {{Mitigation Instruments}} in {{Infrastructure}}}\ \APACbVolEdTR{}{\BTR{}}.
+  \newblock
+  \APACaddressInstitution{}{{World Economic Forum}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Peduzzi%
+  \ \protect \BOthers {.}}{%
+  Peduzzi%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  Peduzzi:2012iq}
+  \APACinsertmetastar {%
+  Peduzzi:2012iq}%
+  \begin{APACrefauthors}%
+  Peduzzi, P.%
+  , Chatenoux, B.%
+  , Dao, H.%
+  , De~Bono, A.%
+  , Herold, C.%
+  , Kossin, J.%
+  \BDBL {}Nordbeck, O.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2012}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Global Trends in Tropical Cyclone Risk} {Global trends
+    in tropical cyclone risk}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Nature Climate Change}{2}{4}{289-294}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1038/nclimate1410} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Powers%
+  }{%
+  Powers%
+  }{%
+  {\protect \APACyear {2003}}%
+  }]{%
+  powers:2003}
+  \APACinsertmetastar {%
+  powers:2003}%
+  \begin{APACrefauthors}%
+  Powers, K.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2003}{}{}.
+  \newblock
+  \APACrefbtitle {Benefit-{{Cost Analysis}} and the {{Discount Rate}} for the
+    {{Corps}} of {{Engineers}}' {{Water Resource Projects}}: {{Theory}} and
+    {{Practice}}} {Benefit-{{Cost Analysis}} and the {{Discount Rate}} for the
+    {{Corps}} of {{Engineers}}' {{Water Resource Projects}}: {{Theory}} and
+    {{Practice}}}\ \APACbVolEdTR{}{\BTR{}\ \BNUM\ RL31976}.
+  \newblock
+  \APACaddressInstitution{Washington, DC}{{Congressional Research Service}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Rabiner%
+  \ \BBA {} Juang%
+  }{%
+  Rabiner%
+  \ \BBA {} Juang%
+  }{%
+  {\protect \APACyear {1986}}%
+  }]{%
+  Rabiner:1986jk}
+  \APACinsertmetastar {%
+  Rabiner:1986jk}%
+  \begin{APACrefauthors}%
+  Rabiner, L.%
+  \BCBT {}\ \BBA {} Juang, B.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1986}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {An {{Introduction}} to {{Hidden Markov Models}}} {An
+    {{Introduction}} to {{Hidden Markov Models}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{IEEE ASSP Magazine}{3}{1}{4-16}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1109/MASSP.1986.1165342} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Ramesh%
+  , Cane%
+  , Seager%
+  \BCBL {}\ \BBA {} Lee%
+  }{%
+  Ramesh%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Ramesh:2016hf}
+  \APACinsertmetastar {%
+  Ramesh:2016hf}%
+  \begin{APACrefauthors}%
+  Ramesh, N.%
+  , Cane, M\BPBI A.%
+  , Seager, R.%
+  \BCBL {}\ \BBA {} Lee, D\BPBI E.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Predictability and Prediction of Persistent Cool States
+    of the {{Tropical Pacific Ocean}}} {Predictability and prediction of
+    persistent cool states of the {{Tropical Pacific Ocean}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Climate Dynamics}{49}{7-8}{2291-2307}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1007/s00382-016-3446-3} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Roesch%
+  \ \BBA {} Schmidbauer%
+  }{%
+  Roesch%
+  \ \BBA {} Schmidbauer%
+  }{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Roesch:wlBQQoIs}
+  \APACinsertmetastar {%
+  Roesch:wlBQQoIs}%
+  \begin{APACrefauthors}%
+  Roesch, A.%
+  \BCBT {}\ \BBA {} Schmidbauer, H.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYear{2016}.
+  \newblock
+  \APACrefbtitle {{{WaveletComp}}: Computational Wavelet Analysis. {{R}} Package
+    Version 1.1} {{{WaveletComp}}: Computational wavelet analysis. {{R}} package
+    version 1.1}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Rojo~Hernandez%
+  , Lall%
+  \BCBL {}\ \BBA {} Mesa%
+  }{%
+  Rojo~Hernandez%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  rojohernandez:2017}
+  \APACinsertmetastar {%
+  rojohernandez:2017}%
+  \begin{APACrefauthors}%
+  Rojo~Hernandez, J\BPBI D.%
+  , Lall, U.%
+  \BCBL {}\ \BBA {} Mesa, O\BPBI J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {{{ENSO Dynamics}} and {{Trends}}, an {{Alternate View}}}
+    {{{ENSO Dynamics}} and {{Trends}}, an {{Alternate View}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{AGU Fall Meeting Abstracts}{41}{}{}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Ropelewski%
+  \ \BBA {} Halpert%
+  }{%
+  Ropelewski%
+  \ \BBA {} Halpert%
+  }{%
+  {\protect \APACyear {1987}}%
+  }]{%
+  Ropelewski:1987do}
+  \APACinsertmetastar {%
+  Ropelewski:1987do}%
+  \begin{APACrefauthors}%
+  Ropelewski, C\BPBI F.%
+  \BCBT {}\ \BBA {} Halpert, M\BPBI S.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1987}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Global and {{Regional Scale Precipitation Patterns
+    Associated}} with the {{El}}-{{Nino Southern Oscillation}}} {Global and
+    {{Regional Scale Precipitation Patterns Associated}} with the {{El}}-{{Nino
+    Southern Oscillation}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Monthly Weather Review}{115}{8}{1606-1626}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/1520-0493(1987)115<1606:GARSPP>2.0.CO;2}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Salas%
+  , Obeysekera%
+  \BCBL {}\ \BBA {} Vogel%
+  }{%
+  Salas%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2018}}%
+  }]{%
+  Salas:2018ge}
+  \APACinsertmetastar {%
+  Salas:2018ge}%
+  \begin{APACrefauthors}%
+  Salas, J\BPBI D.%
+  , Obeysekera, J.%
+  \BCBL {}\ \BBA {} Vogel, R\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2018}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Techniques for Assessing Water Infrastructure for
+    Nonstationary Extreme Events: A Review} {Techniques for assessing water
+    infrastructure for nonstationary extreme events: A review}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Hydrological Sciences Journal}{63}{3}{325-352}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1080/02626667.2018.1426858} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Sarachik%
+  \ \BBA {} Cane%
+  }{%
+  Sarachik%
+  \ \BBA {} Cane%
+  }{%
+  {\protect \APACyear {2009}}%
+  }]{%
+  Sarachik:2009dr}
+  \APACinsertmetastar {%
+  Sarachik:2009dr}%
+  \begin{APACrefauthors}%
+  Sarachik, E\BPBI S.%
+  \BCBT {}\ \BBA {} Cane, M\BPBI A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYear{2009}.
+  \newblock
+  \APACrefbtitle {The {{El Niño}}–{{Southern Oscillation Phenomenon}}} {The
+    {{El Niño}}–{{Southern Oscillation Phenomenon}}}.
+  \newblock
+  \APACaddressPublisher{Cambridge}{{Cambridge University Press}}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1017/CBO9780511817496} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Schreiber%
+  }{%
+  Schreiber%
+  }{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  schreiber:2017}
+  \APACinsertmetastar {%
+  schreiber:2017}%
+  \begin{APACrefauthors}%
+  Schreiber, J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Pomegranate: {{Fast}} and {{Flexible Probabilistic
+    Modeling}} in {{Python}}} {Pomegranate: {{Fast}} and {{Flexible Probabilistic
+    Modeling}} in {{Python}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Machine Learning
+    Research}{18}{1}{5992--5997}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Selvam%
+  }{%
+  Selvam%
+  }{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  selvam:2017}
+  \APACinsertmetastar {%
+  selvam:2017}%
+  \begin{APACrefauthors}%
+  Selvam, A\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Universal {{Inverse Power}}-{{Law Distribution}} for
+    {{Fractal Fluctuations}} in {{Dynamical Systems}}: {{Applications}} for
+    {{Predictability}} of {{Inter}}-{{Annual Variability}} of {{Indian}} and
+    {{USA Region Rainfall}}} {Universal {{Inverse Power}}-{{Law Distribution}}
+    for {{Fractal Fluctuations}} in {{Dynamical Systems}}: {{Applications}} for
+    {{Predictability}} of {{Inter}}-{{Annual Variability}} of {{Indian}} and
+    {{USA Region Rainfall}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Pure and Applied Geophysics}{174}{1}{413-426}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1007/s00024-016-1394-9} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Serinaldi%
+  \ \BBA {} Kilsby%
+  }{%
+  Serinaldi%
+  \ \BBA {} Kilsby%
+  }{%
+  {\protect \APACyear {2015}}%
+  }]{%
+  Serinaldi:2015bq}
+  \APACinsertmetastar {%
+  Serinaldi:2015bq}%
+  \begin{APACrefauthors}%
+  Serinaldi, F.%
+  \BCBT {}\ \BBA {} Kilsby, C\BPBI G.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2015}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Stationarity {{Is Undead}}: {{Uncertainty Dominates}}
+    the {{Distribution}} of {{Extremes}}} {Stationarity {{Is Undead}}:
+    {{Uncertainty Dominates}} the {{Distribution}} of {{Extremes}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Advances in Water Resources}{77}{}{17-36}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1016/j.advwatres.2014.12.013} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Shaw%
+  \ \protect \BOthers {.}}{%
+  Shaw%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Shaw:2016bo}
+  \APACinsertmetastar {%
+  Shaw:2016bo}%
+  \begin{APACrefauthors}%
+  Shaw, T\BPBI A.%
+  , Baldwin, M.%
+  , Barnes, E\BPBI A.%
+  , Caballero, R.%
+  , Garfinkel, C\BPBI I.%
+  , Hwang, Y\BPBI T.%
+  \BDBL {}Voigt, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Storm {{Track Processes}} and the {{Opposing
+    Influences}} of {{Climate Change}}} {Storm {{Track Processes}} and the
+    {{Opposing Influences}} of {{Climate Change}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Nature Geoscience}{9}{9}{656-664}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1038/ngeo2783} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Simonsen%
+  , Hansen%
+  \BCBL {}\ \BBA {} Nes%
+  }{%
+  Simonsen%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {1998}}%
+  }]{%
+  simonsen:1998}
+  \APACinsertmetastar {%
+  simonsen:1998}%
+  \begin{APACrefauthors}%
+  Simonsen, I.%
+  , Hansen, A.%
+  \BCBL {}\ \BBA {} Nes, O\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1998}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Determination of the {{Hurst}} Exponent by Use of
+    Wavelet Transforms} {Determination of the {{Hurst}} exponent by use of
+    wavelet transforms}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Physical Review E}{58}{3}{2779-2787}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1103/PhysRevE.58.2779} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Simpson%
+  , Rue%
+  , Riebler%
+  , Martins%
+  \BCBL {}\ \BBA {} Sørbye%
+  }{%
+  Simpson%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2017}}%
+  }]{%
+  simpson:2017}
+  \APACinsertmetastar {%
+  simpson:2017}%
+  \begin{APACrefauthors}%
+  Simpson, D.%
+  , Rue, H.%
+  , Riebler, A.%
+  , Martins, T\BPBI G.%
+  \BCBL {}\ \BBA {} Sørbye, S\BPBI H.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2017}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Penalising {{Model Component Complexity}}: {{A
+    Principled}}, {{Practical Approach}} to {{Constructing Priors}}} {Penalising
+    {{Model Component Complexity}}: {{A Principled}}, {{Practical Approach}} to
+    {{Constructing Priors}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Statistical Science}{32}{1}{1-28}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1214/16-STS576} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Sodastrom%
+  , Sokolove%
+  \BCBL {}\ \BBA {} Fairfax%
+  }{%
+  Sodastrom%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {1999}}%
+  }]{%
+  sodastrom:1999}
+  \APACinsertmetastar {%
+  sodastrom:1999}%
+  \begin{APACrefauthors}%
+  Sodastrom, E.%
+  , Sokolove, J.%
+  \BCBL {}\ \BBA {} Fairfax, S\BPBI K.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1999}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Federal {{Reserved Water Rights Applied}} to {{School
+    Trust Lands Land}} and {{Water Division}}: {{Article}}} {Federal {{Reserved
+    Water Rights Applied}} to {{School Trust Lands Land}} and {{Water Division}}:
+    {{Article}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Land and Water Law Review}{34}{}{1-38}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Spence%
+  \ \BBA {} Brown%
+  }{%
+  Spence%
+  \ \BBA {} Brown%
+  }{%
+  {\protect \APACyear {2016}}%
+  }]{%
+  Spence:2016ca}
+  \APACinsertmetastar {%
+  Spence:2016ca}%
+  \begin{APACrefauthors}%
+  Spence, C\BPBI M.%
+  \BCBT {}\ \BBA {} Brown, C\BPBI M.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2016}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Nonstationary {{Decision Model}} for {{Flood Risk
+    Decision Scaling}}} {Nonstationary {{Decision Model}} for {{Flood Risk
+    Decision Scaling}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Water Resources Research}{52}{11}{8650-8667}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/2016WR018981} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Swierczynski%
+  \ \protect \BOthers {.}}{%
+  Swierczynski%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2012}}%
+  }]{%
+  Swierczynski:2012km}
+  \APACinsertmetastar {%
+  Swierczynski:2012km}%
+  \begin{APACrefauthors}%
+  Swierczynski, T.%
+  , Brauer, A.%
+  , Lauterbach, S.%
+  , {Martín-Puertas}, C.%
+  , Dulski, P.%
+  , {von Grafenstein}, U.%
+  \BCBL {}\ \BBA {} Rohr, C.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2012}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A 1600 {{Yr Seasonally Resolved Record}} of
+    {{Decadal}}-{{Scale Flood Variability}} from the {{Austrian Pre}}-{{Alps}}}
+    {A 1600 {{Yr Seasonally Resolved Record}} of {{Decadal}}-{{Scale Flood
+    Variability}} from the {{Austrian Pre}}-{{Alps}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Geology}{40}{11}{1047-1050}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1130/G33493.1} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Torrence%
+  \ \BBA {} Compo%
+  }{%
+  Torrence%
+  \ \BBA {} Compo%
+  }{%
+  {\protect \APACyear {1998}}%
+  }]{%
+  Torrence:1998jp}
+  \APACinsertmetastar {%
+  Torrence:1998jp}%
+  \begin{APACrefauthors}%
+  Torrence, C.%
+  \BCBT {}\ \BBA {} Compo, G\BPBI P.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1998}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A {{Practical Guide}} to {{Wavelet Analysis}}} {A
+    {{Practical Guide}} to {{Wavelet Analysis}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Bulletin of the American Meteorological
+    Society}{79}{1}{61-78}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/1520-0477(1998)079<0061:APGTWA>2.0.CO;2}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Trenberth%
+  , Dai%
+  , Rasmussen%
+  \BCBL {}\ \BBA {} Parsons%
+  }{%
+  Trenberth%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2003}}%
+  }]{%
+  Trenberth:2003bj}
+  \APACinsertmetastar {%
+  Trenberth:2003bj}%
+  \begin{APACrefauthors}%
+  Trenberth, K\BPBI E.%
+  , Dai, A.%
+  , Rasmussen, R\BPBI M.%
+  \BCBL {}\ \BBA {} Parsons, D\BPBI B.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2003}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{Changing Character}} of {{Precipitation}}} {The
+    {{Changing Character}} of {{Precipitation}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Bulletin of the American Meteorological
+    Society}{84}{9}{1205-1217}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/BAMS-84-9-1205} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  {USACE}%
+  }{%
+  {USACE}%
+  }{%
+  {\protect \APACyear {2007}}%
+  }]{%
+  USACE:2007ta}
+  \APACinsertmetastar {%
+  USACE:2007ta}%
+  \begin{APACrefauthors}%
+  {USACE}.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2007}{}{}.
+  \newblock
+  \APACrefbtitle {Flood {{Damage Analysis Santa Maria Levee Project}}} {Flood
+    {{Damage Analysis Santa Maria Levee Project}}}\ \APACbVolEdTR{}{\BTR{}}.
+  \newblock
+  \APACaddressInstitution{Los Angeles, CA}{{United States Army Corps of Engineers
+    Los Angeles District}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  {USACE}%
+  }{%
+  {USACE}%
+  }{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  USACE:2014vn}
+  \APACinsertmetastar {%
+  USACE:2014vn}%
+  \begin{APACrefauthors}%
+  {USACE}.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  \APACrefbtitle {Dallas {{Floodway}}, {{Dallas TX}}} {Dallas {{Floodway}},
+    {{Dallas TX}}}\ \APACbVolEdTR{}{\BTR{}}.
+  \newblock
+  \APACaddressInstitution{Fort Worth, TX}{{United States Army Corps of Engineers,
+    Fort Worth District}}.
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  van~der Walt%
+  , Colbert%
+  \BCBL {}\ \BBA {} Varoquaux%
+  }{%
+  van~der Walt%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2011}}%
+  }]{%
+  vanderWalt:2011dp}
+  \APACinsertmetastar {%
+  vanderWalt:2011dp}%
+  \begin{APACrefauthors}%
+  van~der Walt, S.%
+  , Colbert, S\BPBI C.%
+  \BCBL {}\ \BBA {} Varoquaux, G.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2011}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {The {{NumPy Array}}: {{A Structure}} for {{Efficient
+    Numerical Computation}}} {The {{NumPy Array}}: {{A Structure}} for
+    {{Efficient Numerical Computation}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Computing in Science \& Engineering}{13}{2}{22-30}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1109/MCSE.2011.37} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Walker%
+  , Haasnoot%
+  \BCBL {}\ \BBA {} Kwakkel%
+  }{%
+  Walker%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  Walker:2013gi}
+  \APACinsertmetastar {%
+  Walker:2013gi}%
+  \begin{APACrefauthors}%
+  Walker, W.%
+  , Haasnoot, M.%
+  \BCBL {}\ \BBA {} Kwakkel, J.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Adapt or {{Perish}}: {{A Review}} of {{Planning
+    Approaches}} for {{Adaptation}} under {{Deep Uncertainty}}} {Adapt or
+    {{Perish}}: {{A Review}} of {{Planning Approaches}} for {{Adaptation}} under
+    {{Deep Uncertainty}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Sustainability}{5}{3}{955-979}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.3390/su5030955} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Ward%
+  \ \protect \BOthers {.}}{%
+  Ward%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Ward:2014gg}
+  \APACinsertmetastar {%
+  Ward:2014gg}%
+  \begin{APACrefauthors}%
+  Ward, P\BPBI J.%
+  , Jongman, B.%
+  , Kummu, M.%
+  , Dettinger, M\BPBI D.%
+  , Sperna~Weiland, F\BPBI C.%
+  \BCBL {}\ \BBA {} Winsemius, H\BPBI C.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Strong {{Influence}} of {{El Nino Southern Oscillation}}
+    on {{Flood Risk}} around the {{World}}} {Strong {{Influence}} of {{El Nino
+    Southern Oscillation}} on {{Flood Risk}} around the {{World}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Proceedings of the National Academy of Sciences of the
+    United States of America}{111}{44}{15659-15664}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1073/pnas.1409822111} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Woollings%
+  \ \protect \BOthers {.}}{%
+  Woollings%
+  \ \protect \BOthers {.}}{%
+  {\protect \APACyear {2014}}%
+  }]{%
+  Woollings:2014kd}
+  \APACinsertmetastar {%
+  Woollings:2014kd}%
+  \begin{APACrefauthors}%
+  Woollings, T.%
+  , Franzke, C.%
+  , Hodson, D\BPBI L\BPBI R.%
+  , Dong, B.%
+  , Barnes, E\BPBI A.%
+  , Raible, C\BPBI C.%
+  \BCBL {}\ \BBA {} Pinto, J\BPBI G.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2014}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Contrasting {{Interannual}} and {{Multidecadal NAO
+    Variability}}} {Contrasting {{Interannual}} and {{Multidecadal NAO
+    Variability}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Climate Dynamics}{45}{1-2}{539-556}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1007/s00382-014-2237-y} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Zebiak%
+  \ \BBA {} Cane%
+  }{%
+  Zebiak%
+  \ \BBA {} Cane%
+  }{%
+  {\protect \APACyear {1987}}%
+  }]{%
+  Zebiak:1987cl}
+  \APACinsertmetastar {%
+  Zebiak:1987cl}%
+  \begin{APACrefauthors}%
+  Zebiak, S\BPBI E.%
+  \BCBT {}\ \BBA {} Cane, M\BPBI A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1987}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {A {{Model El Niño}}-{{Southern Oscillation}}} {A
+    {{Model El Niño}}-{{Southern Oscillation}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Monthly Weather Review}{115}{10}{2262-2278}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1175/1520-0493(1987)115<2262:AMENO>2.0.CO;2}
+    \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Ångström%
+  }{%
+  Ångström%
+  }{%
+  {\protect \APACyear {1935}}%
+  }]{%
+  Angstrom:1935ej}
+  \APACinsertmetastar {%
+  Angstrom:1935ej}%
+  \begin{APACrefauthors}%
+  Ångström, A.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{1935}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {Teleconnections of {{Climatic Changes}} in {{Present
+    Time}}} {Teleconnections of {{Climatic Changes}} in {{Present Time}}}.{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Geografiska Annaler}{17}{3-4}{242-258}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1080/20014422.1935.11880600} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \bibitem [\protect \citeauthoryear {%
+  Živković%
+  \ \BBA {} Rypdal%
+  }{%
+  Živković%
+  \ \BBA {} Rypdal%
+  }{%
+  {\protect \APACyear {2013}}%
+  }]{%
+  zivkovic:2013}
+  \APACinsertmetastar {%
+  zivkovic:2013}%
+  \begin{APACrefauthors}%
+  Živković, T.%
+  \BCBT {}\ \BBA {} Rypdal, K.%
+  \end{APACrefauthors}%
+  \unskip\
+  \newblock
+  \APACrefYearMonthDay{2013}{}{}.
+  \newblock
+  {\BBOQ}\APACrefatitle {{{ENSO}} Dynamics: {{Low}}-Dimensional-Chaotic or
+    Stochastic?} {{{ENSO}} dynamics: {{Low}}-dimensional-chaotic or
+    stochastic?}{\BBCQ}
+  \newblock
+  \APACjournalVolNumPages{Journal of Geophysical Research:
+    Atmospheres}{118}{5}{2161-2168}.
+  \newblock
+  \begin{APACrefDOI} \doi{10.1002/jgrd.50190} \end{APACrefDOI}
+  \PrintBackRefs{\CurrentBib}
+
+  \end{thebibliography}
 

### 2019-04-08 report-earths-future/si_template_2019.tex
add line to caption noting colorbar nonlinearity
Final version with manu suggestion

@@
 
+\subsection{Implications of Bias and Variance}
+
+Figure \ref{fig:conceptual-bias-variance} shows a schematic representation of the implications of too-low or too-high bias and variance in climate risk projections.
+
 \subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}
@@
 \begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
+  \caption{
+    Consequences of model bias or incorrect model representation of uncertainty.
+    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive.
+    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
+  }\label{fig:conceptual-bias-variance}
+\end{figure}
+
+\begin{figure}
       \includegraphics[width=\textwidth]{enso_wavelet.png}
@@
         (L): wavelet power spectrum.
+        Note that the color bar uses a quantile scale and is thus nonlinear.
         (R): global (average) power spectrum.

### 2019-04-08 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
tweak key point

@@
 \begin{keypoints}
-  \item \add{Quasi-periodic and secular climate signals with different identifiability and predictability control uncertainty and risk into the future}
-  \item \add{Adaptation strategies need to consider how the bias and variance in these risk predictions influences choices and decisions pathways}
-  \item \add{Stylized experiments reveal how bias and variance of climate risk estimates over a finite planning periodinfluence risk mitigation}
+  \item \add{Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk}
+  \item \add{Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways}
+  \item \add{Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period}
 \end{keypoints}

### 2019-04-10 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Add line to describe data location.

@@
   siunitx,            % for SI notation
+  url,                % show a url
   xspace,             % for ie and eg
@@
 \graphicspath{{./figs/}{../fig/}} % folders of figures
+
 \allowdisplaybreaks{} % let the align environment span multiple pages
@@
 JDG thanks the NSF GRFP program (grant DGE 16-44869: ``Understanding \& Predicting Climate Drivers of Extreme, Mid-latitude River Floods'') and SERDP program (grant 2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate'') for support.
+All codes and data used to generate this paper are available in a live repository at \url{https://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk/} and a permanent archive at \url{https://doi.org/10.5281/zenodo.2634249}.
 

### 2019-04-11 slides-IPSWD/presentation.tex
slides for IPSWD

--- 2018-12-13 slides-agu/presentation.tex
@@
 
-\title{H52F-05H: Robust Adaptation to\\Multi-Scale Climate Variability}
-\subtitle{Toward Better Water Planning and Management in an Uncertain World I}
-\date{14 December 2018}
+\title{Evaluating staged investments in critical infrastructure for climate adaptation}
+\subtitle{Interdisciplinary Ph.D. Workshop in Sustainable Development 2019}
+\date{13 April 2019}
 \author{\alert{James Doss-Gollin}$^1$, David J. Farnham$^2$, Scott Steinschneider$^3$, Upmanu Lall$^1$}
@@
 \usepackage{
+  amsmath,amssymb,                    % for more math commands
   array,                              % for custom table widths
@@
 \newacronym{gcm}{GCM}{general circulation model}
-\newacronym{hmm}{HMM}{hidden Markov model}
+\newacronym{hmm}{HMM}{Hidden Markov Model}
 \newacronym{iid}{IID}{independent and identically distributed}
@@
 
-\begin{frame}{Motivating Example}
+\begin{frame}{The really big picture}
+  \begin{figure}
+    \centering
+    \pause
+    \includegraphics[height=0.325\textheight]{powerless.png}~
+    \includegraphics[height=0.325\textheight]{ho-2017-1.png}\\
+    \pause
+    \includegraphics[height=0.325\textheight]{jongman-fig-2.jpg}~
+    \includegraphics[height=0.325\textheight]{wapo-usa-temp.png}
+  \end{figure}
+  \nocite{economist-nigeria:2016,Ho:2017gy,Jongman:2012cr}
+\end{frame}
+
+\begin{frame}{Motivating case study}
   What to do after Sandy? \citep{CityofNewYork:2013uh}
@@
 
+\begin{frame}{Today}
+  My perspective: sequential planning under uncertainty
+  \pause
+  \begin{enumerate}
+    \item How does organized variability in the climate inform this planning?
+    \pause
+    \item Observations $\Rightarrow$ notice three interesting things about the world
+    \pause
+    \item Stylized computational experiments $\Rightarrow$ understand implications in an idealized system
+  \end{enumerate}
+  \pause
+  Paper submitted to Earth's Future; all codes are available at \url{http://github.com/jdossgollin/2018-robust-adaptation-cyclical-risk}
+\end{frame}
+
 \section{Hypotheses}
@@
   \pause
-  Project should be evaluated on climate conditions over this finite planning period:
+  Project success depends on climate conditions over this finite planning period:
   \begin{itemize}
-    \item For ``mega-project'', $M \geq \SI{50}{years}$
-    \item For small, flexible project, $M \leq \SI{5}{years}$
+    \item For dam, storm barrier: $M \geq \SI{50}{years}$
+    \item For cat bond, zoning change: $M \leq \SI{5}{years}$
   \end{itemize}
+  \pause
+  Small $M$: defer large investment and allow some uncertainties to be resolved
 \end{frame}
@@
  \end{alertblock}
-  Let $P^* \def \mathbb{P} \qty(X > X^*)$.
-  Note that the insurance premium (or risk factor) is:
-  \begin{equation*}
-    R = \mathbb{E}\qty[ P^*] + \lambda \mathbb{V} \qty[P^*]
-  \end{equation*}
+  Let $P^* = \mathbb{P} \qty(X > X^*)$.
   \pause
-  \begin{columns}
-    \begin{column}{0.5\textwidth}
+  We consider
+  \begin{itemize}
+    \item 3 \alert{idealized scenarios of climate change}
+    \item 3 \alert{simple} models for projecting risk
+  \end{itemize}
+  Measure bias and variance of $P^*$.
+  \vfill
+  \pause
+  \emph{Don't use these models for actual estimation!}
+\end{frame}
+
+\begin{frame}{How it works}
   \begin{figure}
-        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-Stationary-Trend.pdf}
+    \centering
+    \includegraphics[width=\textwidth,height=0.7\textheight]{Example-NINO3-M100-N50.pdf}
   \end{figure}
-    \end{column}
-    \begin{column}{0.5\textwidth}
-      Systematic, stylized experiments:
-      what happens as we vary $M$, $N$, climate structure, estimating model?
-    \end{column}
-  \end{columns}
 \end{frame}
@@
   \begin{columns}[T]
-    \begin{column}{0.5\textwidth}
-      Assertions:
+    \begin{column}{0.45\textwidth}
       \begin{itemize}
@@
     \end{column}
-    \pause
-    \begin{column}{0.5\textwidth}
-      Implications:
-      \begin{itemize}
-        \pause
-        \item \emph{Ability} to identify and predict different climate signals depends on information available (\eg, $N$)
-        \pause
-        \item \emph{Importance} of predicting different climate signals depends on extrapolation desired (\ie, planning period)
-        \pause
-        \item In general, low risk tolerance and/or limited information favor investments with short planning periods.
-      \end{itemize}
+    \begin{column}{0.55\textwidth}
+      \begin{figure}
+        \includegraphics[width=\textwidth]{M-N-Sketch.pdf}
+      \end{figure}
     \end{column}
@@
 \end{frame}
+
+\begin{frame}{Conclusions}
+  \begin{itemize}
+    \item Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk
+    \pause
+    \item Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways
+    \pause
+    \item Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period
+  \end{itemize}
+\end{frame}
+
+\section{Next steps}
 
@@
 \section{Generating Synthetic Streamflow Sequences}
-
-\begin{frame}{Example Sequences and Fits}
-  \begin{figure}
-    \includegraphics[width=\textwidth]{Example-NINO3-M100-N50.pdf}
-    \caption{Example of sequences generated with $M=100$ and $N=50$}
-  \end{figure}
-\end{frame}
 

### 2019-04-11 slides-IPSWD/presentation.tex
Use biblatex

@@
   booktabs,                           % for better (alternative?) tables
-  natbib,                             % references!
 	physics,                            % for better notation
@@
 \setbeamertemplate{bibliography item}[text] % don't print the symbols
+
+\usepackage[
+  backend=biber,
+  doi=true,
+  url=false,
+  isbn=false,
+  style=authoryear-comp, natbib=true,
+  backref=false,
+  maxbibnames=3,
+  maxcitenames=2,
+  uniquename=false,
+  uniquelist=false
+]{biblatex}
+\renewbibmacro{in:}{}
+\AtEveryBibitem{\clearfield{month}\clearfield{pages}}
+\addbibresource{library.bib}
 
@@
   \end{figure}
-  \nocite{economist-nigeria:2016,Ho:2017gy,Jongman:2012cr}
+  \citep{economist-nigeria:2016,Ho:2017gy,Jongman:2012cr}
 \end{frame}
@@
 
+\renewcommand*{\bibfont}{\small}
 \begin{frame}[allowframebreaks]{References}
-  \renewcommand*{\bibfont}{\scriptsize}
-  \renewcommand{\bibsection}{}
-  \nocite{DossGollin:TjTkb07T}
-	\bibliographystyle{agu}
-  \bibliography{library}
+  \printbibliography[heading=none]
 \end{frame}
@@
 \begin{frame}{Hidden Markov Model}
-  Two-state \gls{hmm} \citep[see][]{Rabiner:1986jk} implemented using pomegranate python package \citep{Schreiber:2017tg}.
+  Two-state \gls{hmm} \citep[see][]{Rabiner:1986jk} implemented using pomegranate python package \citep{schreiber:2017}.
   See package documentation for reference.

### 2019-05-17 report-earths-future/reviewer-response.md
update reviewer response
Final version

--- 2019-03-27 report-earths-future/reviewer-response.md
@@
 
-We thank
+We again thank Dr. Mari, Dr. Montanari, and one anonymous reviewer for their helpful comments on our paper "Robust Adaptation to Multi-Scale Climate Variability".
+Please see below our responses to the minor comments from reviewer #2.
 
-## Reviewer 1
+I am glad to see you added a discussion about robustness.
+To reflect the focus of the article, which is on sources of uncertainty, I would reword title as something like, "Understanding Multi-Scale Climate Variability for Robust Adaption."
 
-### Overall comments
+> Thank you for your helpful comments.
+The two titles are very similar, but we feel that the current wording is the most understandable and direct.
+In particular, we think that the title as-is focuses not about understanding multiscale climate variability in the context of robust adaptation, but how to think abut adaptation given the multiscale variability that exists.
 
-I enjoyed reading this paper, which confirms that uncertainty dominates when predicting climate change over an extended lead time.
+Line 62-63: I appreciate that you reworded this sentence. Minor suggestion: "In this paper we focus more narrowly on the temporal ..." (to avoid using the word question twice in this sentence).
 
-The analysis is carried out through a series of synthetic experiments.
-It is interesting to see what are the limits of predictability in the presence of natural variability and secular change.
+> We thank the reviewer for this suggestion and agree that this wording is clearer. The updated wording is on L63.
 
-I think this paper is timely in that it confirms that particular care should be taken when making long term predictions in the presence of a changing climate.
-As the analysis clearly shows, it is not easy to recognize climate signals.
-Natural variability may interact with climate change therefore masking inherent behaviours.
+The introduction of the Hurst phenomenon is an interesting addition to your paper.
+I suggest being more upfront with a definition, e.g.: The Hurst phenomenon is a well-known mathematical relationship between the magnitude of natural events and their standard deviation of occurrence, ...
 
-I think the paper is well written and organized. I made some comments in the attached pdf file.
-My comments are essentially minor.
+> Thank you for this suggestion.
+We have re-worded this sentence to include a bit more information about the Hurst phenomenon.
+It how reads "The Hurst phenomenon is a well known mathematical relationship which describes the long memory of processes found in in geophysics, physics, biology, medicine, traffic, network dynamics, and finance," (L121).
 
-I would like to point out that I missed in the paper a discussion on the Hurst effect.
-There is an extended literature on it. Such effect implies the presence of low frequency variability (long term cycles) which were observed in several secular time series of environmental variables.
-The Hurst effect may be difficult to distinguish from anthropogenic climate change (ACC).
-This effect is ignored by several climate studies. I believe this paper would be a good opportunity to discuss the interaction of the Hurst effect with ACC.
-To this end, a proper discussion should be included in the manuscript.
+Lines 285-291: I think these sentences would be better presented as an equation.
 
-I would like to congratulate the authors for their interesting work.
+> Thank you for this suggestion.
+We agree that describing these operations in a paragraph is rather convoluted.
+However, describing these operations using equations would actually require several equations: to define the estimated threshold, the bias, and the variance.
+We feel that adding these equations would interrupt the flow of this paragraph and add needless complexity to a paper which already asks the reader to keep track of a lot of ideas at once.
+The openly accessible codes make it easy for anyone who wishes to reproduce or modify this work, and so we would like to leave these sentences as they are.
 
-Best regards,
-Alberto Montanari
+I'm confused by figure 3.
+The variables N and M could be redefined in the caption.
+Isn't M the investment lifespan, i.e.
+occurring in the future, and N is the historical sample period; if so, why does the third column show estimated and observed exceedances for N and M, respectively?
 
-### Line comments
+> Thanks for pointing this out.
+We have updated figure 3 and its caption to improve clarity.
+Specifically, we have re-worded the box which read "Estimated Exceedances per Year" to "Projected Future Exceedances per Year" to emphasize that the estimating model and N-year historical record are used to estimate future risk.
 
-7: It is not clear to me what this key point means
+Line 417: Reword sentence to reflect the more limited illustration in fig. 2.
 
-> We have updated this key point following suggestions from reviewer 2
-
-10: I believe these key points are not much informative. I am not sure of my opinion, but my feeling it that they should reformulated to better explain what the paper discovered.
-
-> Working on it!
-
-34: What do you mean exactly? I don't think mitigation efforts proved to be successful.
-
-> We (unfortunately!) agree that mitigation efforts have not yet proven successful.
-We have re-phrased this sentence.
-It now reads "Even if future mitigation efforts are successful, existing levels of atmospheric CO2 and ocean heat content mean that novel adaptation strategies are needed."
-
-84: I believe the impact of ACC on floods and drought is far from being well documented.
-It is not yet clear whether droughts like the Millenium Drought and the Californian drought have been caused by ACC.
-California is subjected to drought since the XIX century.
-Please see  https://en.wikipedia.org/wiki/Droughts_in_California and references therein.
-
-> We agree with the reviewer that the influence of ACC on hydrological extremes is unclear and complex.
-We have re-worded this sentence to be more clear about this point.
-It now reads "Of these processes, ACC has received the most attention in the climate adaptation literature and its influence on some river floods, droughts, hurricanes, urban flooding, and many other climate hazards has been the subject of substantial investigation..."
-
-87: I fully agree with this sentence.
-
-> Thank you very much
-
-103: LFV is strictly related to the Hurst effect. There is an extended literature. I think a mention to the Hurst effect would be appropriate here.
-
-> We have added a short paragraph here which seeks to join some disparate literatures on paleoclimate, Hurst effects, and ocean-atmosphere dynamics.
-These all explain similar phenomena but use different perspectives.
-
-115: I am not sure it is appropriate to make mention to a hypothetically perfect climate model.
-It should be made clear that this is a VERY ideal assumption.
-Perfect environmental models do not exist and, as hydrologists, we should be clear on that.
-
-> We strongly agree with this assessment.
-In a presentation of this paper at the 2018 AGU Fall Meeting we referred to opening a bottle on a beach and using one of three wishes granted by a magical genie on a climate model to emphasize the ridiculousness of having a perfect model.
-In this slightly more formal setting we have re-worded the sentence to "Even in the idealized, and unrealistic, case of a perfect climate model, these uncertainties will be large."
-
-118: This is why I believe we should not make reference to perfect models.
-
-> See previous comment
-
-231: I believe these details on the software should not be given here, in the body of the paper. These may be part of the user guide of the produced software
-
-> We are deferring here to the editor.
-However, I feel that as a consumer of open source software I have an ethical responsibility to cite the primary packages that I use.
-Further, as a scientist committed to reproducibility I believe that describing software used is as important as describing the experiment design.
-However, making code accessible on GitHub partially alleviates this concern.
-
-## Reviewer 2
-
-### Overall comments
-
-The authors have teased apart climatic risks that are considered for planning purposes across multiple time horizons.
-It is an interesting exploration the interaction of various types of climatic variability, planning horizons, and length of time series, and how this interaction impacts the choice of instruments of climate risk mitigation.
-As this paper uses the word robust in the title, the introduction needs to link the concept of robustness to their questions and methods.
-Suggest citing literature on robust adaptation if you keep ‘robust’ in the title.
-
-> See Line 39 (verify)
-
-In Sections 2.1 and 2.2 you describe the three models of climatic change, and three ways of modeling each climate model.
-Is there a way to clarify the structure of your computational experiment in these two sections, including your intent of using multiple methods of generating these types of climate variability?
-
-> **QUESTION**: I think we just need to modify the beginning of the results section. Is there anything we can make clearer?
-
-Please explicitly link the lognormal model with the LN2 acronym in text before using in figures.
-
-> We have done this in section 2.2
-
-It would be interesting to see whether there is a relationship in the results between bias and variance; e.g.
-can the variance inform or predict the bias in each of the three types of climate variability?
-
-> This is certainly an interesting question.
-However, the precise structure of bias and variance as a function of $N$ and $M$ depends strongly on the specific model used to create nonstationarity and to estimate future climate risk.
-As such, we cannot expect that any answer we obtain to this question would generalize beyond the specific assumptions of our stylized computational experiment.
-For this reason we have determined not to address this question in this work.
-
-The discussion needs to reference the computational results, at minimum to justify the computational experiment as opposed to solely a conceptual exploration.
-
-> We have added some modifications to the discussion to reference our computational results more directly and clearly.
-
-### Key points
-
-To me, planning implies a process rather than the expected lifespan of a project.
-Can you clarify which meaning you use?
-If the latter, perhaps the following would be more readable: Climate risk varies over the planned lifespan of a project.
-This can also be clarified in other locations of the paper, e.g. line 364 in the summary.
-
-### Line comments
-
-15-16: Can you restate phrase copied below, unclear terms in bold:
-**the role of low-frequency variability in modulating climate extremes**
-
-> We agree and have re-phrased this sentence. It now reads "... historical and paleo climate records emphasize interannual to multidecadal modes of variability"
-
-23: The discount rate is given as a decision factor but this is not modeled in the paper. I suggest relegating mention of a discount rate to the introduction or discussion.
-
-> We have removed the distracting mention of a discount rate
-
-24: Suggest because instead of where.
-
-> Our point here is that if variability can be identified and predicted -- which is not a given -- shorter design lives are preferred.
-The sentence now reads "Shorter design lives are preferred for situations where inter-annual to decadal variability can be successfully identified and predicted..."
-
-38-39: Is this formatting of reference correct?
-
-> Per the document "Brief Guide to AGU Style and Grammar" document available at https://publications.agu.org/brief-guide-agu-style-grammar/, "AGU’s latex template and APA style in reference management software do not include this exception.
-You do not need to update your manuscript with this exception; it will be applied by the production vendor during the copy-editing process."
-
-41: perhaps a good place to insert a definition and example of operational and financial instruments?
-An example is given for a structural instrument and that is what people are most likely to understand already.
-
-> This is a helpful suggestion.
-The sentence now reads "Rather, actors such as New York city have turned to a combination of structural (e.g., index insurance), operational (e.g., improved evacuation routes), and financial (eg, index insurance) investments for reducing vulnerability and increasing resilience to climate extremes." We note that all of these examples are contained in the City of New York report referenced.
-We believe that these examples are sufficient since a strict definition of different classes of instruments would substantially disrupt the structure of the paragraph.
-
-50-51: Phrasing could be reworked; at minimum change this to these at the end of the sentence.
-
-> The sentence now reads "In this paper we focus on the more narrow question of how the temporal structure of climate risk, and the uncertainties associated with its estimation, influence the answers to these questions."
-
-53: comma after implications
-
-> Added
-
-67: perhaps a good place to define length of long-term short. Is this the interannual to multidecadal described in the abstract or the one-year term of a financial contract (line 59)?
-
-> This is an important question, as we are certainly a bit vague here.
-The points are clearest in the limits of prediction at time scales less than one year or greater than a century; most real-world centuries are in between.
-To add some clarity, we have added another sentence with an example.
-This sentence reads "For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved." We hope that our wording does not imply that we have performed any analysis for New York City.
-
-90-91: Phrasing is a bit obtuse.
-
-> This sentence now reads "Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk."
-
-122: portend (singular)
-
-> Fixed
-
-Figure 1: Inconsistent capitalization. Can you add units?
-
-> Working on this, but going to follow up RE units of American River
-
-147: Recommend preserving the order you present the climate scenarios throughout the paper.
-
-> This is reasonable. We order them (i) secular change only; (ii) LFV only; (iii) LFV plus secular change. This also mirrors the results.
-
-185: Is the word ‘consider’ necessary in these paragraphs?
-
-> We have changed two sentences.
-The first now reads "In the first case we fit a stationary model to the observed flood record, following...".
-The second now reads "Finally, we explicitly model LFV using a hidden Markov model (HMM)."
-
-198: be consistent with use of acronyms (i.e. just say LFV).
-
-> Sentence changed; see previous comment
-
-205: suggest “log-normal distribution that is conditional only on the unobserved state
-variables” (or S(t) instead of state).
-
-> Thank you for the helpful suggestion
-
-215 and 217: Can you say ‘that’ instead of ‘designed based on’? If these words are meaningful, can you elaborate?
-
-> We agree that the wording was vague. The sentence now reads "An instrument whose design was based on projections with overestimated variance or positive bias will be..."
-
-219: suggest “M, N, and
-
-> This is helpful
+> We agree that based on figure 2 we should restrict our focus to climate uncertainties, rather than initiating a discussion of socio-economic and other uncertainties.
+We have re-worded this sentence to read  "As illustrated schematically in fig 2, large uncertainties (e.g., as to future CO2 concentrations and local climate impacts) lead to large intrinsic uncertainty in projections of future climate risk." (L419).

### 2019-05-17 report-earths-future/figs/flowchart.tex
Final version

@@
 
-  \node [calculated, right=of historical] (pThat) {Estimated Exceedances per Year};
+  \node [calculated, right=of historical] (pThat) {Projected Future Exceedances per Year};
   \node [calculated, right= of future] (pT) {Observed Exceedances per Year};
@@
 
-  \node [result, right=of pThat] (bias) {bias of mean};
-  \node [result, right=of pT] (variance) {bias of variance};
+  \node [result, right=of pThat] (bias) {Bias of Mean};
+  \node [result, right=of pT] (variance) {Variance};
 

### 2019-05-17 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Update paper-2018-robust-adaptation-cyclical-risk.tex
Final version

@@
 \sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
+%\graphicspath{{./}} % folders of figures
+
 \graphicspath{{./figs/}{../fig/}} % folders of figures
@@
 \begin{keypoints}
-  \item \add{Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk}
-  \item \add{Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways}
-  \item \add{Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period}
+  \item Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk
+  \item Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways
+  \item Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period
 \end{keypoints}
@@
   The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
-  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records \add{emphasize} \remove{highlight the role of low-frequency variability (LFV) in modulating climate extremes at} interannual to multidecadal \remove{time scales} \add{modes of variability}, \acrlong{acc} is expected to alter their occurrence at longer time scales.
+  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records emphasizemodes of variability, \acrlong{acc} is expected to alter their occurrence at longer time scales.
   In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
@@
   The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
-  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$\remove{ and the future discount rate for investment}.
-  Shorter design lives are preferred \add{for situations} where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
+  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$.
+  Shorter design lives are preferred for situations where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
 \end{abstract}
@@
 At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
-Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of} novel adaptation strategies\remove{ are needed}.
+Even if future mitigation efforts are successful, existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of novel adaptation strategies.
 
 This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
-\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
-Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently \remove{slowed their implementation}\add{dampened enthusiasm}.
-\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, a catastrophe bond)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
+A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures \citep{lempert:2007,Borgomeo:2018hk} and which fail along non-catastrophic modes \citep{brown:2010}.
+Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently dampened enthusiasm.
+Rather, actors such as New York City have turned to a combination of structural (e.g.\@\xspace, stormwater barrier), operational (e.g.\@\xspace, improved evacuation routes), and financial (e.g.\@\xspace, a catastrophe bond) instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
 These instruments are not typically implemented in isolation or statically.
@@
 Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
-In this paper we focus on \add{the narrower question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
-We continue this section with three specific observations about climate risk which, while \remove{apparently}\add{seemingly} obvious, have important and subtle implications \add{that}\remove{which} we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
+In this paper we focus \add{more narrowly on the temporal}\remove{ on the narrower question of how the temporal} structure of climate risk, and the uncertainties associated with its estimation, influence the answers to these questions.
+We continue this section with three specific observations about climate risk which, while seemingly obvious, have important and subtle implications that we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
 
@@
 This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
-\add{For example, the optimal policy for New York City to manage uncertain hurricane risk in the 21st century might potentially be to keep areas devastated by hurricane Sandy zoned for low-impact development for the next 10 years.
-This would reduce future risk over all climate scenarios while postponing major investments until large uncertainties as to the magnitude of future sea level rise are resolved.}
+For example, the optimal policy for New York City to manage uncertain hurricane risk in the 21st century might potentially be to keep areas devastated by hurricane Sandy zoned for low-impact development for the next 10 years.
+This would reduce future risk over all climate scenarios while postponing major investments until large uncertainties as to the magnitude of future sea level rise are resolved.
 The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.
@@
 Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
-The Hurst phenomenon is well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
+The Hurst phenomenon is \add{a} well known \add{mathematical relationship which describes the} \remove{and identified with} long memory \add{of} processes \add{found in} in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
 The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
@@
 Peaks for the American River time series are apparent at 2.3 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
-This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
+This is illustrated by the blue line in \cref{fig:observed-lfv}(b), which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
 A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
@@
     (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
-    \remove{High numbers}\add{Lower values} indicate more severe drought.
+    Lower values indicate more severe drought.
     A 20-year running mean is also shown in blue.
@@
 As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
-Even \remove{with} \add{in the idealized and unrealistic case of} a perfect climate model, these uncertainties will be large.
+Even in the idealized and unrealistic case of a perfect climate model, these uncertainties will be large.
 By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.
@@
 Since different physical processes control climate risk at different timescales, successful integration of climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
-A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portend\remove{s} for risk mitigation instruments with a planning period ranging from a few years to several decades.
+A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portend for risk mitigation instruments with a planning period ranging from a few years to several decades.
 As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
@@
 
-We consider three scenarios for climate risk, which we define by the structure of the underlying climate \annote{signal}{changed the order}:
+We consider three scenarios for climate risk, which we define by the structure of the underlying climate signal:
 \begin{enumerate*}[label= (\roman*)]
@@
 \end{equation}
-This formulation describes \add{all} scenarios \add{for future climate considered in this paper within} \remove{in} a single equation\remove{;}\add{.}
-\add{To add climate variability to the system,} the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
+This formulation describes all scenarios for future climate considered in this paper within a single equation.
+To add climate variability to the system, the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
 \begin{equation}\label{eq:nino3}
@@
 Supplemental figure S1 shows a wavelet spectrum and time series plot of the resulting annual time series.
-In the online supplemental information we consider an alternative parameterization of $\mu(t)$, \remove{i.e.\@\xspace an alternative to,} which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
+In the online supplemental information we consider an alternative parameterization of $\mu(t)$, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.
 
@@
 For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
-Once a statistical model is fit to these \remove{``}observations\remove{''}, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
-
-\add{In the first case we} \remove{We first consider fitting}\add{fit} a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
+Once a statistical model is fit to these observations, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.
+
+In the first case we fit a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
 In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
-\add{We refer to this model as ``LN2 Stationary.''}
+We refer to this model as ``LN2 Stationary.''
 The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
@@
 We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in supplemental equation (S2).
-\add{We refer to this model as ``LN2 Linear Trend''}
+We refer to this model as ``LN2 Linear Trend.''
 This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.
 
-\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
+Finally, we explicitly model \gls{lfv} using a \gls{hmm}.
 An \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
@@
 We fit  streamflow sequences using a \gls{hmm} with two states.
-The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
+The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution that is conditional only on the unobserved state variables.
 This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
@@
 Both estimation bias and estimation uncertainty affect the utility of a climate risk projection.
-An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
-Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the \remove{user}\add{public}.
+An instrument whose design was based on projections with overestimated variance or positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
+Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the public.
 
 We evaluate both the estimation bias and estimation uncertainty.
-For a given choice of $M$, $N$, \add{and} generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'' and the $K=1000$ posterior simulations of future flows.
+For a given choice of $M$, $N$, and generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'' and the $K=1000$ posterior simulations of future flows.
 The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
@@
 \Cref{fig:flowchart} describes the experimental design.
-We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
+We assess estimation bias and variance for three scenarios of future climate.
 First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped (``secular change only'').
@@
     Parameters are shown in red.
+    \add{$N$ denotes the informational uncertainty (length of historical record) and $M$ the amount of extrapolation (project design life).}
     Calculated quantities are shown in white.
@@
 
-These three \remove{cases}\add{scenarios for future climate considered} are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
-\add{We also show projected future climate risk} with each of the three \remove{fitting}\add{estimating} models \add{described in} \cref{sec:methods-estimating}\remove{are also shown}.
+These three scenarios for future climate considered are illustrated in \cref{fig:example-fit}, which shows a single synthetic streamflow sequence generated with $N=50$ and $M=100$.
+We also show projected future climate risk with each of the three estimating models described in \cref{sec:methods-estimating}.
 This figure highlights that even where projections of average streamflow are unbiased, if the spread is too large then projection of the threshold exceedance probability may be too large.
@@
 
-\Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} \add{show that} for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
+\Cref{fig:lfv-secular-nino3-bias-variance,fig:lfv-secular-nino3-bias-variance,fig:secular-nino3-bias-variance} show that for projects where $M$ is sufficiently short, intrinsic uncertainty is low and cyclical climate variability is dominant over the project planning period \citep{Jain:2001hz,Hodgkins:2017hw}.
 However, one's ability to identify and predict this variability depends on having a model of sufficient complexity to represent the processes that cause \gls{lfv}, and the data to fit the model.
@@
 
-For projects with longer $M$, \add{our results highlight the importance of identifying and predicting} \remove{it becomes necessary to consider the role of} secular change.
-As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
+For projects with longer $M$, our results highlight the importance of identifying and predicting secular change.
+As illustrated schematically in \cref{fig:conceptual-sketch}, \add{large} uncertainties \add{(e.g.\@\xspace, as to} future \remove{population, economic development, renewable energy technology, and} \add{$\text{CO}_2$ concentrations and local climate impacts)} \remove{collectively} lead to \remove{very high}\add{large} intrinsic uncertainty in projections of future climate risk.
 As the physical mechanisms cascade from global (e.g.\@\xspace, global mean surface temperature) to regional  \citep[e.g.\@\xspace, storm track position;][]{Barnes:2015gl} and local (e.g.\@\xspace, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
 With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
-Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation \add{(i.e.\xspace} a ``danger zone''\add{ for planning)}.
-
-These findings were derived \add{conceptually and through} \remove{using simple models} \add{idealized computational experiments} for \add{simulating and predicting climate} risk\remove{ projection}, but the principles are applicable to more complex, physically based methods.
+Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation (i.e.\xspace a ``danger zone'' for planning).
+
+These findings were derived conceptually and through idealized computational experiments for simulating and predicting climate risk, but the principles are applicable to more complex, physically based methods.
 For example, flood frequency analysis may join observations across time and space \citep{Lima:2016kd,Merz:2008eh} or apply model chains based on general circulation models and hydrologic models \citep[see][]{Merz:2014gf} to increase $N$.
@@
 Projects with long planning periods must therefore overcome future discounting, the potential for large bias or variance, and that all estimates are made with informational uncertainty $N$.
-\add{By contrast, the informational uncertainties for a sequence of short-term instruments are $N, N+M, N+2M, \ldots$, potentially yielding improved identifiability and predictability of relevant climate signals.}\remove{rather than a sequence of $N$, $N+\Delta N, \ldots$}.
+By contrast, the informational uncertainties for a sequence of short-term instruments are $N, N+M, N+2M, \ldots$, potentially yielding improved identifiability and predictability of relevant climate signals.
 
@@
 \begin{enumerate*}[label= (\roman*)]
-  \item that \remove{planning decisions are made with finite horizons}\add{different climate risk mitigation instruments have different planned lifespans};
+  \item that different climate risk mitigation instruments have different planned lifespans;
   \item that climate risk varies on many scales; and
@@
 
-The authors thank Dr. Alberto Montanari and one anonymous reviewer for insightful comments on this manuscript.
+The authors thank Celine Mari, Alberto Montanari, and one anonymous reviewer for comments which greatly improved this manuscript.
 The authors thank Nandini Ramesh of Columbia University for providing the synthetic NINO3 index from a \SI{100000}{year} run of the Cane-Zebiak model as described in \citet{Ramesh:2016hf}.

### 2019-05-18 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Final tweaks

@@
 Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
-In this paper we focus \add{more narrowly on the temporal}\remove{ on the narrower question of how the temporal} structure of climate risk, and the uncertainties associated with its estimation, influence the answers to these questions.
+In this paper we focus \add{more narrowly on the temporal}\remove{ on the narrower question of how the temporal} structure of climate risk and \add{how} the uncertainty associated with its estimation\remove{,} influence\add{s} the answers to these questions.
 We continue this section with three specific observations about climate risk which, while seemingly obvious, have important and subtle implications that we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.
@@
 
-Public or private sector investments in climate adaptation require not only the design of each potential structural instrument, but also selecting between instruments with vastly different operational planning periods.
+Public or private sector investments in climate adaptation require not only the design of each potential\remove{structural} instrument, but also selecting between instruments with vastly different operational planning periods.
 This project planning period, which we define as being $M$ years, describes the nominal economic or physical lifespan of the structure or contract.

### 2019-08-25 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
use autoenv and fix word issue

@@
   The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
-  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records emphasizemodes of variability, \acrlong{acc} is expected to alter their occurrence at longer time scales.
+  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records emphasize low-frequency modes of variability, \acrlong{acc} is expected to alter their occurrence at longer time scales.
   In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
